### 2025-03-27 paper.qmd
First commit

---
title: Tail Flexibility in the Degrees of Preferential
 Attachment Networks
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-basic
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    number-sections: true
author:
  - name: Thomas William Boughen
    affiliations:
      - name: Newcastle University
        id: 1
        department: School of Mathematics, Statistics and Physics
  - name: Clement Lee
    affiliations:
      - ref: 1
  - name: Vianey Palacios Ramirez
    affiliations:
      - ref: 1
abstract: |
  Devising the underlying generating mechanism of a real-life network is difficult as, more often than not, only its snapshots are available, but not its full evolution. One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law. Consequently, the growth of real-life networks that roughly display such power-law behaviour is commonly modelled by preferential attachment. However, the validity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the right tail of the degree distribution is often lighter than implied by the body, whilst still being heavy. In this paper, we study a modified version of the model with a flexible preference function that allows super/sub-linear behaviour whilst also guaranteeing that the limiting degree distribution has a heavy tail. We relate the distributions tail heaviness directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
keywords: [networks, extremes]
bibliography: refs.bib  
---
```{r,echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(paletteer)
library(gridExtra)
library(latex2exp)
library(ggpubr)
```

# Introduction {#sec-intro}

The degree distribution of a network can be very informative about the networks structure, giving some understanding about the inequality in the influence of the nodes. Often the degree distribution is one of the simplest places to look when trying to understand this, as the full evolution of the network is usually not available. This makes modelling the degree distribution an important topic in research with contributions such as [ref].

When looking at the degree distributions of real networks, a power law is an attractive option to use when modelling as it is seemingly observed in a lot of networks. Fitting the discrete power law to the degree distribution come with the added benefit of suggesting preferential attachment as the generating mechanism since it has been shown to generate networks with a power law degree distribution.

If the power law is an adequate model to use for real degree distributions has become a heated debate, with alternatives such as [@Broido_2019] seeming to outperform the power law in some cases. Additionally, [ref] find that for many real degree distributions the body may follow a power law, the right tail does not or at least seems to be lighter than is implied by the body whilst still being heavy. One method for modelling the tail of the degree distributions of real networks is to use a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:

$$
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
$$
for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.

The IGPD has been shown to be in the same domain of attraction and has the same tail-index $\xi$ as the regular GPD with the same parameters by using the following quantity introduced by [@shimura12]:

$$
\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1},
$$
the limit of this quantity if it exists and is non-zero is exactly the tail-index of the distribution $F$, and if the limit is zero then the distribution is in the Gumbel maximum domain of attraction. These results will be useful later on when trying to establish the tail behaviour of a network's limiting degree distribution.

These models that seem to outperform the power law do not however come with the added benefit of suggesting a generating mechanism for the networks in the same way the power law does. Efforts to make the preferential attachment model more flexible have not yet remedied this issue with [@krapivsky01] finding that using a sub/super-linear preference function yields a Weibull tail and a degenerate network, where one vertex eventually gains all new edges, respectively.

[@rudas07] introduces theoretical results for a PA model with a general preference function in the cases that the network being generated is a tree; presenting an opportunity to design a preference function that will capture the behaviours observed in the degree distributions of real networks. This paper studies a flexible class of preference functions that allow for a power law body and a tail that is heavy and flexible, and proposes a strategy for fitting this model to real data. With good fits, this method does not only allow the tail heaviness to be quantified, but it also allows for some inference on the preference function assuming a general preferential attachment model is the underlying generating mechanism.

# General Preferential Attachment {#sec-gpa}

The focus of this section is the limiting degree distribution of a preferential attachment network with preference function $b(\cdot)$ where each vertex brings with it $m$ edges. This preference function is subject to the following conditions:

$$
b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
$$

$$
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
$$ {#eq-condb2}

Given these two conditions an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07]. Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:

$$
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
$$
where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.

The limiting survival can be viewed as the limit of the empirical proportion of vertices with degree over a threshold $k$, that is:

$$
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
$$
which by the previously stated theorem can also be written as:

$$
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$
Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:

$$
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
$$

which aligns with the result from [@rudas07], where the expression is derived using the same branching process results.

Using results from [@shimura12] the domain of attraction of this limiting degree distribution can be found.

:::{#prp-omega}
If $b(k) \rightarrow \infty$ as $k\rightarrow \infty$ then
$$
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$
and

$$
\lim_{k\rightarrow\infty}\Omega(F,n) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
$$
:::

@prp-omega implies that the only way to obtain a heavy tailed degree distribution using a non-decreasing preference function is to have that preference function be asymptotically linear. The next subsection focuses on a particular class of these preference functions that provide flexible behaviour but guarantee a  heavy tail.

## Preferential Attachment with flexible heavy tail {#sec-model}

Previously, in [@krapivsky01], a preference function of the form $k^\alpha +\varepsilon$ has been considered. However, it has been noted that super-linear preferential attachment ($\alpha > 1$) leads to a degenerate limiting degree distribution where at some point one vertex gains the connection from every vertex that joins the network, additionally sub-linear preferential attachment ($\alpha <1$) has been shown to lead to a light tailed limiting degree distribution. Consider a preference function of the form:

$$
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.

Using a preference function with guaranteed linear behaviour in the limit, allows for the inclusion of sub/super linear behaviour without losing the heavy tails or ending up with a degenerate degree distribution.

The limiting degree distribution resulting from using a preference function of this form can be found to have survival:

$$
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
\end{cases}
$${#eq-polysurv}

with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where:

$$
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
$$

which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$ since when $\lambda\le \beta$ the infinite sum no longer converges and instead goes to infinity.

Some examples of what the degree distribution looks like are shown below in @fig-polylinsurv:

```{r, echo=FALSE}
source('funcs.R')
```

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical survival distributions of the limiting degree distributions, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed preferential attachment model.
#| label: fig-polylinsurv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = 0:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
}
pars = cbind(pars,surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

ggplot(data = pars) + geom_line(aes(x=(x+1),y=surv, linetype=as.character(b),colour =as.character(b))) +
  scale_x_log10() + scale_y_log10(limits=c(1e-5,1))+theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) + 
  facet_grid(eps~a,labeller = labeller, scales='free')
```

The survival (@eq-polysurv) can be connected to the IGPD mentioned in [@sec-intro] by using Sterling's approximation of the gamma function to obtain:

<!-- $$ -->
<!-- \frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y}, -->
<!-- $$ -->
<!-- and utilising the expression from @eq-polysurv: -->

<!-- \begin{align*} -->
<!-- \bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\ -->
<!-- &\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\ -->
<!-- &=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\ -->
<!-- &=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta} -->
<!-- \end{align*} -->

$$
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
$$ meaning that for $k\ge k_0$ the limiting degree distribution is similar to $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.

To assess how close of an approximation this is the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey. The approximation seems to hold up fairly well even for large degrees.

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical conditional survivals (grey) alongside their IGPD approximations (coloured).
#| label: fig-approx_surv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = k0s:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
gp_surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])/
    S(pars$k0[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
  gp_surv[i] = mev::pgp(pars$x[i],pars$k0[i]-1,((pars$k0[i]+1)^pars$a[i] +pars$eps[i]-1 )/pars$lambda[i],pars$b[i]/pars$lambda[i],lower.tail = F)
}
pars = cbind(pars,surv,gp_surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv','gp_surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

library(ggh4x)

l1 = scale_y_log10(limits=c(1e-1,1))
l2 = scale_y_log10(limits=c(1e-2,1))
l3 = scale_y_log10(limits=c(1e-3,1))
l4 = scale_y_log10(limits=c(1e-4,1))

limlist = list(l1,l2,l3,l4)

ggplot(data = pars) + geom_line(aes(x=(x),y=surv, linetype=as.character(b),colour =as.character(b)), lwd=1) +
  geom_line(data = pars[pars$x>=pars$k0,],aes(x=(x+1),y=gp_surv, linetype=as.character(b)),colour ='grey')+
  scale_x_log10()+
  theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) +
  facet_grid(eps~a,labeller = labeller,scales='free') + facetted_pos_scales(y=limlist)

```

Since $\beta>0$, the shape parameter of the IGPD is positive and thus the distribution is heavy tailed. Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices:

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 12
#| fig-height: 6
#| fig-cap: Heat maps of $\xi$ for various combinations of the parameters of the proposed model.
#| label: fig-polyheat
#| 
rho_optim_ba= Vectorize(function(a,eps, b, k0){
  return(abs(rho(2*b, polylin(a, eps), b, k0)-1))
},vectorize.args = 'a')
find_a_ba = Vectorize(function(b, eps, k0){
  out = optimise(rho_optim_ba, c(0.00001,3), b=b, eps=eps, k0=k0)$minimum
  return(out)
}, vectorize.args = 'b')
N =50
as = seq(0,2,l=N+1)[-1]
bs = seq(0,2,l=N+1)[-1]
eps = c(.01,.1,.5,1)
k0 = c(25,100)
pars = expand.grid(as,bs,eps,k0)
names(pars) = c('a', 'b', 'eps', 'k0')
lambdas = numeric(nrow(pars))
a_for_ba = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
  a_for_ba[i] = find_a_ba(pars$b[i], pars$eps[i], pars$k0[i])
}
pars = cbind(pars,lambdas,a_for_ba)
names(pars) = c('a', 'b', 'eps', 'k0','lambda','ba')
labeller = label_bquote(cols = `epsilon`==.(eps),rows = ~k[0]==.(k0))

ggplot(pars) + geom_raster(aes(x=a,y=b,fill=b/lambda)) +
  geom_line(aes(x=ba, y=b),linetype='dashed', colour='red',lwd=1)+
  scale_fill_paletteer_c(palette='grDevices::Blues',limits=c(0,1),direction=-1)+theme(aspect.ratio = 1)+ylim(min(bs), max(bs))+xlim(min(as), max(as))+
  facet_grid(k0~eps,labeller = labeller) + labs(fill=TeX('\\xi')) + xlab(TeX('\\alpha')) + ylab(TeX('\\beta')) + theme_minimal()

```

The darker regions on the heat maps correspond to a heavier tail and the lighter to a lighter tail, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail heaviness as the Barabási-Albert model, $\xi=0.5$.

### Piecewise Linear Model

When $\alpha = 1$ the preference function simplifies to a piecewise linear function with $\beta$ representing the ratio of the gradients of the two components:

$$
b(k) = \min(k,k_0) + \varepsilon + \mathbb I\{k\ge k_0\}\beta(k-k_0)
$$ for some $\beta, \varepsilon, k_0>0$.

Following @eq-polysurv the limiting degree distribution is:

$$
\bar F(k) = \begin{cases}
\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k+1+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k+1+\lambda^*+\epsilon)},&k<k_0\\
\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times \frac{\Gamma\left(\frac{k_0+\varepsilon+\lambda^*}{\beta}\right)}{\Gamma\left(\frac{k_0+\varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon}{\beta}+1\right)}{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon+\lambda^*}{\beta}+1\right)},&k\ge k_0
\end{cases}
$$ {#eq-linsurv}

Just like the previous model, the survival can be approximated by:

$$
\bar F(k) \approx \begin{cases}
\frac{\Gamma(\lambda^*+\varepsilon)}{\Gamma(\varepsilon)} (1+\varepsilon)^{\lambda^*}\left(\frac{k}{1+\varepsilon} + 1\right)^{-\lambda^*},&k<k_0\\
\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times\left(\frac{\beta(k+1-k_0)}{k_0+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0
\end{cases}
$$

Some examples of the possible degree distributions are shown in the central row of @fig-polylinsurv.

# Recovery

The goal of was to provide a possible preference function that is able to explain the growth of real networks with varied tail behaviour in their degree distributions. This section aims to show that the parameters of the model in @sec-model can be recovered from the degree distribution of a network simulated from it.

The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts in the form of a vector of degrees $\pmb x = (1,2,\ldots,M)$ and the number of vertices with those degrees $\pmb n$ where $M$ is the maximum degree.

Using these and the p.m.f derived in [@rudas07], the likelihood given that $x_i \ge l$ for all $i =1,2,\ldots,M$ is:

\begin{align*}
L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{x_i\ge k_0}n_{x_i}\right)} \prod_{l \le x_i<k_0}\left(\frac{\lambda^*}{\lambda^* +x_i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{x_i\ge k_0}\left(\frac{\text{B}(x_i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}

where $B(y,z)$ is the the beta function.

This likelihood allows inference to be made about the parameters using a Bayesian approach using the priors:

\begin{align*}
\alpha&\sim \text{Ga}(1,0.01),\\
\beta &\sim  \text{Ga}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Ga(1,0.01)},
\end{align*}

to obtain a posterior distribution that can then be used in an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm  to obtain posterior samples. The results of this inference are shown in @fig-rec1 and @fig-rec2. For these simulated networks $l=2$.

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
#| label: fig-rec1
surfit_list = list()
pars = readRDS('recovery_pars.rds')
recovery_list = readRDS('recovery_dat.rds')
selected = c(1,27,14,35)
# selected = 1:36
for(k in 1:length(selected)){
  j=selected[k]
  x = sort(unique(recovery_list[[j]]$degs))
  ls = c()

  y_975 = recovery_list[[j]]$mcmc$surv$CI[2,]
  y_025 = recovery_list[[j]]$mcmc$surv$CI[1,]
  y_50 = recovery_list[[j]]$mcmc$surv$est

  surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
    geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
    scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
    ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
                       ', $\\epsilon = $',pars$eps[j],
                       ', $\\beta = $',pars$b[j])))
}

fig = ggpubr::ggarrange(plotlist = surfit_list,
          nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
          label.x='Degree', label.y = 'Survival')

ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')

```

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
#| label: fig-rec2

library(gridExtra)
plt = readRDS('mcmc_plot.rds')
plot(plt)

```

# Application to Real Data

Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources. [describe networks being used].

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 9
#| fig-height: 6
#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
#| label: fig-real1

plt1 = readRDS('real_survs.rds')

plot(plt1)

```

@fig-real1 displays the posterior estimates of the conditional survival $F(k|k\ge2)$ for several data sets, for many of the data sets the fit is quite good  whilst for other the fit is very bad. This may indicate that the data sets without a good fit are unlikely to have come from a preferential attachment model. However for those with a good fit, it is possible to obtain information about the underlying PA mechanism assuming that the corresponding networks did indeed grow in that fashion.

# Conclusion and Discussion

# Additional Results {.appendix}

## Tail heaviness of general preferential attachment model

Recall the limiting survival function:

$$
\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
$$

The aim is to determine how this distribution behaves for different choices of $b$, specifically what maximum domain of attraction does this distribution belong to and is it affected by the choice of preference function $b$. @shimura12 introduces a quantity that will help in determining what domain of attraction a discrete distribution belongs to, that is:

For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:

$$
\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
$$

[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$. Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.

Consider a preference function $b(\cdot)$ that fulfills:

$$
\lim_{n\rightarrow \infty} b(n) = \infty
$${#eq-condb1}

Substituting in the form of $\bar F(n)$ from @eq-surv and taking the limit, subject to @eq-condb1:

$$
\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
$$ {#eq-omegares} (see appendix for full proof).

Whilst there are many classes of functions that satisfy @eq-condb1 and @eq-condb2, the vast majority of these result in $\Omega(F,n) \rightarrow 0$, meaning that the limiting degree distribution is light tailed. In fact, in order for the limiting degree distribution to be heavy tailed, $b$ must be asymptotically linear ($b(k)\sim k$).

# Proofs {.appendix}

## Tail heaviness of GPA

Taking the form of the GPA degree survival: $$
\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
$$ and substituting into the formula for $\Omega(F,n)$:

```{=tex}
\begin{align*}
\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
\end{align*}
```
Clearly if $b(n)=c$ or $\lim_{n\rightarrow\infty}b(n)=c$ for some $c>0$ then $\Omega(F,n)=0$. Now consider a non-constant $b(n)$ and re-write $\Omega(F,n)$ as:

```{=tex}
\begin{align*}
\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
\end{align*}
```
Then if $\lim_{n\rightarrow\infty}b(n)=\infty$ it follows that:

```{=tex}
\begin{align*}
\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
\end{align*}
```
\newpage

## Removing the infinite sum

For a preference function of the form:

$$
b(k) = \begin{cases}
g(k),&k<k_0\\
g(k_0) + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\beta>0, k_0\in\mathbb N$ we have that

```{=tex}
\begin{align*}
\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
\end{align*}
```
Now using the fact that:

$$
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
$$ and reindexing the product in the second sum

```{=tex}
\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
\end{align*}
```
In order to simplify the infinite sum, consider:

```{=tex}
\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
\end{align*}
```
this infinite sum does not converge when $x\le1$ as each term is $O(n^{-x})$. We can now use this in $\hat\rho(\lambda)$:

```{=tex}
\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
\end{align*}
```

# References
<!-- the pattern below controls the placement of the references -->
:::{#refs}
:::

### 2025-03-27 paper.tex
First commit

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-basic,
]{sn-jnl}

\usepackage{amsmath,amssymb}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math}
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else  
    % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{xcolor}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\setcounter{secnumdepth}{5}
% Make \paragraph and \subparagraph free-standing
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
% Set default figure placement to htbp
\makeatletter
\def\fps@figure{htbp}
\makeatother

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\ifLuaTeX
  \usepackage{selnolig}  % disable illegal ligatures
\fi
\usepackage{bookmark}

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={Tail Flexibility in the Degrees of Preferential Attachment Networks},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
  pdfkeywords={networks, extremes},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Tail Flexibility in the Degrees of Preferential Attachment
Networks]{Tail Flexibility in the Degrees of Preferential Attachment
Networks}

% author setup
\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
% affil setup
\affil[1]{\orgdiv{School of Mathematics, Statistics and
Physics}, \orgname{Newcastle University}}

% abstract

\abstract{Devising the underlying generating mechanism of a real-life
network is difficult as, more often than not, only its snapshots are
available, but not its full evolution. One candidate for the generating
mechanism is preferential attachment which, in its simplest form,
results in a degree distribution that follows the power law.
Consequently, the growth of real-life networks that roughly display such
power-law behaviour is commonly modelled by preferential attachment.
However, the validity of the power law has been challenged by the
presence of alternatives with comparable performance, as well as the
recent findings that the right tail of the degree distribution is often
lighter than implied by the body, whilst still being heavy. In this
paper, we study a modified version of the model with a flexible
preference function that allows super/sub-linear behaviour whilst also
guaranteeing that the limiting degree distribution has a heavy tail. We
relate the distributions tail heaviness directly to the model
parameters, allowing direct inference of the parameters from the degree
distribution alone.}

% keywords
\keywords{networks,  extremes}

\begin{document}
\maketitle

\section{Introduction}\label{sec-intro}

The degree distribution of a network can be very informative about the
networks structure, giving some understanding about the inequality in
the influence of the nodes. Often the degree distribution is one of the
simplest places to look when trying to understand this, as the full
evolution of the network is usually not available. This makes modelling
the degree distribution an important topic in research with
contributions such as {[}ref{]}.

When looking at the degree distributions of real networks, a power law
is an attractive option to use when modelling as it is seemingly
observed in a lot of networks. Fitting the discrete power law to the
degree distribution come with the added benefit of suggesting
preferential attachment as the generating mechanism since it has been
shown to generate networks with a power law degree distribution.

If the power law is an adequate model to use for real degree
distributions has become a heated debate, with alternatives such as
\citep{Broido_2019} seeming to outperform the power law in some cases.
Additionally, {[}ref{]} find that for many real degree distributions the
body may follow a power law, the right tail does not or at least seems
to be lighter than is implied by the body whilst still being heavy. One
method for modelling the tail of the degree distributions of real
networks is to use a discretised variation of the generalised Pareto
distribution (GPD) usually dubbed the Integer GPD (GPD) where
\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:

\[
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).

The IGPD has been shown to be in the same domain of attraction and has
the same tail-index \(\xi\) as the regular GPD with the same parameters
by using the following quantity introduced by \citep{shimura12}:

\[
\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1},
\] the limit of this quantity if it exists and is non-zero is exactly
the tail-index of the distribution \(F\), and if the limit is zero then
the distribution is in the Gumbel maximum domain of attraction. These
results will be useful later on when trying to establish the tail
behaviour of a network's limiting degree distribution.

These models that seem to outperform the power law do not however come
with the added benefit of suggesting a generating mechanism for the
networks in the same way the power law does. Efforts to make the
preferential attachment model more flexible have not yet remedied this
issue with \citep{krapivsky01} finding that using a sub/super-linear
preference function yields a Weibull tail and a degenerate network,
where one vertex eventually gains all new edges, respectively.

\citep{rudas07} introduces theoretical results for a PA model with a
general preference function in the cases that the network being
generated is a tree; presenting an opportunity to design a preference
function that will capture the behaviours observed in the degree
distributions of real networks. This paper studies a flexible class of
preference functions that allow for a power law body and a tail that is
heavy and flexible, and proposes a strategy for fitting this model to
real data. With good fits, this method does not only allow the tail
heaviness to be quantified, but it also allows for some inference on the
preference function assuming a general preferential attachment model is
the underlying generating mechanism.

\section{General Preferential Attachment}\label{sec-gpa}

The focus of this section is the limiting degree distribution of a
preferential attachment network with preference function \(b(\cdot)\)
where each vertex brings with it \(m\) edges. This preference function
is subject to the following conditions:

\[
b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
\]

\begin{equation}\phantomsection\label{eq-condb2}{
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
}\end{equation}

Given these two conditions an expression for the survival of the
limiting degree distribution can be found in the case that \(m=1\);
obtained by considering a branching process that is equivalent to the
growth of the network, as in \citep{rudas07}. Theorem 1 from
\citep{rudas07} states that for the tree \(\Upsilon(t)\) at time \(t\):

\[
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).

The limiting survival can be viewed as the limit of the empirical
proportion of vertices with degree over a threshold \(k\), that is:

\[
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
\] which by the previously stated theorem can also be written as:

\[
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
the probability mass function (p.m.f.) can be shown to be:

\[
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
\]

which aligns with the result from \citep{rudas07}, where the expression
is derived using the same branching process results.

Using results from \citep{shimura12} the domain of attraction of this
limiting degree distribution can be found.

\begin{proposition}[]\protect\hypertarget{prp-omega}{}\label{prp-omega}

If \(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\) then \[
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
\] and

\[
\lim_{k\rightarrow\infty}\Omega(F,n) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
\]

\end{proposition}

Proposition~\ref{prp-omega} implies that the only way to obtain a heavy
tailed degree distribution using a non-decreasing preference function is
to have that preference function be asymptotically linear. The next
subsection focuses on a particular class of these preference functions
that provide flexible behaviour but guarantee a heavy tail.

\subsection{Preferential Attachment with flexible heavy
tail}\label{sec-model}

Previously, in \citep{krapivsky01}, a preference function of the form
\(k^\alpha +\varepsilon\) has been considered. However, it has been
noted that super-linear preferential attachment (\(\alpha > 1\)) leads
to a degenerate limiting degree distribution where at some point one
vertex gains the connection from every vertex that joins the network,
additionally sub-linear preferential attachment (\(\alpha <1\)) has been
shown to lead to a light tailed limiting degree distribution. Consider a
preference function of the form:

\[
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
\] for \(\alpha,\beta, \varepsilon>0\) and \(k_0\in\mathbb N\).

Using a preference function with guaranteed linear behaviour in the
limit, allows for the inclusion of sub/super linear behaviour without
losing the heavy tails or ending up with a degenerate degree
distribution.

The limiting degree distribution resulting from using a preference
function of this form can be found to have survival:

\begin{equation}\phantomsection\label{eq-polysurv}{
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
\end{cases}
}\end{equation}

with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where:

\[
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
\]

which must be solved numerically for most parameter choices. Also, note
that \(\lambda>\beta\) since when \(\lambda\le \beta\) the infinite sum
no longer converges and instead goes to infinity.

Some examples of what the degree distribution looks like are shown below
in Figure~\ref{fig-polylinsurv}:

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}

}

\caption{\label{fig-polylinsurv}Theoretical survival distributions of
the limiting degree distributions, according to various combinations of
\((\alpha, \beta, \varepsilon)\) and \(k_0=20\) of the proposed
preferential attachment model.}

\end{figure}%

The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
mentioned in Section~\ref{sec-intro} by using Sterling's approximation
of the gamma function to obtain:

\[
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
\] meaning that for \(k\ge k_0\) the limiting degree distribution is
similar to
\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).

To assess how close of an approximation this is the theoretical
conditional survivals are shown in Figure~\ref{fig-approx_surv} in
colour and their IGPD approximations are shown in grey. The
approximation seems to hold up fairly well even for large degrees.

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-approx_surv-1.pdf}

}

\caption{\label{fig-approx_surv}Theoretical conditional survivals (grey)
alongside their IGPD approximations (coloured).}

\end{figure}%

Since \(\beta>0\), the shape parameter of the IGPD is positive and thus
the distribution is heavy tailed. Additionally the value of the shape
parameter \(\xi\) is shown in Figure~\ref{fig-polyheat} for various
parameter choices:

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polyheat-1.pdf}

}

\caption{\label{fig-polyheat}Heat maps of \(\xi\) for various
combinations of the parameters of the proposed model.}

\end{figure}%

The darker regions on the heat maps correspond to a heavier tail and the
lighter to a lighter tail, the red dashed line shows combinations of
\(\alpha\) and \(\beta\) that produce a limiting degree distribution
with the same tail heaviness as the Barabási-Albert model, \(\xi=0.5\).

\subsubsection{Piecewise Linear Model}\label{piecewise-linear-model}

When \(\alpha = 1\) the preference function simplifies to a piecewise
linear function with \(\beta\) representing the ratio of the gradients
of the two components:

\[
b(k) = \min(k,k_0) + \varepsilon + \mathbb I\{k\ge k_0\}\beta(k-k_0)
\] for some \(\beta, \varepsilon, k_0>0\).

Following Equation~\ref{eq-polysurv} the limiting degree distribution
is:

\begin{equation}\phantomsection\label{eq-linsurv}{
\bar F(k) = \begin{cases}
\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k+1+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k+1+\lambda^*+\epsilon)},&k<k_0\\
\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times \frac{\Gamma\left(\frac{k_0+\varepsilon+\lambda^*}{\beta}\right)}{\Gamma\left(\frac{k_0+\varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon}{\beta}+1\right)}{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon+\lambda^*}{\beta}+1\right)},&k\ge k_0
\end{cases}
}\end{equation}

Just like the previous model, the survival can be approximated by:

\[
\bar F(k) \approx \begin{cases}
\frac{\Gamma(\lambda^*+\varepsilon)}{\Gamma(\varepsilon)} (1+\varepsilon)^{\lambda^*}\left(\frac{k}{1+\varepsilon} + 1\right)^{-\lambda^*},&k<k_0\\
\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times\left(\frac{\beta(k+1-k_0)}{k_0+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0
\end{cases}
\]

Some examples of the possible degree distributions are shown in the
central row of Figure~\ref{fig-polylinsurv}.

\section{Recovery}\label{recovery}

The goal of was to provide a possible preference function that is able
to explain the growth of real networks with varied tail behaviour in
their degree distributions. This section aims to show that the
parameters of the model in Section~\ref{sec-model} can be recovered from
the degree distribution of a network simulated from it.

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
counts in the form of a vector of degrees \(\pmb x = (1,2,\ldots,M)\)
and the number of vertices with those degrees \(\pmb n\) where \(M\) is
the maximum degree.

Using these and the p.m.f derived in \citep{rudas07}, the likelihood
given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:

\begin{align*}
L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{x_i\ge k_0}n_{x_i}\right)} \prod_{l \le x_i<k_0}\left(\frac{\lambda^*}{\lambda^* +x_i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{x_i\ge k_0}\left(\frac{\text{B}(x_i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}

where \(B(y,z)\) is the the beta function.

This likelihood allows inference to be made about the parameters using a
Bayesian approach using the priors:

\begin{align*}
\alpha&\sim \text{Ga}(1,0.01),\\
\beta &\sim  \text{Ga}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Ga(1,0.01)},
\end{align*}

to obtain a posterior distribution that can then be used in an adaptive
Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm to obtain
posterior samples. The results of this inference are shown in
Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
networks \(l=2\).

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec1-1.pdf}

}

\caption{\label{fig-rec1}Posterior estimates of survival function for
data simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

\end{figure}%

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec2-1.pdf}

}

\caption{\label{fig-rec2}Posterior estimates of paramters for data
simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

\end{figure}%

\section{Application to Real Data}\label{application-to-real-data}

Turning now to real data, the goal is to fit the model to the degree
distributions of real networks from various sources. {[}describe
networks being used{]}.

\begin{figure}

\centering{

\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-real1-1.pdf}

}

\caption{\label{fig-real1}Posterior estimates (solid red) of survival
for several real data sets and their 95\% credible intervals (dotted
red)}

\end{figure}%

Figure~\ref{fig-real1} displays the posterior estimates of the
conditional survival \(F(k|k\ge2)\) for several data sets, for many of
the data sets the fit is quite good whilst for other the fit is very
bad. This may indicate that the data sets without a good fit are
unlikely to have come from a preferential attachment model. However for
those with a good fit, it is possible to obtain information about the
underlying PA mechanism assuming that the corresponding networks did
indeed grow in that fashion.

\section{Conclusion and Discussion}\label{conclusion-and-discussion}

\section{Additional Results}\label{additional-results}

\subsection{Tail heaviness of general preferential attachment
model}\label{tail-heaviness-of-general-preferential-attachment-model}

Recall the limiting survival function:

\[
\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
\]

The aim is to determine how this distribution behaves for different
choices of \(b\), specifically what maximum domain of attraction does
this distribution belong to and is it affected by the choice of
preference function \(b\). \citet{shimura12} introduces a quantity that
will help in determining what domain of attraction a discrete
distribution belongs to, that is:

For a distribution \(F\) with survival function \(\bar F\) and some
\(n\in\mathbb Z^+\) let:

\[
\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
\]

\citep{shimura12} then states that if
\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
distribution is light tailed.

Consider a preference function \(b(\cdot)\) that fulfills:

\begin{equation}\phantomsection\label{eq-condb1}{
\lim_{n\rightarrow \infty} b(n) = \infty
}\end{equation}

Substituting in the form of \(\bar F(n)\) from \textbf{?@eq-surv} and
taking the limit, subject to Equation~\ref{eq-condb1}:

\begin{equation}\phantomsection\label{eq-omegares}{
\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
}\end{equation} (see appendix for full proof).

Whilst there are many classes of functions that satisfy
Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}, the vast majority
of these result in \(\Omega(F,n) \rightarrow 0\), meaning that the
limiting degree distribution is light tailed. In fact, in order for the
limiting degree distribution to be heavy tailed, \(b\) must be
asymptotically linear (\(b(k)\sim k\)).

\section{Proofs}\label{proofs}

\subsection{Tail heaviness of GPA}\label{tail-heaviness-of-gpa}

Taking the form of the GPA degree survival: \[
\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
\] and substituting into the formula for \(\Omega(F,n)\):

\begin{align*}
\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
\end{align*}

Clearly if \(b(n)=c\) or \(\lim_{n\rightarrow\infty}b(n)=c\) for some
\(c>0\) then \(\Omega(F,n)=0\). Now consider a non-constant \(b(n)\) and
re-write \(\Omega(F,n)\) as:

\begin{align*}
\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
\end{align*}

Then if \(\lim_{n\rightarrow\infty}b(n)=\infty\) it follows that:

\begin{align*}
\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
\end{align*}

\newpage

\subsection{Removing the infinite sum}\label{removing-the-infinite-sum}

For a preference function of the form:

\[
b(k) = \begin{cases}
g(k),&k<k_0\\
g(k_0) + \beta(k-k_0), &k\ge k_0
\end{cases}
\] for \(\beta>0, k_0\in\mathbb N\) we have that

\begin{align*}
\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
\end{align*}

Now using the fact that:

\[
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
\] and reindexing the product in the second sum

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
\end{align*}

In order to simplify the infinite sum, consider:

\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
\end{align*}

this infinite sum does not converge when \(x\le1\) as each term is
\(O(n^{-x})\). We can now use this in \(\hat\rho(\lambda)\):

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
\end{align*}

\section{References}\label{references}

\renewcommand{\bibsection}{}
\bibliography{refs.bib}

\end{document}

### 2025-03-27 template.tex
First commit

--- 2025-03-27 paper.tex
@@
 \documentclass[
-  sn-basic,
+  sn-nature,
 ]{sn-jnl}
@@
 \setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
+\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
 % Make \paragraph and \subparagraph free-standing
@@
 \newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
-\usepackage{amsthm}
-\theoremstyle{plain}
-\newtheorem{proposition}{Proposition}[section]
-\theoremstyle{remark}
-\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
-\newtheorem{refremark}{Remark}[section]
-\newtheorem{refsolution}{Solution}[section]
 \makeatother
@@
 \hypersetup{
-  pdftitle={Tail Flexibility in the Degrees of Preferential Attachment Networks},
-  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
-  pdfkeywords={networks, extremes},
+  pdftitle={Quarto Template for Springer Nature},
+  pdfauthor={Author One; Author Two; Author Three},
+  pdfkeywords={template, demo},
   colorlinks=true,
@@
 
-\title[Tail Flexibility in the Degrees of Preferential Attachment
-Networks]{Tail Flexibility in the Degrees of Preferential Attachment
-Networks}
+\title[Quarto Template for Springer Nature]{Quarto Template for Springer
+Nature}
 
 % author setup
-\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
+\author[1,2]{\fnm{Author} \sur{One}}\equalcont{These authors contributed equally to this work.}\author*[3]{\fnm{Author} \sur{Two}}\email{corresponding@email.com}\equalcont{These authors contributed equally to this work.}\author[1]{\fnm{Author} \sur{Three}}
 % affil setup
-\affil[1]{\orgdiv{School of Mathematics, Statistics and
-Physics}, \orgname{Newcastle University}}
+\affil[1]{\orgdiv{Department of Government}, \orgname{Harvard
+University}, \orgaddress{\street{1737 Cambridge
+Street}, \city{Cambridge}, \postcode{2138}}}
+\affil[2]{\orgdiv{Department of Statistics}, \orgname{Harvard
+University}, \orgaddress{\street{1 Oxford
+Street}, \city{Cambridge}, \postcode{2138}}}
+\affil[3]{\orgdiv{Department of Political Science}, \orgname{Yale
+University}, \orgaddress{\street{115 Prospect Street}, \city{New
+Haven}, \postcode{3401}}}
 
@@
 
-\abstract{Devising the underlying generating mechanism of a real-life
-network is difficult as, more often than not, only its snapshots are
-available, but not its full evolution. One candidate for the generating
-mechanism is preferential attachment which, in its simplest form,
-results in a degree distribution that follows the power law.
-Consequently, the growth of real-life networks that roughly display such
-power-law behaviour is commonly modelled by preferential attachment.
-However, the validity of the power law has been challenged by the
-presence of alternatives with comparable performance, as well as the
-recent findings that the right tail of the degree distribution is often
-lighter than implied by the body, whilst still being heavy. In this
-paper, we study a modified version of the model with a flexible
-preference function that allows super/sub-linear behaviour whilst also
-guaranteeing that the limiting degree distribution has a heavy tail. We
-relate the distributions tail heaviness directly to the model
-parameters, allowing direct inference of the parameters from the degree
-distribution alone.}
+\abstract{The abstract serves both as a general introduction to the
+topic and as a brief, non-technical summary of the main results and
+their implications. Authors are advised to check the author instructions
+for the journal they are submitting to for word limits and if structural
+elements like subheadings, citations, or equations are permitted.}
 
 % keywords
-\keywords{networks,  extremes}
+\keywords{template,  demo}
 
@@
 
-The degree distribution of a network can be very informative about the
-networks structure, giving some understanding about the inequality in
-the influence of the nodes. Often the degree distribution is one of the
-simplest places to look when trying to understand this, as the full
-evolution of the network is usually not available. This makes modelling
-the degree distribution an important topic in research with
-contributions such as {[}ref{]}.
+The Introduction section, of referenced text \citep{greenwade93} expands
+on the background of the work (some overlap with the Abstract is
+acceptable). The introduction should not include subheadings.
 
-When looking at the degree distributions of real networks, a power law
-is an attractive option to use when modelling as it is seemingly
-observed in a lot of networks. Fitting the discrete power law to the
-degree distribution come with the added benefit of suggesting
-preferential attachment as the generating mechanism since it has been
-shown to generate networks with a power law degree distribution.
+Springer Nature does not impose a strict layout as standard however
+authors are advised to check the individual requirements for the journal
+they are planning to submit to as there may be journal-level
+preferences. When preparing your text please also be aware that some
+stylistic choices are not supported in full text XML (publication
+version), including coloured font. These will not be replicated in the
+typeset article if it is accepted.
 
-If the power law is an adequate model to use for real degree
-distributions has become a heated debate, with alternatives such as
-\citep{Broido_2019} seeming to outperform the power law in some cases.
-Additionally, {[}ref{]} find that for many real degree distributions the
-body may follow a power law, the right tail does not or at least seems
-to be lighter than is implied by the body whilst still being heavy. One
-method for modelling the tail of the degree distributions of real
-networks is to use a discretised variation of the generalised Pareto
-distribution (GPD) usually dubbed the Integer GPD (GPD) where
-\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:
+\section{Using this template}\label{sec-template}
 
-\[
-\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).
+This Quarto template is unofficial and built out of Springer Nature's
+template. Some examples of commonly used commands and features are
+listed below, to help you get started.
 
-The IGPD has been shown to be in the same domain of attraction and has
-the same tail-index \(\xi\) as the regular GPD with the same parameters
-by using the following quantity introduced by \citep{shimura12}:
+As seen below, you can mix markdown and Latex with each other, though
+it's likely best to mostly use markdown.
 
-\[
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1},
-\] the limit of this quantity if it exists and is non-zero is exactly
-the tail-index of the distribution \(F\), and if the limit is zero then
-the distribution is in the Gumbel maximum domain of attraction. These
-results will be useful later on when trying to establish the tail
-behaviour of a network's limiting degree distribution.
+\subsection{Cross Referencing}\label{cross-referencing}
 
-These models that seem to outperform the power law do not however come
-with the added benefit of suggesting a generating mechanism for the
-networks in the same way the power law does. Efforts to make the
-preferential attachment model more flexible have not yet remedied this
-issue with \citep{krapivsky01} finding that using a sub/super-linear
-preference function yields a Weibull tail and a degenerate network,
-where one vertex eventually gains all new edges, respectively.
+To reference a figure with example label ``plot'', use
+\texttt{@fig-plot}. Analogously, to reference a table with example label
+``data'', use \texttt{@tbl-data}. To reference a section, such as the
+Introduction (Section~\ref{sec-intro}), use \texttt{@sec-intro}.
 
-\citep{rudas07} introduces theoretical results for a PA model with a
-general preference function in the cases that the network being
-generated is a tree; presenting an opportunity to design a preference
-function that will capture the behaviours observed in the degree
-distributions of real networks. This paper studies a flexible class of
-preference functions that allow for a power law body and a tail that is
-heavy and flexible, and proposes a strategy for fitting this model to
-real data. With good fits, this method does not only allow the tail
-heaviness to be quantified, but it also allows for some inference on the
-preference function assuming a general preferential attachment model is
-the underlying generating mechanism.
+For complete information on cross referencing with Quarto, see
+\url{https://quarto.org/docs/authoring/cross-references.html}.
 
-\section{General Preferential Attachment}\label{sec-gpa}
+\subsection{Citations}\label{citations}
 
-The focus of this section is the limiting degree distribution of a
-preferential attachment network with preference function \(b(\cdot)\)
-where each vertex brings with it \(m\) edges. This preference function
-is subject to the following conditions:
+Quarto formats citations and references automatically using the
+bibliography records in your .bib file. For a citation in parentheses
+use \citep{greenwade93}. Multiple citations can be given as
+\citep{greenwade93, knuth1984texbook}. If the tex output is to be
+included in a submission to a preprint server or publisher, the default
+citation method, \texttt{citeproc}, may not always produce
+bibliographies compatible with the provided \texttt{bst} styles. In this
+case, you can set \texttt{cite-method:\ natbib}. See
+\url{https://quarto.org/docs/authoring/footnotes-and-citations.html\#sec-biblatex}.
+In this case, you may either restrict usage to the default bracketed key
+citation formats supported by pandoc, such as \citep{greenwade93} or
+\citep{greenwade93, knuth1984texbook}. Or, use a CSL file, several of
+which are vendored in
+\href{https://github.com/christopherkenny/nature/tree/main/_extensions/nature/csl}{\_extensions/nature/csl}.
+The advantage of the latter is that references will be consistent across
+html, docx, and pdf outputs. See
+\url{https://pandoc.org/MANUAL.html\#citation-syntax} for additional
+details on pandoc citation styles.
 
-\[
-b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
-\]
+\section{Tables and Figures}\label{tables-and-figures}
 
-\begin{equation}\phantomsection\label{eq-condb2}{
-\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
-}\end{equation}
-
-Given these two conditions an expression for the survival of the
-limiting degree distribution can be found in the case that \(m=1\);
-obtained by considering a branching process that is equivalent to the
-growth of the network, as in \citep{rudas07}. Theorem 1 from
-\citep{rudas07} states that for the tree \(\Upsilon(t)\) at time \(t\):
-
-\[
-\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).
-
-The limiting survival can be viewed as the limit of the empirical
-proportion of vertices with degree over a threshold \(k\), that is:
-
-\[
-\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
-\] which by the previously stated theorem can also be written as:
-
-\[
-\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
-the probability mass function (p.m.f.) can be shown to be:
-
-\[
-f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
-\]
-
-which aligns with the result from \citep{rudas07}, where the expression
-is derived using the same branching process results.
-
-Using results from \citep{shimura12} the domain of attraction of this
-limiting degree distribution can be found.
-
-\begin{proposition}[]\protect\hypertarget{prp-omega}{}\label{prp-omega}
-
-If \(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\) then \[
-\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-\] and
-
-\[
-\lim_{k\rightarrow\infty}\Omega(F,n) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
-\]
-
-\end{proposition}
-
-Proposition~\ref{prp-omega} implies that the only way to obtain a heavy
-tailed degree distribution using a non-decreasing preference function is
-to have that preference function be asymptotically linear. The next
-subsection focuses on a particular class of these preference functions
-that provide flexible behaviour but guarantee a heavy tail.
-
-\subsection{Preferential Attachment with flexible heavy
-tail}\label{sec-model}
-
-Previously, in \citep{krapivsky01}, a preference function of the form
-\(k^\alpha +\varepsilon\) has been considered. However, it has been
-noted that super-linear preferential attachment (\(\alpha > 1\)) leads
-to a degenerate limiting degree distribution where at some point one
-vertex gains the connection from every vertex that joins the network,
-additionally sub-linear preferential attachment (\(\alpha <1\)) has been
-shown to lead to a light tailed limiting degree distribution. Consider a
-preference function of the form:
-
-\[
-b(k) = \begin{cases}
-k^\alpha + \varepsilon,&k<k_0\\
-k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
-\end{cases}
-\] for \(\alpha,\beta, \varepsilon>0\) and \(k_0\in\mathbb N\).
-
-Using a preference function with guaranteed linear behaviour in the
-limit, allows for the inclusion of sub/super linear behaviour without
-losing the heavy tails or ending up with a degenerate degree
-distribution.
-
-The limiting degree distribution resulting from using a preference
-function of this form can be found to have survival:
-
-\begin{equation}\phantomsection\label{eq-polysurv}{
-\bar F(k) = \begin{cases}
-\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
-\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
-\end{cases}
-}\end{equation}
-
-with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where:
-
-\[
-\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}
-\]
-
-which must be solved numerically for most parameter choices. Also, note
-that \(\lambda>\beta\) since when \(\lambda\le \beta\) the infinite sum
-no longer converges and instead goes to infinity.
-
-Some examples of what the degree distribution looks like are shown below
-in Figure~\ref{fig-polylinsurv}:
+To include figures, you can use Quarto syntax.
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
+\includegraphics{fig.eps}
 
@@
 
-\caption{\label{fig-polylinsurv}Theoretical survival distributions of
-the limiting degree distributions, according to various combinations of
-\((\alpha, \beta, \varepsilon)\) and \(k_0=20\) of the proposed
-preferential attachment model.}
+\caption{\label{fig-plot}An example figure (an empty plot)}
 
@@
 
-The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
-mentioned in Section~\ref{sec-intro} by using Sterling's approximation
-of the gamma function to obtain:
+For both figures and tables, you can use LaTeX syntax if you need
+additional customizability.
 
-\[
-\bar F(k) 
-\begin{cases}
-=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
-\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
-\end{cases}
-\] meaning that for \(k\ge k_0\) the limiting degree distribution is
-similar to
-\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).
+For example, to use footnotes within a table, you may want to use LaTeX.
 
-To assess how close of an approximation this is the theoretical
-conditional survivals are shown in Figure~\ref{fig-approx_surv} in
-colour and their IGPD approximations are shown in grey. The
-approximation seems to hold up fairly well even for large degrees.
+\begin{table}
 
-\begin{figure}
+\caption{\label{tbl-data}LaTeX caption text}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
+[h]
+%
+\begin{tabular}{@{}llll@{}}
+\toprule
+Column 1 & Column 2  & Column 3 & Column 4\\
+\midrule
+row 1    & data 1   & data 2  & data 3  \\
+row 2    & data 4   & data 5\footnotemark[1]  & data 6  \\
+row 3    & data 7   & data 8  & data 9\footnotemark[2]  \\
+\botrule
+\end{tabular}
+\footnotetext{Source: This is an example of table footnote. This is an example of table footnote.}
+\footnotetext[1]{Example for a first table footnote. This is an example of table footnote.}
+\footnotetext[2]{Example for a second table footnote. This is an example of table footnote.}
 
@@
 
-\caption{\label{fig-approx_surv}Theoretical conditional survivals (grey)
-alongside their IGPD approximations (coloured).}
-
-\end{figure}%
-
-Since \(\beta>0\), the shape parameter of the IGPD is positive and thus
-the distribution is heavy tailed. Additionally the value of the shape
-parameter \(\xi\) is shown in Figure~\ref{fig-polyheat} for various
-parameter choices:
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polyheat-1.pdf}
-
-}
-
-\caption{\label{fig-polyheat}Heat maps of \(\xi\) for various
-combinations of the parameters of the proposed model.}
-
-\end{figure}%
-
-The darker regions on the heat maps correspond to a heavier tail and the
-lighter to a lighter tail, the red dashed line shows combinations of
-\(\alpha\) and \(\beta\) that produce a limiting degree distribution
-with the same tail heaviness as the Barabási-Albert model, \(\xi=0.5\).
-
-\subsubsection{Piecewise Linear Model}\label{piecewise-linear-model}
-
-When \(\alpha = 1\) the preference function simplifies to a piecewise
-linear function with \(\beta\) representing the ratio of the gradients
-of the two components:
-
-\[
-b(k) = \min(k,k_0) + \varepsilon + \mathbb I\{k\ge k_0\}\beta(k-k_0)
-\] for some \(\beta, \varepsilon, k_0>0\).
-
-Following Equation~\ref{eq-polysurv} the limiting degree distribution
-is:
-
-\begin{equation}\phantomsection\label{eq-linsurv}{
-\bar F(k) = \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k+1+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k+1+\lambda^*+\epsilon)},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times \frac{\Gamma\left(\frac{k_0+\varepsilon+\lambda^*}{\beta}\right)}{\Gamma\left(\frac{k_0+\varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon}{\beta}+1\right)}{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon+\lambda^*}{\beta}+1\right)},&k\ge k_0
-\end{cases}
-}\end{equation}
-
-Just like the previous model, the survival can be approximated by:
-
-\[
-\bar F(k) \approx \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)}{\Gamma(\varepsilon)} (1+\varepsilon)^{\lambda^*}\left(\frac{k}{1+\varepsilon} + 1\right)^{-\lambda^*},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times\left(\frac{\beta(k+1-k_0)}{k_0+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0
-\end{cases}
-\]
-
-Some examples of the possible degree distributions are shown in the
-central row of Figure~\ref{fig-polylinsurv}.
-
-\section{Recovery}\label{recovery}
-
-The goal of was to provide a possible preference function that is able
-to explain the growth of real networks with varied tail behaviour in
-their degree distributions. This section aims to show that the
-parameters of the model in Section~\ref{sec-model} can be recovered from
-the degree distribution of a network simulated from it.
-
-The procedure for recovering the parameters begins with simulating a
-network from the model with \(N=100,000\) vertices and \(m=1\) given
-some set of parameters
-\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
-counts in the form of a vector of degrees \(\pmb x = (1,2,\ldots,M)\)
-and the number of vertices with those degrees \(\pmb n\) where \(M\) is
-the maximum degree.
-
-Using these and the p.m.f derived in \citep{rudas07}, the likelihood
-given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:
-
-\begin{align*}
-L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{x_i\ge k_0}n_{x_i}\right)} \prod_{l \le x_i<k_0}\left(\frac{\lambda^*}{\lambda^* +x_i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{x_i\ge k_0}\left(\frac{\text{B}(x_i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
-\end{align*}
-
-where \(B(y,z)\) is the the beta function.
-
-This likelihood allows inference to be made about the parameters using a
-Bayesian approach using the priors:
-
-\begin{align*}
-\alpha&\sim \text{Ga}(1,0.01),\\
-\beta &\sim  \text{Ga}(1,0.01),\\
-k_0 &\sim \text{U}(1,10,000),\\
-\varepsilon &\sim \text{Ga(1,0.01)},
-\end{align*}
-
-to obtain a posterior distribution that can then be used in an adaptive
-Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm to obtain
-posterior samples. The results of this inference are shown in
-Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
-networks \(l=2\).
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec1-1.pdf}
-
-}
-
-\caption{\label{fig-rec1}Posterior estimates of survival function for
-data simulated from the proposed model with various combinations of
-(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
-
-\end{figure}%
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec2-1.pdf}
-
-}
-
-\caption{\label{fig-rec2}Posterior estimates of paramters for data
-simulated from the proposed model with various combinations of
-(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
-
-\end{figure}%
-
-\section{Application to Real Data}\label{application-to-real-data}
-
-Turning now to real data, the goal is to fit the model to the degree
-distributions of real networks from various sources. {[}describe
-networks being used{]}.
-
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-real1-1.pdf}
-
-}
-
-\caption{\label{fig-real1}Posterior estimates (solid red) of survival
-for several real data sets and their 95\% credible intervals (dotted
-red)}
-
-\end{figure}%
-
-Figure~\ref{fig-real1} displays the posterior estimates of the
-conditional survival \(F(k|k\ge2)\) for several data sets, for many of
-the data sets the fit is quite good whilst for other the fit is very
-bad. This may indicate that the data sets without a good fit are
-unlikely to have come from a preferential attachment model. However for
-those with a good fit, it is possible to obtain information about the
-underlying PA mechanism assuming that the corresponding networks did
-indeed grow in that fashion.
-
-\section{Conclusion and Discussion}\label{conclusion-and-discussion}
-
-\section{Additional Results}\label{additional-results}
-
-\subsection{Tail heaviness of general preferential attachment
-model}\label{tail-heaviness-of-general-preferential-attachment-model}
-
-Recall the limiting survival function:
-
-\[
-\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
-\]
-
-The aim is to determine how this distribution behaves for different
-choices of \(b\), specifically what maximum domain of attraction does
-this distribution belong to and is it affected by the choice of
-preference function \(b\). \citet{shimura12} introduces a quantity that
-will help in determining what domain of attraction a discrete
-distribution belongs to, that is:
-
-For a distribution \(F\) with survival function \(\bar F\) and some
-\(n\in\mathbb Z^+\) let:
-
-\[
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
-\]
-
-\citep{shimura12} then states that if
-\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
-Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
-distribution is light tailed.
-
-Consider a preference function \(b(\cdot)\) that fulfills:
-
-\begin{equation}\phantomsection\label{eq-condb1}{
-\lim_{n\rightarrow \infty} b(n) = \infty
-}\end{equation}
-
-Substituting in the form of \(\bar F(n)\) from \textbf{?@eq-surv} and
-taking the limit, subject to Equation~\ref{eq-condb1}:
-
-\begin{equation}\phantomsection\label{eq-omegares}{
-\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
-}\end{equation} (see appendix for full proof).
-
-Whilst there are many classes of functions that satisfy
-Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}, the vast majority
-of these result in \(\Omega(F,n) \rightarrow 0\), meaning that the
-limiting degree distribution is light tailed. In fact, in order for the
-limiting degree distribution to be heavy tailed, \(b\) must be
-asymptotically linear (\(b(k)\sim k\)).
-
-\section{Proofs}\label{proofs}
-
-\subsection{Tail heaviness of GPA}\label{tail-heaviness-of-gpa}
-
-Taking the form of the GPA degree survival: \[
-\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
-\] and substituting into the formula for \(\Omega(F,n)\):
-
-\begin{align*}
-\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
-&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
-&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
-\end{align*}
-
-Clearly if \(b(n)=c\) or \(\lim_{n\rightarrow\infty}b(n)=c\) for some
-\(c>0\) then \(\Omega(F,n)=0\). Now consider a non-constant \(b(n)\) and
-re-write \(\Omega(F,n)\) as:
-
-\begin{align*}
-\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
-&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
-\end{align*}
-
-Then if \(\lim_{n\rightarrow\infty}b(n)=\infty\) it follows that:
-
-\begin{align*}
-\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
-&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
-&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
-&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
-\end{align*}
-
-\newpage
-
-\subsection{Removing the infinite sum}\label{removing-the-infinite-sum}
-
-For a preference function of the form:
-
-\[
-b(k) = \begin{cases}
-g(k),&k<k_0\\
-g(k_0) + \beta(k-k_0), &k\ge k_0
-\end{cases}
-\] for \(\beta>0, k_0\in\mathbb N\) we have that
-
-\begin{align*}
-\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
-\end{align*}
-
-Now using the fact that:
-
-\[
-\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
-\] and reindexing the product in the second sum
-
-\begin{align*}
-\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
-&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
-\end{align*}
-
-In order to simplify the infinite sum, consider:
-
-\begin{align*}
-\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
-&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
-&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
-&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
-&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
-&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
-&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
-\end{align*}
-
-this infinite sum does not converge when \(x\le1\) as each term is
-\(O(n^{-x})\). We can now use this in \(\hat\rho(\lambda)\):
-
-\begin{align*}
-\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
-\end{align*}
+\end{table}%
 
@@
 \renewcommand{\bibsection}{}
-\bibliography{refs.bib}
+\bibliography{bibliography.bib}
+
+\section{Appendix}\label{appendix}
+
+{{< lipsum 5 >}}
 

### 2025-03-27 data_fit_background.R

@@ -1,0 +1,62 @@
+source('funcs.R')
+library(ggplot2)
+data_raw = read.csv('degs.csv')
+nms = unique(data_raw$name)
+selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
+fits = list()
+plots = list()
+for(i in 1:length(selected)){
+  print(i)
+  fits[[i]] = fit_model(data_raw[data_raw$name==nms[selected[i]],2:3],N=5e4,quiet=F,
+                        trunc.at = 5,k0_jump = 1)
+}
+
+
+saveRDS(fits, 'real_mcmc.rds')
+
+
+
+
+fits = readRDS('real_mcmc.rds')
+
+
+plots = list()
+for(i in 1:length(fits)){
+  plots[[i]] = ggplot() + scale_x_log10() + scale_y_log10() +
+    geom_point(data = twbfn::deg_surv(counts_to_degs(fits[[i]]$dat)),aes(x=degree, y=surv))+
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y=!!fits[[i]]$surv$est),colour='red') + 
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[1,]),linetype=2, colour='red') + 
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[2,]),linetype=2,colour='red')+
+    geom_vline(aes(xintercept = !!fits[[i]]$smps$k0), colour='blue',alpha=0.1)+
+    ggtitle(nms[selected[i]])
+}
+
+
+real_survs = ggpubr::ggarrange(plotlist = plots)
+
+real_survs
+
+
+saveRDS(real_survs,'real_survs.rds')
+
+
+
+
+
+
+
+max()
+
+
+
+
+
+
+
+fits[[9]]$surv$CI[2,]
+
+
+
+
+
+

### 2025-03-27 funcs.R

@@ -1,0 +1,306 @@
+f = Vectorize(function(x, g, lambda, k0, b){
+  if(x>=k0){
+    k0s = 0:(k0-1)
+    phi_g = lambda*gamma((lambda+g(k0))/b)*prod(g(k0s)/(lambda+g(k0s)))/
+      (b*gamma(lambda/b + 1)*gamma(g(k0)/b))
+    return(phi_g*beta(g(k0)/b + x-k0, lambda/b + 1))
+  }
+  if(x>0){
+    ks = 0:(x-1)
+    return(lambda/(lambda+g(x))*prod(g(ks)/(lambda+g(ks))))
+  }
+  if(x==0){
+    return(lambda/(lambda+g(x)))
+  }
+}, vectorize.args = 'x')
+S = Vectorize(function(x, g, lambda, k0, b){
+  if(x<0){
+    return(1)
+  }
+  if(x>=k0){
+    k0s = 0:(k0-1)
+    
+    phi_g = lgamma((lambda+g(k0))/b)  + sum(log(g(k0s)/(lambda+g(k0s))))-
+      (lgamma(lambda/b)+lgamma(g(k0)/b))
+    
+    return(exp(phi_g)*beta(g(k0)/b + x-k0+1, lambda/b))
+  }
+  if(x>0){
+    ks = 0:(x)
+    return(prod(g(ks)/(lambda+g(ks))))
+  }
+  if(x==0){
+    return(g(x)/(lambda+g(x)))
+  }
+  
+},vectorize.args = 'x')
+sample_gpa_tree = function(n, g,b,k0,m=1){
+  degs = c(0)
+  prefs = c(g(0))
+  for(i in 1:n){
+    # message(i)
+    selected = sample(1:length(degs), min(m, length(degs)), prob=prefs)
+    degs = c(degs, 0)
+    prefs = c(prefs, g(0))
+    degs[selected] = degs[selected]+1
+    prefs[selected] = ifelse(degs[selected]<k0, g(degs[selected]), g(k0) + b*(degs[selected]-k0))
+  }
+  return(degs)
+}
+rho = Vectorize(function(lambda, g, b, k0){
+  n1 = 1:k0
+  n2 = 0:(k0-1)
+  out1 = sum(exp(cumsum(log(1-lambda/(lambda+g(n1-1))))))
+  out2 = exp(sum(log(1-lambda/(lambda+g(n2)))))
+  return(out1 + g(k0)*out2/(lambda-b))
[…]

### 2025-03-27 gpd_to_pref.R

@@ -1,0 +1,297 @@
+logdiff <- function(l1, l2) { l1 + VGAM::log1mexp(l1-l2); }
+digp_log = function(x, shape, scale, threshold){
+  out=x
+  v=threshold
+  sig=scale
+  xi=shape
+  p1 = pmax(x[x>v]*0, 1 + xi * (x[x>v]+1-v)/(sig))
+  p2 = pmax(x[x>v]*0, 1 + xi * (x[x>v]-v)/(sig))
+  pows2 = pows1 = x[x>v]*0 + 1
+  pows1[p1>0] = -1/xi
+  pows2[p2>0] = -1/xi
+  # out[x>v] =  ( p2^pows2 - p1^pows1)
+  out[x>v] = logdiff(p2^pows2,p1^pows1)
+  out[x<=v] = 0
+  return(out)
+}
+
+igp_ll_neg = function(pars,dat,threshold=0){
+  shape = pars[1]
+  scale = pars[2]
+  if(shape==0){
+    return(1e9)
+  }
+  if(shape<0 & max(dat[,1]+1)<(threshold - scale/shape)){
+    return(1e9)
+  }
+  out = -sum(dat[,2]*log(digp(dat[,1], shape, scale, threshold)))
+  if(is.infinite(out)){
+    return(1e9)
+  }else{
+    return(out)
+  }
+}
+
+fit.igp = function(dat, threshold=0){
+  return(optim(c(0.5, 1), fn = igp_ll_neg,dat=dat[dat[,1]>threshold,], threshold=threshold,
+               lower=c(-Inf,0.001), upper=c(Inf, Inf),method='L-BFGS-B'))
+}
+
+ft = fit.igp(fit$dat,10)
+
+
+degs = counts_to_degs(fit$dat)
+
+
+p = digp(degs[degs>10], -0.001, 1, 10)
+
+max(p)
+min(p)
+
+
+which.min(p)
+
+degs[degs>10][2645]
+
+
+# -------------------------------------------------------------------------
+
+
+igpd_rho = function(pars, shape, scale, threshold, exceedance){
+  p2 = shape*exceedance / (1-shape)
+  x = 1:(threshold+1)
+  p1 = sum(cumprod(
+    (x^pars[1] - (threshold+1)^pars[1] + pars[2]*scale)/
+      (x^pars[1] - (threshold+1)^pars[1] + pars[2]*(scale+1))
+  ))
+  return(p1 + p2)
+}
+
[…]

### 2025-03-27 real_data_fitting.R

@@ -1,0 +1,174 @@
+source('funcs.R')
+library(ggplot2)
+data_raw = read.csv('degs.csv')
+nms = unique(data_raw$name)
+selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
+selected=11
+fits = list()
+plots = list()
+for(i in 1:length(selected)){
+  print(i)
+  fits[[i]] = fit_model(data_raw[data_raw$name==nms[selected[i]],2:3],N=5e4,quiet=F,
+                        trunc.at = 5,k0_jump = 1)
+}
+
+
+saveRDS(fits, 'real_mcmc.rds')
+
+
+for(i in 1:length(fits)){
+  plots[[i]] = ggplot() + scale_x_log10() + scale_y_log10() +
+    geom_point(data = twbfn::deg_surv(counts_to_degs(fits[[i]]$dat)),aes(x=degree, y=surv))+
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y=!!fits[[i]]$surv$est),colour='red') + 
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[1,]),linetype=2, colour='red') + 
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[2,]),,linetype=2,colour='red')+
+    ggtitle(nms[selected[i]])
+}
+
+
+real_survs = ggpubr::ggarrange(plotlist = plots)
+
+saveRDS(real_survs,'real_survs.rds')
+
+
+
+
+real_survs
+# igpd --------------------------------------------------------------------
+
+
+fits = readRDS('real_mcmc.rds')
+
+gpfit = mev::fit.gpd(counts_to_degs(fits[[1]]$dat),threshold=6)
+
+scale = gpfit$estimate[1]
+shape = gpfit$estimate[2]
+
+x = fits[[1]]$dat[,1]
+y = mev::pgp(x,loc=6,scale=scale,shape=shape,lower.tail = F)
+degs = counts_to_degs(fits[[1]]$dat)
+
+plot(x,y,log='xy',type='l')
+points(twbfn::deg_surv(degs[degs>6]))
+
+# -------------------------------------------------------------------------
+
+data_raw = read.csv('degs.csv')
[…]

### 2025-03-27 gpd_to_pref.R
Update files

@@ -4,8 +4,8 @@ digp_log = function(x, shape, scale, threshold){
   v=threshold
   sig=scale
   xi=shape
-  p1 = pmax(x[x>v]*0, 1 + xi * (x[x>v]+1-v)/(sig))
-  p2 = pmax(x[x>v]*0, 1 + xi * (x[x>v]-v)/(sig))
+  p1 = pmax(x[x>v]*0, 1 + xi * (x[x>v]+1-v)/sig)
+  p2 = pmax(x[x>v]*0, 1 + xi * (x[x>v]-v)/sig)
   pows2 = pows1 = x[x>v]*0 + 1
   pows1[p1>0] = -1/xi
   pows2[p2>0] = -1/xi
@@ -14,7 +14,6 @@ digp_log = function(x, shape, scale, threshold){
   out[x<=v] = 0
   return(out)
 }
-
 igp_ll_neg = function(pars,dat,threshold=0){
   shape = pars[1]
   scale = pars[2]
@@ -24,14 +23,13 @@ igp_ll_neg = function(pars,dat,threshold=0){
   if(shape<0 & max(dat[,1]+1)<(threshold - scale/shape)){
     return(1e9)
   }
-  out = -sum(dat[,2]*log(digp(dat[,1], shape, scale, threshold)))
+  out = -sum(dat[,2]*digp_log(dat[,1], shape, scale, threshold))
   if(is.infinite(out)){
     return(1e9)
   }else{
     return(out)
   }
 }
-
 fit.igp = function(dat, threshold=0){
   return(optim(c(0.5, 1), fn = igp_ll_neg,dat=dat[dat[,1]>threshold,], threshold=threshold,
                lower=c(-Inf,0.001), upper=c(Inf, Inf),method='L-BFGS-B'))
@@ -60,10 +58,10 @@ degs[degs>10][2645]
 igpd_rho = function(pars, shape, scale, threshold, exceedance){
   p2 = shape*exceedance / (1-shape)
   x = 1:(threshold+1)
-  p1 = sum(cumprod(
-    (x^pars[1] - (threshold+1)^pars[1] + pars[2]*scale)/
-      (x^pars[1] - (threshold+1)^pars[1] + pars[2]*(scale+1))
-  ))
+  p1 = sum(exp(cumsum(
+    log(x^pars[1] - (threshold+1)^pars[1] + pars[2]*scale)-
+      log(x^pars[1] - (threshold+1)^pars[1] + pars[2]*(scale+1))
+  )))
[…]

### 2025-03-27 real_data_fitting.R

@@ -3,7 +3,6 @@ library(ggplot2)
 data_raw = read.csv('degs.csv')
 nms = unique(data_raw$name)
 selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
-selected=11
 fits = list()
 plots = list()
 for(i in 1:length(selected)){
@@ -77,7 +76,6 @@ fit = fit.gpd(degs, threshold=15)
 fits = readRDS('real_mcmc.rds')
 
 
-plot(coda::as.mcmc(fits[[7]]$smps))
 
 
 
@@ -95,53 +93,34 @@ for(j in 1:length(fits)){
   shapes<- numeric()
   xs <- numeric()
   for(i in 1:(nrow(fit$dat)-3)){
-    if(fit$dat[i,1]>max(fit$mcmc$pars$k0[-(1:2e4)])){
+    if(fit$dat[i,1]>max(fit$mcmc$pars$k0[seq(2e4, 5e4,by=10)])){
       break;
     }
-    igp = fit.igp(fit$dat,threshold=fit$dat[i,1])
-    shapes = c(shapes,igp$par[1])
+    igp = mev::fit.gpd(counts_to_degs(fit$dat),threshold=fit$dat[i,1])
+    shapes = c(shapes,igp$estimate[2])
     xs = c(xs,fit$dat[i,1])
   }
   plts[[j]] = ggplot() +
-    geom_point(aes(x=!!fit$mcmc$pars$k0[-(1:2e4)],y=!!(fit$mcmc$pars$b[-(1:2e4)]/fit$mcmc$lambdas[-(1:2e4)])),colour='blue', alpha=0.0025) + 
-    geom_point(aes(x=!!xs[xs<=max(fit$mcmc$pars$k0[-(1:2e4)])],y=!!shapes[xs<=max(fit$mcmc$pars$k0[-(1:2e4)])]), colour='red') +
+    geom_point(aes(x=!!fit$mcmc$pars$k0[seq(2e4, 5e4,by=10)],y=!!(fit$mcmc$pars$b[seq(2e4, 5e4,by=10)]/fit$mcmc$lambdas[seq(2e4, 5e4,by=10)])),colour='blue', alpha=0.0025) + 
[…]

### 2025-03-27 paper.qmd
Updating figures

@@
 
-The IGPD has been shown to be in the same domain of attraction and has the same tail-index $\xi$ as the regular GPD with the same parameters by using the following quantity introduced by [@shimura12]:
+@shimura12 introduces a quantity that will help in determining what domain of attraction a discrete distribution belongs to, that is:
+
+For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:
 
 $$
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1},
+\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
 $$
-the limit of this quantity if it exists and is non-zero is exactly the tail-index of the distribution $F$, and if the limit is zero then the distribution is in the Gumbel maximum domain of attraction.
-These results will be useful later on when trying to establish the tail behaviour of a network's limiting degree distribution.
+
+[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$.
+Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.
 
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
-#| fig-width: 8
-#| fig-height: 8
-#| fig-cap: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
-#| label: fig-rec2
-
-library(gridExtra)
-plt = readRDS('mcmc_plot.rds')
-plot(plt)
-
-```
+:::{#fig-rec2 out.width="80%"}
+![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](mcmc_plot.jpg)
+:::
 
@@
 However for those with a good fit, it is possible to obtain information about the underlying PA mechanism assuming that the corresponding networks did indeed grow in that fashion.
+
+To demonstrate that the model is capturing the tail behaviour of the degree distributions @fig-gp_comp shows the tail heaviness based on samples from the MCMC for each network as well as the tail heaviness as estimated by fitting the GPD to the degrees for various thresholds.
+
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+#| fig-width: 9
+#| fig-height: 6
+#| fig-cap: Posterior samples of shape parameter from model (blue) and GPD shape parameter (red) at various thresholds
+#| label: fig-gp_comp
+
+plt2 = readRDS('gp_plt.rds')
+
+plot(plt2)
+
+```
 

### 2025-03-27 paper.tex
Updating figures

@@
 
-The IGPD has been shown to be in the same domain of attraction and has
-the same tail-index \(\xi\) as the regular GPD with the same parameters
-by using the following quantity introduced by \citep{shimura12}:
+\citet{shimura12} introduces a quantity that will help in determining
+what domain of attraction a discrete distribution belongs to, that is:
+
+For a distribution \(F\) with survival function \(\bar F\) and some
+\(n\in\mathbb Z^+\) let:
 
 \[
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1},
-\] the limit of this quantity if it exists and is non-zero is exactly
-the tail-index of the distribution \(F\), and if the limit is zero then
-the distribution is in the Gumbel maximum domain of attraction. These
-results will be useful later on when trying to establish the tail
-behaviour of a network's limiting degree distribution.
+\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
+\]
+
+\citep{shimura12} then states that if
+\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
+then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
+Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
+distribution is light tailed.
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec2-1.pdf}
+\captionsetup{labelsep=none}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{mcmc_plot.jpg}
 
@@
 
-\caption{\label{fig-rec2}Posterior estimates of paramters for data
-simulated from the proposed model with various combinations of
+\caption{: Posterior estimates of paramters for data simulated from the
+proposed model with various combinations of
 (\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
+
+\end{figure}%
+
+}
+
+\caption{\label{fig-rec2}}
 
@@
 indeed grow in that fashion.
+
+To demonstrate that the model is capturing the tail behaviour of the
+degree distributions Figure~\ref{fig-gp_comp} shows the tail heaviness
+based on samples from the MCMC for each network as well as the tail
+heaviness as estimated by fitting the GPD to the degrees for various
+thresholds.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-gp_comp-1.pdf}
+
+}
+
+\caption{\label{fig-gp_comp}Posterior samples of shape parameter from
+model (blue) and GPD shape parameter (red) at various thresholds}
+
+\end{figure}%
 

### 2025-04-11 paper.qmd
Updating draft

@@
   nature-pdf:
+    fig-pos: 'H'
     # journal.cite-style is included in the tex file but ignored by pandoc if 
@@
 library(ggpubr)
+library(Rcpp)
+library(RcppArmadillo)
 ```
@@
 The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts in the form of a vector of degrees $\pmb x = (1,2,\ldots,M)$ and the number of vertices with those degrees $\pmb n$ where $M$ is the maximum degree.
-
-Using these and the p.m.f derived in [@rudas07], the likelihood given that $x_i \ge l$ for all $i =1,2,\ldots,M$ is:
+However, when it comes to fitting this model to real data in the next section, small degrees are extremely influential and will often cause the model to have a bad fit even though the rest of the data looks similar to that of the simulations; truncating the data above some value, say $l<k_0$, should allow the model to be more effectively fitted.
+This accounts in some way for the fact that this model only applies to trees, and could not possibly capture the behaviour for nodes with small degrees.
+Using the p.m.f derived in [@rudas07], the likelihood given that $x_i \ge l$ for all $i =1,2,\ldots,M$ is:
 
@@
 The results of this inference are shown in @fig-rec1 and @fig-rec2.
-For these simulated networks $l=2$.
+For these simulated networks $l=0$.
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
 #| fig-width: 8
@@
 #| label: fig-rec1
-surfit_list = list()
-pars = readRDS('recovery_pars.rds')
-recovery_list = readRDS('recovery_dat.rds')
-selected = c(1,27,14,35)
-# selected = 1:36
-for(k in 1:length(selected)){
-  j=selected[k]
-  x = sort(unique(recovery_list[[j]]$degs))
-  ls = c()
-
-  y_975 = recovery_list[[j]]$mcmc$surv$CI[2,]
-  y_025 = recovery_list[[j]]$mcmc$surv$CI[1,]
-  y_50 = recovery_list[[j]]$mcmc$surv$est
-
-  surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
-    geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
-    geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
-    geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
-    scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
-    ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
-                       ', $\\epsilon = $',pars$eps[j],
-                       ', $\\beta = $',pars$b[j])))
-}
-
-fig = ggpubr::ggarrange(plotlist = surfit_list,
-          nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
-          label.x='Degree', label.y = 'Survival')
-
-ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')
+# surfit_list = list()
+# pars = readRDS('recovery_pars.rds')
+# recovery_list = readRDS('recovery_dat.rds')
+# selected = c(1,27,14,35)
+# # selected = 1:36
+# for(k in 1:length(selected)){
+#   j=selected[k]
+#   x = sort(unique(recovery_list[[j]]$degs))
+#   ls = c()
+#   
+#   y_975 = recovery_list[[j]]$mcmc$surv$CI[2,]
+#   y_025 = recovery_list[[j]]$mcmc$surv$CI[1,]
+#   y_50 = recovery_list[[j]]$mcmc$surv$est
+#   
+#   surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
+#     geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
+#     geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
+#     geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
+#     scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
+#     ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
+#                        ', $\\epsilon = $',pars$eps[j],
+#                        ', $\\beta = $',pars$b[j])))
+# }
+# 
+# fig = ggpubr::ggarrange(plotlist = surfit_list,
+#           nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
+#           label.x='Degree', label.y = 'Survival')
+# 
+# 
+# ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')
 
@@
 
-Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources. [describe networks being used].
+Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources.
+[describe networks being used].
+Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in [clement] and is similar to others that have been used to model degree distributions [others].
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
 #| fig-width: 9
-#| fig-height: 6
+#| fig-height: 9
 #| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
@@
 
-plt1 = readRDS('real_survs.rds')
+plt1 = readRDS('comp_plot.rds')
 
@@
 
-@fig-real1 displays the posterior estimates of the conditional survival $F(k|k\ge2)$ for several data sets, for many of the data sets the fit is quite good  whilst for other the fit is very bad.
-This may indicate that the data sets without a good fit are unlikely to have come from a preferential attachment model.
-However for those with a good fit, it is possible to obtain information about the underlying PA mechanism assuming that the corresponding networks did indeed grow in that fashion.
+@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model.
+In most cases, the GPA model does not necessarily provide an improvement in fit when compared to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme.
+@fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model.
 
-To demonstrate that the model is capturing the tail behaviour of the degree distributions @fig-gp_comp shows the tail heaviness based on samples from the MCMC for each network as well as the tail heaviness as estimated by fitting the GPD to the degrees for various thresholds.
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+#| fig-width: 9
+#| fig-height: 9
+#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
+#| label: fig-shapes
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
-#| fig-width: 9
-#| fig-height: 6
-#| fig-cap: Posterior samples of shape parameter from model (blue) and GPD shape parameter (red) at various thresholds
-#| label: fig-gp_comp
-
-plt2 = readRDS('gp_plt.rds')
+plt2 = readRDS('shape_plt.rds')
 

### 2025-04-11 paper.tex
Updating draft

@@
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 and the number of vertices with those degrees \(\pmb n\) where \(M\) is
-the maximum degree.
-
-Using these and the p.m.f derived in \citep{rudas07}, the likelihood
-given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:
+the maximum degree. However, when it comes to fitting this model to real
+data in the next section, small degrees are extremely influential and
+will often cause the model to have a bad fit even though the rest of the
+data looks similar to that of the simulations; truncating the data above
+some value, say \(l<k_0\), should allow the model to be more effectively
+fitted. This accounts in some way for the fact that this model only
+applies to trees, and could not possibly capture the behaviour for nodes
+with small degrees. Using the p.m.f derived in \citep{rudas07}, the
+likelihood given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:
 
@@
 posterior samples. The results of this inference are shown in
-Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
-networks \(l=2\).
+\textbf{?@fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
+networks \(l=0\).
 
-\begin{figure}
-
-\centering{
-
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec1-1.pdf}
-
-}
-
-\caption{\label{fig-rec1}Posterior estimates of survival function for
-data simulated from the proposed model with various combinations of
-(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
-
-\end{figure}%
-
-\begin{figure}
+\begin{figure}[H]
 
@@
 distributions of real networks from various sources. {[}describe
-networks being used{]}.
+networks being used{]}. Alongside fitting this model to the degree
+distributions, we then compare the fit to that of an mixture
+distribution that was used in {[}clement{]} and is similar to others
+that have been used to model degree distributions {[}others{]}.
 
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-Figure~\ref{fig-real1} displays the posterior estimates of the
-conditional survival \(F(k|k\ge2)\) for several data sets, for many of
-the data sets the fit is quite good whilst for other the fit is very
-bad. This may indicate that the data sets without a good fit are
-unlikely to have come from a preferential attachment model. However for
-those with a good fit, it is possible to obtain information about the
-underlying PA mechanism assuming that the corresponding networks did
-indeed grow in that fashion.
+Figure~\ref{fig-real1} displays the posterior estimates of the survival
+function for various data sets, obtained from fitting the GPA model and
+the Zipf-IGP mixture model. In most cases, the GPA model does not
+necessarily provide an improvement in fit when compared to the Zipf-IGP
+model but where the GPA model fits well we gain additional information
+about the preference function assuming that the network evolved
+according the the GPA scheme. Figure~\ref{fig-shapes} shows the
+posterior of the shape parameter \(\xi\) obtained from the Zipf-IGP
+model alongside the posterior of the equivalent shape parameter
+\(\beta/\lambda^*\) obtained from fitting the GPA model.
 
-To demonstrate that the model is capturing the tail behaviour of the
-degree distributions Figure~\ref{fig-gp_comp} shows the tail heaviness
-based on samples from the MCMC for each network as well as the tail
-heaviness as estimated by fitting the GPD to the degrees for various
-thresholds.
-
-\begin{figure}
+\begin{figure}[H]
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-gp_comp-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-shapes-1.pdf}
 
@@
 
-\caption{\label{fig-gp_comp}Posterior samples of shape parameter from
-model (blue) and GPD shape parameter (red) at various thresholds}
+\caption{\label{fig-shapes}Posterior estimates (solid red) of survival
+for several real data sets and their 95\% credible intervals (dotted
+red)}
 

### 2025-04-11 data_fit_background.R

@@ -1,62 +1,128 @@
 source('funcs.R')
 library(ggplot2)
+library(crandep)
+library(data.table)
 data_raw = read.csv('degs.csv')
 nms = unique(data_raw$name)
 selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
 fits = list()
+mixfits = list()
 plots = list()
+dat_list = list()
+trunc.at = 3
+
+
+
+data.files = list.files('./data')
+nms = tstrsplit(data.files,'out.',keep=2)[[1]]
+
+saveRDS(nms, 'nms.rds')
+
+
+i=2
+message(nms[i])
+df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
+df[,1] = as.numeric(as.character(df[,1]))
+dat_list[[i]] = as.matrix(df)
+fits[[i]] = fit_model(dat_list[[i]],N=1e5,quiet=F,
+                      burn_in=1e4,thin.by=5,trunc.at = 5,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
+
+
+for(i in 1:length(data.files)){
+  message(nms[i])
+  df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
+  df[,1] = as.numeric(as.character(df[,1]))
+  dat_list[[i]] = as.matrix(df)
+  fits[[i]] = fit_model(dat_list[[i]],N=1e5,quiet=F,
+                        burn_in=1e4,thin.by=5,trunc.at = trunc.at,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
+}
+
+plots = list()
+
+saveRDS(fits,'fits.rds')
+for(i in 1:length(fits)){
+  udf = as.data.frame(table(fits[[i]]$smps$k0))
+  udf$Var1 = as.numeric(as.character(udf$Var1))
+  sdf = twbfn::deg_surv(counts_to_degs(dat_list[[i]]))
+  ptrunc = sdf$surv[sdf$degree==min(fits[[i]]$dat[,1])]
+  plots[[i]] = ggplot() + scale_x_log10(limits = c(1, max(dat_list[[i]][,1]))) + scale_y_log10(limits=c(1/sum(dat_list[[i]][,2]),1)) +
[…]

### 2025-04-11 funcs.R

@@ -1,3 +1,8 @@
+require(Rcpp)
+require(RcppArmadillo)
+
+sourceCpp('cpp_funcs.cpp')
+
 f = Vectorize(function(x, g, lambda, k0, b){
   if(x>=k0){
     k0s = 0:(k0-1)
@@ -23,7 +28,7 @@ S = Vectorize(function(x, g, lambda, k0, b){
     phi_g = lgamma((lambda+g(k0))/b)  + sum(log(g(k0s)/(lambda+g(k0s))))-
       (lgamma(lambda/b)+lgamma(g(k0)/b))
     
-    return(exp(phi_g)*beta(g(k0)/b + x-k0+1, lambda/b))
+    return(exp(phi_g+lbeta(g(k0)/b + x-k0+1, lambda/b)))
   }
   if(x>0){
     ks = 0:(x)
@@ -34,19 +39,47 @@ S = Vectorize(function(x, g, lambda, k0, b){
   }
   
 },vectorize.args = 'x')
-sample_gpa_tree = function(n, g,b,k0,m=1){
-  degs = c(0)
-  prefs = c(g(0))
+# sample_gpa_tree = function(n, g,b,k0,m=1){
+#   degs = c(0)
+#   prefs = c(g(0))
+#   for(i in 1:n){
+#     # message(i)
+#     selected = sample(1:length(degs), min(m, length(degs)), prob=prefs)
+#     degs = c(degs, 0)
+#     prefs = c(prefs, g(0))
+#     degs[selected] = degs[selected]+1
+#     prefs[selected] = ifelse(degs[selected]<k0, g(degs[selected]), g(k0) + b*(degs[selected]-k0))
+#   }
+#   return(degs)
+# }
+sample_gpa_tree = function(n, g, b, k0, m=1,quiet=T){
+  df = data.frame(deg = c(0),
+                  count = c(1),
+                  pref = c(g(0)))
   for(i in 1:n){
-    # message(i)
-    selected = sample(1:length(degs), min(m, length(degs)), prob=prefs)
-    degs = c(degs, 0)
-    prefs = c(prefs, g(0))
-    degs[selected] = degs[selected]+1
-    prefs[selected] = ifelse(degs[selected]<k0, g(degs[selected]), g(k0) + b*(degs[selected]-k0))
+    if(!quiet){
+      message(i)
+    }
+    selected = sample(1:nrow(df), min(m, sum(df$count)),prob = df$pref*df$count,replace=T)
+    for(j in 1:length(selected)){
[…]

### 2025-04-11 real_data_fitting.R

@@ -73,16 +73,20 @@ fit = fit.gpd(degs, threshold=15)
 # -------------------------------------------------------------------------
 
 
-fits = readRDS('real_mcmc.rds')
+fits = readRDS('modelfits.rds')
+
+
+
+plt = readRDS('real_survs.rds')
 
 
 
 
+library(coda)
+plot(as.mcmc(fits[[10]]$smps), type='l')
 
 
-
-# -------------------------------------------------------------------------
-
+# gpd  vs model -----------------------------------------------------------
 
 
 plts = list()
@@ -90,10 +94,17 @@ plts = list()
 for(j in 1:length(fits)){
   print(j)
   fit = fits[[j]]
+  fit$mcmc$pars  = as.data.frame(fit$mcmc$pars)
+  names(fit$mcmc$pars) = c('a','eps','k0','b')
+  N = nrow(fit$mcmc$pars)
   shapes<- numeric()
   xs <- numeric()
+  
   for(i in 1:(nrow(fit$dat)-3)){
-    if(fit$dat[i,1]>max(fit$mcmc$pars$k0[seq(2e4, 5e4,by=10)])){
+    if(fit$dat[i,1]< min(fit$mcmc$pars$k0[seq(5e3, N, by=10)])){
+      next;
+    }
+    if(fit$dat[i,1]>max(fit$mcmc$pars$k0[seq(5e3, N, by=10)] )){
       break;
     }
     igp = mev::fit.gpd(counts_to_degs(fit$dat),threshold=fit$dat[i,1])
@@ -101,19 +112,23 @@ for(j in 1:length(fits)){
     xs = c(xs,fit$dat[i,1])
   }
   plts[[j]] = ggplot() +
-    geom_point(aes(x=!!fit$mcmc$pars$k0[seq(2e4, 5e4,by=10)],y=!!(fit$mcmc$pars$b[seq(2e4, 5e4,by=10)]/fit$mcmc$lambdas[seq(2e4, 5e4,by=10)])),colour='blue', alpha=0.0025) + 
-    geom_point(aes(x=!!xs[xs<=max(fit$mcmc$pars$k0[seq(2e4, 5e4,by=10)])],y=!!shapes[xs<=max(fit$mcmc$pars$k0[seq(2e4, 5e4,by=10)])]), colour='red') +
[…]

### 2025-04-11 testing.R

@@ -1,0 +1,14 @@
+source('funcs.R')
+
+
+out = recover_params(1.2, 1.0, 0.5, 25.0,quiet=F,n.iter=1e4)
+
+
+
+
+
+
+
+
+
+

### 2025-04-27 paper.qmd
Updating

@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
 #| fig-width: 9
@@
 plot(plt2)
+
+```
+
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+#| fig-width: 9
+#| fig-height: 9
+#| fig-cap: df
+#| label: fig-pa
+
+plt3 = readRDS('PA_plot.rds')
+plot(plt3)
+
+```
+
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+#| fig-width: 9
+#| fig-height: 9
+#| fig-cap: df
+#| label: fig-pars
+
+plt5 = readRDS('pars_plot.rds')
+plot(plt5)
 

### 2025-04-27 data_fit_background.R

@@ -13,22 +13,23 @@ trunc.at = 3
 
 
 
-data.files = list.files('./data')
+data.files = list.files('./data') 
 nms = tstrsplit(data.files,'out.',keep=2)[[1]]
 
 saveRDS(nms, 'nms.rds')
 
 
-i=2
+i=4
 message(nms[i])
 df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
 df[,1] = as.numeric(as.character(df[,1]))
 dat_list[[i]] = as.matrix(df)
 fits[[i]] = fit_model(dat_list[[i]],N=1e5,quiet=F,
-                      burn_in=1e4,thin.by=5,trunc.at = 5,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
+                      burn_in=1e4,thin.by=5,trunc.at = 4,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
 
 
-for(i in 1:length(data.files)){
+
+for(i in 2:2){
   message(nms[i])
   df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
   df[,1] = as.numeric(as.character(df[,1]))
@@ -37,9 +38,13 @@ for(i in 1:length(data.files)){
                         burn_in=1e4,thin.by=5,trunc.at = trunc.at,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
 }
 
+
+saveRDS(dat_list, 'dat_list.rds')
+
+dat_list = readRDS('dat_list.rds')
+
 plots = list()
 
-saveRDS(fits,'fits.rds')
 for(i in 1:length(fits)){
   udf = as.data.frame(table(fits[[i]]$smps$k0))
   udf$Var1 = as.numeric(as.character(udf$Var1))
@@ -53,11 +58,28 @@ for(i in 1:length(fits)){
     geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[2,]*!!ptrunc/fits[[i]]$surv$CI[2,1]),linetype=2, colour='red')+
     ggtitle(nms[i]) + theme(legend.position = 'NULL',aspect.ratio = 1)
 }
+
+tmp = fits[[4]]
+
+
+fits = readRDS('modelfits.rds')
+
+fits[[4]] = tmp
+
+saveRDS(fits, 'modelfits.rds')
+
+
[…]

### 2025-04-27 funcs.R

@@ -62,14 +62,14 @@ sample_gpa_tree = function(n, g, b, k0, m=1,quiet=T){
     }
     selected = sample(1:nrow(df), min(m, sum(df$count)),prob = df$pref*df$count,replace=T)
     for(j in 1:length(selected)){
-      df$count[selected] = df$count[selected]-1
-      if((df$deg[selected]+1) %in% df$deg){
-        df$count[df$deg==df$deg[selected]+1] = df$count[df$deg==df$deg[selected]+1]+1
+      df$count[selected[j]] = df$count[selected[j]]-1
+      if((df$deg[selected[j]]+1) %in% df$deg){
+        df$count[df$deg==df$deg[selected[j]]+1] = df$count[df$deg==df$deg[selected[j]]+1]+1
       }else{
-        df = rbind(df, c(df$deg[selected]+1, 1,
-                         ifelse((df$deg[selected]+1)<k0,
-                                g(df$deg[selected]+1),
-                                g(k0) + b*(df$deg[selected]+1-k0))))
+        df = rbind(df, c(df$deg[selected[j]]+1, 1,
+                         ifelse((df$deg[selected[j]]+1)<k0,
+                                g(df$deg[selected[j]]+1),
+                                g(k0) + b*(df$deg[selected[j]]+1-k0))))
       }
       df$count[1] = df$count[1] + 1
       df = df[df$count!=0,]

### 2025-04-27 real_data_fitting.R

@@ -165,6 +165,264 @@ saveRDS(shape_plt, 'shape_plt.rds')
 
 comp_plot
 
+# -------------------------------------------------------------------------
+library(ggplot2)
+
+
+
+fits = readRDS('fits.rds')
+nms = readRDS('nms.rds')
+plts = list()
+
+
+for(i in 1:length(fits)){
+  plts[[i]] = ggplot()  +
+    geom_point(aes(x=!!fits[[i]]$dat$x, y=!!fits[[i]]$PA$est),size=0.9)+
+    geom_line(aes(x=!!fits[[i]]$dat$x, y=!!fits[[i]]$PA$CI[1,]), lty=2)+
+    geom_line(aes(x=!!fits[[i]]$dat$x, y=!!fits[[i]]$PA$CI[2,]), lty=2)+
+    scale_x_log10() + scale_y_log10() + ggtitle(nms[i])+
+    xlab('Degree') + ylab('P') + theme_bw()
+}
+
+paplot = ggpubr::ggarrange(plotlist = plts)
+
+compplot =  readRDS('comp_plot.rds')
+
+dev.new()
+paplot
+
+
+# -------------------------------------------------------------------------
+
+
+fits = readRDS('fits.rds')
+nms = readRDS('nms.rds')
+
+plts = list()
+
+
+
+pref = function(pars, x){
+  pars = unlist(pars)
+  out = x
+  out[x<pars[3]] = x[x<pars[3]]^pars[1] + pars[2]
+  out[x>=pars[3]] = pars[3]^pars[1] + pars[2] + pars[4]*(x[x>=pars[3]]-pars[3])
+  return(out)
+}
+
+for(i in 1:length(fits)){
+  message(i)
+  x = 1:max(fits[[i]]$dat$x)
+  prefmat = apply(as.matrix(fits[[i]]$smps), 1, pref, x=x)
+  prefCI = apply(prefmat, 1, quantile, prob = c(0.025, 0.5, 0.975))
+  
+  plts[[i]] = ggplot() + geom_line(aes(x = !!x, y= !!prefCI[1,]),lty=2) +
+    geom_line(aes(x = !!x, y= !!prefCI[2,])) +
+    geom_line(aes(x = !!x, y= !!prefCI[3,]),lty=2) +
+     ggtitle(nms[i])
+  
+}
+
+ggpubr::ggarrange(plotlist = plts)
+# -------------------------------------------------------------------------
+
[…]

### 2025-04-27 recreating.R

@@ -1,0 +1,226 @@
+sample_rnd = function(n, pref, gamma_g,quiet = T){
+  degs = c(0)
+  gammas = c(gamma_g(1))
+  prefs = c(pref(degs[1], gammas[1]))
+  for(i in 2:n){
+    if(!quiet) message(i)
+    selected = sample(1:(i-1), 1, prob = prefs)
+    degs[selected] = degs[selected]+1
+    prefs[selected] = pref(degs[selected], gammas[selected])
+    degs = c(degs, 0)
+    gammas = c(gammas, gamma_g(1))
+    prefs  = c(prefs, pref(degs[i], gammas[i]))
+  }
+  return(data.frame(degree=degs, gamma = gammas, pref = prefs))
+}
+
+
+# -------------------------------------------------------------------------
+library(ggplot2)
+
+k0 = 20
+a = 1
+eps = 1
+
+
+pref = function(x, gamma){
+  return(ifelse(x<=k0, x^a+gamma[,2], k0^a + eps + gamma[,1]*(x-k0)))
+}
+gamma_g = function(n){
+  return(rgamma(n*2, 1, 10))
+}
+
+n=1e4
+G = sample_rnd(n, pref, gamma_g, quiet = F)
+degs =  G$degree
+par(mfrow=c(2,2))
+plot(twbfn::deg_dist(degs[degs>0]), log='xy',pch=20,ylim = c(1/n,1), xlim=c(1,n+1))
+plot(twbfn::deg_surv(degs[degs>0]), log='xy',pch=20,ylim = c(1/n,1), xlim=c(1,n+1))
+# plot(density(G$gamma))
+plot(G$degree+1, G$pref,pch=20, col = scales::alpha('blue', 0.3), log='xy', ylim=c(0.01,1e3))
+
+# mcmc network ------------------------------------------------------------
+
+source('funcs.R')
+
+fits = readRDS('modelfits.rds')
+
+
+# -------------------------------------------------------------------------
+
+j = 1
+mcmcdat = fits[[j]]$smps
+
+N = 1e4
+par_rows = sample(1:nrow(mcmcdat),N,replace=T)
+par_mat = mcmcdat[par_rows, ]
[…]

### 2025-04-27 data_fit_background.R
Fixing uploads

@@ -169,6 +169,7 @@ for(i in 1:length(fits)){
 
 shape_plot = ggpubr::ggarrange(plotlist = plots, common.legend = T,label.x='', label.y='\\xi', legend='right')
 shape_plot
+saveRDS(shape_plot, 'shape_plot.rds')
 
 # PA plot -----------------------------------------------------------------
 
@@ -189,8 +190,8 @@ for(i in 1:length(fits)){
   plots[[i]] = ggplot() + geom_line(aes(x=!!x, y = !!pref_CI[1,]), lty=2)+
     geom_line(aes(x=!!x, y = !!pref_CI[2,]), lty=1)+
     geom_line(aes(x=!!x, y = !!pref_CI[3,]), lty=2) + xlab('Degree') + ylab('Preference')+
-    theme(aspect.ratio = 1, axis.title.x = element_blank(), axis.title.y =element_blank()) + ggtitle(nms[i])
-  PA_overlay = PA_overlay + geom_line(aes(!!x, y =!! pref_CI[2,]/sum(pref_CI[2,]), colour=!!as.character(nms[i])))
+    theme(aspect.ratio = 1, axis.title.x = element_blank(), axis.title.y =element_blank()) + ggtitle(nms[i]) + scale_x_log10() + scale_y_log10()
+  PA_overlay = PA_overlay + geom_line(aes(!!x, y =!! pref_CI[2,]/sum(!!pref_CI[2,]), colour=!!as.character(nms[i])))
 }
 
 PA_plot = ggpubr::ggarrange(plotlist = plots,common.legend = T,label.x = 'Degree', label.y='Preference')
@@ -232,7 +233,7 @@ saveRDS(parsplot, 'pars_plot.rds')
 
 
 
-
+ggsave('pars_plot.png', parsplot)
 
 
 

### 2025-04-27 paper.qmd
Fixing uploads
Updating

@@
 bibliography: refs.bib
+
 ---
@@
 
-[@rudas07] introduces theoretical results for a PA model with a general preference function in the cases that the network being generated is a tree; presenting an opportunity to design a preference function that will capture the behaviours observed in the degree distributions of real networks.
-This paper studies a flexible class of preference functions that allow for a power law body and a tail that is heavy and flexible, and proposes a strategy for fitting this model to real data.
-With good fits, this method does not only allow the tail heaviness to be quantified, but it also allows for some inference on the preference function assuming a general preferential attachment model is the underlying generating mechanism.
+The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:
+
+Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some function $b$ of their degree.
+
+Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with index 2 and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution in the Gumbel maximum domain of attraction.
+
+Estimating the preference function used by network generative models like this is of great interest, with the potential to provide a deeper understanding of the relationships between the vertices in the network.
+Several papers and at least one R package (`PAfit`) have been dedicated to estimating the preference function that may have been used when a network was growing; one thing all these have in common is requiring access to a networks evolution data, something that is not always readily available or feasible to access.
+A method for estimating a preference function without this data available would be extremely useful.
+
+[@rudas07] introduces theoretical results for this model with a general preference function in the cases that the network being generated is a tree ($m=1$); presenting an opportunity to design a preference function that will capture the behaviours observed in the degree distributions of real networks.
+Since they provide theoretical results for the limiting degree distribution, a model of this form would be able to be fitted to the degree distributions of real networks where the full evolution is not available, only a snapshot.
+
+## Structure of this paper
+
+@sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from [@rudas07], going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body.
+@sec-rec shows that the preference function parameters can be fairly well recovered from the degree distributions of networks simulated from the GPA model using various functions of the class introduced in @sec-gpa.
+Having shown that for simulated data the model parameters are recoverable from the degree distribution alone, @sec-real builds upon this and attempts to fit the model to degree distributions of real networks; modelling not only the degree distribution itself but also the estimates for the parameters of the preference function assuming the networks evolved according to the GPA scheme.
+Ultimately, @sec-conc discusses the main results of the paper, pitfalls of this method and future work.
 
@@
 
-# Recovery
+# Recovery {#sec-rec}
 
@@
 #| label: fig-rec1
-# surfit_list = list()
-# pars = readRDS('recovery_pars.rds')
-# recovery_list = readRDS('recovery_dat.rds')
-# selected = c(1,27,14,35)
-# # selected = 1:36
-# for(k in 1:length(selected)){
-#   j=selected[k]
-#   x = sort(unique(recovery_list[[j]]$degs))
-#   ls = c()
-#   
-#   y_975 = recovery_list[[j]]$mcmc$surv$CI[2,]
-#   y_025 = recovery_list[[j]]$mcmc$surv$CI[1,]
-#   y_50 = recovery_list[[j]]$mcmc$surv$est
-#   
-#   surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
-#     geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
-#     geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
-#     geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
-#     scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
-#     ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
-#                        ', $\\epsilon = $',pars$eps[j],
-#                        ', $\\beta = $',pars$b[j])))
-# }
-# 
-# fig = ggpubr::ggarrange(plotlist = surfit_list,
-#           nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
-#           label.x='Degree', label.y = 'Survival')
-# 
-# 
-# ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')
+
+surfit_list = list()
+pars = readRDS('recovery_pars.rds')
+recovery_list = readRDS('recovery_dat.rds')
+selected = c(1,27,14,35)
+# selected = 1:36
+for(k in 1:length(selected)){
+  j=selected[k]
+  x = recovery_list[[j]]$mcmc$dat[,1]
+  ls = c()
+
+  y_975 = recovery_list[[j]]$mcmc$surv$CI[2,][order(x)]
+  y_025 = recovery_list[[j]]$mcmc$surv$CI[1,][order(x)]
+  y_50 = recovery_list[[j]]$mcmc$surv$est[order(x)]
+  x = sort(x)
+  surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
+    geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
+    geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
+    geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
+    scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
+    ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
+                       ', $\\epsilon = $',pars$eps[j],
+                       ', $\\beta = $',pars$b[j])))
+}
+
+fig = ggpubr::ggarrange(plotlist = surfit_list,
+          nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
+          label.x='Degree', label.y = 'Survival')
+
+ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')
 
@@
 
-# Application to Real Data
+@fig-rec1 and @fig-rec2 demonstrate that using this methodology it is possible to recover the model parameters fairly well from only the final degree distribution of a simulated network.
+This indicates that the method may also be able to be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.
+
+# Application to Real Data {#sec-real}
 
@@
 @fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model.
+Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution and where it doesn't it appears to either be because it is fitting better at the tail than the Zipf-IGP model or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA.
+This again shows the effects that small degrees have on this model, which is somewhat expected as the theory used for this model is for trees and none of these real networks (nor many real networks) are.
 
@@
 
-plt2 = readRDS('shape_plt.rds')
+plt2 = readRDS('shape_plot.rds')
 
@@
 
+@fig-pa shows the estimated preference function $b$ alongside the 95% credible interval  on a log-log plot.
+Although the credible interval becomes very large for the largest degrees, this is expected as not all of these networks had data in that region, for those that do the credible interval is much narrower as is the case for `sx-mathoverflow`.
+The most insightful conclusion we can draw from these plots come from the shape of the preference function.
+There appears to be two distinct shapes of preference function.
+The first appears mostly flat (similar to UA) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`.
+The second distinct shape appears provide some clear PA behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.
+
 ```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
@@
 #| fig-height: 9
-#| fig-cap: df
+#| fig-cap: Posterior estimate for preference function (solid) with 95% credible interval (dashed) on log-log scale.
 #| label: fig-pa
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
-#| fig-width: 9
-#| fig-height: 9
-#| fig-cap: df
-#| label: fig-pars
+:::{#fig-rec2 out.width="80%"}
+![: Posterior estimates of paramters for real data.](pars_plot.png)
+:::
 
-plt5 = readRDS('pars_plot.rds')
-plot(plt5)
+# Conclusion and Discussion {#sec-conc}
 
-```
+In this paper we introduced a flexible class of preference function that when used under the GPA scheme (in the tree setting) is guaranteed to generate a network with a heavy tailed degree distribution whilst remaining flexible in the body.
+Using simulations from networks using this class of preference function we showed that the model parameters are quite easily estimated from the degrees alone at a snapshot in time.
+Seeing this we applied this method to the degree distributions of real networks, estimating their model parameters assuming they evolved in the same way.
+Not only did this yield fairly good fits for the degree distribution, similar to that of the Zipf-IGP, it came with the added benefit of giving a posterior estimate for a preference function.
 
-# Conclusion and Discussion
+Obviously this method had its flaws in that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model as a result of using theory developed for trees and applying it to general networks.
+Hopefully, as the field progresses we will be able to apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.
 
-# Additional Results {.appendix}
+GPA is also a fairly simplistic one on its own despite having large flexibility in the preference function allowing for edges to only be added at times when a new node is introduced into the network and not allowing for anything like vertex/edge death, reciprocal edges and rewiring steps.
+[@deijfen2015] introduces some theory for GPA with vertex death based on the work by [@rudas07], but is unable to come to an expression for the degree distribution.
+However, adding something like fitness scores into the GPA model seems fairly simple to do even going so far as making each of the parameters used in the class of functions here node-wise instead of universal.
+For example having the preference weight for a vertex be given by $b(k_i, \underline{\gamma_i})$  where $\underline{\gamma_i}\sim G$ is the vertex's fitness and $k_i$ is the vertex's degree, it is reasonable to expect that the degree distribution for a model like this to be given by:
 
-## Tail heaviness of general preferential attachment model
+$$
+\bar F(k) = \mathbb E_G\left[\prod_{i=0}^k \frac{b(i, \gamma)}{\lambda^* + b(i,\gamma)}\right],\qquad \gamma\sim G,\, k=0,1,2,\ldots
+$$
+
+# Appendix {.appendix}
+
+## Additional Results 
+### Tail heaviness of general preferential attachment model
 
@@
 
-# Proofs {.appendix}
+## Proofs
 
-## Tail heaviness of GPA
+### Tail heaviness of GPA
 
@@
 ```
-\newpage
 
-## Removing the infinite sum
+### Removing the infinite sum
 

### 2025-04-27 paper.tex
Updating
Fixing uploads

@@
 
-\citep{rudas07} introduces theoretical results for a PA model with a
+The network generative model that we will be focussing on in this paper
+is dubbed General Preferential Attachment in \citep{rudas07} and is
+defined as follows:
+
+Starting at time \(t=0\) with an initial network of \(m\) vertices that
+each have no edges, at times \(t=1,2,\ldots\) a new vertex is added to
+the network bringing with it \(m\) directed edges (with the new vertex
+as the source); the target for each of these edges are selected from the
+vertices already in the network with weights proportional to some
+function \(b\) of their degree.
+
+Special cases of this model include the Barabási-Albert (BA) model when
+\(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
+with index 2 and the Uniform Attachment (UA) model where \(b(k)=c\)
+leading to a degree distribution in the Gumbel maximum domain of
+attraction.
+
+Estimating the preference function used by network generative models
+like this is of great interest, with the potential to provide a deeper
+understanding of the relationships between the vertices in the network.
+Several papers and at least one R package (\texttt{PAfit}) have been
+dedicated to estimating the preference function that may have been used
+when a network was growing; one thing all these have in common is
+requiring access to a networks evolution data, something that is not
+always readily available or feasible to access. A method for estimating
+a preference function without this data available would be extremely
+useful.
+
+\citep{rudas07} introduces theoretical results for this model with a
 general preference function in the cases that the network being
-generated is a tree; presenting an opportunity to design a preference
-function that will capture the behaviours observed in the degree
-distributions of real networks. This paper studies a flexible class of
-preference functions that allow for a power law body and a tail that is
-heavy and flexible, and proposes a strategy for fitting this model to
-real data. With good fits, this method does not only allow the tail
-heaviness to be quantified, but it also allows for some inference on the
-preference function assuming a general preferential attachment model is
-the underlying generating mechanism.
+generated is a tree (\(m=1\)); presenting an opportunity to design a
+preference function that will capture the behaviours observed in the
+degree distributions of real networks. Since they provide theoretical
+results for the limiting degree distribution, a model of this form would
+be able to be fitted to the degree distributions of real networks where
+the full evolution is not available, only a snapshot.
+
+\subsection{Structure of this paper}\label{structure-of-this-paper}
+
+Section~\ref{sec-gpa} looks at the theoretical limiting behaviour of a
+GPA model when \(m=1\) according to the results from \citep{rudas07},
+going on to introduce a flexible class of preference functions that can
+guarantee a heavy tailed degree distribution while still being flexible
+in the body. Section~\ref{sec-rec} shows that the preference function
+parameters can be fairly well recovered from the degree distributions of
+networks simulated from the GPA model using various functions of the
+class introduced in Section~\ref{sec-gpa}. Having shown that for
+simulated data the model parameters are recoverable from the degree
+distribution alone, Section~\ref{sec-real} builds upon this and attempts
+to fit the model to degree distributions of real networks; modelling not
+only the degree distribution itself but also the estimates for the
+parameters of the preference function assuming the networks evolved
+according to the GPA scheme. Ultimately, Section~\ref{sec-conc}
+discusses the main results of the paper, pitfalls of this method and
+future work.
 
@@
 
-\section{Recovery}\label{recovery}
+\section{Recovery}\label{sec-rec}
 
@@
 posterior samples. The results of this inference are shown in
-\textbf{?@fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
+Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
 networks \(l=0\).
+
+\begin{figure}[H]
+
+\centering{
+
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec1-1.pdf}
+
+}
+
+\caption{\label{fig-rec1}Posterior estimates of survival function for
+data simulated from the proposed model with various combinations of
+(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
+
+\end{figure}%
 
@@
 
-\section{Application to Real Data}\label{application-to-real-data}
+Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} demonstrate that using
+this methodology it is possible to recover the model parameters fairly
+well from only the final degree distribution of a simulated network.
+This indicates that the method may also be able to be applied to the
+degree distributions of real networks, estimating the model parameters
+assuming they evolved according to the GPA scheme.
+
+\section{Application to Real Data}\label{sec-real}
 
@@
 model alongside the posterior of the equivalent shape parameter
-\(\beta/\lambda^*\) obtained from fitting the GPA model.
+\(\beta/\lambda^*\) obtained from fitting the GPA model. Generally, the
+GPA model performs similarly to the Zipf-IGP when estimating the tail
+behaviour of the degree distribution and where it doesn't it appears to
+either be because it is fitting better at the tail than the Zipf-IGP
+model or because of the threshold being estimated as too low forcing
+almost all of the data to be modelled by the linear part of the GPA.
+This again shows the effects that small degrees have on this model,
+which is somewhat expected as the theory used for this model is for
+trees and none of these real networks (nor many real networks) are.
 
@@
 
-\section{Conclusion and Discussion}\label{conclusion-and-discussion}
+Figure~\ref{fig-pa} shows the estimated preference function \(b\)
+alongside the 95\% credible interval on a log-log plot. Although the
+credible interval becomes very large for the largest degrees, this is
+expected as not all of these networks had data in that region, for those
+that do the credible interval is much narrower as is the case for
+\texttt{sx-mathoverflow}. The most insightful conclusion we can draw
+from these plots come from the shape of the preference function. There
+appears to be two distinct shapes of preference function. The first
+appears mostly flat (similar to UA) for the smallest degrees and then
+after a threshold PA kicks in, some with this shape are
+\texttt{pajek-erdos} and \texttt{sx-mathoverflow}. The second distinct
+shape appears provide some clear PA behaviour that then slows down after
+a certain point, examples of this are seen in the two infrastructure
+networks \texttt{opsahl-openflights} and \texttt{topology}.
 
-\section{Additional Results}\label{additional-results}
+\begin{figure}[H]
 
-\subsection{Tail heaviness of general preferential attachment
+\centering{
+
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pa-1.pdf}
+
+}
+
+\caption{\label{fig-pa}Posterior estimate for preference function
+(solid) with 95\% credible interval (dashed) on log-log scale.}
+
+\end{figure}%
+
+\begin{figure}[H]
+
+\centering{
+
+\captionsetup{labelsep=none}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{pars_plot.png}
+
+}
+
+\caption{: Posterior estimates of paramters for real data.}
+
+\end{figure}%
+
+}
+
+\caption{\label{fig-rec2}}
+
+\end{figure}%
+
+\section{Conclusion and Discussion}\label{sec-conc}
+
+In this paper we introduced a flexible class of preference function that
+when used under the GPA scheme (in the tree setting) is guaranteed to
+generate a network with a heavy tailed degree distribution whilst
+remaining flexible in the body. Using simulations from networks using
+this class of preference function we showed that the model parameters
+are quite easily estimated from the degrees alone at a snapshot in time.
+Seeing this we applied this method to the degree distributions of real
+networks, estimating their model parameters assuming they evolved in the
+same way. Not only did this yield fairly good fits for the degree
+distribution, similar to that of the Zipf-IGP, it came with the added
+benefit of giving a posterior estimate for a preference function.
+
+Obviously this method had its flaws in that the lowest degrees needed to
+be truncated as they had a very large effect on the fit of the model as
+a result of using theory developed for trees and applying it to general
+networks. Hopefully, as the field progresses we will be able to apply
+theory developed for general networks using a similar method to this,
+allowing us to compare the results here something that is more accurate.
+
+GPA is also a fairly simplistic one on its own despite having large
+flexibility in the preference function allowing for edges to only be
+added at times when a new node is introduced into the network and not
+allowing for anything like vertex/edge death, reciprocal edges and
+rewiring steps. \citep{deijfen2015} introduces some theory for GPA with
+vertex death based on the work by \citep{rudas07}, but is unable to come
+to an expression for the degree distribution. However, adding something
+like fitness scores into the GPA model seems fairly simple to do even
+going so far as making each of the parameters used in the class of
+functions here node-wise instead of universal. For example having the
+preference weight for a vertex be given by
+\(b(k_i, \underline{\gamma_i})\) where \(\underline{\gamma_i}\sim G\) is
+the vertex's fitness and \(k_i\) is the vertex's degree, it is
+reasonable to expect that the degree distribution for a model like this
+to be given by:
+
+\[
+\bar F(k) = \mathbb E_G\left[\prod_{i=0}^k \frac{b(i, \gamma)}{\lambda^* + b(i,\gamma)}\right],\qquad \gamma\sim G,\, k=0,1,2,\ldots
+\]
+
+\section{Appendix}\label{appendix}
+
+\subsection{Additional Results}\label{additional-results}
+
+\subsubsection{Tail heaviness of general preferential attachment
 model}\label{tail-heaviness-of-general-preferential-attachment-model}
@@
 
-\section{Proofs}\label{proofs}
+\subsection{Proofs}\label{proofs}
 
-\subsection{Tail heaviness of GPA}\label{tail-heaviness-of-gpa}
+\subsubsection{Tail heaviness of GPA}\label{tail-heaviness-of-gpa}
 
@@
 
-\newpage
-
-\subsection{Removing the infinite sum}\label{removing-the-infinite-sum}
+\subsubsection{Removing the infinite
+sum}\label{removing-the-infinite-sum}
 

### 2025-04-30 paper.qmd
1st repro; rm & untrack tmp*tex, paper_cache/* paper_files/*

@@
 
-:::{#fig-rec2 out.width="80%"}
-![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](mcmc_plot.jpg)
-:::
+```{r fig-rec2, out.width="80%"}
+#| echo: false
+#| fig.cap: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
+knitr::include_graphics("mcmc_plot.jpg")
+```
 
@@
 
-:::{#fig-rec2 out.width="80%"}
-![: Posterior estimates of paramters for real data.](pars_plot.png)
-:::
+```{r fig-rec3, out.width="80%"}
+#| fig-cap: Posterior estimates of paramters for real data.
+#| echo: false
+knitr::include_graphics("pars_plot.png")
+```
 

### 2025-04-30 paper.tex
1st repro; rm & untrack tmp*tex, paper_cache/* paper_files/*

@@
+% Options for packages loaded elsewhere
 % Options for packages loaded elsewhere
@@
 
+\usepackage{xcolor}
 \usepackage{amsmath,amssymb}
+\setcounter{secnumdepth}{5}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math}
+  \usepackage{unicode-math} % this also loads fontspec
   \defaultfontfeatures{Scale=MatchLowercase}
@@
 \makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
+\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}{
+    \@ifstar
+      \xxxParagraphStar
+      \xxxParagraphNoStar
+  }
+  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
+  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}{
+    \@ifstar
+      \xxxSubParagraphStar
+      \xxxSubParagraphNoStar
+  }
+  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
+  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
+\makeatother
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \makeatletter
-\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
-\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
-\makeatother
-% Scale images if necessary, so that they will not overflow the page
-% margins by default, and it is still possible to overwrite the defaults
-% using explicit options in \includegraphics[width, height, ...]{}
-\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
+\newsavebox\pandoc@box
+\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
+  \sbox\pandoc@box{#1}%
+  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
+  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
+  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
+  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
+  \else\usebox{\pandoc@box}%
+  \fi%
+}
 % Set default figure placement to htbp
-\makeatletter
 \def\fps@figure{htbp}
 \makeatother
+
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \makeatother
-\ifLuaTeX
-  \usepackage{selnolig}  % disable illegal ligatures
-\fi
 \usepackage{bookmark}
-
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polyheat-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec1-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
 
@@
 
-\captionsetup{labelsep=none}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{mcmc_plot.jpg}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{mcmc_plot.jpg}
 
@@
 
-\caption{: Posterior estimates of paramters for data simulated from the
-proposed model with various combinations of
+\caption{\label{fig-rec2}Posterior estimates of paramters for data
+simulated from the proposed model with various combinations of
 (\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
-
-\end{figure}%
-
-}
-
-\caption{\label{fig-rec2}}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-real1-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-shapes-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
 
@@
 
-\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pa-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pa-1.pdf}
 
@@
 
-\captionsetup{labelsep=none}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{pars_plot.png}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{pars_plot.png}
 
@@
 
-\caption{: Posterior estimates of paramters for real data.}
-
-\end{figure}%
-
-}
-
-\caption{\label{fig-rec2}}
+\caption{\label{fig-rec3}Posterior estimates of paramters for real
+data.}
 

### 2025-04-30 paper.qmd
mv *rds results/; edit *R accordingly; pdf still reproduced

@@
 surfit_list = list()
-pars = readRDS('recovery_pars.rds')
-recovery_list = readRDS('recovery_dat.rds')
+pars = readRDS('results/recovery_pars.rds')
+recovery_list = readRDS('results/recovery_dat.rds')
 selected = c(1,27,14,35)
@@
 
-plt1 = readRDS('comp_plot.rds')
+plt1 = readRDS('results/comp_plot.rds')
 
@@
 
-plt2 = readRDS('shape_plot.rds')
+plt2 = readRDS('results/shape_plot.rds')
 
@@
 
-plt3 = readRDS('PA_plot.rds')
+plt3 = readRDS('results/PA_plot.rds')
 plot(plt3)

### 2025-04-30 data_fit_background.R

@@ -16,7 +16,7 @@ trunc.at = 3
 data.files = list.files('./data') 
 nms = tstrsplit(data.files,'out.',keep=2)[[1]]
 
-saveRDS(nms, 'nms.rds')
+saveRDS(nms, 'results/nms.rds')
 
 
 i=4
@@ -39,9 +39,9 @@ for(i in 2:2){
 }
 
 
-saveRDS(dat_list, 'dat_list.rds')
+saveRDS(dat_list, 'results/dat_list.rds')
 
-dat_list = readRDS('dat_list.rds')
+dat_list = readRDS('results/dat_list.rds')
 
 plots = list()
 
@@ -62,22 +62,22 @@ for(i in 1:length(fits)){
 tmp = fits[[4]]
 
 
-fits = readRDS('modelfits.rds')
+fits = readRDS('results/modelfits.rds')
 
 fits[[4]] = tmp
 
-saveRDS(fits, 'modelfits.rds')
+saveRDS(fits, 'results/modelfits.rds')
 
 
 # -------------------------------------------------------------------------
 
 
 # saveRDS(fits,'modelfits.rds')
-mixfits = readRDS('mixfits.rds')
-dat_list = readRDS('dat_list.rds')
+mixfits = readRDS('results/mixfits.rds')
+dat_list = readRDS('results/dat_list.rds')
 real_survs = ggpubr::ggarrange(plotlist = plots)
 real_survs
-saveRDS(real_survs,'real_survs.rds')
+saveRDS(real_survs,'results/real_survs.rds')
 igpplots = list()
 
 for(i in 1:length(selected)){
@@ -102,7 +102,7 @@ mixfits[[i]] = crandep::mcmc_mix2_wrapper(df,seed=123L,burn=1e4L)
 # -------------------------------------------------------------------------
 
 
-mixfits  =readRDS('mixfits.rds')
+mixfits  =readRDS('results/mixfits.rds')
 
 for(i in 1:length(mixfits)){
   udf = as.data.frame(table(mixfits[[i]]$pars$u))
@@ -121,17 +121,17 @@ for(i in 1:length(mixfits)){
 mix_survs = ggpubr::ggarrange(plotlist = igpplots)
 mix_survs
 
-saveRDS(mix_survs,'mix_survs.rds')
+saveRDS(mix_survs,'results/mix_survs.rds')
 
-saveRDS(mixfits,'mixfits.rds')
+saveRDS(mixfits,'results/mixfits.rds')
 
 
[…]

### 2025-04-30 real_data_fitting.R

@@ -12,7 +12,7 @@ for(i in 1:length(selected)){
 }
 
 
-saveRDS(fits, 'real_mcmc.rds')
+saveRDS(fits, 'results/real_mcmc.rds')
 
 
 for(i in 1:length(fits)){
@@ -27,7 +27,7 @@ for(i in 1:length(fits)){
 
 real_survs = ggpubr::ggarrange(plotlist = plots)
 
-saveRDS(real_survs,'real_survs.rds')
+saveRDS(real_survs,'results/real_survs.rds')
 
 
 
@@ -36,7 +36,7 @@ real_survs
 # igpd --------------------------------------------------------------------
 
 
-fits = readRDS('real_mcmc.rds')
+fits = readRDS('results/real_mcmc.rds')
 
 gpfit = mev::fit.gpd(counts_to_degs(fits[[1]]$dat),threshold=6)
 
@@ -73,11 +73,11 @@ fit = fit.gpd(degs, threshold=15)
 # -------------------------------------------------------------------------
 
 
-fits = readRDS('modelfits.rds')
+fits = readRDS('results/modelfits.rds')
 
 
 
-plt = readRDS('real_survs.rds')
+plt = readRDS('results/real_survs.rds')
 
 
 
@@ -125,7 +125,7 @@ igp_plt = ggpubr::ggarrange(plotlist = plts,nrow=3,ncol=4)
 plot(igp_plt)
 
 
-saveRDS(igp_plt, 'gp_plt.rds')
+saveRDS(igp_plt, 'results/gp_plt.rds')
 
 
 
@@ -135,13 +135,13 @@ df = data.frame(k0 = fit$mcmc$pars$k0,shape = fit$mcmc$pars$b/fit$mcmc$lambdas)
 df$real = shapes[match(df$k0,xs)]
 
 
-plt = readRDS('real_survs.rds')
+plt = readRDS('results/real_survs.rds')
 
 # -------------------------------------------------------------------------
 
-mixfits = readRDS('mixfits.rds')
-fits = readRDS('fits.rds')
-nms = readRDS('nms.rds')
+mixfits = readRDS('results/mixfits.rds')
+fits = readRDS('results/fits.rds')
+nms = readRDS('results/nms.rds')
 plts = list()
 for(j in 1:length(fits)){
   fit = fits[[j]]
@@ -160,7 +160,7 @@ for(j in 1:length(fits)){
 
   
[…]

### 2025-04-30 recovery.R

@@ -223,12 +223,12 @@ plot(recovery_list[[28]]$mcmc$mcmc$pars[1:10,4],type='l')
 
 recovery_list[[28]]$mcmc$mcmc$sig
 
-saveRDS(recovery_list, 'recovery_dat.rds')
+saveRDS(recovery_list, 'results/recovery_dat.rds')
 
-saveRDS(pars, 'recovery_pars.rds')
+saveRDS(pars, 'results/recovery_pars.rds')
 
-recovery_list  =readRDS('recovery_dat.rds')
-recover_pars = readRDS('recovery_pars.rds')
+recovery_list  =readRDS('results/recovery_dat.rds')
+recover_pars = readRDS('results/recovery_pars.rds')
 thin.by = 5
 full_pars = recovery_list[[1]]$mcmc$smps
 for(i in 2:length(recovery_list)){
@@ -269,7 +269,7 @@ full_plot = ggpubr::ggarrange(pa,peps,pk0,pb,common.legend = T, nrow=2,ncol=2,le
 plot(full_plot)
 
 
-saveRDS(full_plot, 'mcmc_plot.rds')
+saveRDS(full_plot, 'results/mcmc_plot.rds')
 
 
 ggplot(data=full_pars) + geom_point(aes(x=a, y=eps, colour = as.character(true_b)),alpha=0.1) + geom_point(aes(x=true_a,y=true_eps)) + facet_grid(true_a~true_eps)

### 2025-04-30 recreating.R

@@ -43,7 +43,7 @@ plot(G$degree+1, G$pref,pch=20, col = scales::alpha('blue', 0.3), log='xy', ylim
 
 source('funcs.R')
 
-fits = readRDS('modelfits.rds')
+fits = readRDS('results/modelfits.rds')
 
 
 # -------------------------------------------------------------------------
@@ -80,7 +80,7 @@ degs_mn = mcmc_network(N,par_mat_mn)$degs
 g = function(x){
   return(x^par_vec[1] + par_vec[2])
 }
-dat_list = readRDS('dat_list.rds')
+dat_list = readRDS('results/dat_list.rds')
 par(mfrow = c(2,2))
 
 

### 2025-04-30 paper.qmd
edit qmd w/ added ref (in bib); rm (untrack) paper.pdf

@@
 
-If the power law is an adequate model to use for real degree distributions has become a heated debate, with alternatives such as [@Broido_2019] seeming to outperform the power law in some cases.
-Additionally, [ref] find that for many real degree distributions the body may follow a power law, the right tail does not or at least seems to be lighter than is implied by the body whilst still being heavy.
+If the power law is an adequate model to use for real degree distributions has become a heated debate, with alternatives such as @Broido_2019 seeming to outperform the power law in some cases.
+Additionally, @lef24 find that for many real degree distributions the body may follow a power law, the right tail does not or at least seems to be lighter than is implied by the body whilst still being heavy.
 One method for modelling the tail of the degree distributions of real networks is to use a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:
@@
 for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.
+
+[CL: Is the precise form of the IGPD needed here?
+The big picture is probably more important: extreme value methods have not been used in analysing degree distributions until recently, such as @vvvk19 and @lef24.
+Not using these methods essentially downweights the deviation of the vertices with large degrees, which have far more influence on the network than those with small degrees, from the power law behaviour.]
 
@@
 
-[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$.
+@shimura12 then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$.
 Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.
 
+[CL: The results of @shimura12 could be moved closer to Proposition 2.1.
+In the introduction, if you want to include @shimura12, you can mention the paper in passing, when outlining your approach, e.g.
+"Using the results by @shimura12, we derive the tail index / heaviness of the limiting degree distribution of the GPA model [@rudas07].
+This informs us the precise class of preference functions that result in flexible heavy tail.
+"]
+
 These models that seem to outperform the power law do not however come with the added benefit of suggesting a generating mechanism for the networks in the same way the power law does.
-Efforts to make the preferential attachment model more flexible have not yet remedied this issue with [@krapivsky01] finding that using a sub/super-linear preference function yields a Weibull tail and a degenerate network, where one vertex eventually gains all new edges, respectively.
+Efforts to make the preferential attachment model more flexible have not yet remedied this issue with @krapivsky01 finding that using a sub/super-linear preference function yields a Weibull tail and a degenerate network, where one vertex eventually gains all new edges, respectively.
 
-The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:
+The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in @rudas07 and is defined as follows:
 
@@
 #| echo: false
-#| fig.cap: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
+#| fig.cap: Posterior estimates of parameters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
 knitr::include_graphics("mcmc_plot.jpg")
@@
 [describe networks being used].
-Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in [clement] and is similar to others that have been used to model degree distributions [others].
+Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @lef24 and is similar to others that have been used to model degree distributions [others].
 
@@
 ```{r fig-rec3, out.width="80%"}
-#| fig-cap: Posterior estimates of paramters for real data.
+#| fig-cap: Posterior estimates of parameters for real data.
 #| echo: false

### 2025-04-30 paper.tex
edit qmd w/ added ref (in bib); rm (untrack) paper.pdf

@@
 distributions has become a heated debate, with alternatives such as
-\citep{Broido_2019} seeming to outperform the power law in some cases.
-Additionally, {[}ref{]} find that for many real degree distributions the
-body may follow a power law, the right tail does not or at least seems
-to be lighter than is implied by the body whilst still being heavy. One
-method for modelling the tail of the degree distributions of real
-networks is to use a discretised variation of the generalised Pareto
-distribution (GPD) usually dubbed the Integer GPD (GPD) where
+\citet{Broido_2019} seeming to outperform the power law in some cases.
+Additionally, \citet{lef24} find that for many real degree distributions
+the body may follow a power law, the right tail does not or at least
+seems to be lighter than is implied by the body whilst still being
+heavy. One method for modelling the tail of the degree distributions of
+real networks is to use a discretised variation of the generalised
+Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where
 \(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:
@@
 \] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).
+
+\citep[CL: Is the precise form of the IGPD needed here? The big picture
+is probably more important: extreme value methods have not been used in
+analysing degree distributions until recently, such as][ and
+\citet{lef24}. Not using these methods essentially downweights the
+deviation of the vertices with large degrees, which have far more
+influence on the network than those with small degrees, from the power
+law behaviour.]{vvvk19}
 
@@
 
-\citep{shimura12} then states that if
+\citet{shimura12} then states that if
 \(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
@@
 
+\citep[CL: The results of][ could be moved closer to Proposition 2.1. In
+the introduction, if you want to include \citet{shimura12}, you can
+mention the paper in passing, when outlining your approach, e.g.~``Using
+the results by \citet{shimura12}, we derive the tail index / heaviness
+of the limiting degree distribution of the GPA model \citep{rudas07}.
+This informs us the precise class of preference functions that result in
+flexible heavy tail.'']{shimura12}
+
 These models that seem to outperform the power law do not however come
@@
 preferential attachment model more flexible have not yet remedied this
-issue with \citep{krapivsky01} finding that using a sub/super-linear
+issue with \citet{krapivsky01} finding that using a sub/super-linear
 preference function yields a Weibull tail and a degenerate network,
@@
 The network generative model that we will be focussing on in this paper
-is dubbed General Preferential Attachment in \citep{rudas07} and is
+is dubbed General Preferential Attachment in \citet{rudas07} and is
 defined as follows:
@@
 
-\caption{\label{fig-rec2}Posterior estimates of paramters for data
+\caption{\label{fig-rec2}Posterior estimates of parameters for data
 simulated from the proposed model with various combinations of
@@
 distributions, we then compare the fit to that of an mixture
-distribution that was used in {[}clement{]} and is similar to others
+distribution that was used in \citet{lef24} and is similar to others
 that have been used to model degree distributions {[}others{]}.
@@
 
-\caption{\label{fig-rec3}Posterior estimates of paramters for real
+\caption{\label{fig-rec3}Posterior estimates of parameters for real
 data.}

### 2025-05-07 paper.qmd
make equations & figures larger in qmd

@@
 We relate the distributions tail heaviness directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
-keywords: [networks, extremes]
+keywords: [networks, discrete extremes, degree distribution]
 bibliography: refs.bib
@@
 
-## Structure of this paper
-
-@sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from [@rudas07], going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body.
+The rest of this paper is as follows: @sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from [@rudas07], going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body.
 @sec-rec shows that the preference function parameters can be fairly well recovered from the degree distributions of networks simulated from the GPA model using various functions of the class introduced in @sec-gpa.
@@
 
-which aligns with the result from [@rudas07], where the expression is derived using the same branching process results.
+which aligns with the result from @rudas07, where the expression is derived using the same branching process results.
 
-Using results from [@shimura12] the domain of attraction of this limiting degree distribution can be found.
+Using results from @shimura12, the domain of attraction of this limiting degree distribution can be found.
 
@@
 
-Previously, in [@krapivsky01], a preference function of the form $k^\alpha +\varepsilon$ has been considered.
+Previously, in @krapivsky01, a preference function of the form $k^\alpha +\varepsilon$ has been considered.
 However, it has been noted that super-linear preferential attachment ($\alpha > 1$) leads to a degenerate limiting degree distribution where at some point one vertex gains the connection from every vertex that joins the network, additionally sub-linear preferential attachment ($\alpha <1$) has been shown to lead to a light tailed limiting degree distribution.
@@
 \bar F(k) = \begin{cases}
-\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
-\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
+\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
+\displaystyle\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
 \end{cases}
@@
 
-with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where:
+with $\lambda^*$ satisfying
 
 $$
-\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}
+\hat\rho(\lambda^{*}) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda^{*}+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda^{*}-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^{*} + i^\alpha + \varepsilon} = 1,
 $$
 
-which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$ since when $\lambda\le \beta$ the infinite sum no longer converges and instead goes to infinity.
+which has to be solved numerically for most parameter choices. Note that $\lambda>\beta$ since when $\lambda\le \beta$ the infinite sum no longer converges and instead goes to infinity.
 
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="95%"}
 #| fig-width: 8
@@
 k0s = 20
-x = 0:200
+x = seq(0, 2000, by = 10)
 pars = expand.grid(as,bs,eps,k0s,x)
@@
 ggplot(data = pars) + geom_line(aes(x=(x+1),y=surv, linetype=as.character(b),colour =as.character(b))) +
-  scale_x_log10() + scale_y_log10(limits=c(1e-5,1))+theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) + 
+  scale_x_log10() + scale_y_log10(limits=c(1e-8,1))+theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) + 
   facet_grid(eps~a,labeller = labeller, scales='free')
@@
 
-The survival (@eq-polysurv) can be connected to the IGPD mentioned in [@sec-intro] by using Sterling's approximation of the gamma function to obtain:
+Applying Stirling's approximation of the gamma function to the second case of @eq-polysurv, we obtain, for $k\geq k_0$,
 
@@
 \bar F(k) 
-\begin{cases}
-=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
-\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
-\end{cases}
-$$ meaning that for $k\ge k_0$ the limiting degree distribution is similar to $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
+%\begin{cases}
+%=\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
+\displaystyle\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},%&k\ge k_0,
+%\end{cases}
+$$ which is the survival function of the $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$ distribution.
+This means that, above the threshold of the piecewise preference function, the limiting degree distribution can be approximated by the IGPD with shape parameter $\beta/\lambda^{*}$.
 
-To assess how close of an approximation this is the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey.
+To assess the quality of the approximation, the conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey.
 The approximation seems to hold up fairly well even for large degrees.
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="95%"}
 #| fig-width: 8
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="95%"}
 #| fig-width: 12
 #| fig-height: 6
-#| fig-cap: Heat maps of $\xi$ for various combinations of the parameters of the proposed model.
+#| fig-cap: Heatmaps of $\xi$ for various combinations of the parameters of the proposed model.
 #| label: fig-polyheat
@@
   scale_fill_paletteer_c(palette='grDevices::Blues',limits=c(0,1),direction=-1)+theme(aspect.ratio = 1)+ylim(min(bs), max(bs))+xlim(min(as), max(as))+
-  facet_grid(k0~eps,labeller = labeller) + labs(fill=TeX('\\xi')) + xlab(TeX('\\alpha')) + ylab(TeX('\\beta')) + theme_minimal()
+  facet_grid(k0~eps,labeller = labeller) + labs(fill=TeX('\\xi')) + xlab(TeX('\\alpha')) + ylab(TeX('\\beta')) + theme_minimal(15)
 
@@
 \bar F(k) = \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k+1+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k+1+\lambda^*+\epsilon)},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times \frac{\Gamma\left(\frac{k_0+\varepsilon+\lambda^*}{\beta}\right)}{\Gamma\left(\frac{k_0+\varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon}{\beta}+1\right)}{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon+\lambda^*}{\beta}+1\right)},&k\ge k_0
+\displaystyle\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k+1+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k+1+\lambda^*+\epsilon)},&k<k_0,\\
+\displaystyle\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times \frac{\Gamma\left(\frac{k_0+\varepsilon+\lambda^*}{\beta}\right)}{\Gamma\left(\frac{k_0+\varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon}{\beta}+1\right)}{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon+\lambda^*}{\beta}+1\right)},&k\ge k_0.
 \end{cases}
@@
 
-Just like the previous model, the survival can be approximated by:
+Similar to the general case, the survival function can be approximated by:
 
@@
 \bar F(k) \approx \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)}{\Gamma(\varepsilon)} (1+\varepsilon)^{\lambda^*}\left(\frac{k}{1+\varepsilon} + 1\right)^{-\lambda^*},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times\left(\frac{\beta(k+1-k_0)}{k_0+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0
+\displaystyle\frac{\Gamma(\lambda^*+\varepsilon)}{\Gamma(\varepsilon)} (1+\varepsilon)^{\lambda^*}\left(\frac{k}{1+\varepsilon} + 1\right)^{-\lambda^*},&k<k_0,\\
+\displaystyle\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times\left(\frac{\beta(k+1-k_0)}{k_0+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0.
 \end{cases}
@@
 
-# Recovery {#sec-rec}
+# Simulation study {#sec-rec}
 
-The goal of was to provide a possible preference function that is able to explain the growth of real networks with varied tail behaviour in their degree distributions.
+<!-- The goal of was to provide a possible preference function that is able to explain the growth of real networks with varied tail behaviour in their degree distributions. -->
 This section aims to show that the parameters of the model in @sec-model can be recovered from the degree distribution of a network simulated from it.

### 2025-05-07 paper.tex
make equations & figures larger in qmd

@@
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
-  pdfkeywords={networks, extremes},
+  pdfkeywords={networks, discrete extremes, degree distribution},
   colorlinks=true,
@@
 % keywords
-\keywords{networks,  extremes}
+\keywords{networks,  discrete extremes,  degree distribution}
 
@@
 
-\subsection{Structure of this paper}\label{structure-of-this-paper}
-
-Section~\ref{sec-gpa} looks at the theoretical limiting behaviour of a
-GPA model when \(m=1\) according to the results from \citep{rudas07},
-going on to introduce a flexible class of preference functions that can
-guarantee a heavy tailed degree distribution while still being flexible
-in the body. Section~\ref{sec-rec} shows that the preference function
-parameters can be fairly well recovered from the degree distributions of
-networks simulated from the GPA model using various functions of the
-class introduced in Section~\ref{sec-gpa}. Having shown that for
-simulated data the model parameters are recoverable from the degree
-distribution alone, Section~\ref{sec-real} builds upon this and attempts
-to fit the model to degree distributions of real networks; modelling not
-only the degree distribution itself but also the estimates for the
-parameters of the preference function assuming the networks evolved
-according to the GPA scheme. Ultimately, Section~\ref{sec-conc}
-discusses the main results of the paper, pitfalls of this method and
-future work.
+The rest of this paper is as follows: Section~\ref{sec-gpa} looks at the
+theoretical limiting behaviour of a GPA model when \(m=1\) according to
+the results from \citep{rudas07}, going on to introduce a flexible class
+of preference functions that can guarantee a heavy tailed degree
+distribution while still being flexible in the body.
+Section~\ref{sec-rec} shows that the preference function parameters can
+be fairly well recovered from the degree distributions of networks
+simulated from the GPA model using various functions of the class
+introduced in Section~\ref{sec-gpa}. Having shown that for simulated
+data the model parameters are recoverable from the degree distribution
+alone, Section~\ref{sec-real} builds upon this and attempts to fit the
+model to degree distributions of real networks; modelling not only the
+degree distribution itself but also the estimates for the parameters of
+the preference function assuming the networks evolved according to the
+GPA scheme. Ultimately, Section~\ref{sec-conc} discusses the main
+results of the paper, pitfalls of this method and future work.
 
@@
 
-which aligns with the result from \citep{rudas07}, where the expression
+which aligns with the result from \citet{rudas07}, where the expression
 is derived using the same branching process results.
 
-Using results from \citep{shimura12} the domain of attraction of this
+Using results from \citet{shimura12}, the domain of attraction of this
 limiting degree distribution can be found.
@@
 
-Previously, in \citep{krapivsky01}, a preference function of the form
+Previously, in \citet{krapivsky01}, a preference function of the form
 \(k^\alpha +\varepsilon\) has been considered. However, it has been
@@
 
-with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where:
+with \(\lambda^*\) satisfying
 
 \[
-\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}
+\hat\rho(\lambda^{*}) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda^{*}+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda^{*}-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^{*} + i^\alpha + \varepsilon} = 1,
 \]
 
-which must be solved numerically for most parameter choices. Also, note
-that \(\lambda>\beta\) since when \(\lambda\le \beta\) the infinite sum
-no longer converges and instead goes to infinity.
+which has to be solved numerically for most parameter choices. Note that
+\(\lambda>\beta\) since when \(\lambda\le \beta\) the infinite sum no
+longer converges and instead goes to infinity.
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
+\includegraphics[width=0.95\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
 
@@
 
-The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
-mentioned in Section~\ref{sec-intro} by using Sterling's approximation
-of the gamma function to obtain:
+Applying Stirling's approximation of the gamma function to the second
+case of Equation~\ref{eq-polysurv}, we obtain, for \(k\geq k_0\),
 
@@
 \bar F(k) 
-\begin{cases}
-=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
-\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
-\end{cases}
-\] meaning that for \(k\ge k_0\) the limiting degree distribution is
-similar to
-\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).
+%\begin{cases}
+%=\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
+\displaystyle\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},%&k\ge k_0,
+%\end{cases}
+\] which is the survival function of the
+\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\)
+distribution. This means that, above the threshold of the piecewise
+preference function, the limiting degree distribution can be
+approximated by the IGPD with shape parameter \(\beta/\lambda^{*}\).
 
-To assess how close of an approximation this is the theoretical
-conditional survivals are shown in Figure~\ref{fig-approx_surv} in
-colour and their IGPD approximations are shown in grey. The
-approximation seems to hold up fairly well even for large degrees.
+To assess the quality of the approximation, the conditional survivals
+are shown in Figure~\ref{fig-approx_surv} in colour and their IGPD
+approximations are shown in grey. The approximation seems to hold up
+fairly well even for large degrees.
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
+\includegraphics[width=0.95\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}
+\includegraphics[width=0.95\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}
 
@@
 
-\caption{\label{fig-polyheat}Heat maps of \(\xi\) for various
+\caption{\label{fig-polyheat}Heatmaps of \(\xi\) for various
 combinations of the parameters of the proposed model.}
@@
 
-Just like the previous model, the survival can be approximated by:
+Similar to the general case, the survival function can be approximated
+by:
 
@@
 
-\section{Recovery}\label{sec-rec}
+\section{Simulation study}\label{sec-rec}
 
-The goal of was to provide a possible preference function that is able
-to explain the growth of real networks with varied tail behaviour in
-their degree distributions. This section aims to show that the
-parameters of the model in Section~\ref{sec-model} can be recovered from
-the degree distribution of a network simulated from it.
+This section aims to show that the parameters of the model in
+Section~\ref{sec-model} can be recovered from the degree distribution of
+a network simulated from it.
 

### 2025-05-08 paper.qmd
Add temporary section of comments w/ some minor edits in qmd

@@
     affiliations:
-      - name: Newcastle University
+      - name: Newcastle University, UK
         id: 1
@@
 
+# CL: Overall comments {-}
+
+1. The introduction section is not quite clear and needs a lot of improvement, so here are my suggestions on the outline, in a hopefully more logical flow:
+   a.
+Start with the broad picture e.g.
+that networks exist in many fields (name some for which you can cite applications), and that statistical methods are have been used to study networks.
+Some are for revealing the latent / underlying structures of the networks (stochastic block models), some are for explaining some response variables by network properties (ERGMs), and some are for postulating how the network grows and evolves (mechanistic models).
+Find some references, either the groundbreaking ones or comprehensive reviews to back this up.
+   b.
+Introduce the general PA model as one of the mechanistic models.
+This will be a natural point to mention the rules (in terms as simple as possible) and introduce the preference function.
+Mention the special case when $b(k)=k$ and cite @Barabasi99.
+   c.
+Then talk about PA model and mechanistic models in general being useful to **simulate** networks but not be fitted to real data because usually only snapshots are available.
+This leads well into the point on the degree distribution being informative.
+   d. Then in the paragraph starting "When looking at ...", point out that it is the BA model [@Barabasi99] that leads to a power law degree distribution.
+   e.
+Then point out that while it's attempting to attribute to the BA model when degrees seemingly follow the power law, there are two main concerns: first, a general preference function doesn't lead to the power law [@krapivsky01]; second, some claim that a lot of real data don't actually follow the power law.
+Mention not only @Broido_2019 but also responses to them e.g.
+@vvvk19.
+   f.
+@vvvk19 leads naturally to the point that extreme value methods have not been used in analysing degree distributions until recently, such as @vvvk19 and @lef24.
+Not using these methods essentially downweights the deviation of the vertices with large degrees, which have far more influence on the network than those with small degrees, from the power law behaviour.
+   g. Also mention Resnick's works (and others) on the crossroads of networks and extremes. Summarise what these papers do.
+   h.
+Come back to the focus on the degree distribution, and point out that most methods, whether extremes are incorporated or not, are descriptive in the sense that no information about the preference function $b(\cdot)$ is revealed, even if PA model was indeed the underlying mechanism.
+   i.
+Then mention that this paper addresses that gap.
+Specifically, **given the degree distribution of the network assumed to come from the PA model, can we directly infer the model parameters?** While @rudas07 have derived the limiting degree distribution in terms of $b(\cdot)$ and its parameters, we build upon it by finding the class of preference functions that result in realistic tails (heavy but not as heavy as implied by the power law).
+   j.
+Then go straight to the division of the rest of the paper, and conclude the introduction section.
+The remaining materials should mostly go to the next section, but you can retain some materials if you think they complement (or already address) the above points.
+2. Rough outline of Section 2, which should be on the PA model and deriving the tail heaviness of the limiting degree distribution in terms of $b(\cdot)$:
+   a.
+Start with the general description of the PA model.
+You can move here the paragraphs in the introduction section, starting with "The network generative model ...", "Starting at time ...", and "Special cases of this model ...".
+Most of what you have written up till the pmf can be kept.
+   b.
+Next, first state that of interest is the tail heaviness (which here refers to the shape parameter of the (integer) GPD approximation of the tail).
+Then move here the paragraphs on @shimura12 from the introduction.
+   c. Then state proposition 2.1 and point to the Appendix for the proof.
+   d.
+Next, we need a bit more discussion and explanation before stating that $b(\cdot)$ has to be asymptotically linear for a heavy tail.
+First point out that this result covers and aligns with the previously obtained result i.e.
+@krapivsky01.
+This is pretty much the paragraph currently under Section 2.1, but with a better flow.
+You might even include the special case of the special case i.e.
+$b(k)=k+\epsilon$, to briefly illustrate how a tail heaviness of $0.5$ is obtained for the BA model.
+   e.
+Now, argue that $b(\cdot)$ has to be (asymptotically) linear above a threshold, in order to have a positive tail heaviness, then introduce the proposed preference function as our model, with tail heaviness $\beta/\lambda^{*}$.
+   f. Conclude the section by saying that we will examine its property in the next Section (see next point).
+3. The section could be called "Preferential attachment with flexible heavy tail", essentially making the subsection a section itself.
+   a.
+Most of the materials can be kept, and the form of the IGPD should be inserted when approximating the survival using Stirling's (not Sterling's) approximation.
+The survival of the IGPD was in the introduction but I suggest moving it here.
+   b. The piecewise linear model doesn't add much, so I would consider excluding it unless there's a point e.g. computational simplicity?
+   c.
+I would conclude the section by moving the likelihood expression here, and say that the parameters can be inferred in a standard way, via the frequentist or Bayesian approach, as the likelihood is an explicit function of the parameters.
+4.
+For simplicity, it's easier to use $m=1$ throughout and only mention the general case $m>1$ as future work in the discussion section.
+Also, it might be worth just calling the general preferential attachment model without the word "general" throughout, as it is unambiguous that it is the model with a flexible / general $b(\cdot)$, while the BA model is the one with $\b(k)=k+\epsilon$.
+5. There are inconsistencies and duplications in various places, probably because you edited at one place and didn't update associated instances elsewhere. Do go back and forth when making edits.
+6. The presentation can use many incremental improvements:
+   a. use $k$ or $n$ for the integer but not both, unless necessary and in a consistent way. There were places where $n$ was used as the subject initially but it changed to $k$.
+   b. ensure that the text in plots are legible.
+   c. when citing, use the parentheses correctly, as they have been used inappropriately; in Quarto syntax, the square brackets will automatically put the parentheses (similar to citep in LaTeX).
+   d. ensure that the maths don't over the linewidth.
+   e. introduce the full term with the acronym in parentheses when using it for the first time, e.g. GPA.
+   f. some references of equations are not working e.g. eq-surv?
+   g. unless necessary, avoid strong adverbs / adjectives.
+7.
+It seems that you version control all the files, which I won't recommend.
+Only track the files that you manually change, not the generated ones (cache, pdf, and tex unless you manually created one).
+This will facilitate the reproducibility.
+I have removed the generated files in my branch.
+8.
+Consider creating subfolders to store the results, data, figures, etc.
+This will streamline your file structure.
+I have accordingly edited the R scripts so that if you run them, the rds files will be stored in results/.
+9. Accents in names in bib / tex / qmd might not work on some computers. Here's the conversion chart: https://www.bibtex.org/SpecialSymbols/
+
 # Introduction {#sec-intro}
@@
 
-If the power law is an adequate model to use for real degree distributions has become a heated debate, with alternatives such as @Broido_2019 seeming to outperform the power law in some cases.
+Whether the power law is adequate for real degree distributions has become a heated debate, with alternatives such as @Broido_2019 seeming to outperform the power law in some cases.
 Additionally, @lef24 find that for many real degree distributions the body may follow a power law, the right tail does not or at least seems to be lighter than is implied by the body whilst still being heavy.
@@
 
-[CL: Is the precise form of the IGPD needed here?
-The big picture is probably more important: extreme value methods have not been used in analysing degree distributions until recently, such as @vvvk19 and @lef24.
-Not using these methods essentially downweights the deviation of the vertices with large degrees, which have far more influence on the network than those with small degrees, from the power law behaviour.]
+[CL: Is the precise form of the IGPD needed here? The big picture is probably more important (point 3f in overall comments)]
 
@@
 Estimating the preference function used by network generative models like this is of great interest, with the potential to provide a deeper understanding of the relationships between the vertices in the network.
-Several papers and at least one R package (`PAfit`) have been dedicated to estimating the preference function that may have been used when a network was growing; one thing all these have in common is requiring access to a networks evolution data, something that is not always readily available or feasible to access.
+Several papers and an R package (`PAfit`) have been dedicated to estimating the preference function that may have been used when a network was growing; one thing all these have in common is requiring access to a networks evolution data, something that is not always readily available or feasible to access.
 A method for estimating a preference function without this data available would be extremely useful.
 
-[@rudas07] introduces theoretical results for this model with a general preference function in the cases that the network being generated is a tree ($m=1$); presenting an opportunity to design a preference function that will capture the behaviours observed in the degree distributions of real networks.
+@rudas07 introduces theoretical results for this model with a general preference function in the cases that the network being generated is a tree ($m=1$); presenting an opportunity to design a preference function that will capture the behaviours observed in the degree distributions of real networks.
 Since they provide theoretical results for the limiting degree distribution, a model of this form would be able to be fitted to the degree distributions of real networks where the full evolution is not available, only a snapshot.
 
-The rest of this paper is as follows: @sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from [@rudas07], going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body.
+The rest of this paper is as follows: @sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from @rudas07, going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body.
 @sec-rec shows that the preference function parameters can be fairly well recovered from the degree distributions of networks simulated from the GPA model using various functions of the class introduced in @sec-gpa.
 Having shown that for simulated data the model parameters are recoverable from the degree distribution alone, @sec-real builds upon this and attempts to fit the model to degree distributions of real networks; modelling not only the degree distribution itself but also the estimates for the parameters of the preference function assuming the networks evolved according to the GPA scheme.
-Ultimately, @sec-conc discusses the main results of the paper, pitfalls of this method and future work.
+@sec-conc discusses the main results, proposes future work and concludes the paper.
 
-# General Preferential Attachment {#sec-gpa}
+# General Preferential Attachment Model {#sec-gpa}
 
-The focus of this section is the limiting degree distribution of a preferential attachment network with preference function $b(\cdot)$ where each vertex brings with it $m$ edges.
+In this section, we introduce a general preferential attachment model, with the focus on the limiting degree distribution as a function of the preference function $b(\cdot)$ where each vertex brings with it $m$ edges.
 This preference function is subject to the following conditions:
@@
 
-Given these two conditions an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07].
-Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:
+Given these two conditions an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in @rudas07.
+Theorem 1 from @rudas07 states that for the tree $\Upsilon(t)$ at time $t$:
 
@@
 $$
-where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.
+where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$. (CL: $\hat{\rho}$ has not been defined yet.)
 
@@
 $$
-\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
+\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
 $$
-Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:
+Using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (pmf) is
 
 $$
-f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
+f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N,
 $$
@@
 $$
-\lim_{k\rightarrow\infty}\Omega(F,n) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
+\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
 $$
@@
 
-Previously, in @krapivsky01, a preference function of the form $k^\alpha +\varepsilon$ has been considered.
-However, it has been noted that super-linear preferential attachment ($\alpha > 1$) leads to a degenerate limiting degree distribution where at some point one vertex gains the connection from every vertex that joins the network, additionally sub-linear preferential attachment ($\alpha <1$) has been shown to lead to a light tailed limiting degree distribution.
+In @krapivsky01, a preference function of the form $k^\alpha +\varepsilon$ has been considered.
+However, it has been noted that super-linear preferential attachment ($\alpha > 1$) leads to a degenerate limiting degree distribution where at some point one vertex gains the connection (CL: edge?) from every vertex that joins the network, additionally sub-linear preferential attachment ($\alpha <1$) has been shown to lead to a light tailed limiting degree distribution.
 Consider a preference function of the form:
@@
 b(k) = \begin{cases}
-k^\alpha + \varepsilon,&k<k_0\\
-k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
+k^\alpha + \varepsilon,&k<k_0,\\
+k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0,
 \end{cases}
@@
 \bar F(k) = \begin{cases}
-\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
+\displaystyle\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
 \displaystyle\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
@@
 
-with $\lambda^*$ satisfying
+with $\lambda^*$ satisfying $\hat{\rho}(\lambda^{*})=1$, where
 
 $$
-\hat\rho(\lambda^{*}) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda^{*}+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda^{*}-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^{*} + i^\alpha + \varepsilon} = 1,
+\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} = 1,
 $$
 
-which has to be solved numerically for most parameter choices. Note that $\lambda>\beta$ since when $\lambda\le \beta$ the infinite sum no longer converges and instead goes to infinity.
+which has to be solved numerically for most parameter choices.
+Note that $\lambda$ has to be greater than $\beta$ since when $\lambda\le \beta$ the infinite sum diverges.
+(CL: Where is the infinite sum?)
 
@@
 %\end{cases}
-$$ which is the survival function of the $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$ distribution.
+$$ which is the survival function of the $\text{IGPD}\displaystyle\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$ distribution.
 This means that, above the threshold of the piecewise preference function, the limiting degree distribution can be approximated by the IGPD with shape parameter $\beta/\lambda^{*}$.
@@
 
-Some examples of the possible degree distributions are shown in the central row of @fig-polylinsurv.
+Some examples of the limiting degree distributions are shown in the central column of @fig-polylinsurv.
 
@@
 This accounts in some way for the fact that this model only applies to trees, and could not possibly capture the behaviour for nodes with small degrees.
-Using the p.m.f derived in [@rudas07], the likelihood given that $x_i \ge l$ for all $i =1,2,\ldots,M$ is:
+Using the p.m.f derived in @rudas07, the likelihood given that $x_i \ge l$ for all $i =1,2,\ldots,M$ is:
 
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="95%"}
 #| fig-width: 8
 #| fig-height: 8
-#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
+#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of ($\alpha,\beta,\varepsilon$) and $k_0=20$.
 #| label: fig-rec1

### 2025-05-08 paper.tex
Add temporary section of comments w/ some minor edits in qmd

@@
 \affil[1]{\orgdiv{School of Mathematics, Statistics and
-Physics}, \orgname{Newcastle University}}
+Physics}, \orgname{Newcastle University, UK}}
 
@@
 
+\section*{CL: Overall comments}\label{cl-overall-comments}
+\addcontentsline{toc}{section}{CL: Overall comments}
+
+\begin{enumerate}
+\def\labelenumi{\arabic{enumi}.}
+\tightlist
+\item
+  The introduction section is not quite clear and needs a lot of
+  improvement, so here are my suggestions on the outline, in a hopefully
+  more logical flow:
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    Start with the broad picture e.g.~that networks exist in many fields
+    (name some for which you can cite applications), and that
+    statistical methods are have been used to study networks. Some are
+    for revealing the latent / underlying structures of the networks
+    (stochastic block models), some are for explaining some response
+    variables by network properties (ERGMs), and some are for
+    postulating how the network grows and evolves (mechanistic models).
+    Find some references, either the groundbreaking ones or
+    comprehensive reviews to back this up.
+  \item
+    Introduce the general PA model as one of the mechanistic models.
+    This will be a natural point to mention the rules (in terms as
+    simple as possible) and introduce the preference function. Mention
+    the special case when \(b(k)=k\) and cite \citet{Barabasi99}.
+  \item
+    Then talk about PA model and mechanistic models in general being
+    useful to \textbf{simulate} networks but not be fitted to real data
+    because usually only snapshots are available. This leads well into
+    the point on the degree distribution being informative.
+  \item
+    Then in the paragraph starting ``When looking at \ldots{}'', point
+    out that it is the BA model \citep{Barabasi99} that leads to a power
+    law degree distribution.
+  \item
+    Then point out that while it's attempting to attribute to the BA
+    model when degrees seemingly follow the power law, there are two
+    main concerns: first, a general preference function doesn't lead to
+    the power law \citep{krapivsky01}; second, some claim that a lot of
+    real data don't actually follow the power law. Mention not only
+    \citet{Broido_2019} but also responses to them e.g. \citet{vvvk19}.
+  \item
+    \citet{vvvk19} leads naturally to the point that extreme value
+    methods have not been used in analysing degree distributions until
+    recently, such as \citet{vvvk19} and \citet{lef24}. Not using these
+    methods essentially downweights the deviation of the vertices with
+    large degrees, which have far more influence on the network than
+    those with small degrees, from the power law behaviour.
+  \item
+    Also mention Resnick's works (and others) on the crossroads of
+    networks and extremes. Summarise what these papers do.
+  \item
+    Come back to the focus on the degree distribution, and point out
+    that most methods, whether extremes are incorporated or not, are
+    descriptive in the sense that no information about the preference
+    function \(b(\cdot)\) is revealed, even if PA model was indeed the
+    underlying mechanism.
+  \item
+    Then mention that this paper addresses that gap. Specifically,
+    \textbf{given the degree distribution of the network assumed to come
+    from the PA model, can we directly infer the model parameters?}
+    While \citet{rudas07} have derived the limiting degree distribution
+    in terms of \(b(\cdot)\) and its parameters, we build upon it by
+    finding the class of preference functions that result in realistic
+    tails (heavy but not as heavy as implied by the power law).
+  \item
+    Then go straight to the division of the rest of the paper, and
+    conclude the introduction section. The remaining materials should
+    mostly go to the next section, but you can retain some materials if
+    you think they complement (or already address) the above points.
+  \end{enumerate}
+\item
+  Rough outline of Section 2, which should be on the PA model and
+  deriving the tail heaviness of the limiting degree distribution in
+  terms of \(b(\cdot)\):
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    Start with the general description of the PA model. You can move
+    here the paragraphs in the introduction section, starting with ``The
+    network generative model \ldots{}'', ``Starting at time \ldots{}'',
+    and ``Special cases of this model \ldots{}''. Most of what you have
+    written up till the pmf can be kept.
+  \item
+    Next, first state that of interest is the tail heaviness (which here
+    refers to the shape parameter of the (integer) GPD approximation of
+    the tail). Then move here the paragraphs on \citet{shimura12} from
+    the introduction.
+  \item
+    Then state proposition 2.1 and point to the Appendix for the proof.
+  \item
+    Next, we need a bit more discussion and explanation before stating
+    that \(b(\cdot)\) has to be asymptotically linear for a heavy tail.
+    First point out that this result covers and aligns with the
+    previously obtained result i.e. \citet{krapivsky01}. This is pretty
+    much the paragraph currently under Section 2.1, but with a better
+    flow. You might even include the special case of the special case
+    i.e.~\(b(k)=k+\epsilon\), to briefly illustrate how a tail heaviness
+    of \(0.5\) is obtained for the BA model.
+  \item
+    Now, argue that \(b(\cdot)\) has to be (asymptotically) linear above
+    a threshold, in order to have a positive tail heaviness, then
+    introduce the proposed preference function as our model, with tail
+    heaviness \(\beta/\lambda^{*}\).
+  \item
+    Conclude the section by saying that we will examine its property in
+    the next Section (see next point).
+  \end{enumerate}
+\item
+  The section could be called ``Preferential attachment with flexible
+  heavy tail'', essentially making the subsection a section itself.
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    Most of the materials can be kept, and the form of the IGPD should
+    be inserted when approximating the survival using Stirling's (not
+    Sterling's) approximation. The survival of the IGPD was in the
+    introduction but I suggest moving it here.
+  \item
+    The piecewise linear model doesn't add much, so I would consider
+    excluding it unless there's a point e.g.~computational simplicity?
+  \item
+    I would conclude the section by moving the likelihood expression
+    here, and say that the parameters can be inferred in a standard way,
+    via the frequentist or Bayesian approach, as the likelihood is an
+    explicit function of the parameters.
+  \end{enumerate}
+\item
+  For simplicity, it's easier to use \(m=1\) throughout and only mention
+  the general case \(m>1\) as future work in the discussion section.
+  Also, it might be worth just calling the general preferential
+  attachment model without the word ``general'' throughout, as it is
+  unambiguous that it is the model with a flexible / general
+  \(b(\cdot)\), while the BA model is the one with \(\b(k)=k+\epsilon\).
+\item
+  There are inconsistencies and duplications in various places, probably
+  because you edited at one place and didn't update associated instances
+  elsewhere. Do go back and forth when making edits.
+\item
+  The presentation can use many incremental improvements:
+
+  \begin{enumerate}
+  \def\labelenumii{\alph{enumii}.}
+  \tightlist
+  \item
+    use \(k\) or \(n\) for the integer but not both, unless necessary
+    and in a consistent way. There were places where \(n\) was used as
+    the subject initially but it changed to \(k\).
+  \item
+    ensure that the text in plots are legible.
+  \item
+    when citing, use the parentheses correctly, as they have been used
+    inappropriately; in Quarto syntax, the square brackets will
+    automatically put the parentheses (similar to citep in LaTeX).
+  \item
+    ensure that the maths don't over the linewidth.
+  \item
+    introduce the full term with the acronym in parentheses when using
+    it for the first time, e.g.~GPA.
+  \item
+    some references of equations are not working e.g.~eq-surv?
+  \item
+    unless necessary, avoid strong adverbs / adjectives.
+  \end{enumerate}
+\item
+  It seems that you version control all the files, which I won't
+  recommend. Only track the files that you manually change, not the
+  generated ones (cache, pdf, and tex unless you manually created one).
+  This will facilitate the reproducibility. I have removed the generated
+  files in my branch.
+\item
+  Consider creating subfolders to store the results, data, figures, etc.
+  This will streamline your file structure. I have accordingly edited
+  the R scripts so that if you run them, the rds files will be stored in
+  results/.
+\item
+  Accents in names in bib / tex / qmd might not work on some computers.
+  Here's the conversion chart: https://www.bibtex.org/SpecialSymbols/
+\end{enumerate}
+
 \section{Introduction}\label{sec-intro}
@@
 
-If the power law is an adequate model to use for real degree
-distributions has become a heated debate, with alternatives such as
-\citet{Broido_2019} seeming to outperform the power law in some cases.
-Additionally, \citet{lef24} find that for many real degree distributions
-the body may follow a power law, the right tail does not or at least
-seems to be lighter than is implied by the body whilst still being
-heavy. One method for modelling the tail of the degree distributions of
-real networks is to use a discretised variation of the generalised
-Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where
+Whether the power law is adequate for real degree distributions has
+become a heated debate, with alternatives such as \citet{Broido_2019}
+seeming to outperform the power law in some cases. Additionally,
+\citet{lef24} find that for many real degree distributions the body may
+follow a power law, the right tail does not or at least seems to be
+lighter than is implied by the body whilst still being heavy. One method
+for modelling the tail of the degree distributions of real networks is
+to use a discretised variation of the generalised Pareto distribution
+(GPD) usually dubbed the Integer GPD (GPD) where
 \(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:
@@
 
-\citep[CL: Is the precise form of the IGPD needed here? The big picture
-is probably more important: extreme value methods have not been used in
-analysing degree distributions until recently, such as][ and
-\citet{lef24}. Not using these methods essentially downweights the
-deviation of the vertices with large degrees, which have far more
-influence on the network than those with small degrees, from the power
-law behaviour.]{vvvk19}
+{[}CL: Is the precise form of the IGPD needed here? The big picture is
+probably more important (point 3f in overall comments){]}
 
@@
 understanding of the relationships between the vertices in the network.
-Several papers and at least one R package (\texttt{PAfit}) have been
-dedicated to estimating the preference function that may have been used
-when a network was growing; one thing all these have in common is
-requiring access to a networks evolution data, something that is not
-always readily available or feasible to access. A method for estimating
-a preference function without this data available would be extremely
+Several papers and an R package (\texttt{PAfit}) have been dedicated to
+estimating the preference function that may have been used when a
+network was growing; one thing all these have in common is requiring
+access to a networks evolution data, something that is not always
+readily available or feasible to access. A method for estimating a
+preference function without this data available would be extremely
 useful.
 
-\citep{rudas07} introduces theoretical results for this model with a
+\citet{rudas07} introduces theoretical results for this model with a
 general preference function in the cases that the network being
@@
 theoretical limiting behaviour of a GPA model when \(m=1\) according to
-the results from \citep{rudas07}, going on to introduce a flexible class
+the results from \citet{rudas07}, going on to introduce a flexible class
 of preference functions that can guarantee a heavy tailed degree
@@
 the preference function assuming the networks evolved according to the
-GPA scheme. Ultimately, Section~\ref{sec-conc} discusses the main
-results of the paper, pitfalls of this method and future work.
-
-\section{General Preferential Attachment}\label{sec-gpa}
-
-The focus of this section is the limiting degree distribution of a
-preferential attachment network with preference function \(b(\cdot)\)
-where each vertex brings with it \(m\) edges. This preference function
-is subject to the following conditions:
+GPA scheme. Section~\ref{sec-conc} discusses the main results, proposes
+future work and concludes the paper.
+
+\section{General Preferential Attachment Model}\label{sec-gpa}
+
+In this section, we introduce a general preferential attachment model,
+with the focus on the limiting degree distribution as a function of the
+preference function \(b(\cdot)\) where each vertex brings with it \(m\)
+edges. This preference function is subject to the following conditions:
 
@@
 obtained by considering a branching process that is equivalent to the
-growth of the network, as in \citep{rudas07}. Theorem 1 from
-\citep{rudas07} states that for the tree \(\Upsilon(t)\) at time \(t\):
+growth of the network, as in \citet{rudas07}. Theorem 1 from
+\citet{rudas07} states that for the tree \(\Upsilon(t)\) at time \(t\):
 
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).
+\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\). (CL:
+\(\hat{\rho}\) has not been defined yet.)
 
@@
 \[
-\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
-the probability mass function (p.m.f.) can be shown to be:
-
-\[
-f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
+\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
+\] Using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\) the
+probability mass function (pmf) is
+
+\[
+f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N,
 \]
@@
 \[
-\lim_{k\rightarrow\infty}\Omega(F,n) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
+\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
 \]
@@
 
-Previously, in \citet{krapivsky01}, a preference function of the form
+In \citet{krapivsky01}, a preference function of the form
 \(k^\alpha +\varepsilon\) has been considered. However, it has been
@@
 to a degenerate limiting degree distribution where at some point one
-vertex gains the connection from every vertex that joins the network,
-additionally sub-linear preferential attachment (\(\alpha <1\)) has been
-shown to lead to a light tailed limiting degree distribution. Consider a
-preference function of the form:
+vertex gains the connection (CL: edge?) from every vertex that joins the
+network, additionally sub-linear preferential attachment (\(\alpha <1\))
+has been shown to lead to a light tailed limiting degree distribution.
+Consider a preference function of the form:
 
@@
 
-with \(\lambda^*\) satisfying
-
-\[
-\hat\rho(\lambda^{*}) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda^{*}+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda^{*}-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^{*} + i^\alpha + \varepsilon} = 1,
+with \(\lambda^*\) satisfying \(\hat{\rho}(\lambda^{*})=1\), where
+
+\[
+\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} = 1,
 \]
@@
 which has to be solved numerically for most parameter choices. Note that
-\(\lambda>\beta\) since when \(\lambda\le \beta\) the infinite sum no
-longer converges and instead goes to infinity.
+\(\lambda\) has to be greater than \(\beta\) since when
+\(\lambda\le \beta\) the infinite sum diverges. (CL: Where is the
+infinite sum?)
 
@@
 \] which is the survival function of the
-\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\)
+\(\text{IGPD}\displaystyle\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\)
 distribution. This means that, above the threshold of the piecewise
@@
 
-Some examples of the possible degree distributions are shown in the
-central row of Figure~\ref{fig-polylinsurv}.
+Some examples of the limiting degree distributions are shown in the
+central column of Figure~\ref{fig-polylinsurv}.
 
@@
 applies to trees, and could not possibly capture the behaviour for nodes
-with small degrees. Using the p.m.f derived in \citep{rudas07}, the
+with small degrees. Using the p.m.f derived in \citet{rudas07}, the
 likelihood given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
+\includegraphics[width=0.95\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
 
@@
 data simulated from the proposed model with various combinations of
-(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
+(\(\alpha,\beta,\varepsilon\)) and \(k_0=20\).}
 

### 2025-05-19 paper.qmd
Restructuring paper

--- 2025-04-30 paper.qmd
@@
 
-The degree distribution of a network can be very informative about the networks structure, giving some understanding about the inequality in the influence of the nodes.
-Often the degree distribution is one of the simplest places to look when trying to understand this, as the full evolution of the network is usually not available.
-This makes modelling the degree distribution an important topic in research with contributions such as [ref].
+Networks appear in many fields such as sociology, politics, epidemiology, and economics.
+Statistical methods have been and continue to be used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17].
+One such mechanistic model is the general preferential attachment (PA) model where as a network grows and new vertices join they connect to those already present in the network with weights proportional to some preference function $b$ of their degree.
+The well-known Barabási-Albert (BA) model is a special case of this when $b\sim k$ [BA99].
 
-When looking at the degree distributions of real networks, a power law is an attractive option to use when modelling as it is seemingly observed in a lot of networks.
-Fitting the discrete power law to the degree distribution come with the added benefit of suggesting preferential attachment as the generating mechanism since it has been shown to generate networks with a power law degree distribution.
+Mechanistic models, like the PA model, are in general only particularly useful when simulating networks and not fitting to the real data as usually only snapshots of real networks are available and not the entire evolution.
+As such, one of the most informative statistics when regarding a network is the degree distribution.
+When looking at the degree distributions of real networks it is attractive to many to use the power law to model the degree distribution, as the BA model generates degrees that follow a power law, and therefore claim that since the degree distribution is roughly power law then the underlying growth mechanism may be preferential attachment such as the BA model.
+However, it has been shown that a general preference function will not always lead to a power law degree distribution [@krapivsky01] and additionally it has been debated for decades whether most real data actually follows a power law [@Voitalov_2019] or not[@Broido_2019].
 
-If the power law is an adequate model to use for real degree distributions has become a heated debate, with alternatives such as @Broido_2019 seeming to outperform the power law in some cases.
-Additionally, @lef24 find that for many real degree distributions the body may follow a power law, the right tail does not or at least seems to be lighter than is implied by the body whilst still being heavy.
-One method for modelling the tail of the degree distributions of real networks is to use a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:
+Throughout this debate the use of methods from extremes has been absent until relatively recently with the papers such as [@Voitalov_2019] and [Lee].
+Forgoing the use of extreme value methods essentially downplays how deviations from the power law in the largest degrees may affect the network given that they are often much more influential than the vertices with smaller degrees.
+
+Noting this, many have now started to use these extreme value methods when analysing networks.
+[@wan19] showed that extreme value estimation methods prove more robust than existing parametric approaches especially when the data are corrupted or the model has been misspecified, demonstrating that extreme value methods can be a valuable tool when studying networks.
+[@wang2022random] investigates a PA model with a affine preference function (similar to the BA model) with the added behaviour of heterogeneous reciprocal edges, they show that the distribution of in-degrees and out-degrees produced by such a model is multivariate regularly varying and they also have hidden regular variation.
+[Lee] using methods from discrete extremes modelled the right tail of network degrees with a  discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:
 
@@
 
-[CL: Is the precise form of the IGPD needed here?
-The big picture is probably more important: extreme value methods have not been used in analysing degree distributions until recently, such as @vvvk19 and @lef24.
-Not using these methods essentially downweights the deviation of the vertices with large degrees, which have far more influence on the network than those with small degrees, from the power law behaviour.]
-
-@shimura12 introduces a quantity that will help in determining what domain of attraction a discrete distribution belongs to, that is:
-
-For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:
-
-$$
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
-$$
-
-@shimura12 then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$.
-Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.
-
-[CL: The results of @shimura12 could be moved closer to Proposition 2.1.
-In the introduction, if you want to include @shimura12, you can mention the paper in passing, when outlining your approach, e.g.
-"Using the results by @shimura12, we derive the tail index / heaviness of the limiting degree distribution of the GPA model [@rudas07].
-This informs us the precise class of preference functions that result in flexible heavy tail.
-"]
-
-These models that seem to outperform the power law do not however come with the added benefit of suggesting a generating mechanism for the networks in the same way the power law does.
-Efforts to make the preferential attachment model more flexible have not yet remedied this issue with @krapivsky01 finding that using a sub/super-linear preference function yields a Weibull tail and a degenerate network, where one vertex eventually gains all new edges, respectively.
-
-The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in @rudas07 and is defined as follows:
-
-Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some function $b$ of their degree.
-
-Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with index 2 and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution in the Gumbel maximum domain of attraction.
-
-Estimating the preference function used by network generative models like this is of great interest, with the potential to provide a deeper understanding of the relationships between the vertices in the network.
-Several papers and at least one R package (`PAfit`) have been dedicated to estimating the preference function that may have been used when a network was growing; one thing all these have in common is requiring access to a networks evolution data, something that is not always readily available or feasible to access.
-A method for estimating a preference function without this data available would be extremely useful.
-
-[@rudas07] introduces theoretical results for this model with a general preference function in the cases that the network being generated is a tree ($m=1$); presenting an opportunity to design a preference function that will capture the behaviours observed in the degree distributions of real networks.
-Since they provide theoretical results for the limiting degree distribution, a model of this form would be able to be fitted to the degree distributions of real networks where the full evolution is not available, only a snapshot.
+The study of degree distributions, whether using extremes or not, usually does not reveal any information about the preference function even if PA was the underlying generating mechanism of the network.
+Here we address this gap in the literature, more specifically investigating if given the degree distribution of a network assumed to come from the PA model can we directly infer the model parameters?
+We build upon the foundation provided by [@rudas07] deriving the limiting degree distribution in terms of the preference function and suggest a class of preference functions that result in realistic tails that are heavy but are lighter than implied by the power law.
 
@@
 
-# General Preferential Attachment {#sec-gpa}
+# PA model
 
-The focus of this section is the limiting degree distribution of a preferential attachment network with preference function $b(\cdot)$ where each vertex brings with it $m$ edges.
-This preference function is subject to the following conditions:
+The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:
+
+Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some function $b$ of their degree where the preference function $b$ is chosen such that:
 
@@
 b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
-$$
+$$ {#eq-condb1}
 
@@
 
-Given these two conditions an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07].
+Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with index 2 and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution in the Gumbel maximum domain of attraction.
+
+Given conditions [@eq-condb1] and [@eq-condb2] an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07].
 Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:
@@
 $$
+We are specifically interested in how the tail heaviness ($\xi$ in the IGPD) is affected by the preference function $b$, [@shimura12] introduces a quantity that will help in determining what domain of attraction this discrete distribution belongs to, that is:
 
-which aligns with the result from [@rudas07], where the expression is derived using the same branching process results.
+For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:
 
-Using results from [@shimura12] the domain of attraction of this limiting degree distribution can be found.
+$$
+\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
+$$
+
+[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$.
+Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.
+This allows us to show the following
 
@@
 :::
+See [appendix] for the details of the proof.
 
-@prp-omega implies that the only way to obtain a heavy tailed degree distribution using a non-decreasing preference function is to have that preference function be asymptotically linear.
-The next subsection focuses on a particular class of these preference functions that provide flexible behaviour but guarantee a  heavy tail.
+@prp-omega aligns with the result from [@krapivsky01] demonstrating that a sub-linear preference function will lead to a light tailed distribution.
+So, in order for the degree distribution to be heavy tailed we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
+We can in fact show that BA model produces a heavy tailed degree distribution with index $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$,  noting that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ is simply $\lim_{k\rightarrow\infty} 1 = 1$ leaving the tail heaviness to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
 
-## Preferential Attachment with flexible heavy tail {#sec-model}
+We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e. $\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$.
 
-Previously, in [@krapivsky01], a preference function of the form $k^\alpha +\varepsilon$ has been considered.
-However, it has been noted that super-linear preferential attachment ($\alpha > 1$) leads to a degenerate limiting degree distribution where at some point one vertex gains the connection from every vertex that joins the network, additionally sub-linear preferential attachment ($\alpha <1$) has been shown to lead to a light tailed limiting degree distribution.
-Consider a preference function of the form:
+Consider the definition of heavy tails:
+
+\begin{align*}
+\lim_{k\rightarrow\infty}\Omega(F,n) &= c>0\\
+\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} &= c\\
+\lim_{k\rightarrow\infty}[b(k+1)-b(k)] &= \lambda^*c\\
+\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{(k+1)-(k)} &= \lambda^*c\\
+\end{align*}
+
+by Stolz-Cesaro Theorem:
+
+$$
+\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \lambda^*c > 0.
+$$
+
+Since $b$ must be asymptotically linear in order to have positive tail heaviness we propose the following preference function:
 
@@
 $$ for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.
+
+This preference function, as per @prp-omega, will produce a degree distribution with tail heaviness $\beta/\lambda^*$ guaranteeing a heavy tail.
+We study this preference function further in the next section.
+
+# Preferential Attachment with flexible heavy tail {#sec-model}
 
@@
 
-The survival (@eq-polysurv) can be connected to the IGPD mentioned in [@sec-intro] by using Sterling's approximation of the gamma function to obtain:
+The survival (@eq-polysurv) can be connected to the IGPD mentioned in [@sec-intro] by using Stirling's approximation of the gamma function to obtain:
 
-<!-- $$ -->
-<!-- \frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y}, -->
-<!-- $$ -->
-<!-- and utilising the expression from @eq-polysurv: -->
+$$
+\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
+$$
+and utilising the expression from @eq-polysurv:
 
-<!-- \begin{align*} -->
-<!-- \bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\ -->
-<!-- &\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\ -->
-<!-- &=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\ -->
-<!-- &=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta} -->
-<!-- \end{align*} -->
+\begin{align*}
+\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
+&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
+&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
+&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
+\end{align*}
 
@@
 
-### Piecewise Linear Model
+For a given network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$ and maximum degree $M$ the likelihood is given by:
 
-When $\alpha = 1$ the preference function simplifies to a piecewise linear function with $\beta$ representing the ratio of the gradients of the two components:
+\begin{align*}
+L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
+\end{align*}
 
-$$
-b(k) = \min(k,k_0) + \varepsilon + \mathbb I\{k\ge k_0\}\beta(k-k_0)
-$$ for some $\beta, \varepsilon, k_0>0$.
+where $B(y,z)$ is the the beta function and $l\ge0$ variable that allows truncating the data such that the minimum degree is $l$.
 
-Following @eq-polysurv the limiting degree distribution is:
-
-$$
-\bar F(k) = \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k+1+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k+1+\lambda^*+\epsilon)},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times \frac{\Gamma\left(\frac{k_0+\varepsilon+\lambda^*}{\beta}\right)}{\Gamma\left(\frac{k_0+\varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon}{\beta}+1\right)}{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon+\lambda^*}{\beta}+1\right)},&k\ge k_0
-\end{cases}
-$$ {#eq-linsurv}
-
-Just like the previous model, the survival can be approximated by:
-
-$$
-\bar F(k) \approx \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)}{\Gamma(\varepsilon)} (1+\varepsilon)^{\lambda^*}\left(\frac{k}{1+\varepsilon} + 1\right)^{-\lambda^*},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times\left(\frac{\beta(k+1-k_0)}{k_0+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0
-\end{cases}
-$$
-
-Some examples of the possible degree distributions are shown in the central row of @fig-polylinsurv.
+This likelihood allows for the parameters to be inferred using either a frequentist or a Bayesian approach.
 
@@
 surfit_list = list()
-pars = readRDS('results/recovery_pars.rds')
-recovery_list = readRDS('results/recovery_dat.rds')
+pars = readRDS('recovery_pars.rds')
+recovery_list = readRDS('recovery_dat.rds')
 selected = c(1,27,14,35)
@@
 
-```{r fig-rec2, out.width="80%"}
-#| echo: false
-#| fig.cap: Posterior estimates of parameters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
-knitr::include_graphics("mcmc_plot.jpg")
-```
+:::{#fig-rec2 out.width="80%"}
+![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](mcmc_plot.jpg)
+:::
 
@@
 [describe networks being used].
-Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @lef24 and is similar to others that have been used to model degree distributions [others].
+Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in [clement] and is similar to others that have been used to model degree distributions [others].
 
@@
 
-plt1 = readRDS('results/comp_plot.rds')
+plt1 = readRDS('comp_plot.rds')
 
@@
 
-plt2 = readRDS('results/shape_plot.rds')
+plt2 = readRDS('shape_plot.rds')
 
@@
 
-plt3 = readRDS('results/PA_plot.rds')
+plt3 = readRDS('PA_plot.rds')
 plot(plt3)
@@
 
-```{r fig-rec3, out.width="80%"}
-#| fig-cap: Posterior estimates of parameters for real data.
-#| echo: false
-knitr::include_graphics("pars_plot.png")
-```
+:::{#fig-rec2 out.width="80%"}
+![: Posterior estimates of paramters for real data.](pars_plot.png)
+:::
 
@@
 $$
+
+If this were the case, the posterior distributions of parameters from @sec-real could be interpreted as the approximate distributions from which a vertex's fitness values are drawn instead of providing estimates of the universal parameter values.
+Additionally, it allows the ability to set some parameters as universal and some as vertex-wise for example one could let $\alpha, \beta, k_0$ be the same for all vertices but let each vertex in the network have its own "zero appeal" $\varepsilon_i\overset{\text{iid}}{\sim} \text{Ga}(\theta_1, \theta_2)$; moving the model to being more realistic and accounting for the fact that in many real networks the objects vertices represent have some value instrinsic to themselves other than the number of connections they have.
 

### 2025-05-19 paper.tex
Restructuring paper

--- 2025-04-30 paper.tex
@@
-% Options for packages loaded elsewhere
 % Options for packages loaded elsewhere
@@
 
-\usepackage{xcolor}
 \usepackage{amsmath,amssymb}
-\setcounter{secnumdepth}{5}
 \usepackage{iftex}
@@
 \else % if luatex or xetex
-  \usepackage{unicode-math} % this also loads fontspec
+  \usepackage{unicode-math}
   \defaultfontfeatures{Scale=MatchLowercase}
@@
 \makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
-\makeatletter
 \ifx\paragraph\undefined\else
   \let\oldparagraph\paragraph
-  \renewcommand{\paragraph}{
-    \@ifstar
-      \xxxParagraphStar
-      \xxxParagraphNoStar
-  }
-  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
-  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
+  \renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
 \fi
@@
   \let\oldsubparagraph\subparagraph
-  \renewcommand{\subparagraph}{
-    \@ifstar
-      \xxxSubParagraphStar
-      \xxxSubParagraphNoStar
-  }
-  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
-  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
+  \renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
 \fi
-\makeatother
 
-\usepackage{longtable,booktabs,array}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \makeatletter
-\newsavebox\pandoc@box
-\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
-  \sbox\pandoc@box{#1}%
-  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
-  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
-  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
-  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
-  \else\usebox{\pandoc@box}%
-  \fi%
-}
+\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
+\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
+\makeatother
+% Scale images if necessary, so that they will not overflow the page
+% margins by default, and it is still possible to overwrite the defaults
+% using explicit options in \includegraphics[width, height, ...]{}
+\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
 % Set default figure placement to htbp
+\makeatletter
 \def\fps@figure{htbp}
 \makeatother
-
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \makeatother
+\ifLuaTeX
+  \usepackage{selnolig}  % disable illegal ligatures
+\fi
 \usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same}
+\urlstyle{same} % disable monospaced font for URLs
 \hypersetup{
@@
 
-The degree distribution of a network can be very informative about the
-networks structure, giving some understanding about the inequality in
-the influence of the nodes. Often the degree distribution is one of the
-simplest places to look when trying to understand this, as the full
-evolution of the network is usually not available. This makes modelling
-the degree distribution an important topic in research with
-contributions such as {[}ref{]}.
+Networks appear in many fields such as sociology, politics,
+epidemiology, and economics. Statistical methods have been and continue
+to be used to study networks that appear in these fields and beyond,
+from the use of stochastic block models to detect communities within the
+French political blogosphere \citep{Latouche11}, to using Exponential
+Random Graph Models (ERGMs) to analyse the global trade network
+\citep{Setayesh22}, and the use of mechanistic models in neuroscience to
+investigate the patterns of connections in neural systems
+\citep{Betzel17}. One such mechanistic model is the general preferential
+attachment (PA) model where as a network grows and new vertices join
+they connect to those already present in the network with weights
+proportional to some preference function \(b\) of their degree. The
+well-known Barabási-Albert (BA) model is a special case of this when
+\(b\sim k\) {[}BA99{]}.
 
-When looking at the degree distributions of real networks, a power law
-is an attractive option to use when modelling as it is seemingly
-observed in a lot of networks. Fitting the discrete power law to the
-degree distribution come with the added benefit of suggesting
-preferential attachment as the generating mechanism since it has been
-shown to generate networks with a power law degree distribution.
+Mechanistic models, like the PA model, are in general only particularly
+useful when simulating networks and not fitting to the real data as
+usually only snapshots of real networks are available and not the entire
+evolution. As such, one of the most informative statistics when
+regarding a network is the degree distribution. When looking at the
+degree distributions of real networks it is attractive to many to use
+the power law to model the degree distribution, as the BA model
+generates degrees that follow a power law, and therefore claim that
+since the degree distribution is roughly power law then the underlying
+growth mechanism may be preferential attachment such as the BA model.
+However, it has been shown that a general preference function will not
+always lead to a power law degree distribution \citep{krapivsky01} and
+additionally it has been debated for decades whether most real data
+actually follows a power law \citep{Voitalov_2019} or
+not\citep{Broido_2019}.
 
-If the power law is an adequate model to use for real degree
-distributions has become a heated debate, with alternatives such as
-\citet{Broido_2019} seeming to outperform the power law in some cases.
-Additionally, \citet{lef24} find that for many real degree distributions
-the body may follow a power law, the right tail does not or at least
-seems to be lighter than is implied by the body whilst still being
-heavy. One method for modelling the tail of the degree distributions of
-real networks is to use a discretised variation of the generalised
-Pareto distribution (GPD) usually dubbed the Integer GPD (GPD) where
-\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by the survival:
+Throughout this debate the use of methods from extremes has been absent
+until relatively recently with the papers such as \citep{Voitalov_2019}
+and {[}Lee{]}. Forgoing the use of extreme value methods essentially
+downplays how deviations from the power law in the largest degrees may
+affect the network given that they are often much more influential than
+the vertices with smaller degrees.
+
+Noting this, many have now started to use these extreme value methods
+when analysing networks. \citep{wan19} showed that extreme value
+estimation methods prove more robust than existing parametric approaches
+especially when the data are corrupted or the model has been
+misspecified, demonstrating that extreme value methods can be a valuable
+tool when studying networks. \citep{wang2022random} investigates a PA
+model with a affine preference function (similar to the BA model) with
+the added behaviour of heterogeneous reciprocal edges, they show that
+the distribution of in-degrees and out-degrees produced by such a model
+is multivariate regularly varying and they also have hidden regular
+variation. {[}Lee{]} using methods from discrete extremes modelled the
+right tail of network degrees with a discretised variation of the
+generalised Pareto distribution (GPD) usually dubbed the Integer GPD
+(IGPD) where \(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by
+the survival:
 
@@
 
-\citep[CL: Is the precise form of the IGPD needed here? The big picture
-is probably more important: extreme value methods have not been used in
-analysing degree distributions until recently, such as][ and
-\citet{lef24}. Not using these methods essentially downweights the
-deviation of the vertices with large degrees, which have far more
-influence on the network than those with small degrees, from the power
-law behaviour.]{vvvk19}
+The study of degree distributions, whether using extremes or not,
+usually does not reveal any information about the preference function
+even if PA was the underlying generating mechanism of the network. Here
+we address this gap in the literature, more specifically investigating
+if given the degree distribution of a network assumed to come from the
+PA model can we directly infer the model parameters? We build upon the
+foundation provided by \citep{rudas07} deriving the limiting degree
+distribution in terms of the preference function and suggest a class of
+preference functions that result in realistic tails that are heavy but
+are lighter than implied by the power law.
 
-\citet{shimura12} introduces a quantity that will help in determining
-what domain of attraction a discrete distribution belongs to, that is:
+\subsection{Structure of this paper}\label{structure-of-this-paper}
 
-For a distribution \(F\) with survival function \(\bar F\) and some
-\(n\in\mathbb Z^+\) let:
+\textbf{?@sec-gpa} looks at the theoretical limiting behaviour of a GPA
+model when \(m=1\) according to the results from \citep{rudas07}, going
+on to introduce a flexible class of preference functions that can
+guarantee a heavy tailed degree distribution while still being flexible
+in the body. Section~\ref{sec-rec} shows that the preference function
+parameters can be fairly well recovered from the degree distributions of
+networks simulated from the GPA model using various functions of the
+class introduced in \textbf{?@sec-gpa}. Having shown that for simulated
+data the model parameters are recoverable from the degree distribution
+alone, Section~\ref{sec-real} builds upon this and attempts to fit the
+model to degree distributions of real networks; modelling not only the
+degree distribution itself but also the estimates for the parameters of
+the preference function assuming the networks evolved according to the
+GPA scheme. Ultimately, Section~\ref{sec-conc} discusses the main
+results of the paper, pitfalls of this method and future work.
 
-\[
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
-\]
-
-\citet{shimura12} then states that if
-\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
-Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
-distribution is light tailed.
-
-\citep[CL: The results of][ could be moved closer to Proposition 2.1. In
-the introduction, if you want to include \citet{shimura12}, you can
-mention the paper in passing, when outlining your approach, e.g.~``Using
-the results by \citet{shimura12}, we derive the tail index / heaviness
-of the limiting degree distribution of the GPA model \citep{rudas07}.
-This informs us the precise class of preference functions that result in
-flexible heavy tail.'']{shimura12}
-
-These models that seem to outperform the power law do not however come
-with the added benefit of suggesting a generating mechanism for the
-networks in the same way the power law does. Efforts to make the
-preferential attachment model more flexible have not yet remedied this
-issue with \citet{krapivsky01} finding that using a sub/super-linear
-preference function yields a Weibull tail and a degenerate network,
-where one vertex eventually gains all new edges, respectively.
+\section{PA model}\label{pa-model}
 
 The network generative model that we will be focussing on in this paper
-is dubbed General Preferential Attachment in \citet{rudas07} and is
+is dubbed General Preferential Attachment in \citep{rudas07} and is
 defined as follows:
@@
 vertices already in the network with weights proportional to some
-function \(b\) of their degree.
+function \(b\) of their degree where the preference function \(b\) is
+chosen such that:
+
+\begin{equation}\phantomsection\label{eq-condb1}{
+b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
+}\end{equation}
+
+\begin{equation}\phantomsection\label{eq-condb2}{
+\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
+}\end{equation}
 
@@
 
-Estimating the preference function used by network generative models
-like this is of great interest, with the potential to provide a deeper
-understanding of the relationships between the vertices in the network.
-Several papers and at least one R package (\texttt{PAfit}) have been
-dedicated to estimating the preference function that may have been used
-when a network was growing; one thing all these have in common is
-requiring access to a networks evolution data, something that is not
-always readily available or feasible to access. A method for estimating
-a preference function without this data available would be extremely
-useful.
-
-\citep{rudas07} introduces theoretical results for this model with a
-general preference function in the cases that the network being
-generated is a tree (\(m=1\)); presenting an opportunity to design a
-preference function that will capture the behaviours observed in the
-degree distributions of real networks. Since they provide theoretical
-results for the limiting degree distribution, a model of this form would
-be able to be fitted to the degree distributions of real networks where
-the full evolution is not available, only a snapshot.
-
-\subsection{Structure of this paper}\label{structure-of-this-paper}
-
-Section~\ref{sec-gpa} looks at the theoretical limiting behaviour of a
-GPA model when \(m=1\) according to the results from \citep{rudas07},
-going on to introduce a flexible class of preference functions that can
-guarantee a heavy tailed degree distribution while still being flexible
-in the body. Section~\ref{sec-rec} shows that the preference function
-parameters can be fairly well recovered from the degree distributions of
-networks simulated from the GPA model using various functions of the
-class introduced in Section~\ref{sec-gpa}. Having shown that for
-simulated data the model parameters are recoverable from the degree
-distribution alone, Section~\ref{sec-real} builds upon this and attempts
-to fit the model to degree distributions of real networks; modelling not
-only the degree distribution itself but also the estimates for the
-parameters of the preference function assuming the networks evolved
-according to the GPA scheme. Ultimately, Section~\ref{sec-conc}
-discusses the main results of the paper, pitfalls of this method and
-future work.
-
-\section{General Preferential Attachment}\label{sec-gpa}
-
-The focus of this section is the limiting degree distribution of a
-preferential attachment network with preference function \(b(\cdot)\)
-where each vertex brings with it \(m\) edges. This preference function
-is subject to the following conditions:
-
-\[
-b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
-\]
-
-\begin{equation}\phantomsection\label{eq-condb2}{
-\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
-}\end{equation}
-
-Given these two conditions an expression for the survival of the
-limiting degree distribution can be found in the case that \(m=1\);
-obtained by considering a branching process that is equivalent to the
-growth of the network, as in \citep{rudas07}. Theorem 1 from
-\citep{rudas07} states that for the tree \(\Upsilon(t)\) at time \(t\):
+Given conditions Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}
+an expression for the survival of the limiting degree distribution can
+be found in the case that \(m=1\); obtained by considering a branching
+process that is equivalent to the growth of the network, as in
+\citep{rudas07}. Theorem 1 from \citep{rudas07} states that for the tree
+\(\Upsilon(t)\) at time \(t\):
 
@@
 f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
+\] We are specifically interested in how the tail heaviness (\(\xi\) in
+the IGPD) is affected by the preference function \(b\),
+\citep{shimura12} introduces a quantity that will help in determining
+what domain of attraction this discrete distribution belongs to, that
+is:
+
+For a distribution \(F\) with survival function \(\bar F\) and some
+\(n\in\mathbb Z^+\) let:
+
+\[
+\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
 \]
 
-which aligns with the result from \citep{rudas07}, where the expression
-is derived using the same branching process results.
-
-Using results from \citep{shimura12} the domain of attraction of this
-limiting degree distribution can be found.
+\citep{shimura12} then states that if
+\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
+then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
+Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
+distribution is light tailed. This allows us to show the following
 
@@
 
-Proposition~\ref{prp-omega} implies that the only way to obtain a heavy
-tailed degree distribution using a non-decreasing preference function is
-to have that preference function be asymptotically linear. The next
-subsection focuses on a particular class of these preference functions
-that provide flexible behaviour but guarantee a heavy tail.
+See \hyperref[appendix]{appendix} for the details of the proof.
 
-\subsection{Preferential Attachment with flexible heavy
-tail}\label{sec-model}
+Proposition~\ref{prp-omega} aligns with the result from
+\citep{krapivsky01} demonstrating that a sub-linear preference function
+will lead to a light tailed distribution. So, in order for the degree
+distribution to be heavy tailed we need that the limit
+\(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is positive. We can
+in fact show that BA model produces a heavy tailed degree distribution
+with index \(\xi=0.5\) by considering the preference function
+\(b(k) = k + \varepsilon\), noting that the limit
+\(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) is simply
+\(\lim_{k\rightarrow\infty} 1 = 1\) leaving the tail heaviness to be
+\(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\).
 
-Previously, in \citep{krapivsky01}, a preference function of the form
-\(k^\alpha +\varepsilon\) has been considered. However, it has been
-noted that super-linear preferential attachment (\(\alpha > 1\)) leads
-to a degenerate limiting degree distribution where at some point one
-vertex gains the connection from every vertex that joins the network,
-additionally sub-linear preferential attachment (\(\alpha <1\)) has been
-shown to lead to a light tailed limiting degree distribution. Consider a
-preference function of the form:
+We can show that in order for the degree distribution to be heavy
+tailed, the preference function must be asymptotically linear
+i.e.~\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0\).
+
+Consider the definition of heavy tails:
+
+\begin{align*}
+\lim_{k\rightarrow\infty}\Omega(F,n) &= c>0\\
+\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} &= c\\
+\lim_{k\rightarrow\infty}[b(k+1)-b(k)] &= \lambda^*c\\
+\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{(k+1)-(k)} &= \lambda^*c\\
+\end{align*}
+
+by Stolz-Cesaro Theorem:
+
+\[
+\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \lambda^*c > 0.
+\]
+
+Since \(b\) must be asymptotically linear in order to have positive tail
+heaviness we propose the following preference function:
 
@@
 \] for \(\alpha,\beta, \varepsilon>0\) and \(k_0\in\mathbb N\).
+
+This preference function, as per Proposition~\ref{prp-omega}, will
+produce a degree distribution with tail heaviness \(\beta/\lambda^*\)
+guaranteeing a heavy tail. We study this preference function further in
+the next section.
+
+\section{Preferential Attachment with flexible heavy
+tail}\label{sec-model}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
 
@@
 The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
-mentioned in Section~\ref{sec-intro} by using Sterling's approximation
+mentioned in Section~\ref{sec-intro} by using Stirling's approximation
 of the gamma function to obtain:
+
+\[
+\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
+\] and utilising the expression from Equation~\ref{eq-polysurv}:
+
+\begin{align*}
+\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
+&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
+&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
+&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
+\end{align*}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-polyheat-1.pdf}
 
@@
 
-\subsubsection{Piecewise Linear Model}\label{piecewise-linear-model}
+For a given network with degree count vector
+\(\pmb n = (n_0, n_1, \ldots, n_M)\) and maximum degree \(M\) the
+likelihood is given by:
 
-When \(\alpha = 1\) the preference function simplifies to a piecewise
-linear function with \(\beta\) representing the ratio of the gradients
-of the two components:
+\begin{align*}
+L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
+\end{align*}
 
-\[
-b(k) = \min(k,k_0) + \varepsilon + \mathbb I\{k\ge k_0\}\beta(k-k_0)
-\] for some \(\beta, \varepsilon, k_0>0\).
+where \(B(y,z)\) is the the beta function and \(l\ge0\) variable that
+allows truncating the data such that the minimum degree is \(l\).
 
-Following Equation~\ref{eq-polysurv} the limiting degree distribution
-is:
-
-\begin{equation}\phantomsection\label{eq-linsurv}{
-\bar F(k) = \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k+1+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k+1+\lambda^*+\epsilon)},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times \frac{\Gamma\left(\frac{k_0+\varepsilon+\lambda^*}{\beta}\right)}{\Gamma\left(\frac{k_0+\varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon}{\beta}+1\right)}{\Gamma\left(k-k_0 + \frac{k_0+\varepsilon+\lambda^*}{\beta}+1\right)},&k\ge k_0
-\end{cases}
-}\end{equation}
-
-Just like the previous model, the survival can be approximated by:
-
-\[
-\bar F(k) \approx \begin{cases}
-\frac{\Gamma(\lambda^*+\varepsilon)}{\Gamma(\varepsilon)} (1+\varepsilon)^{\lambda^*}\left(\frac{k}{1+\varepsilon} + 1\right)^{-\lambda^*},&k<k_0\\
-\frac{\Gamma(\lambda^*+\varepsilon)\Gamma(k_0+\varepsilon)}{\Gamma(\varepsilon)\Gamma(k_0+\lambda^*+\varepsilon)}\times\left(\frac{\beta(k+1-k_0)}{k_0+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0
-\end{cases}
-\]
-
-Some examples of the possible degree distributions are shown in the
-central row of Figure~\ref{fig-polylinsurv}.
+This likelihood allows for the parameters to be inferred using either a
+frequentist or a Bayesian approach.
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-rec1-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{mcmc_plot.jpg}
+\captionsetup{labelsep=none}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{mcmc_plot.jpg}
 
@@
 
-\caption{\label{fig-rec2}Posterior estimates of parameters for data
-simulated from the proposed model with various combinations of
+\caption{: Posterior estimates of paramters for data simulated from the
+proposed model with various combinations of
 (\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
+
+\end{figure}%
+
+}
+
+\caption{\label{fig-rec2}}
 
@@
 distributions, we then compare the fit to that of an mixture
-distribution that was used in \citet{lef24} and is similar to others
+distribution that was used in {[}clement{]} and is similar to others
 that have been used to model degree distributions {[}others{]}.
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-real1-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-shapes-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pa-1.pdf}
+\includegraphics[width=0.8\textwidth,height=\textheight]{paper_files/figure-pdf/fig-pa-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{pars_plot.png}
+\captionsetup{labelsep=none}
+
+\begin{figure}[H]
+
+{\centering \includegraphics{pars_plot.png}
 
@@
 
-\caption{\label{fig-rec3}Posterior estimates of parameters for real
-data.}
+\caption{: Posterior estimates of paramters for real data.}
+
+\end{figure}%
+
+}
+
+\caption{\label{fig-rec2}}
 
@@
 \]
+
+If this were the case, the posterior distributions of parameters from
+Section~\ref{sec-real} could be interpreted as the approximate
+distributions from which a vertex's fitness values are drawn instead of
+providing estimates of the universal parameter values. Additionally, it
+allows the ability to set some parameters as universal and some as
+vertex-wise for example one could let \(\alpha, \beta, k_0\) be the same
+for all vertices but let each vertex in the network have its own ``zero
+appeal''
+\(\varepsilon_i\overset{\text{iid}}{\sim} \text{Ga}(\theta_1, \theta_2)\);
+moving the model to being more realistic and accounting for the fact
+that in many real networks the objects vertices represent have some
+value instrinsic to themselves other than the number of connections they
+have.
 

### 2025-05-21 paper.qmd
Updating draft

@@
   nature-pdf:
-    fig-pos: 'H'
     # journal.cite-style is included in the tex file but ignored by pandoc if 
@@
 bibliography: refs.bib  
+---
 
----
+\newpage
+
 ```{r,echo=FALSE, warning=FALSE, message=FALSE}
@@
 
-# Introduction {#sec-intro}
+# Introduction
 
@@
 Statistical methods have been and continue to be used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17].
-One such mechanistic model is the general preferential attachment (PA) model where as a network grows and new vertices join they connect to those already present in the network with weights proportional to some preference function $b$ of their degree.
-The well-known Barabási-Albert (BA) model is a special case of this when $b\sim k$ [BA99].
 
-Mechanistic models, like the PA model, are in general only particularly useful when simulating networks and not fitting to the real data as usually only snapshots of real networks are available and not the entire evolution.
-As such, one of the most informative statistics when regarding a network is the degree distribution.
-When looking at the degree distributions of real networks it is attractive to many to use the power law to model the degree distribution, as the BA model generates degrees that follow a power law, and therefore claim that since the degree distribution is roughly power law then the underlying growth mechanism may be preferential attachment such as the BA model.
-However, it has been shown that a general preference function will not always lead to a power law degree distribution [@krapivsky01] and additionally it has been debated for decades whether most real data actually follows a power law [@Voitalov_2019] or not[@Broido_2019].
+One example of a mechanistic model that may be considered is the preferential attachment (PA) model where as a network grows and new vertices join they connect to those already present in the network with weights proportional to some preference function $b(\cdot)$ of their in-degree.
+The well-known Barabási-Albert (BA) model is a special case of this when $b(k) = k+1$ [@Barabasi99].
 
-Throughout this debate the use of methods from extremes has been absent until relatively recently with the papers such as [@Voitalov_2019] and [Lee].
-Forgoing the use of extreme value methods essentially downplays how deviations from the power law in the largest degrees may affect the network given that they are often much more influential than the vertices with smaller degrees.
+These mechanistic models are generally useful when simulating networks, but not for fitting to real data as usually only snapshots of a network are available and not the full evolution.
+However, one statistic that is usually available for snapshots of networks is the degree distribution making it valuable to study the degrees in a network.
+Often, when looking at the degree distributions of real networks it can be attractive to use the power law when modelling as it is seemingly observed in much of the data.
+It also comes with the added benefit of offering a connection to preferential attachment since the BA model results in a power law degree distribution [@Barabasi99], perhaps suggesting that if the power law fits well then the network may have grown according to similar simple rules.
+However, attempting to explain the networks growth using preferential attachment when the degrees seemingly follow a power law may not be the best course of action as [@krapivsky01] shows that a general preference function does not lead to power law degrees and additionally, it is still being argued whether or not a lot of real data actually follows a power law.
+@Broido_2019 provides some evidence that suggests that real world networks rarely have degrees that follow a power law through the use of statistical analysis of almost 1000 networks across a wide variety of fields; @Voitalov_2019, on the other hand, disagrees and claims that these kinds of networks are not nearly as rare and only appear so as a result of an unrealistic expectation of power law without deviations or noise.
 
-Noting this, many have now started to use these extreme value methods when analysing networks.
-[@wan19] showed that extreme value estimation methods prove more robust than existing parametric approaches especially when the data are corrupted or the model has been misspecified, demonstrating that extreme value methods can be a valuable tool when studying networks.
-[@wang2022random] investigates a PA model with a affine preference function (similar to the BA model) with the added behaviour of heterogeneous reciprocal edges, they show that the distribution of in-degrees and out-degrees produced by such a model is multivariate regularly varying and they also have hidden regular variation.
-[Lee] using methods from discrete extremes modelled the right tail of network degrees with a  discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:
+Extreme value methods have been scarcely used when it comes to analysis of degree distributions and whether they follow the power law.
+The absence of these methods essentially down weights the deviation of vertices with the largest degrees, which are usually the most influential, from the power law.
+Somewhat recent publications [@Voitalov_2019; @Lee24; @wang2022random] have addressed this gap in the literature.
+@Voitalov_2019 proposes estimating the power law exponent of a degree distribution using three estimators based in extreme value theory, in contrast to the estimators generally used in network science, the degree distributions are then split into one of three categories (not power-law , hardly power-law or power-law) depending on the values of the estimates obtained from each of the estimators.
+@Lee24 proposes a framework for modelling entire degree distributions using a combination of two generalisations of the power law including a model selection step that is capable of suggesting whether or not the data is adequately modelled by a power law, with one conclusion being that most data seems to have a tail lighter than what is implied by the body whilst still being heavy tailed.
+@wang2022random considers a preferential attachment model with an affine preference function and reciprocal edges, they study the limiting degree distribution of this model for both the in-degrees and out-degrees - finding that the joint limiting distribution is multivariate regularly varying and has the property of hidden regular variation.
 
-$$
-\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-$$
-for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.
+Evidently, the study of degree distributions is a very active but the vast majority of research in the field (whether using extremes or not) are purely descriptive in the sense that no information about the preference function is revealed even if preferential attachment was the underlying mechanism.
+This paper aims to address this gap - asking if given the degree distribution of a network assumed to come from the preferential attachment model, can we directly infer the model parameters?
 
-The study of degree distributions, whether using extremes or not, usually does not reveal any information about the preference function even if PA was the underlying generating mechanism of the network.
-Here we address this gap in the literature, more specifically investigating if given the degree distribution of a network assumed to come from the PA model can we directly infer the model parameters?
-We build upon the foundation provided by [@rudas07] deriving the limiting degree distribution in terms of the preference function and suggest a class of preference functions that result in realistic tails that are heavy but are lighter than implied by the power law.
+@rudas07 derives the limiting degree distribution in terms of the preference function $b(\cdot)$ and its parameters, we build upon this by finding a class of preference functions that result in realistic tails i.e.
+heavy but not as heavy as implied by the power law.
 
-## Structure of this paper
+The remainder of this paper begins with a more detailed description of the PA model alongside the theoretical results for the limiting degree distribution using results from @rudas07, this section (@sec-tail) then goes on to analyse the tail behaviour of this limiting distribution in terms of the preference function $b(\cdot)$ arguing that a heavy tailed degree distribution can only be the result of an asymptotically linear preference function - ending with a suggestion for a preference function that can guarantee heavy tails while remaining flexible.
+@sec-model continues on from the previous section by considering the limiting distribution obtained when using a preference function of the form stated at the end of @sec-tail providing examples of how the shape of the distribution varies with respect to the model parameters as well as visualisations for how the tail heaviness changes when these parameters are varied - continued study of this distribution then also reveals a connection to the general Pareto distribution (GPD).
+@sec-sim consists of a simulation study where networks are simulated from the PA model for various parameter combinations and then the model specified in @sec-model is fitted to the resulting degree distributions using Bayesian inference, demonstrating that (for simulated data) the model parameters can be recovered from only the degree distribution.
+Finally, @sec-real fits the model to some real data and provides posterior estimates for the preference function assuming the networks evolved according to the PA model.
 
-@sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from [@rudas07], going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body.
-@sec-rec shows that the preference function parameters can be fairly well recovered from the degree distributions of networks simulated from the GPA model using various functions of the class introduced in @sec-gpa.
-Having shown that for simulated data the model parameters are recoverable from the degree distribution alone, @sec-real builds upon this and attempts to fit the model to degree distributions of real networks; modelling not only the degree distribution itself but also the estimates for the parameters of the preference function assuming the networks evolved according to the GPA scheme.
-Ultimately, @sec-conc discusses the main results of the paper, pitfalls of this method and future work.
+\newpage
 
-# PA model
+# Tail Behaviour of Preferential Attachment Model {#sec-tail}
 
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-$$
-where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.
+$$ where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.
 
@@
 \bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
-$$
-which by the previously stated theorem can also be written as:
+$$ which by the previously stated theorem can also be written as:
 
@@
 \bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-$$
-Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:
+$$ Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:
 
@@
 $$
+
 We are specifically interested in how the tail heaviness ($\xi$ in the IGPD) is affected by the preference function $b$, [@shimura12] introduces a quantity that will help in determining what domain of attraction this discrete distribution belongs to, that is:
 
-For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:
+For a distribution $F$ with survival function $\bar F$ and some $k\in\mathbb Z^+$ let:
 
 $$
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
+\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
 $$
 
-[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$.
-Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.
-This allows us to show the following
+[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(k) \sim k^{-\alpha}$.
+Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is light tailed.
+This allows us to show the following:
 
-:::{#prp-omega}
-If $b(k) \rightarrow \infty$ as $k\rightarrow \infty$ then
-$$
+::: {#prp-omega}
+If $b(k) \rightarrow \infty$ as $k\rightarrow \infty$ then $$
 \bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-$$
-and
+$$ and
 
 $$
-\lim_{k\rightarrow\infty}\Omega(F,n) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
+\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
 $$
 :::
-See [appendix] for the details of the proof.
+
+See appendix [-@sec-proofs] for the details of the proof.
 
@@
 
-We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e. $\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$.
+We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e.
+$\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$.
+First we must consider the following theorem:
 
-Consider the definition of heavy tails:
+::: {#thm-stolz}
+## Stolz-Cesàro Theorem [@cesaro]
 
-\begin{align*}
-\lim_{k\rightarrow\infty}\Omega(F,n) &= c>0\\
-\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} &= c\\
-\lim_{k\rightarrow\infty}[b(k+1)-b(k)] &= \lambda^*c\\
-\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{(k+1)-(k)} &= \lambda^*c\\
-\end{align*}
-
-by Stolz-Cesaro Theorem:
+Let $(a_n)_{n\ge1}$ and $(b_n)_{n\ge1}$ be two sequence of real numbers. Assume $(a_n)_{n\ge1}$ is a strictly monotone and divergent sequence and the following limit exists:
 
 $$
-\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \lambda^*c > 0.
+\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
+$$ Then,
+
 $$
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
+$$
+:::
 
-Since $b$ must be asymptotically linear in order to have positive tail heaviness we propose the following preference function:
+From @prp-omega we know that in order for the limiting degree distribution to have heavy tails, we must have that:
+
+$$
+\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
+$$ Now, setting $b_k = b(k)$ and $a_k = k$:
+
+$$
+\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
+$$ meaning by @thm-stolz:
+
+$$
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0.
+$$ Therefore, in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear.
+With this under consideration we propose the following preference function:
 
@@
 
-# Preferential Attachment with flexible heavy tail {#sec-model}
+\newpage
+
+# Preferential Attachment with Flexible Heavy Tail {#sec-model}
 
@@
 \end{cases}
-$${#eq-polysurv}
+$$ {#eq-polysurv}
 
@@
 \hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}
-$$
+$$ {#eq-rho}
 
-which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$ since when $\lambda\le \beta$ the infinite sum no longer converges and instead goes to infinity.
+which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$.
 
@@
 ```{r, echo=FALSE}
-source('funcs.R')
+source('scripts/funcs.R')
 ```
@@
 
-The survival (@eq-polysurv) can be connected to the IGPD mentioned in [@sec-intro] by using Stirling's approximation of the gamma function to obtain:
+The survival (@eq-polysurv) can be connected to the IGPD by using Stirling's approximation of the gamma function, where the IGPD is a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ defined by the survival: $$
+\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
+$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.
 
-$$
+Using Stirling's approximation we have that: $$
 \frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
-$$
-and utilising the expression from @eq-polysurv:
+$$ now utilising the expression from @eq-polysurv:
 
@@
 
-# Recovery {#sec-rec}
+# Simulation Study {#sec-sim}
 
-The goal of was to provide a possible preference function that is able to explain the growth of real networks with varied tail behaviour in their degree distributions.
 This section aims to show that the parameters of the model in @sec-model can be recovered from the degree distribution of a network simulated from it.
 
-The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts in the form of a vector of degrees $\pmb x = (1,2,\ldots,M)$ and the number of vertices with those degrees $\pmb n$ where $M$ is the maximum degree.
-However, when it comes to fitting this model to real data in the next section, small degrees are extremely influential and will often cause the model to have a bad fit even though the rest of the data looks similar to that of the simulations; truncating the data above some value, say $l<k_0$, should allow the model to be more effectively fitted.
-This accounts in some way for the fact that this model only applies to trees, and could not possibly capture the behaviour for nodes with small degrees.
-Using the p.m.f derived in [@rudas07], the likelihood given that $x_i \ge l$ for all $i =1,2,\ldots,M$ is:
-
-\begin{align*}
-L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{x_i\ge k_0}n_{x_i}\right)} \prod_{l \le x_i<k_0}\left(\frac{\lambda^*}{\lambda^* +x_i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{x_i\ge k_0}\left(\frac{\text{B}(x_i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
-\end{align*}
-
-where $B(y,z)$ is the the beta function.
-
-This likelihood allows inference to be made about the parameters using a Bayesian approach using the priors:
+The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongisde the priors:
 
@@
 
-:::{#fig-rec2 out.width="80%"}
-![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](mcmc_plot.jpg)
-:::
+![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2}
 
@@
 Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources.
-[describe networks being used].
-Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in [clement] and is similar to others that have been used to model degree distributions [others].
+Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @Lee24.
 
@@
 
-:::{#fig-rec2 out.width="80%"}
-![: Posterior estimates of paramters for real data.](pars_plot.png)
-:::
+![: Posterior estimates of paramters for real data.](images/pars_plot.png)
 
-# Conclusion and Discussion {#sec-conc}
+# Conclusion and Discussion
 
@@
 
-GPA is also a fairly simplistic one on its own despite having large flexibility in the preference function allowing for edges to only be added at times when a new node is introduced into the network and not allowing for anything like vertex/edge death, reciprocal edges and rewiring steps.
-[@deijfen2015] introduces some theory for GPA with vertex death based on the work by [@rudas07], but is unable to come to an expression for the degree distribution.
-However, adding something like fitness scores into the GPA model seems fairly simple to do even going so far as making each of the parameters used in the class of functions here node-wise instead of universal.
-For example having the preference weight for a vertex be given by $b(k_i, \underline{\gamma_i})$  where $\underline{\gamma_i}\sim G$ is the vertex's fitness and $k_i$ is the vertex's degree, it is reasonable to expect that the degree distribution for a model like this to be given by:
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
 
-$$
-\bar F(k) = \mathbb E_G\left[\prod_{i=0}^k \frac{b(i, \gamma)}{\lambda^* + b(i,\gamma)}\right],\qquad \gamma\sim G,\, k=0,1,2,\ldots
-$$
+\newpage
 
-If this were the case, the posterior distributions of parameters from @sec-real could be interpreted as the approximate distributions from which a vertex's fitness values are drawn instead of providing estimates of the universal parameter values.
-Additionally, it allows the ability to set some parameters as universal and some as vertex-wise for example one could let $\alpha, \beta, k_0$ be the same for all vertices but let each vertex in the network have its own "zero appeal" $\varepsilon_i\overset{\text{iid}}{\sim} \text{Ga}(\theta_1, \theta_2)$; moving the model to being more realistic and accounting for the fact that in many real networks the objects vertices represent have some value instrinsic to themselves other than the number of connections they have.
+# Proofs and Derivations {#sec-proofs .appendix}
 
-# Appendix {.appendix}
-
-## Additional Results 
-### Tail heaviness of general preferential attachment model
-
-Recall the limiting survival function:
-
-$$
-\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
-$$
-
-The aim is to determine how this distribution behaves for different choices of $b$, specifically what maximum domain of attraction does this distribution belong to and is it affected by the choice of preference function $b$.
-@shimura12 introduces a quantity that will help in determining what domain of attraction a discrete distribution belongs to, that is:
-
-For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:
-
-$$
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
-$$
-
-[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$.
-Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.
-
-Consider a preference function $b(\cdot)$ that fulfills:
-
-$$
-\lim_{n\rightarrow \infty} b(n) = \infty
-$${#eq-condb1}
-
-Substituting in the form of $\bar F(n)$ from @eq-surv and taking the limit, subject to @eq-condb1:
-
-$$
-\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
-$$ {#eq-omegares} (see appendix for full proof).
-
-Whilst there are many classes of functions that satisfy @eq-condb1 and @eq-condb2, the vast majority of these result in $\Omega(F,n) \rightarrow 0$, meaning that the limiting degree distribution is light tailed.
-In fact, in order for the limiting degree distribution to be heavy tailed, $b$ must be asymptotically linear ($b(k)\sim k$).
-
-## Proofs
-
-### Tail heaviness of GPA
+## Proof of @prp-omega
 
@@
 
-```{=tex}
 \begin{align*}
@@
 \end{align*}
-```
+
 Clearly if $b(n)=c$ or $\lim_{n\rightarrow\infty}b(n)=c$ for some $c>0$ then $\Omega(F,n)=0$. Now consider a non-constant $b(n)$ and re-write $\Omega(F,n)$ as:
 
-```{=tex}
 \begin{align*}
-\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
-&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
+\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}\\&+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
+&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\&+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
 \end{align*}
-```
+
 Then if $\lim_{n\rightarrow\infty}b(n)=\infty$ it follows that:
 
-```{=tex}
 \begin{align*}
-\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
+\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} \\&- \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
 &\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
@@
 \end{align*}
-```
 
-### Removing the infinite sum
+## Derivation of @eq-rho
 
@@
 \end{align*}
-```
+
 Now using the fact that:
@@
 \end{align*}
-```
+
 In order to simplify the infinite sum, consider:
 
-```{=tex}
 \begin{align*}
-\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
+\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{B}(n+x,y)\\
 &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
@@
 \end{align*}
-```
+
 this infinite sum does not converge when $x\le1$ as each term is $O(n^{-x})$. We can now use this in $\hat\rho(\lambda)$:
 
-```{=tex}
 \begin{align*}
@@
 \end{align*}
-```
 
-# References
+\newpage
+
+# Additional Tables and Figures {.appendix}
+
+\newpage
+
+# References {.unnumbered}
+
 <!-- the pattern below controls the placement of the references -->
-:::{#refs}
+
+::: {#refs}
 :::

### 2025-05-21 cpp_funcs.cpp

@@ -1,587 +1,0 @@
-#include <RcppArmadillo.h>
-
-// [[Rcpp::depends(RcppArmadillo)]]
-
-
-using namespace Rcpp;
-using namespace arma;
-
-
-NumericVector subsetNumVec(NumericVector x, IntegerVector index) {  
-  // Length of the index vector
-  int n = index.size();
-  // Initialize output vector
-  NumericVector out(n);
-  
-  // Subtract 1 from index as C++ starts to count at 0
-  // Loop through index vector and extract values of x at the given positions
-  for (int i = 0; i < n; i++) {
-    out[i] = x[index[i]];
-  }
-  
-  // Return output
-  return out;
-}
-
-
-double rho(double lambda, NumericVector pars){
-  double a = pars[0];
-  double eps = pars[1];
-  double k0 = pars[2];
-  double b = pars[3];
-  
-  IntegerVector k0_vec_int = seq_len(k0);
-  NumericVector k0_vec = as<NumericVector>(k0_vec_int);
-  NumericVector logfracs = log(pow(k0_vec-1,a) + eps) -
-    log(lambda + pow(k0_vec-1, a)+eps);
-  NumericVector plogsums = cumsum(logfracs);
-  
-  double out = sum(exp(plogsums)) + (pow(k0,a)+eps)*exp(plogsums(k0-1))/(lambda-b);
-  return(out);
-}
-
-
-double rho_adj(double lambda, NumericVector pars){
-  return 1.0-rho(lambda, pars);
-}
-
-double find_lambda(NumericVector pars){
-  
-  double upper = 20;
-  double tol = 1e-7;
-  double lower = pars[3]+tol;
-  int max_iter=100;
-  double mid;
-  
-  for(int i;i<=max_iter;i++){
-    mid= (lower+upper)/2.0;
-    double f_mid = rho_adj(mid, pars);
-    if(std::abs(f_mid)<tol){
-      break;
-    }else if(f_mid>0){
-      upper=mid;
-    }else{
-      lower=mid;
-    }
-  }
-  
-  return mid;
-}
-
-
-
-
-
-
-// [[Rcpp::export]]
-Rcpp::NumericMatrix rmvnorm_rcpp(int n, Rcpp::NumericVector mean, Rcpp::NumericMatrix sigma) {
-  int p = mean.size();
-  Rcpp::NumericMatrix result(n, p);
-  
[…]

### 2025-05-21 data_fit_background.R

@@ -1,295 +1,0 @@
-source('funcs.R')
-library(ggplot2)
-library(crandep)
-library(data.table)
-data_raw = read.csv('degs.csv')
-nms = unique(data_raw$name)
-selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
-fits = list()
-mixfits = list()
-plots = list()
-dat_list = list()
-trunc.at = 3
-
-
-
-data.files = list.files('./data') 
-nms = tstrsplit(data.files,'out.',keep=2)[[1]]
-
-saveRDS(nms, 'nms.rds')
-
-
-i=4
-message(nms[i])
-df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
-df[,1] = as.numeric(as.character(df[,1]))
-dat_list[[i]] = as.matrix(df)
-fits[[i]] = fit_model(dat_list[[i]],N=1e5,quiet=F,
-                      burn_in=1e4,thin.by=5,trunc.at = 4,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
-
-
-
-for(i in 2:2){
-  message(nms[i])
-  df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
-  df[,1] = as.numeric(as.character(df[,1]))
-  dat_list[[i]] = as.matrix(df)
-  fits[[i]] = fit_model(dat_list[[i]],N=1e5,quiet=F,
-                        burn_in=1e4,thin.by=5,trunc.at = trunc.at,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
-}
-
-
-saveRDS(dat_list, 'dat_list.rds')
-
-dat_list = readRDS('dat_list.rds')
-
-plots = list()
-
-for(i in 1:length(fits)){
-  udf = as.data.frame(table(fits[[i]]$smps$k0))
-  udf$Var1 = as.numeric(as.character(udf$Var1))
-  sdf = twbfn::deg_surv(counts_to_degs(dat_list[[i]]))
-  ptrunc = sdf$surv[sdf$degree==min(fits[[i]]$dat[,1])]
-  plots[[i]] = ggplot() + scale_x_log10(limits = c(1, max(dat_list[[i]][,1]))) + scale_y_log10(limits=c(1/sum(dat_list[[i]][,2]),1)) +
[…]

### 2025-05-21 funcs.R

@@ -1,263 +1,0 @@
-require(Rcpp)
-require(RcppArmadillo)
-
-sourceCpp('cpp_funcs.cpp')
-
-f = Vectorize(function(x, g, lambda, k0, b){
-  if(x>=k0){
-    k0s = 0:(k0-1)
-    phi_g = lambda*gamma((lambda+g(k0))/b)*prod(g(k0s)/(lambda+g(k0s)))/
-      (b*gamma(lambda/b + 1)*gamma(g(k0)/b))
-    return(phi_g*beta(g(k0)/b + x-k0, lambda/b + 1))
-  }
-  if(x>0){
-    ks = 0:(x-1)
-    return(lambda/(lambda+g(x))*prod(g(ks)/(lambda+g(ks))))
-  }
-  if(x==0){
-    return(lambda/(lambda+g(x)))
-  }
-}, vectorize.args = 'x')
-S = Vectorize(function(x, g, lambda, k0, b){
-  if(x<0){
-    return(1)
-  }
-  if(x>=k0){
-    k0s = 0:(k0-1)
-    
-    phi_g = lgamma((lambda+g(k0))/b)  + sum(log(g(k0s)/(lambda+g(k0s))))-
-      (lgamma(lambda/b)+lgamma(g(k0)/b))
-    
-    return(exp(phi_g+lbeta(g(k0)/b + x-k0+1, lambda/b)))
-  }
-  if(x>0){
-    ks = 0:(x)
-    return(prod(g(ks)/(lambda+g(ks))))
-  }
-  if(x==0){
-    return(g(x)/(lambda+g(x)))
-  }
-  
-},vectorize.args = 'x')
-# sample_gpa_tree = function(n, g,b,k0,m=1){
-#   degs = c(0)
-#   prefs = c(g(0))
-#   for(i in 1:n){
-#     # message(i)
-#     selected = sample(1:length(degs), min(m, length(degs)), prob=prefs)
-#     degs = c(degs, 0)
-#     prefs = c(prefs, g(0))
-#     degs[selected] = degs[selected]+1
-#     prefs[selected] = ifelse(degs[selected]<k0, g(degs[selected]), g(k0) + b*(degs[selected]-k0))
-#   }
-#   return(degs)
-# }
-sample_gpa_tree = function(n, g, b, k0, m=1,quiet=T){
-  df = data.frame(deg = c(0),
-                  count = c(1),
-                  pref = c(g(0)))
-  for(i in 1:n){
-    if(!quiet){
-      message(i)
-    }
[…]

### 2025-05-21 gpd_to_pref.R

@@ -1,405 +1,0 @@
-logdiff <- function(l1, l2) { l1 + VGAM::log1mexp(l1-l2); }
-digp_log = function(x, shape, scale, threshold){
-  out=x
-  v=threshold
-  sig=scale
-  xi=shape
-  p1 = pmax(x[x>v]*0, 1 + xi * (x[x>v]+1-v)/sig)
-  p2 = pmax(x[x>v]*0, 1 + xi * (x[x>v]-v)/sig)
-  pows2 = pows1 = x[x>v]*0 + 1
-  pows1[p1>0] = -1/xi
-  pows2[p2>0] = -1/xi
-  # out[x>v] =  ( p2^pows2 - p1^pows1)
-  out[x>v] = logdiff(p2^pows2,p1^pows1)
-  out[x<=v] = 0
-  return(out)
-}
-igp_ll_neg = function(pars,dat,threshold=0){
-  shape = pars[1]
-  scale = pars[2]
-  if(shape==0){
-    return(1e9)
-  }
-  if(shape<0 & max(dat[,1]+1)<(threshold - scale/shape)){
-    return(1e9)
-  }
-  out = -sum(dat[,2]*digp_log(dat[,1], shape, scale, threshold))
-  if(is.infinite(out)){
-    return(1e9)
-  }else{
-    return(out)
-  }
-}
-fit.igp = function(dat, threshold=0){
-  return(optim(c(0.5, 1), fn = igp_ll_neg,dat=dat[dat[,1]>threshold,], threshold=threshold,
-               lower=c(-Inf,0.001), upper=c(Inf, Inf),method='L-BFGS-B'))
-}
-
-ft = fit.igp(fit$dat,10)
-
-
-degs = counts_to_degs(fit$dat)
-
-
-p = digp(degs[degs>10], -0.001, 1, 10)
-
-max(p)
-min(p)
-
-
-which.min(p)
-
-degs[degs>10][2645]
-
-
-# -------------------------------------------------------------------------
-
-
-igpd_rho = function(pars, shape, scale, threshold, exceedance){
-  p2 = shape*exceedance / (1-shape)
-  x = 1:(threshold+1)
-  p1 = sum(exp(cumsum(
-    log(x^pars[1] - (threshold+1)^pars[1] + pars[2]*scale)-
-      log(x^pars[1] - (threshold+1)^pars[1] + pars[2]*(scale+1))
-  )))
-  return(p1 + p2)
-}
-
[…]

### 2025-05-21 real_data_fitting.R

@@ -1,435 +1,0 @@
-source('funcs.R')
-library(ggplot2)
-data_raw = read.csv('degs.csv')
-nms = unique(data_raw$name)
-selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
-fits = list()
-plots = list()
-for(i in 1:length(selected)){
-  print(i)
-  fits[[i]] = fit_model(data_raw[data_raw$name==nms[selected[i]],2:3],N=5e4,quiet=F,
-                        trunc.at = 5,k0_jump = 1)
-}
-
-
-saveRDS(fits, 'real_mcmc.rds')
-
-
-for(i in 1:length(fits)){
-  plots[[i]] = ggplot() + scale_x_log10() + scale_y_log10() +
-    geom_point(data = twbfn::deg_surv(counts_to_degs(fits[[i]]$dat)),aes(x=degree, y=surv))+
-    geom_line(aes(x=!!fits[[i]]$dat[,1],y=!!fits[[i]]$surv$est),colour='red') + 
-    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[1,]),linetype=2, colour='red') + 
-    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[2,]),,linetype=2,colour='red')+
-    ggtitle(nms[selected[i]])
-}
-
-
-real_survs = ggpubr::ggarrange(plotlist = plots)
-
-saveRDS(real_survs,'real_survs.rds')
-
-
-
-
-real_survs
-# igpd --------------------------------------------------------------------
-
-
-fits = readRDS('real_mcmc.rds')
-
-gpfit = mev::fit.gpd(counts_to_degs(fits[[1]]$dat),threshold=6)
-
-scale = gpfit$estimate[1]
-shape = gpfit$estimate[2]
-
-x = fits[[1]]$dat[,1]
-y = mev::pgp(x,loc=6,scale=scale,shape=shape,lower.tail = F)
-degs = counts_to_degs(fits[[1]]$dat)
-
-plot(x,y,log='xy',type='l')
-points(twbfn::deg_surv(degs[degs>6]))
-
-# -------------------------------------------------------------------------
-
-data_raw = read.csv('degs.csv')
-nms = unique(data_raw$name)
-selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
-
[…]

### 2025-05-21 recovery.R

@@ -1,718 +1,0 @@
-f = Vectorize(function(x, g, lambda, k0, b){
-  if(x>=k0){
-    k0s = 0:(k0-1)
-    phi_g = lambda*gamma((lambda+g(k0))/b)*prod(g(k0s)/(lambda+g(k0s)))/
-      (b*gamma(lambda/b + 1)*gamma(g(k0)/b))
-    return(phi_g*beta(g(k0)/b + x-k0, lambda/b + 1))
-  }
-  if(x>0){
-    ks = 0:(x-1)
-    return(lambda/(lambda+g(x))*prod(g(ks)/(lambda+g(ks))))
-  }
-  if(x==0){
-    return(lambda/(lambda+g(x)))
-  }
-}, vectorize.args = 'x')
-S = Vectorize(function(x, g, lambda, k0, b){
-  if(x>=k0){
-    k0s = 0:(k0-1)
-    
-    phi_g = lgamma((lambda+g(k0))/b)  + sum(log(g(k0s)/(lambda+g(k0s))))-
-      (lgamma(lambda/b)+lgamma(g(k0)/b))
-    
-    return(exp(phi_g)*beta(g(k0)/b + x-k0+1, lambda/b))
-  }
-  if(x>0){
-    ks = 0:(x)
-    return(prod(g(ks)/(lambda+g(ks))))
-  }
-  if(x==0){
-    return(g(x)/(lambda+g(x)))
-  }
-  
-},vectorize.args = 'x')
-sample_gpa_tree = function(n, g,b,k0,m=1){
-  degs = c(0)
-  prefs = c(g(0))
-  for(i in 1:n){
-    # message(i)
-    selected = sample(1:length(degs), min(m, length(degs)), prob=prefs)
-    degs = c(degs, 0)
-    prefs = c(prefs, g(0))
-    degs[selected] = degs[selected]+1
-    prefs[selected] = ifelse(degs[selected]<k0, g(degs[selected]), g(k0) + b*(degs[selected]-k0))
-  }
-  return(degs)
-}
-rho = Vectorize(function(lambda, g, b, k0){
-  n1 = 1:k0
-  n2 = 0:(k0-1)
-  out1 = sum(exp(cumsum(log(1-lambda/(lambda+g(n1-1))))))
-  out2 = exp(sum(log(1-lambda/(lambda+g(n2)))))
-  return(out1 + g(k0)*out2/(lambda-b))
-}, vectorize.args = 'lambda')
[…]

### 2025-05-21 recreating.R

@@ -1,226 +1,0 @@
-sample_rnd = function(n, pref, gamma_g,quiet = T){
-  degs = c(0)
-  gammas = c(gamma_g(1))
-  prefs = c(pref(degs[1], gammas[1]))
-  for(i in 2:n){
-    if(!quiet) message(i)
-    selected = sample(1:(i-1), 1, prob = prefs)
-    degs[selected] = degs[selected]+1
-    prefs[selected] = pref(degs[selected], gammas[selected])
-    degs = c(degs, 0)
-    gammas = c(gammas, gamma_g(1))
-    prefs  = c(prefs, pref(degs[i], gammas[i]))
-  }
-  return(data.frame(degree=degs, gamma = gammas, pref = prefs))
-}
-
-
-# -------------------------------------------------------------------------
-library(ggplot2)
-
-k0 = 20
-a = 1
-eps = 1
-
-
-pref = function(x, gamma){
-  return(ifelse(x<=k0, x^a+gamma[,2], k0^a + eps + gamma[,1]*(x-k0)))
-}
-gamma_g = function(n){
-  return(rgamma(n*2, 1, 10))
-}
-
-n=1e4
-G = sample_rnd(n, pref, gamma_g, quiet = F)
-degs =  G$degree
-par(mfrow=c(2,2))
-plot(twbfn::deg_dist(degs[degs>0]), log='xy',pch=20,ylim = c(1/n,1), xlim=c(1,n+1))
-plot(twbfn::deg_surv(degs[degs>0]), log='xy',pch=20,ylim = c(1/n,1), xlim=c(1,n+1))
-# plot(density(G$gamma))
-plot(G$degree+1, G$pref,pch=20, col = scales::alpha('blue', 0.3), log='xy', ylim=c(0.01,1e3))
-
-# mcmc network ------------------------------------------------------------
-
-source('funcs.R')
-
-fits = readRDS('modelfits.rds')
-
-
-# -------------------------------------------------------------------------
-
-j = 1
-mcmcdat = fits[[j]]$smps
-
-N = 1e4
-par_rows = sample(1:nrow(mcmcdat),N,replace=T)
-par_mat = mcmcdat[par_rows, ]
[…]

### 2025-05-21 scripts/cpp_funcs.cpp

@@ -1,0 +1,587 @@
+#include <RcppArmadillo.h>
+
+// [[Rcpp::depends(RcppArmadillo)]]
+
+
+using namespace Rcpp;
+using namespace arma;
+
+
+NumericVector subsetNumVec(NumericVector x, IntegerVector index) {  
+  // Length of the index vector
+  int n = index.size();
+  // Initialize output vector
+  NumericVector out(n);
+  
+  // Subtract 1 from index as C++ starts to count at 0
+  // Loop through index vector and extract values of x at the given positions
+  for (int i = 0; i < n; i++) {
+    out[i] = x[index[i]];
+  }
+  
+  // Return output
+  return out;
+}
+
+
+double rho(double lambda, NumericVector pars){
+  double a = pars[0];
+  double eps = pars[1];
+  double k0 = pars[2];
+  double b = pars[3];
+  
+  IntegerVector k0_vec_int = seq_len(k0);
+  NumericVector k0_vec = as<NumericVector>(k0_vec_int);
+  NumericVector logfracs = log(pow(k0_vec-1,a) + eps) -
+    log(lambda + pow(k0_vec-1, a)+eps);
+  NumericVector plogsums = cumsum(logfracs);
+  
+  double out = sum(exp(plogsums)) + (pow(k0,a)+eps)*exp(plogsums(k0-1))/(lambda-b);
+  return(out);
+}
+
+
+double rho_adj(double lambda, NumericVector pars){
+  return 1.0-rho(lambda, pars);
+}
+
+double find_lambda(NumericVector pars){
+  
+  double upper = 20;
+  double tol = 1e-7;
+  double lower = pars[3]+tol;
+  int max_iter=100;
+  double mid;
+  
+  for(int i;i<=max_iter;i++){
+    mid= (lower+upper)/2.0;
+    double f_mid = rho_adj(mid, pars);
+    if(std::abs(f_mid)<tol){
+      break;
+    }else if(f_mid>0){
+      upper=mid;
+    }else{
+      lower=mid;
+    }
+  }
+  
+  return mid;
+}
+
+
+
+
+
+
+// [[Rcpp::export]]
+Rcpp::NumericMatrix rmvnorm_rcpp(int n, Rcpp::NumericVector mean, Rcpp::NumericMatrix sigma) {
+  int p = mean.size();
+  Rcpp::NumericMatrix result(n, p);
+  
[…]

### 2025-05-21 scripts/data_fit_background.R

@@ -1,0 +1,295 @@
+source('funcs.R')
+library(ggplot2)
+library(crandep)
+library(data.table)
+data_raw = read.csv('degs.csv')
+nms = unique(data_raw$name)
+selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
+fits = list()
+mixfits = list()
+plots = list()
+dat_list = list()
+trunc.at = 3
+
+
+
+data.files = list.files('./data') 
+nms = tstrsplit(data.files,'out.',keep=2)[[1]]
+
+saveRDS(nms, 'nms.rds')
+
+
+i=4
+message(nms[i])
+df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
+df[,1] = as.numeric(as.character(df[,1]))
+dat_list[[i]] = as.matrix(df)
+fits[[i]] = fit_model(dat_list[[i]],N=1e5,quiet=F,
+                      burn_in=1e4,thin.by=5,trunc.at = 4,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
+
+
+
+for(i in 2:2){
+  message(nms[i])
+  df  =as.data.frame(table(table(fread(paste0('./data/', data.files[i]))[,2])))
+  df[,1] = as.numeric(as.character(df[,1]))
+  dat_list[[i]] = as.matrix(df)
+  fits[[i]] = fit_model(dat_list[[i]],N=1e5,quiet=F,
+                        burn_in=1e4,thin.by=5,trunc.at = trunc.at,k0_jump = 2,lookback = 1e3, cov_scale=0.05)
+}
+
+
+saveRDS(dat_list, 'dat_list.rds')
+
+dat_list = readRDS('dat_list.rds')
+
+plots = list()
+
+for(i in 1:length(fits)){
+  udf = as.data.frame(table(fits[[i]]$smps$k0))
+  udf$Var1 = as.numeric(as.character(udf$Var1))
+  sdf = twbfn::deg_surv(counts_to_degs(dat_list[[i]]))
+  ptrunc = sdf$surv[sdf$degree==min(fits[[i]]$dat[,1])]
+  plots[[i]] = ggplot() + scale_x_log10(limits = c(1, max(dat_list[[i]][,1]))) + scale_y_log10(limits=c(1/sum(dat_list[[i]][,2]),1)) +
[…]

### 2025-05-21 scripts/funcs.R

@@ -1,0 +1,263 @@
+require(Rcpp)
+require(RcppArmadillo)
+
+sourceCpp('scripts/cpp_funcs.cpp')
+
+f = Vectorize(function(x, g, lambda, k0, b){
+  if(x>=k0){
+    k0s = 0:(k0-1)
+    phi_g = lambda*gamma((lambda+g(k0))/b)*prod(g(k0s)/(lambda+g(k0s)))/
+      (b*gamma(lambda/b + 1)*gamma(g(k0)/b))
+    return(phi_g*beta(g(k0)/b + x-k0, lambda/b + 1))
+  }
+  if(x>0){
+    ks = 0:(x-1)
+    return(lambda/(lambda+g(x))*prod(g(ks)/(lambda+g(ks))))
+  }
+  if(x==0){
+    return(lambda/(lambda+g(x)))
+  }
+}, vectorize.args = 'x')
+S = Vectorize(function(x, g, lambda, k0, b){
+  if(x<0){
+    return(1)
+  }
+  if(x>=k0){
+    k0s = 0:(k0-1)
+    
+    phi_g = lgamma((lambda+g(k0))/b)  + sum(log(g(k0s)/(lambda+g(k0s))))-
+      (lgamma(lambda/b)+lgamma(g(k0)/b))
+    
+    return(exp(phi_g+lbeta(g(k0)/b + x-k0+1, lambda/b)))
+  }
+  if(x>0){
+    ks = 0:(x)
+    return(prod(g(ks)/(lambda+g(ks))))
+  }
+  if(x==0){
+    return(g(x)/(lambda+g(x)))
+  }
+  
+},vectorize.args = 'x')
+# sample_gpa_tree = function(n, g,b,k0,m=1){
+#   degs = c(0)
+#   prefs = c(g(0))
+#   for(i in 1:n){
+#     # message(i)
+#     selected = sample(1:length(degs), min(m, length(degs)), prob=prefs)
+#     degs = c(degs, 0)
+#     prefs = c(prefs, g(0))
+#     degs[selected] = degs[selected]+1
+#     prefs[selected] = ifelse(degs[selected]<k0, g(degs[selected]), g(k0) + b*(degs[selected]-k0))
+#   }
+#   return(degs)
+# }
+sample_gpa_tree = function(n, g, b, k0, m=1,quiet=T){
+  df = data.frame(deg = c(0),
+                  count = c(1),
+                  pref = c(g(0)))
+  for(i in 1:n){
+    if(!quiet){
+      message(i)
[…]

### 2025-05-21 scripts/gpd_to_pref.R

@@ -1,0 +1,405 @@
+logdiff <- function(l1, l2) { l1 + VGAM::log1mexp(l1-l2); }
+digp_log = function(x, shape, scale, threshold){
+  out=x
+  v=threshold
+  sig=scale
+  xi=shape
+  p1 = pmax(x[x>v]*0, 1 + xi * (x[x>v]+1-v)/sig)
+  p2 = pmax(x[x>v]*0, 1 + xi * (x[x>v]-v)/sig)
+  pows2 = pows1 = x[x>v]*0 + 1
+  pows1[p1>0] = -1/xi
+  pows2[p2>0] = -1/xi
+  # out[x>v] =  ( p2^pows2 - p1^pows1)
+  out[x>v] = logdiff(p2^pows2,p1^pows1)
+  out[x<=v] = 0
+  return(out)
+}
+igp_ll_neg = function(pars,dat,threshold=0){
+  shape = pars[1]
+  scale = pars[2]
+  if(shape==0){
+    return(1e9)
+  }
+  if(shape<0 & max(dat[,1]+1)<(threshold - scale/shape)){
+    return(1e9)
+  }
+  out = -sum(dat[,2]*digp_log(dat[,1], shape, scale, threshold))
+  if(is.infinite(out)){
+    return(1e9)
+  }else{
+    return(out)
+  }
+}
+fit.igp = function(dat, threshold=0){
+  return(optim(c(0.5, 1), fn = igp_ll_neg,dat=dat[dat[,1]>threshold,], threshold=threshold,
+               lower=c(-Inf,0.001), upper=c(Inf, Inf),method='L-BFGS-B'))
+}
+
+ft = fit.igp(fit$dat,10)
+
+
+degs = counts_to_degs(fit$dat)
+
+
+p = digp(degs[degs>10], -0.001, 1, 10)
+
+max(p)
+min(p)
+
+
+which.min(p)
+
+degs[degs>10][2645]
+
+
+# -------------------------------------------------------------------------
+
+
+igpd_rho = function(pars, shape, scale, threshold, exceedance){
+  p2 = shape*exceedance / (1-shape)
+  x = 1:(threshold+1)
+  p1 = sum(exp(cumsum(
+    log(x^pars[1] - (threshold+1)^pars[1] + pars[2]*scale)-
+      log(x^pars[1] - (threshold+1)^pars[1] + pars[2]*(scale+1))
+  )))
+  return(p1 + p2)
+}
+
[…]

### 2025-05-21 scripts/real_data_fitting.R

@@ -1,0 +1,435 @@
+source('funcs.R')
+library(ggplot2)
+data_raw = read.csv('degs.csv')
+nms = unique(data_raw$name)
+selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
+fits = list()
+plots = list()
+for(i in 1:length(selected)){
+  print(i)
+  fits[[i]] = fit_model(data_raw[data_raw$name==nms[selected[i]],2:3],N=5e4,quiet=F,
+                        trunc.at = 5,k0_jump = 1)
+}
+
+
+saveRDS(fits, 'real_mcmc.rds')
+
+
+for(i in 1:length(fits)){
+  plots[[i]] = ggplot() + scale_x_log10() + scale_y_log10() +
+    geom_point(data = twbfn::deg_surv(counts_to_degs(fits[[i]]$dat)),aes(x=degree, y=surv))+
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y=!!fits[[i]]$surv$est),colour='red') + 
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[1,]),linetype=2, colour='red') + 
+    geom_line(aes(x=!!fits[[i]]$dat[,1],y= !!fits[[i]]$surv$CI[2,]),,linetype=2,colour='red')+
+    ggtitle(nms[selected[i]])
+}
+
+
+real_survs = ggpubr::ggarrange(plotlist = plots)
+
+saveRDS(real_survs,'real_survs.rds')
+
+
+
+
+real_survs
+# igpd --------------------------------------------------------------------
+
+
+fits = readRDS('real_mcmc.rds')
+
+gpfit = mev::fit.gpd(counts_to_degs(fits[[1]]$dat),threshold=6)
+
+scale = gpfit$estimate[1]
+shape = gpfit$estimate[2]
+
+x = fits[[1]]$dat[,1]
+y = mev::pgp(x,loc=6,scale=scale,shape=shape,lower.tail = F)
+degs = counts_to_degs(fits[[1]]$dat)
+
+plot(x,y,log='xy',type='l')
+points(twbfn::deg_surv(degs[degs>6]))
+
+# -------------------------------------------------------------------------
+
+data_raw = read.csv('degs.csv')
+nms = unique(data_raw$name)
+selected = c(2,3,5,6,7,8,9,11,12,13,17,22)
+
[…]

### 2025-05-21 scripts/recovery.R

@@ -1,0 +1,718 @@
+f = Vectorize(function(x, g, lambda, k0, b){
+  if(x>=k0){
+    k0s = 0:(k0-1)
+    phi_g = lambda*gamma((lambda+g(k0))/b)*prod(g(k0s)/(lambda+g(k0s)))/
+      (b*gamma(lambda/b + 1)*gamma(g(k0)/b))
+    return(phi_g*beta(g(k0)/b + x-k0, lambda/b + 1))
+  }
+  if(x>0){
+    ks = 0:(x-1)
+    return(lambda/(lambda+g(x))*prod(g(ks)/(lambda+g(ks))))
+  }
+  if(x==0){
+    return(lambda/(lambda+g(x)))
+  }
+}, vectorize.args = 'x')
+S = Vectorize(function(x, g, lambda, k0, b){
+  if(x>=k0){
+    k0s = 0:(k0-1)
+    
+    phi_g = lgamma((lambda+g(k0))/b)  + sum(log(g(k0s)/(lambda+g(k0s))))-
+      (lgamma(lambda/b)+lgamma(g(k0)/b))
+    
+    return(exp(phi_g)*beta(g(k0)/b + x-k0+1, lambda/b))
+  }
+  if(x>0){
+    ks = 0:(x)
+    return(prod(g(ks)/(lambda+g(ks))))
+  }
+  if(x==0){
+    return(g(x)/(lambda+g(x)))
+  }
+  
+},vectorize.args = 'x')
+sample_gpa_tree = function(n, g,b,k0,m=1){
+  degs = c(0)
+  prefs = c(g(0))
+  for(i in 1:n){
+    # message(i)
+    selected = sample(1:length(degs), min(m, length(degs)), prob=prefs)
+    degs = c(degs, 0)
+    prefs = c(prefs, g(0))
+    degs[selected] = degs[selected]+1
+    prefs[selected] = ifelse(degs[selected]<k0, g(degs[selected]), g(k0) + b*(degs[selected]-k0))
+  }
+  return(degs)
+}
+rho = Vectorize(function(lambda, g, b, k0){
+  n1 = 1:k0
+  n2 = 0:(k0-1)
+  out1 = sum(exp(cumsum(log(1-lambda/(lambda+g(n1-1))))))
+  out2 = exp(sum(log(1-lambda/(lambda+g(n2)))))
+  return(out1 + g(k0)*out2/(lambda-b))
+}, vectorize.args = 'lambda')
[…]

### 2025-05-21 testing.R

@@ -1,14 +1,0 @@
-source('funcs.R')
-
-
-out = recover_params(1.2, 1.0, 0.5, 25.0,quiet=F,n.iter=1e4)
-
-
-
-
-
-
-
-
-
-

### 2025-05-21 paper.tex
Updating draft

@@
 \newtheorem{proposition}{Proposition}[section]
+\theoremstyle{plain}
+\newtheorem{theorem}{Theorem}[section]
 \theoremstyle{remark}
@@
 
-\section{Introduction}\label{sec-intro}
+\newpage
+
+\section{Introduction}\label{introduction}
 
@@
 investigate the patterns of connections in neural systems
-\citep{Betzel17}. One such mechanistic model is the general preferential
-attachment (PA) model where as a network grows and new vertices join
-they connect to those already present in the network with weights
-proportional to some preference function \(b\) of their degree. The
-well-known Barabási-Albert (BA) model is a special case of this when
-\(b\sim k\) {[}BA99{]}.
+\citep{Betzel17}.
 
-Mechanistic models, like the PA model, are in general only particularly
-useful when simulating networks and not fitting to the real data as
-usually only snapshots of real networks are available and not the entire
-evolution. As such, one of the most informative statistics when
-regarding a network is the degree distribution. When looking at the
-degree distributions of real networks it is attractive to many to use
-the power law to model the degree distribution, as the BA model
-generates degrees that follow a power law, and therefore claim that
-since the degree distribution is roughly power law then the underlying
-growth mechanism may be preferential attachment such as the BA model.
-However, it has been shown that a general preference function will not
-always lead to a power law degree distribution \citep{krapivsky01} and
-additionally it has been debated for decades whether most real data
-actually follows a power law \citep{Voitalov_2019} or
-not\citep{Broido_2019}.
+One example of a mechanistic model that may be considered is the
+preferential attachment (PA) model where as a network grows and new
+vertices join they connect to those already present in the network with
+weights proportional to some preference function \(b(\cdot)\) of their
+in-degree. The well-known Barabási-Albert (BA) model is a special case
+of this when \(b(k) = k+1\) \citep{Barabasi99}.
 
-Throughout this debate the use of methods from extremes has been absent
-until relatively recently with the papers such as \citep{Voitalov_2019}
-and {[}Lee{]}. Forgoing the use of extreme value methods essentially
-downplays how deviations from the power law in the largest degrees may
-affect the network given that they are often much more influential than
-the vertices with smaller degrees.
+These mechanistic models are generally useful when simulating networks,
+but not for fitting to real data as usually only snapshots of a network
+are available and not the full evolution. However, one statistic that is
+usually available for snapshots of networks is the degree distribution
+making it valuable to study the degrees in a network. Often, when
+looking at the degree distributions of real networks it can be
+attractive to use the power law when modelling as it is seemingly
+observed in much of the data. It also comes with the added benefit of
+offering a connection to preferential attachment since the BA model
+results in a power law degree distribution \citep{Barabasi99}, perhaps
+suggesting that if the power law fits well then the network may have
+grown according to similar simple rules. However, attempting to explain
+the networks growth using preferential attachment when the degrees
+seemingly follow a power law may not be the best course of action as
+\citep{krapivsky01} shows that a general preference function does not
+lead to power law degrees and additionally, it is still being argued
+whether or not a lot of real data actually follows a power law.
+\citet{Broido_2019} provides some evidence that suggests that real world
+networks rarely have degrees that follow a power law through the use of
+statistical analysis of almost 1000 networks across a wide variety of
+fields; \citet{Voitalov_2019}, on the other hand, disagrees and claims
+that these kinds of networks are not nearly as rare and only appear so
+as a result of an unrealistic expectation of power law without
+deviations or noise.
 
-Noting this, many have now started to use these extreme value methods
-when analysing networks. \citep{wan19} showed that extreme value
-estimation methods prove more robust than existing parametric approaches
-especially when the data are corrupted or the model has been
-misspecified, demonstrating that extreme value methods can be a valuable
-tool when studying networks. \citep{wang2022random} investigates a PA
-model with a affine preference function (similar to the BA model) with
-the added behaviour of heterogeneous reciprocal edges, they show that
-the distribution of in-degrees and out-degrees produced by such a model
-is multivariate regularly varying and they also have hidden regular
-variation. {[}Lee{]} using methods from discrete extremes modelled the
-right tail of network degrees with a discretised variation of the
-generalised Pareto distribution (GPD) usually dubbed the Integer GPD
-(IGPD) where \(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by
-the survival:
+Extreme value methods have been scarcely used when it comes to analysis
+of degree distributions and whether they follow the power law. The
+absence of these methods essentially down weights the deviation of
+vertices with the largest degrees, which are usually the most
+influential, from the power law. Somewhat recent publications
+\citep{Voitalov_2019, Lee24, wang2022random} have addressed this gap in
+the literature. \citet{Voitalov_2019} proposes estimating the power law
+exponent of a degree distribution using three estimators based in
+extreme value theory, in contrast to the estimators generally used in
+network science, the degree distributions are then split into one of
+three categories (not power-law , hardly power-law or power-law)
+depending on the values of the estimates obtained from each of the
+estimators. \citet{Lee24} proposes a framework for modelling entire
+degree distributions using a combination of two generalisations of the
+power law including a model selection step that is capable of suggesting
+whether or not the data is adequately modelled by a power law, with one
+conclusion being that most data seems to have a tail lighter than what
+is implied by the body whilst still being heavy tailed.
+\citet{wang2022random} considers a preferential attachment model with an
+affine preference function and reciprocal edges, they study the limiting
+degree distribution of this model for both the in-degrees and
+out-degrees - finding that the joint limiting distribution is
+multivariate regularly varying and has the property of hidden regular
+variation.
 
-\[
-\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).
+Evidently, the study of degree distributions is a very active but the
+vast majority of research in the field (whether using extremes or not)
+are purely descriptive in the sense that no information about the
+preference function is revealed even if preferential attachment was the
+underlying mechanism. This paper aims to address this gap - asking if
+given the degree distribution of a network assumed to come from the
+preferential attachment model, can we directly infer the model
+parameters?
 
-The study of degree distributions, whether using extremes or not,
-usually does not reveal any information about the preference function
-even if PA was the underlying generating mechanism of the network. Here
-we address this gap in the literature, more specifically investigating
-if given the degree distribution of a network assumed to come from the
-PA model can we directly infer the model parameters? We build upon the
-foundation provided by \citep{rudas07} deriving the limiting degree
-distribution in terms of the preference function and suggest a class of
-preference functions that result in realistic tails that are heavy but
-are lighter than implied by the power law.
+\citet{rudas07} derives the limiting degree distribution in terms of the
+preference function \(b(\cdot)\) and its parameters, we build upon this
+by finding a class of preference functions that result in realistic
+tails i.e.~heavy but not as heavy as implied by the power law.
 
-\subsection{Structure of this paper}\label{structure-of-this-paper}
+The remainder of this paper begins with a more detailed description of
+the PA model alongside the theoretical results for the limiting degree
+distribution using results from \citet{rudas07}, this section
+(Section~\ref{sec-tail}) then goes on to analyse the tail behaviour of
+this limiting distribution in terms of the preference function
+\(b(\cdot)\) arguing that a heavy tailed degree distribution can only be
+the result of an asymptotically linear preference function - ending with
+a suggestion for a preference function that can guarantee heavy tails
+while remaining flexible. Section~\ref{sec-model} continues on from the
+previous section by considering the limiting distribution obtained when
+using a preference function of the form stated at the end of
+Section~\ref{sec-tail} providing examples of how the shape of the
+distribution varies with respect to the model parameters as well as
+visualisations for how the tail heaviness changes when these parameters
+are varied - continued study of this distribution then also reveals a
+connection to the general Pareto distribution (GPD).
+Section~\ref{sec-sim} consists of a simulation study where networks are
+simulated from the PA model for various parameter combinations and then
+the model specified in Section~\ref{sec-model} is fitted to the
+resulting degree distributions using Bayesian inference, demonstrating
+that (for simulated data) the model parameters can be recovered from
+only the degree distribution. Finally, Section~\ref{sec-real} fits the
+model to some real data and provides posterior estimates for the
+preference function assuming the networks evolved according to the PA
+model.
 
-\textbf{?@sec-gpa} looks at the theoretical limiting behaviour of a GPA
-model when \(m=1\) according to the results from \citep{rudas07}, going
-on to introduce a flexible class of preference functions that can
-guarantee a heavy tailed degree distribution while still being flexible
-in the body. Section~\ref{sec-rec} shows that the preference function
-parameters can be fairly well recovered from the degree distributions of
-networks simulated from the GPA model using various functions of the
-class introduced in \textbf{?@sec-gpa}. Having shown that for simulated
-data the model parameters are recoverable from the degree distribution
-alone, Section~\ref{sec-real} builds upon this and attempts to fit the
-model to degree distributions of real networks; modelling not only the
-degree distribution itself but also the estimates for the parameters of
-the preference function assuming the networks evolved according to the
-GPA scheme. Ultimately, Section~\ref{sec-conc} discusses the main
-results of the paper, pitfalls of this method and future work.
+\newpage
 
-\section{PA model}\label{pa-model}
+\section{Tail Behaviour of Preferential Attachment
+Model}\label{sec-tail}
 
@@
 f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
-\] We are specifically interested in how the tail heaviness (\(\xi\) in
-the IGPD) is affected by the preference function \(b\),
-\citep{shimura12} introduces a quantity that will help in determining
-what domain of attraction this discrete distribution belongs to, that
-is:
+\]
+
+We are specifically interested in how the tail heaviness (\(\xi\) in the
+IGPD) is affected by the preference function \(b\), \citep{shimura12}
+introduces a quantity that will help in determining what domain of
+attraction this discrete distribution belongs to, that is:
 
 For a distribution \(F\) with survival function \(\bar F\) and some
-\(n\in\mathbb Z^+\) let:
+\(k\in\mathbb Z^+\) let:
 
 \[
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
+\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
 \]
@@
 \citep{shimura12} then states that if
-\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
-Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
-distribution is light tailed. This allows us to show the following
+\(\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
+then \(F\) is heavy tailed with \(\bar F(k) \sim k^{-\alpha}\).
+Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,k) = 0\) then the
+distribution is light tailed. This allows us to show the following:
 
@@
 
-See \hyperref[appendix]{appendix} for the details of the proof.
+See appendix \ref{sec-proofs} for the details of the proof.
 
@@
 tailed, the preference function must be asymptotically linear
-i.e.~\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0\).
+i.e.~\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0\). First we must
+consider the following theorem:
 
-Consider the definition of heavy tails:
+\begin{theorem}[Stolz-Cesàro Theorem
+\citep{cesaro}]\protect\hypertarget{thm-stolz}{}\label{thm-stolz}
 
-\begin{align*}
-\lim_{k\rightarrow\infty}\Omega(F,n) &= c>0\\
-\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} &= c\\
-\lim_{k\rightarrow\infty}[b(k+1)-b(k)] &= \lambda^*c\\
-\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{(k+1)-(k)} &= \lambda^*c\\
-\end{align*}
-
-by Stolz-Cesaro Theorem:
+Let \((a_n)_{n\ge1}\) and \((b_n)_{n\ge1}\) be two sequence of real
+numbers. Assume \((a_n)_{n\ge1}\) is a strictly monotone and divergent
+sequence and the following limit exists:
 
 \[
-\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \lambda^*c > 0.
+\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
+\] Then,
+
+\[
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
 \]
 
-Since \(b\) must be asymptotically linear in order to have positive tail
-heaviness we propose the following preference function:
+\end{theorem}
+
+From Proposition~\ref{prp-omega} we know that in order for the limiting
+degree distribution to have heavy tails, we must have that:
+
+\[
+\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
+\] Now, setting \(b_k = b(k)\) and \(a_k = k\):
+
+\[
+\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
+\] meaning by Theorem~\ref{thm-stolz}:
+
+\[
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0.
+\] Therefore, in order for the degree distribution to be heavy tailed,
+the preference function must be asymptotically linear. With this under
+consideration we propose the following preference function:
 
@@
 
-\section{Preferential Attachment with flexible heavy
-tail}\label{sec-model}
+\newpage
+
+\section{Preferential Attachment with Flexible Heavy
+Tail}\label{sec-model}
 
@@
 
-\[
+\begin{equation}\phantomsection\label{eq-rho}{
 \hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}
-\]
+}\end{equation}
 
 which must be solved numerically for most parameter choices. Also, note
-that \(\lambda>\beta\) since when \(\lambda\le \beta\) the infinite sum
-no longer converges and instead goes to infinity.
+that \(\lambda>\beta\).
 
@@
 
-\begin{figure}[H]
+\begin{figure}
 
@@
 The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
-mentioned in Section~\ref{sec-intro} by using Stirling's approximation
-of the gamma function to obtain:
+by using Stirling's approximation of the gamma function, where the IGPD
+is a discretised variation of the generalised Pareto distribution (GPD)
+usually dubbed the Integer GPD (IGPD) where
+\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) defined by the survival: \[
+\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
+\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).
 
-\[
+Using Stirling's approximation we have that: \[
 \frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
-\] and utilising the expression from Equation~\ref{eq-polysurv}:
+\] now utilising the expression from Equation~\ref{eq-polysurv}:
 
@@
 
-\section{Recovery}\label{sec-rec}
+\section{Simulation Study}\label{sec-sim}
 
-The goal of was to provide a possible preference function that is able
-to explain the growth of real networks with varied tail behaviour in
-their degree distributions. This section aims to show that the
-parameters of the model in Section~\ref{sec-model} can be recovered from
-the degree distribution of a network simulated from it.
+This section aims to show that the parameters of the model in
+Section~\ref{sec-model} can be recovered from the degree distribution of
+a network simulated from it.
 
@@
 \(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
-counts in the form of a vector of degrees \(\pmb x = (1,2,\ldots,M)\)
-and the number of vertices with those degrees \(\pmb n\) where \(M\) is
-the maximum degree. However, when it comes to fitting this model to real
-data in the next section, small degrees are extremely influential and
-will often cause the model to have a bad fit even though the rest of the
-data looks similar to that of the simulations; truncating the data above
-some value, say \(l<k_0\), should allow the model to be more effectively
-fitted. This accounts in some way for the fact that this model only
-applies to trees, and could not possibly capture the behaviour for nodes
-with small degrees. Using the p.m.f derived in \citep{rudas07}, the
-likelihood given that \(x_i \ge l\) for all \(i =1,2,\ldots,M\) is:
-
-\begin{align*}
-L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{x_i\ge k_0}n_{x_i}\right)} \prod_{l \le x_i<k_0}\left(\frac{\lambda^*}{\lambda^* +x_i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{x_i\ge k_0}\left(\frac{\text{B}(x_i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
-\end{align*}
-
-where \(B(y,z)\) is the the beta function.
-
-This likelihood allows inference to be made about the parameters using a
-Bayesian approach using the priors:
+counts and using the likelihood from the previous section alongisde the
+priors:
 
@@
 
-\captionsetup{labelsep=none}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{mcmc_plot.jpg}
+\pandocbounded{\includegraphics[keepaspectratio]{images/mcmc_plot.jpg}}
 
@@
 
-\caption{: Posterior estimates of paramters for data simulated from the
-proposed model with various combinations of
+\caption{\label{fig-rec2}: Posterior estimates of paramters for data
+simulated from the proposed model with various combinations of
 (\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
-
-\end{figure}%
-
-}
-
-\caption{\label{fig-rec2}}
 
@@
 Turning now to real data, the goal is to fit the model to the degree
-distributions of real networks from various sources. {[}describe
-networks being used{]}. Alongside fitting this model to the degree
-distributions, we then compare the fit to that of an mixture
-distribution that was used in {[}clement{]} and is similar to others
-that have been used to model degree distributions {[}others{]}.
+distributions of real networks from various sources. Alongside fitting
+this model to the degree distributions, we then compare the fit to that
+of an mixture distribution that was used in \citet{Lee24}.
 
-\begin{figure}[H]
+\begin{figure}
 
@@
 
-\centering{
-
-\captionsetup{labelsep=none}
-
-\begin{figure}[H]
-
-{\centering \includegraphics{pars_plot.png}
+{\centering \pandocbounded{\includegraphics[keepaspectratio]{images/pars_plot.png}}
 
@@
 
-}
-
-\caption{\label{fig-rec2}}
-
-\end{figure}%
-
-\section{Conclusion and Discussion}\label{sec-conc}
+\section{Conclusion and Discussion}\label{conclusion-and-discussion}
 
@@
 
-GPA is also a fairly simplistic one on its own despite having large
-flexibility in the preference function allowing for edges to only be
-added at times when a new node is introduced into the network and not
-allowing for anything like vertex/edge death, reciprocal edges and
-rewiring steps. \citep{deijfen2015} introduces some theory for GPA with
-vertex death based on the work by \citep{rudas07}, but is unable to come
-to an expression for the degree distribution. However, adding something
-like fitness scores into the GPA model seems fairly simple to do even
-going so far as making each of the parameters used in the class of
-functions here node-wise instead of universal. For example having the
-preference weight for a vertex be given by
-\(b(k_i, \underline{\gamma_i})\) where \(\underline{\gamma_i}\sim G\) is
-the vertex's fitness and \(k_i\) is the vertex's degree, it is
-reasonable to expect that the degree distribution for a model like this
-to be given by:
+\setcounter{section}{0}
+\renewcommand{\thesection}{\Alph{section}}
+\setcounter{table}{0}
+\renewcommand{\thetable}{A\arabic{table}}
+\setcounter{figure}{0}
+\renewcommand{\thefigure}{A\arabic{figure}}
 
-\[
-\bar F(k) = \mathbb E_G\left[\prod_{i=0}^k \frac{b(i, \gamma)}{\lambda^* + b(i,\gamma)}\right],\qquad \gamma\sim G,\, k=0,1,2,\ldots
-\]
+\newpage
 
-If this were the case, the posterior distributions of parameters from
-Section~\ref{sec-real} could be interpreted as the approximate
-distributions from which a vertex's fitness values are drawn instead of
-providing estimates of the universal parameter values. Additionally, it
-allows the ability to set some parameters as universal and some as
-vertex-wise for example one could let \(\alpha, \beta, k_0\) be the same
-for all vertices but let each vertex in the network have its own ``zero
-appeal''
-\(\varepsilon_i\overset{\text{iid}}{\sim} \text{Ga}(\theta_1, \theta_2)\);
-moving the model to being more realistic and accounting for the fact
-that in many real networks the objects vertices represent have some
-value instrinsic to themselves other than the number of connections they
-have.
+\section{Proofs and Derivations}\label{sec-proofs}
 
-\section{Appendix}\label{appendix}
-
-\subsection{Additional Results}\label{additional-results}
-
-\subsubsection{Tail heaviness of general preferential attachment
-model}\label{tail-heaviness-of-general-preferential-attachment-model}
-
-Recall the limiting survival function:
-
-\[
-\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
-\]
-
-The aim is to determine how this distribution behaves for different
-choices of \(b\), specifically what maximum domain of attraction does
-this distribution belong to and is it affected by the choice of
-preference function \(b\). \citet{shimura12} introduces a quantity that
-will help in determining what domain of attraction a discrete
-distribution belongs to, that is:
-
-For a distribution \(F\) with survival function \(\bar F\) and some
-\(n\in\mathbb Z^+\) let:
-
-\[
-\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
-\]
-
-\citep{shimura12} then states that if
-\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
-Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
-distribution is light tailed.
-
-Consider a preference function \(b(\cdot)\) that fulfills:
-
-\begin{equation}\phantomsection\label{eq-condb1}{
-\lim_{n\rightarrow \infty} b(n) = \infty
-}\end{equation}
-
-Substituting in the form of \(\bar F(n)\) from \textbf{?@eq-surv} and
-taking the limit, subject to Equation~\ref{eq-condb1}:
-
-\begin{equation}\phantomsection\label{eq-omegares}{
-\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
-}\end{equation} (see appendix for full proof).
-
-Whilst there are many classes of functions that satisfy
-Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}, the vast majority
-of these result in \(\Omega(F,n) \rightarrow 0\), meaning that the
-limiting degree distribution is light tailed. In fact, in order for the
-limiting degree distribution to be heavy tailed, \(b\) must be
-asymptotically linear (\(b(k)\sim k\)).
-
-\subsection{Proofs}\label{proofs}
-
-\subsubsection{Tail heaviness of GPA}\label{tail-heaviness-of-gpa}
+\subsection{\texorpdfstring{Proof of
+Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{proof-of-prp-omega}
 
@@
 \begin{align*}
-\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
-&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
+\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}\\&+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
+&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\&+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
 \end{align*}
@@
 \begin{align*}
-\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
+\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} \\&- \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
 &\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
@@
 
-\subsubsection{Removing the infinite
-sum}\label{removing-the-infinite-sum}
+\subsection{\texorpdfstring{Derivation of
+Equation~\ref{eq-rho}}{Derivation of Equation~}}\label{derivation-of-eq-rho}
 
@@
 \begin{align*}
-\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
+\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{B}(n+x,y)\\
 &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
@@
 
-\section{References}\label{references}
+\newpage
+
+\section{Additional Tables and
+Figures}\label{additional-tables-and-figures}
+
+\newpage
+
+\section*{References}\label{references}
+\addcontentsline{toc}{section}{References}
 

### 2025-06-01 paper.qmd
Implementing changes

@@
 We relate the distributions tail heaviness directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
-keywords: [networks, extremes]
+keywords: [networks, discrete extremes,power law, preferential attachment]
 bibliography: refs.bib  
@@
 
-These mechanistic models are generally useful when simulating networks, but not for fitting to real data as usually only snapshots of a network are available and not the full evolution.
+These mechanistic models are generally useful when simulating networks, but not for fitting to real data as usually only snapshots of a network are available and not the full evolution - without this conclusions about the growth are difficult to make.
 However, one statistic that is usually available for snapshots of networks is the degree distribution making it valuable to study the degrees in a network.
-Often, when looking at the degree distributions of real networks it can be attractive to use the power law when modelling as it is seemingly observed in much of the data.
+Often, when looking at the degree distributions of real networks it can be attractive to use the power law [^1] when modelling as it is seemingly observed in much of the data.
 It also comes with the added benefit of offering a connection to preferential attachment since the BA model results in a power law degree distribution [@Barabasi99], perhaps suggesting that if the power law fits well then the network may have grown according to similar simple rules.
-However, attempting to explain the networks growth using preferential attachment when the degrees seemingly follow a power law may not be the best course of action as [@krapivsky01] shows that a general preference function does not lead to power law degrees and additionally, it is still being argued whether or not a lot of real data actually follows a power law.
+However, attempting to explain the networks growth using preferential attachment when the degrees seemingly follow a power law may not be the best course of action as [@krapivsky01] shows that a general preference function does not necessarily lead to power law degrees, showing that a preference function of form $b(k) = k^\alpha$ leads to light tails when $\alpha<1$, a degenerate network (a finite number of vertices end up with all of the added edges) when $\alpha>1$, and only yielding heavy tails when $\alpha=1$.
+Additionally, it is still being argued whether or not a lot of real data actually follows a power law.
 @Broido_2019 provides some evidence that suggests that real world networks rarely have degrees that follow a power law through the use of statistical analysis of almost 1000 networks across a wide variety of fields; @Voitalov_2019, on the other hand, disagrees and claims that these kinds of networks are not nearly as rare and only appear so as a result of an unrealistic expectation of power law without deviations or noise.
+@rudas07 derives the limiting degree distribution in terms of the preference function $b(\cdot)$ and its parameters, we build upon this by finding a class of preference functions that result in realistic tails i.e.
+heavy but not as heavy as implied by the power law.
+
+[^1]: Power law: $\bar F(k) = L(k)\cdot k^{-1/\xi}$, where $\xi>0$ and $L(k)$ is slowly varying in $k$.
 
@@
 
-@rudas07 derives the limiting degree distribution in terms of the preference function $b(\cdot)$ and its parameters, we build upon this by finding a class of preference functions that result in realistic tails i.e.
-heavy but not as heavy as implied by the power law.
-
-The remainder of this paper begins with a more detailed description of the PA model alongside the theoretical results for the limiting degree distribution using results from @rudas07, this section (@sec-tail) then goes on to analyse the tail behaviour of this limiting distribution in terms of the preference function $b(\cdot)$ arguing that a heavy tailed degree distribution can only be the result of an asymptotically linear preference function - ending with a suggestion for a preference function that can guarantee heavy tails while remaining flexible.
-@sec-model continues on from the previous section by considering the limiting distribution obtained when using a preference function of the form stated at the end of @sec-tail providing examples of how the shape of the distribution varies with respect to the model parameters as well as visualisations for how the tail heaviness changes when these parameters are varied - continued study of this distribution then also reveals a connection to the general Pareto distribution (GPD).
+The remainder of this paper begins with a more detailed description of the PA model alongside the theoretical results for the limiting degree distribution using results from @rudas07, (@sec-tail) then goes on to analyse the tail behaviour of this limiting distribution in terms of the preference function $b(\cdot)$ arguing that a heavy tailed degree distribution can only be the result of an asymptotically linear preference function shown by @prp-omega - ending with a suggestion for a preference function that can guarantee heavy tails while remaining flexible.
+@sec-model builds on the results from @sec-tail and considers the limiting distribution obtained when using a preference function of the form stated at the end of @sec-tail providing examples of how the shape of the distribution varies with respect to the model parameters as well as visualisations for how the tail heaviness changes when these parameters are varied - continued study of this distribution then also reveals a connection to the general Pareto distribution (GPD).
 @sec-sim consists of a simulation study where networks are simulated from the PA model for various parameter combinations and then the model specified in @sec-model is fitted to the resulting degree distributions using Bayesian inference, demonstrating that (for simulated data) the model parameters can be recovered from only the degree distribution.
@@
 
-Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with index 2 and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution in the Gumbel maximum domain of attraction.
+Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with $\xi=1/2$ and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is light tailed.
 
-Given conditions [@eq-condb1] and [@eq-condb2] an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07].
+Given conditions [-@eq-condb1] and [-@eq-condb2] an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07].
 Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-$$ where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.
+$$ {#eq-survlim} where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.
 
@@
 \bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
-$$ which by the previously stated theorem can also be written as:
+$$ which can also be written using @eq-survlim as:
 
@@
 \bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-$$ Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:
+$$ Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the limiting degree distribution can be shown to be:
 
@@
 
-We are specifically interested in how the tail heaviness ($\xi$ in the IGPD) is affected by the preference function $b$, [@shimura12] introduces a quantity that will help in determining what domain of attraction this discrete distribution belongs to, that is:
+We are interested in how the tail behaviour of the limiting degrees is affected by the preference function $b$.
+
+[@shimura12] introduces a quantity that will help in determining this for this discrete distribution, that is:
 
@@
 
-We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e.
+<!-- We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e.
 $\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$.
-First we must consider the following theorem:
+First we must consider the following theorem: -->
 
@@
 \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
-$$ Then,
+$$ 
+Then,
 
@@
 
-From @prp-omega we know that in order for the limiting degree distribution to have heavy tails, we must have that:
+:::{#prp-omega2}
 
+Consider a GPA model with preference function $b(\cdot)$ satisfying @prp-omega.
+If $\lim_{k\rightarrow\infty}\frac{b(k)}{k} = c$, then the limiting degree distribution has a heavy tail with index $c\over\lambda^*$ and this heavy tail may only be the result of a preference function of this form (asymptotically linear).
+
+*Proof*
+
+From @prp-omega, we have that:
 $$
@@
 $$
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0.
-$$ Therefore, in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear.
-With this under consideration we propose the following preference function:
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0\qquad \square.
+$$
+
+:::
+
+Considering @prp-omega2 we propose the following preference function:
 
@@
 
-Some examples of what the degree distribution looks like are shown below in @fig-polylinsurv:
+Some examples of what the degree distribution looks like for various parameters are shown below in @fig-polylinsurv:
 
@@
 
-The survival (@eq-polysurv) can be connected to the IGPD by using Stirling's approximation of the gamma function, where the IGPD is a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ defined by the survival: $$
+The survival function  (-@eq-polysurv) can be connected to a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) with survival: $$
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.
+$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$, denoted as $X|X>u \sim  \mathrm {IGPD}(\xi, \sigma, u)$ where $\xi$ is the shape parameter controlling th tail heaviness.
 
-Using Stirling's approximation we have that: $$
-\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
-$$ now utilising the expression from @eq-polysurv:
+By @eq-polysurv and using Stirling's approximation:
 
@@
 \end{cases}
-$$ meaning that for $k\ge k_0$ the limiting degree distribution is similar to $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
+$$ meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
 
-To assess how close of an approximation this is the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey.
+To assess how close of an approximation this is, the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey.
 The approximation seems to hold up fairly well even for large degrees.
@@
 
-Since $\beta>0$, the shape parameter of the IGPD is positive and thus the distribution is heavy tailed.
-Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices:
+In agreement with @prp-omega2, $\beta>0$, the shape parameter of the IGPD is positive and thus the distribution is heavy tailed.
+Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices.
+The darker regions on the heat maps correspond to a heavier tail and the lighter to a lighter tail, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail heaviness as the Barabási-Albert model, $\xi=0.5$.
 
@@
 
-The darker regions on the heat maps correspond to a heavier tail and the lighter to a lighter tail, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail heaviness as the Barabási-Albert model, $\xi=0.5$.
-
-For a given network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$ and maximum degree $M$ the likelihood is given by:
+Finally to be able to perform inference of the model parameters, we consider a network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$ and maximum degree $M$.
+We can then write the likelihood of this model for this network as:
 
@@
 
-where $B(y,z)$ is the the beta function and $l\ge0$ variable that allows truncating the data such that the minimum degree is $l$.
-
-This likelihood allows for the parameters to be inferred using either a frequentist or a Bayesian approach.
+where $B(y,z)$ is the the beta function and $l\ge0$ is a variable that allows truncating the data such that the minimum degree is $l$, which will allow the model to be fitted whilst ignoring the influence of the lower degrees (those less than $l$) as the model does not capture the behaviour at the lower degrees since @rudas07 only provides results for the case of a preferential attachment tree.
 
@@
 
-@fig-rec1 and @fig-rec2 demonstrate that using this methodology it is possible to recover the model parameters fairly well from only the final degree distribution of a simulated network.
+@fig-rec1 and @fig-rec2 shows the estimations demonstrates that using this methodology it is possible to recover the model parameters fairly well from only the final degree distribution of a simulated network.
 This indicates that the method may also be able to be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.
@@
 
-Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources.
-Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @Lee24.
+Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources and learn about the mechanics of their growth.
+Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @Lee24, but this method has the additional benefit of learning about a network's growth.
+The data consists of 12 networks sourced from [KONECT](konect.cc) and the [Network Data Repository](https://networkrepository.com)[@nr]:
+
+- `as-caida20071105`: network of autonomous systems of the Internet connected with each other from the CAIDA project
+- `dimacs10-astro-ph` : co-authorship network from the "astrophysics" section (astro-ph) of arXiv
+- `ego-twitter`: network of twitter followers
+- `facebook-wosn-wall`: subset of network of Facebook wall pasts
+- `maayan-faa`: USA FAA (Federal Aviation Administration) preferred routes as recommended by the NFDC (National Flight Data Center)
+- `maayan-Stelzl`: network representing interacting pairs of proteins in Humans
+- `moreno-blogs-blogs`: network of URLs found on the
+first pages of individual blogs
+- `opsahl−openflights`: network containing flights between airports of the world.
+- `pajek-erdos`: co-authorship network around Paul Erdős
+- `reactome`: network of protein–protein interactions in Humans
+- `sx-mathoverflow`: interactions from the StackExchange site [MathOverflow](https://mathoverflow.net/)
+- `topology`: network of connections between autonomous systems of the Internet
 
@@
 There appears to be two distinct shapes of preference function.
-The first appears mostly flat (similar to UA) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`.
+The first appears mostly flat (similar to UA) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`, maybe suggesting that most vertices are just as liekly to gain connections as each other but after gaining a number of connections become much more attractive gaining more influence based on the number of connections they have.
 The second distinct shape appears provide some clear PA behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.
+This slowing down could be viewed as a kind of diminishing returns on the degree of a vertex i.e.
+as a vertex gets larger gaining more connections has less of an effect than it did before some threshold $k_0$.
 
@@
 
-Obviously this method had its flaws in that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model as a result of using theory developed for trees and applying it to general networks.
-Hopefully, as the field progresses we will be able to apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.
+One limitation of this method is that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model as a result of using theory developed for trees and applying it to general networks.
+Future work could apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.
 

### 2025-06-01 paper.tex
Implementing changes

@@
 \makeatother
+\usepackage{xcolor}
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 
-\usepackage{longtable,booktabs,array}
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \makeatother
-
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \makeatother
+
 \usepackage{bookmark}
+
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same}
+\urlstyle{same} % disable monospaced font for URLs
 \hypersetup{
@@
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
-  pdfkeywords={networks, extremes},
+  pdfkeywords={networks, discrete extremes, power law, preferential
+attachment},
   colorlinks=true,
@@
 % keywords
-\keywords{networks,  extremes}
+\keywords{networks,  discrete extremes,  power law,  preferential
+attachment}
 
@@
 but not for fitting to real data as usually only snapshots of a network
-are available and not the full evolution. However, one statistic that is
+are available and not the full evolution - without this conclusions
+about the growth are difficult to make. However, one statistic that is
 usually available for snapshots of networks is the degree distribution
@@
 looking at the degree distributions of real networks it can be
-attractive to use the power law when modelling as it is seemingly
-observed in much of the data. It also comes with the added benefit of
-offering a connection to preferential attachment since the BA model
-results in a power law degree distribution \citep{Barabasi99}, perhaps
-suggesting that if the power law fits well then the network may have
-grown according to similar simple rules. However, attempting to explain
-the networks growth using preferential attachment when the degrees
-seemingly follow a power law may not be the best course of action as
+attractive to use the power law \footnote{Power law:
+  \(\bar F(k) = L(k)\cdot k^{-1/\xi}\), where \(\xi>0\) and \(L(k)\) is
+  slowly varying in \(k\).} when modelling as it is seemingly observed
+in much of the data. It also comes with the added benefit of offering a
+connection to preferential attachment since the BA model results in a
+power law degree distribution \citep{Barabasi99}, perhaps suggesting
+that if the power law fits well then the network may have grown
+according to similar simple rules. However, attempting to explain the
+networks growth using preferential attachment when the degrees seemingly
+follow a power law may not be the best course of action as
 \citep{krapivsky01} shows that a general preference function does not
-lead to power law degrees and additionally, it is still being argued
-whether or not a lot of real data actually follows a power law.
+necessarily lead to power law degrees, showing that a preference
+function of form \(b(k) = k^\alpha\) leads to light tails when
+\(\alpha<1\), a degenerate network (a finite number of vertices end up
+with all of the added edges) when \(\alpha>1\), and only yielding heavy
+tails when \(\alpha=1\). Additionally, it is still being argued whether
+or not a lot of real data actually follows a power law.
 \citet{Broido_2019} provides some evidence that suggests that real world
@@
 as a result of an unrealistic expectation of power law without
-deviations or noise.
+deviations or noise. \citet{rudas07} derives the limiting degree
+distribution in terms of the preference function \(b(\cdot)\) and its
+parameters, we build upon this by finding a class of preference
+functions that result in realistic tails i.e.~heavy but not as heavy as
+implied by the power law.
 
@@
 
-\citet{rudas07} derives the limiting degree distribution in terms of the
-preference function \(b(\cdot)\) and its parameters, we build upon this
-by finding a class of preference functions that result in realistic
-tails i.e.~heavy but not as heavy as implied by the power law.
-
 The remainder of this paper begins with a more detailed description of
 the PA model alongside the theoretical results for the limiting degree
-distribution using results from \citet{rudas07}, this section
+distribution using results from \citet{rudas07},
 (Section~\ref{sec-tail}) then goes on to analyse the tail behaviour of
@@
 \(b(\cdot)\) arguing that a heavy tailed degree distribution can only be
-the result of an asymptotically linear preference function - ending with
-a suggestion for a preference function that can guarantee heavy tails
-while remaining flexible. Section~\ref{sec-model} continues on from the
-previous section by considering the limiting distribution obtained when
-using a preference function of the form stated at the end of
+the result of an asymptotically linear preference function shown by
+Proposition~\ref{prp-omega} - ending with a suggestion for a preference
+function that can guarantee heavy tails while remaining flexible.
+Section~\ref{sec-model} builds on the results from
+Section~\ref{sec-tail} and considers the limiting distribution obtained
+when using a preference function of the form stated at the end of
 Section~\ref{sec-tail} providing examples of how the shape of the
@@
 \(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
-with index 2 and the Uniform Attachment (UA) model where \(b(k)=c\)
-leading to a degree distribution in the Gumbel maximum domain of
-attraction.
+with \(\xi=1/2\) and the Uniform Attachment (UA) model where \(b(k)=c\)
+leading to a degree distribution that is light tailed.
 
-Given conditions Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}
-an expression for the survival of the limiting degree distribution can
-be found in the case that \(m=1\); obtained by considering a branching
-process that is equivalent to the growth of the network, as in
-\citep{rudas07}. Theorem 1 from \citep{rudas07} states that for the tree
-\(\Upsilon(t)\) at time \(t\):
+Given conditions \ref{eq-condb1} and \ref{eq-condb2} an expression for
+the survival of the limiting degree distribution can be found in the
+case that \(m=1\); obtained by considering a branching process that is
+equivalent to the growth of the network, as in \citep{rudas07}. Theorem
+1 from \citep{rudas07} states that for the tree \(\Upsilon(t)\) at time
+\(t\):
 
-\[
+\begin{equation}\phantomsection\label{eq-survlim}{
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).
+}\end{equation} where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).
 
@@
 \bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
-\] which by the previously stated theorem can also be written as:
+\] which can also be written using Equation~\ref{eq-survlim} as:
 
@@
 \] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
-the probability mass function (p.m.f.) can be shown to be:
+the limiting degree distribution can be shown to be:
 
@@
 
-We are specifically interested in how the tail heaviness (\(\xi\) in the
-IGPD) is affected by the preference function \(b\), \citep{shimura12}
-introduces a quantity that will help in determining what domain of
-attraction this discrete distribution belongs to, that is:
+We are interested in how the tail behaviour of the limiting degrees is
+affected by the preference function \(b\).
+
+\citep{shimura12} introduces a quantity that will help in determining
+this for this discrete distribution, that is:
 
@@
 
-We can show that in order for the degree distribution to be heavy
-tailed, the preference function must be asymptotically linear
-i.e.~\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0\). First we must
-consider the following theorem:
-
 \begin{theorem}[Stolz-Cesàro Theorem
@@
 
-From Proposition~\ref{prp-omega} we know that in order for the limiting
-degree distribution to have heavy tails, we must have that:
+\begin{proposition}[]\protect\hypertarget{prp-omega2}{}\label{prp-omega2}
 
-\[
+Consider a GPA model with preference function \(b(\cdot)\) satisfying
+Proposition~\ref{prp-omega}. If
+\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} = c\), then the limiting
+degree distribution has a heavy tail with index \(c\over\lambda^*\) and
+this heavy tail may only be the result of a preference function of this
+form (asymptotically linear).
+
+\emph{Proof}
+
+From Proposition~\ref{prp-omega}, we have that: \[
 \lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
@@
 \[
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0.
-\] Therefore, in order for the degree distribution to be heavy tailed,
-the preference function must be asymptotically linear. With this under
-consideration we propose the following preference function:
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0\qquad \square.
+\]
+
+\end{proposition}
+
+Considering Proposition~\ref{prp-omega2} we propose the following
+preference function:
 
@@
 
-Some examples of what the degree distribution looks like are shown below
-in Figure~\ref{fig-polylinsurv}:
+Some examples of what the degree distribution looks like for various
+parameters are shown below in Figure~\ref{fig-polylinsurv}:
 
@@
 
-The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
-by using Stirling's approximation of the gamma function, where the IGPD
-is a discretised variation of the generalised Pareto distribution (GPD)
-usually dubbed the Integer GPD (IGPD) where
-\(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) defined by the survival: \[
+The survival function (\ref{eq-polysurv}) can be connected to a
+discretised variation of the generalised Pareto distribution (GPD)
+usually dubbed the Integer GPD (IGPD) with survival: \[
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).
+\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\), denoted as
+\(X|X>u \sim  \mathrm {IGPD}(\xi, \sigma, u)\) where \(\xi\) is the
+shape parameter controlling th tail heaviness.
 
-Using Stirling's approximation we have that: \[
-\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
-\] now utilising the expression from Equation~\ref{eq-polysurv}:
+By Equation~\ref{eq-polysurv} and using Stirling's approximation:
 
@@
 \end{cases}
-\] meaning that for \(k\ge k_0\) the limiting degree distribution is
-similar to
+\] meaning that for \(k\ge k_0\) the limiting degree distribution (for
+large \(k_0^\alpha\)) is approximated by
 \(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).
 
-To assess how close of an approximation this is the theoretical
+To assess how close of an approximation this is, the theoretical
 conditional survivals are shown in Figure~\ref{fig-approx_surv} in
@@
 
-Since \(\beta>0\), the shape parameter of the IGPD is positive and thus
-the distribution is heavy tailed. Additionally the value of the shape
-parameter \(\xi\) is shown in Figure~\ref{fig-polyheat} for various
-parameter choices:
+In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\), the shape
+parameter of the IGPD is positive and thus the distribution is heavy
+tailed. Additionally the value of the shape parameter \(\xi\) is shown
+in Figure~\ref{fig-polyheat} for various parameter choices. The darker
+regions on the heat maps correspond to a heavier tail and the lighter to
+a lighter tail, the red dashed line shows combinations of \(\alpha\) and
+\(\beta\) that produce a limiting degree distribution with the same tail
+heaviness as the Barabási-Albert model, \(\xi=0.5\).
 
@@
 
-The darker regions on the heat maps correspond to a heavier tail and the
-lighter to a lighter tail, the red dashed line shows combinations of
-\(\alpha\) and \(\beta\) that produce a limiting degree distribution
-with the same tail heaviness as the Barabási-Albert model, \(\xi=0.5\).
-
-For a given network with degree count vector
-\(\pmb n = (n_0, n_1, \ldots, n_M)\) and maximum degree \(M\) the
-likelihood is given by:
+Finally to be able to perform inference of the model parameters, we
+consider a network with degree count vector
+\(\pmb n = (n_0, n_1, \ldots, n_M)\) and maximum degree \(M\). We can
+then write the likelihood of this model for this network as:
 
@@
 
-where \(B(y,z)\) is the the beta function and \(l\ge0\) variable that
-allows truncating the data such that the minimum degree is \(l\).
-
-This likelihood allows for the parameters to be inferred using either a
-frequentist or a Bayesian approach.
+where \(B(y,z)\) is the the beta function and \(l\ge0\) is a variable
+that allows truncating the data such that the minimum degree is \(l\),
+which will allow the model to be fitted whilst ignoring the influence of
+the lower degrees (those less than \(l\)) as the model does not capture
+the behaviour at the lower degrees since \citet{rudas07} only provides
+results for the case of a preferential attachment tree.
 
@@
 
-Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} demonstrate that using
-this methodology it is possible to recover the model parameters fairly
-well from only the final degree distribution of a simulated network.
-This indicates that the method may also be able to be applied to the
-degree distributions of real networks, estimating the model parameters
-assuming they evolved according to the GPA scheme.
+Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} shows the estimations
+demonstrates that using this methodology it is possible to recover the
+model parameters fairly well from only the final degree distribution of
+a simulated network. This indicates that the method may also be able to
+be applied to the degree distributions of real networks, estimating the
+model parameters assuming they evolved according to the GPA scheme.
 
@@
 Turning now to real data, the goal is to fit the model to the degree
-distributions of real networks from various sources. Alongside fitting
-this model to the degree distributions, we then compare the fit to that
-of an mixture distribution that was used in \citet{Lee24}.
+distributions of real networks from various sources and learn about the
+mechanics of their growth. Alongside fitting this model to the degree
+distributions, we then compare the fit to that of an mixture
+distribution that was used in \citet{Lee24}, but this method has the
+additional benefit of learning about a network's growth. The data
+consists of 12 networks sourced from \href{konect.cc}{KONECT} and the
+\href{https://networkrepository.com}{Network Data Repository}\citep{nr}:
+
+\begin{itemize}
+\tightlist
+\item
+  \texttt{as-caida20071105}: network of autonomous systems of the
+  Internet connected with each other from the CAIDA project
+\item
+  \texttt{dimacs10-astro-ph} : co-authorship network from the
+  ``astrophysics'' section (astro-ph) of arXiv
+\item
+  \texttt{ego-twitter}: network of twitter followers
+\item
+  \texttt{facebook-wosn-wall}: subset of network of Facebook wall pasts
+\item
+  \texttt{maayan-faa}: USA FAA (Federal Aviation Administration)
+  preferred routes as recommended by the NFDC (National Flight Data
+  Center)
+\item
+  \texttt{maayan-Stelzl}: network representing interacting pairs of
+  proteins in Humans
+\item
+  \texttt{moreno-blogs-blogs}: network of URLs found on the first pages
+  of individual blogs
+\item
+  \texttt{opsahl−openflights}: network containing flights between
+  airports of the world.
+\item
+  \texttt{pajek-erdos}: co-authorship network around Paul Erdős
+\item
+  \texttt{reactome}: network of protein--protein interactions in Humans
+\item
+  \texttt{sx-mathoverflow}: interactions from the StackExchange site
+  \href{https://mathoverflow.net/}{MathOverflow}
+\item
+  \texttt{topology}: network of connections between autonomous systems
+  of the Internet
+\end{itemize}
 

### 2025-06-02 paper.tex

@@
 \theoremstyle{plain}
+\newtheorem{theorem}{Theorem}[section]
+\theoremstyle{plain}
 \newtheorem{proposition}{Proposition}[section]
-\theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
 \theoremstyle{remark}
@@
 after a threshold PA kicks in, some with this shape are
-\texttt{pajek-erdos} and \texttt{sx-mathoverflow}. The second distinct
-shape appears provide some clear PA behaviour that then slows down after
-a certain point, examples of this are seen in the two infrastructure
-networks \texttt{opsahl-openflights} and \texttt{topology}.
+\texttt{pajek-erdos} and \texttt{sx-mathoverflow}, maybe suggesting that
+most vertices are just as liekly to gain connections as each other but
+after gaining a number of connections become much more attractive
+gaining more influence based on the number of connections they have. The
+second distinct shape appears provide some clear PA behaviour that then
+slows down after a certain point, examples of this are seen in the two
+infrastructure networks \texttt{opsahl-openflights} and
+\texttt{topology}. This slowing down could be viewed as a kind of
+diminishing returns on the degree of a vertex i.e.~as a vertex gets
+larger gaining more connections has less of an effect than it did before
+some threshold \(k_0\).
 
@@
 
-Obviously this method had its flaws in that the lowest degrees needed to
-be truncated as they had a very large effect on the fit of the model as
-a result of using theory developed for trees and applying it to general
-networks. Hopefully, as the field progresses we will be able to apply
-theory developed for general networks using a similar method to this,
-allowing us to compare the results here something that is more accurate.
+One limitation of this method is that the lowest degrees needed to be
+truncated as they had a very large effect on the fit of the model as a
+result of using theory developed for trees and applying it to general
+networks. Future work could apply theory developed for general networks
+using a similar method to this, allowing us to compare the results here
+something that is more accurate.
 

### 2025-06-23 paper.qmd
Updating paper draft

@@
 In this paper, we study a modified version of the model with a flexible preference function that allows super/sub-linear behaviour whilst also guaranteeing that the limiting degree distribution has a heavy tail.
-We relate the distributions tail heaviness directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
+We relate the distributions tail index directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
 keywords: [networks, discrete extremes,power law, preferential attachment]
@@
 Networks appear in many fields such as sociology, politics, epidemiology, and economics.
-Statistical methods have been and continue to be used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17].
+Statistical methods have been used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17].
 
-One example of a mechanistic model that may be considered is the preferential attachment (PA) model where as a network grows and new vertices join they connect to those already present in the network with weights proportional to some preference function $b(\cdot)$ of their in-degree.
+One example of mechanistic models is the preferential attachment (PA) model, where as a network grows and new vertices join, they connect to those already present in the network with probabilities proportional to some preference function of their in-degrees.
 The well-known Barabási-Albert (BA) model is a special case of this when $b(k) = k+1$ [@Barabasi99].
@@
 These mechanistic models are generally useful when simulating networks, but not for fitting to real data as usually only snapshots of a network are available and not the full evolution - without this conclusions about the growth are difficult to make.
-However, one statistic that is usually available for snapshots of networks is the degree distribution making it valuable to study the degrees in a network.
-Often, when looking at the degree distributions of real networks it can be attractive to use the power law [^1] when modelling as it is seemingly observed in much of the data.
-It also comes with the added benefit of offering a connection to preferential attachment since the BA model results in a power law degree distribution [@Barabasi99], perhaps suggesting that if the power law fits well then the network may have grown according to similar simple rules.
-However, attempting to explain the networks growth using preferential attachment when the degrees seemingly follow a power law may not be the best course of action as [@krapivsky01] shows that a general preference function does not necessarily lead to power law degrees, showing that a preference function of form $b(k) = k^\alpha$ leads to light tails when $\alpha<1$, a degenerate network (a finite number of vertices end up with all of the added edges) when $\alpha>1$, and only yielding heavy tails when $\alpha=1$.
-Additionally, it is still being argued whether or not a lot of real data actually follows a power law.
-@Broido_2019 provides some evidence that suggests that real world networks rarely have degrees that follow a power law through the use of statistical analysis of almost 1000 networks across a wide variety of fields; @Voitalov_2019, on the other hand, disagrees and claims that these kinds of networks are not nearly as rare and only appear so as a result of an unrealistic expectation of power law without deviations or noise.
+However, one statistic that is usually available for snapshots of networks is the degree distribution, making it valuable to study the degrees in a network.
+Often, it is appealing to model the degree distributions of real networks using the power law [^1], as it is seemingly observed in much of the data.
+Furthermore, as the BA model results in a power law degree distribution [@Barabasi99], that the power law fits well offers the interpretation that the network may have grown according to PA or similar simple rules.
+However, attempting to explain the networks growth using preferential attachment when the degrees seemingly follow a power law may not be the best course of action as @krapivsky01 show that a general preference function does not necessarily lead to power law degrees.
+Specifically, showing that a preference function of form $b(k) = k^\alpha$ leads to light tails when $\alpha<1$, a degenerate network (a finite number of vertices end up with all of the added edges) when $\alpha>1$, and only yields heavy tails when $\alpha=1$.
 @rudas07 derives the limiting degree distribution in terms of the preference function $b(\cdot)$ and its parameters, we build upon this by finding a class of preference functions that result in realistic tails i.e.
 heavy but not as heavy as implied by the power law.
+On the other hand, it is still being argued whether or not many real data actually follow the power law.
+@Broido_2019 provides some evidence that real world networks rarely have degrees that follow a power law, through the use of statistical analysis of almost 1000 networks across a wide variety of fields; @Voitalov_2019, on the other hand, disagrees and claims that these kinds of networks are not nearly as rare and only appear so as a result of an unrealistic expectation of power law without deviations or noise.
 
-[^1]: Power law: $\bar F(k) = L(k)\cdot k^{-1/\xi}$, where $\xi>0$ and $L(k)$ is slowly varying in $k$.
+[^1]: Power law: $\bar F(k) = L(k)\times k^{-1/\xi}$, where $\xi>0$ and $L(k)$ is slowly varying in $k$.
 
 Extreme value methods have been scarcely used when it comes to analysis of degree distributions and whether they follow the power law.
-The absence of these methods essentially down weights the deviation of vertices with the largest degrees, which are usually the most influential, from the power law.
-Somewhat recent publications [@Voitalov_2019; @Lee24; @wang2022random] have addressed this gap in the literature.
-@Voitalov_2019 proposes estimating the power law exponent of a degree distribution using three estimators based in extreme value theory, in contrast to the estimators generally used in network science, the degree distributions are then split into one of three categories (not power-law , hardly power-law or power-law) depending on the values of the estimates obtained from each of the estimators.
-@Lee24 proposes a framework for modelling entire degree distributions using a combination of two generalisations of the power law including a model selection step that is capable of suggesting whether or not the data is adequately modelled by a power law, with one conclusion being that most data seems to have a tail lighter than what is implied by the body whilst still being heavy tailed.
-@wang2022random considers a preferential attachment model with an affine preference function and reciprocal edges, they study the limiting degree distribution of this model for both the in-degrees and out-degrees - finding that the joint limiting distribution is multivariate regularly varying and has the property of hidden regular variation.
+The absence of these methods essentially downweights the deviation of vertices with the largest degrees, which are usually the most influential, from the power law.
+@Voitalov_2019 proposes estimating the power law exponent of a degree distribution using three estimators based in extreme value theory, in contrast to the estimators generally used in network science, the degree distributions are then split into one of three categories (not power-law, hardly power-law or power-law) depending on the values of the estimates obtained from each of the estimators.
+@Lee24 proposes a framework for modelling entire degree distributions using a combination of two generalisations of the power law with a model selection step that is capable of suggesting whether or not the data is adequately modelled by a power law, with one conclusion being that most data seems to have a tail lighter than what is implied by the power law in body whilst still being heavy tailed.
+One of the models used by @Lee24 is the Zipf-IGP model where up until a certain threshold the data is modelled using the Zipf disribution (a discrete power law) and beyond it a discrete version of the Generalised Pareto distribution is used, we will later compare our model to this as they share some similarities.
+@wang2022random considers a preferential attachment model with an affine preference function and reciprocal edges, finding that the joint limiting distribution of the in-degrees and out-degrees is multivariate regularly varying and has the property of hidden regular variation.
 
 Evidently, the study of degree distributions is a very active but the vast majority of research in the field (whether using extremes or not) are purely descriptive in the sense that no information about the preference function is revealed even if preferential attachment was the underlying mechanism.
-This paper aims to address this gap - asking if given the degree distribution of a network assumed to come from the preferential attachment model, can we directly infer the model parameters?
+This paper aims to address this gap, asking if the degree distribution of a network is assumed to come from the preferential attachment model, can we directly infer the model parameters?
 
-The remainder of this paper begins with a more detailed description of the PA model alongside the theoretical results for the limiting degree distribution using results from @rudas07, (@sec-tail) then goes on to analyse the tail behaviour of this limiting distribution in terms of the preference function $b(\cdot)$ arguing that a heavy tailed degree distribution can only be the result of an asymptotically linear preference function shown by @prp-omega - ending with a suggestion for a preference function that can guarantee heavy tails while remaining flexible.
-@sec-model builds on the results from @sec-tail and considers the limiting distribution obtained when using a preference function of the form stated at the end of @sec-tail providing examples of how the shape of the distribution varies with respect to the model parameters as well as visualisations for how the tail heaviness changes when these parameters are varied - continued study of this distribution then also reveals a connection to the general Pareto distribution (GPD).
-@sec-sim consists of a simulation study where networks are simulated from the PA model for various parameter combinations and then the model specified in @sec-model is fitted to the resulting degree distributions using Bayesian inference, demonstrating that (for simulated data) the model parameters can be recovered from only the degree distribution.
-Finally, @sec-real fits the model to some real data and provides posterior estimates for the preference function assuming the networks evolved according to the PA model.
+The remainder of this paper is as follows: Section 2 gives a detailed description of the PA model alongside the theoretical results for the limiting degree distribution with a focus on its tail behaviour in terms of the preference function $b(\cdot)$.
+It is shown that a heavy tailed degree distribution can only be the result of an asymptotically linear preference function ending with a suggestion for a preference function that can guarantee heavy tails while remaining flexible.
+@sec-model utilises the preference function proposed at the end of Section 2, and illustrates how the tail index of the degree distribution varies with the model parameters.
+@sec-sim consists of a simulation study where networks are simulated from the proposed model for various parameter combinations, demonstrating that the parameters can be recovered from fitting the model to only the degree distribution.
+@sec-real fits the model to some real data and provides posterior estimates for the preference function.
+@sec-conc concludes the article.
 
@@
 
-The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:
+The network generative model that we will focus on in this paper is the General Preferential Attachment model in @rudas07 and is defined as follows:
 
-Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some function $b$ of their degree where the preference function $b$ is chosen such that:
-
-$$
-b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
-$$ {#eq-condb1}
+Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some preference function $b(\cdot)$ of their degree, where $b: N \mapsto \mathbb R^+\setminus\{0\}$ is such that:
 
@@
 
-Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with $\xi=1/2$ and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is light tailed.
+Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with $\xi=1/2$, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is light tailed.
 
-Given conditions [-@eq-condb1] and [-@eq-condb2] an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07].
-Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:
+Given condition [-@eq-condb2] the survival function of the limiting degree distribution, called the limiting survival hereafter, can be analytically derived in the case where $m=1$, obtained by considering a branching process that is equivalent to the growth of the network, as in @rudas07.
+Theorem 1 from @rudas07 states that for the tree $\Upsilon(t)$ at time $t$:
 
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-$$ {#eq-survlim} where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.
+$$ {#eq-survlim} where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to a nodes individual growth, that is:
 
-The limiting survival can be viewed as the limit of the empirical proportion of vertices with degree over a threshold $k$, that is:
+$$
+\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t.
+$$
+
+The limiting survival can be viewed as the limit of the empirical proportion of vertices with degree over a threshold $k\in\mathbb N$, that is:
 
@@
 $$
-\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-$$ Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the limiting degree distribution can be shown to be:
+\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
+$$ Therefore, the corresponding probability mass function of the degree distribution $f(k) = \bar F(k-1) - \bar F(k)$ is
 
 $$
-f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
+f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)}.
 $$
 
-We are interested in how the tail behaviour of the limiting degrees is affected by the preference function $b$.
+We are interested in how the tail behaviour of the discrete limiting degree distribution is affected by the preference function $b$, which can be determined by the results of @shimura12.
 
-[@shimura12] introduces a quantity that will help in determining this for this discrete distribution, that is:
-
-For a distribution $F$ with survival function $\bar F$ and some $k\in\mathbb Z^+$ let:
+For a distribution $F$ with survival function $\bar F$ and some $k\in\mathbb Z^+$ let
 
 $$
-\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
+\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}.
 $$
 
-[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(k) \sim k^{-\alpha}$.
-Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is light tailed.
+@shimura12 states that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy-tailed with $\bar F(k) \sim k^{-\alpha}$.
+On the other hand, if $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is light-tailed.
 This allows us to show the following:
@@
 ::: {#prp-omega}
-If $b(k) \rightarrow \infty$ as $k\rightarrow \infty$ then $$
-\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-$$ and
-
+If $\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}$ and  $b(k) \rightarrow \infty$ as $k\rightarrow \infty$, then
 $$
@@
 
-See appendix [-@sec-proofs] for the details of the proof.
+See Appendix [-@sec-proofs] for the details of the proof.
 
-@prp-omega aligns with the result from [@krapivsky01] demonstrating that a sub-linear preference function will lead to a light tailed distribution.
-So, in order for the degree distribution to be heavy tailed we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
-We can in fact show that BA model produces a heavy tailed degree distribution with index $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$, noting that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ is simply $\lim_{k\rightarrow\infty} 1 = 1$ leaving the tail heaviness to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
+@prp-omega aligns with the result from @krapivsky01 demonstrating that a sub-linear preference function will lead to a light-tailed distribution, as $\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0$ if $b(k)=k^\alpha$ where $\alpha < 1$.
+Proposition 2.1 also aligns with the fact that BA model produces a heavy-tailed degree distribution with index $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$ leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
+So, in order for the degree distribution to be heavy-tailed we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
+To determine the class of functions that will result in heavy-tailed limiting degree distributions, we use the following result:
 
@@
 
-Let $(a_n)_{n\ge1}$ and $(b_n)_{n\ge1}$ be two sequence of real numbers. Assume $(a_n)_{n\ge1}$ is a strictly monotone and divergent sequence and the following limit exists:
+Let $(a_k)_{k\ge1}$ and $(b_k)_{k\ge1}$ be two sequence of real numbers. Assume $(a_k)_{k\ge1}$ is a strictly monotone and divergent sequence and the following limit exists:
 
@@
 Consider a GPA model with preference function $b(\cdot)$ satisfying @prp-omega.
-If $\lim_{k\rightarrow\infty}\frac{b(k)}{k} = c$, then the limiting degree distribution has a heavy tail with index $c\over\lambda^*$ and this heavy tail may only be the result of a preference function of this form (asymptotically linear).
+Then the limiting degree distribution has a heavy tail with index $c/\lambda^{*}$ if and only if $\lim_{k\rightarrow\infty}b(k)/k = c > 0$.
 
@@
 b(k) = \begin{cases}
-k^\alpha + \varepsilon,&k<k_0\\
+k^\alpha + \varepsilon,&k<k_0,\\
 k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
 \end{cases}
-$$ for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.
+$${#eq-pref} for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.
 
-This preference function, as per @prp-omega, will produce a degree distribution with tail heaviness $\beta/\lambda^*$ guaranteeing a heavy tail.
+This preference function, as per @prp-omega, will produce a degree distribution with tail index $\beta/\lambda^*$ guaranteeing a heavy tail.
 We study this preference function further in the next section.
@@
 
-Using a preference function with guaranteed linear behaviour in the limit, allows for the inclusion of sub/super linear behaviour without losing the heavy tails or ending up with a degenerate degree distribution.
-
-The limiting degree distribution resulting from using a preference function of this form can be found to have survival:
+In the previous section, we found that using a preference function with linearity in the limit allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing a heavy tail.
+In this section, we study how the limiting degree distribution varies with the preference function according to @eq-pref.
+Its survival is
 
@@
 \bar F(k) = \begin{cases}
-\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
-\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
+\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
+\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
 \end{cases}
@@
 
-with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where:
+with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where
 
@@
 
-Some examples of what the degree distribution looks like for various parameters are shown below in @fig-polylinsurv:
+Some examples of the limiting degree distribution for various parameter combinations are shown below on log-log scale in @fig-polylinsurv:
 
@@
 
+source('scripts/funcs.R')
 as = c(0.5,1,1.5)
@@
 
+pars[which.max(pars$b/lambdas),]
+max(pars$b/lambdas)
+
 ggplot(data = pars) + geom_line(aes(x=(x+1),y=surv, linetype=as.character(b),colour =as.character(b))) +
@@
 ```
+@fig-polylinsurv demonstrates that this model can capture a wide variety of shapes for the survival function, including a large range of possible tail indices ranging from  0.035 ($\alpha=1.5, \beta=0.1, \varepsilon=1$) to 0.999  ($\alpha=0.5, \beta=1.5, \varepsilon=0.01$).
 
-The survival function  (-@eq-polysurv) can be connected to a discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) with survival: $$
+The survival function  (-@eq-polysurv) can be connected to the discrete version of the generalised Pareto distribution (GP), called the Integer GP (IGP) [@Rohrbeck_2018] with conditional survival:
+$$
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$, denoted as $X|X>u \sim  \mathrm {IGPD}(\xi, \sigma, u)$ where $\xi$ is the shape parameter controlling th tail heaviness.
+$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$, denoted as $X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)$ where $\xi$ is the shape parameter controlling the tail index
 
@@
 
+Therefore, 
 $$
@@
 \end{cases}
-$$ meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
+$$ meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
 
-To assess how close of an approximation this is, the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey.
-The approximation seems to hold up fairly well even for large degrees.
+To assess how close of an approximation this is, the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGP approximations are shown in grey.
+The approximation seems to hold up fairly well even for large degrees and more so when $\alpha$ is larger.
 
@@
 #| fig-height: 4
-#| fig-cap: Theoretical conditional survivals (grey) alongside their IGPD approximations (coloured).
+#| fig-cap: Theoretical conditional survivals (grey) alongside their IGP approximations (coloured).
 #| label: fig-approx_surv
@@
 
-In agreement with @prp-omega2, $\beta>0$, the shape parameter of the IGPD is positive and thus the distribution is heavy tailed.
+In agreement with @prp-omega2, $\beta>0$ ensures that the shape parameter of the IGP is positive and thus the distribution is heavy-tailed.
 Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices.
-The darker regions on the heat maps correspond to a heavier tail and the lighter to a lighter tail, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail heaviness as the Barabási-Albert model, $\xi=0.5$.
+The darker and lighter regions on the heat maps correspond to a heavier and a lighter tail, respectively, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail index as the Barabási-Albert model, that is, $\xi=0.5$.
 
@@
 
-Finally to be able to perform inference of the model parameters, we consider a network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$ and maximum degree $M$.
-We can then write the likelihood of this model for this network as:
+\newpage
+
+To perform inference of the model parameters, we consider a network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$, where $M$ is the maximum degree. We can then write the likelihood as:
 
 \begin{align*}
-L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
-\end{align*}
+L(\pmb x,\pmb n | \pmb \theta) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
+\end{align*}\label{eq-lh}
 
-where $B(y,z)$ is the the beta function and $l\ge0$ is a variable that allows truncating the data such that the minimum degree is $l$, which will allow the model to be fitted whilst ignoring the influence of the lower degrees (those less than $l$) as the model does not capture the behaviour at the lower degrees since @rudas07 only provides results for the case of a preferential attachment tree.
+where $B(y,z)$ is the the beta function, and $l\ge0$ is a variable that allows truncating the data such that the minimum degree is $l$.
+This will allow the model to be fitted whilst ignoring the influence of the lower degrees (those less than $l$) as the model does not capture the behaviour well at the lower degrees since @rudas07 only provides results for the case of a preferential attachment tree.
 
-# Simulation Study {#sec-sim}
+# Applications
 
-This section aims to show that the parameters of the model in @sec-model can be recovered from the degree distribution of a network simulated from it.
+## Simulated Data {#sec-sim}
+
+This first subsection aims to show that the parameters of the model in @sec-model can be recovered from simulating a network from the model, and fitting it to the observed degree distribution, using the likelihood in \eqref{eq-lh}.
 
@@
 
-to obtain a posterior distribution that can then be used in an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm to obtain posterior samples.
-The results of this inference are shown in @fig-rec1 and @fig-rec2.
+where Ga(a, b) is the Gamma distribution with shape a and rate b, and U(a, b) is uniform distribution with lower and upper bounds a and b, to obtain a posterior distribution, up to the proportionality constant.
+Posterior samples can then be obtained by an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm.
 For these simulated networks $l=0$.
@@
 
-![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2}
+![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2 width=80%}
 
-@fig-rec1 and @fig-rec2 shows the estimations demonstrates that using this methodology it is possible to recover the model parameters fairly well from only the final degree distribution of a simulated network.
-This indicates that the method may also be able to be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.
+@fig-rec1 and @fig-rec2 shows the estimations demonstrates the usefulness of the model, as we can recover the model parameters fairly well from only the final degree distribution of the simulated network.
+This indicates that the method may also be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.
 
-# Application to Real Data {#sec-real}
+## Real Data {#sec-real}
 
-Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources and learn about the mechanics of their growth.
-Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in @Lee24, but this method has the additional benefit of learning about a network's growth.
+In this section, we fit the proposed model to the degree distributions of various real networks and learn about the mechanics of their growth.
+While we also compare the fit to that of the mixture distribution by @Lee24 we note that the proposed method has the additional benefit of learning directly about the growth of a network from the inference results.
 The data consists of 12 networks sourced from [KONECT](konect.cc) and the [Network Data Repository](https://networkrepository.com)[@nr]:
@@
 #| fig-height: 9
-#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
+#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red).
 #| label: fig-real1
@@
 
-@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model.
+@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model from @Lee24.
 In most cases, the GPA model does not necessarily provide an improvement in fit when compared to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme.
+
 @fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model.
-Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution and where it doesn't it appears to either be because it is fitting better at the tail than the Zipf-IGP model or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA.
+Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution.
+In the cases of substantial discrepancies, it is either because the GPA model fits the tail better than the Zipf-IGP model does, or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA.
 This again shows the effects that small degrees have on this model, which is somewhat expected as the theory used for this model is for trees and none of these real networks (nor many real networks) are.
@@
 #| fig-height: 9
-#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
+#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red).
 #| label: fig-shapes
@@
 
-@fig-pa shows the estimated preference function $b$ alongside the 95% credible interval on a log-log plot.
-Although the credible interval becomes very large for the largest degrees, this is expected as not all of these networks had data in that region, for those that do the credible interval is much narrower as is the case for `sx-mathoverflow`.
-The most insightful conclusion we can draw from these plots come from the shape of the preference function.
-There appears to be two distinct shapes of preference function.
-The first appears mostly flat (similar to UA) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`, maybe suggesting that most vertices are just as liekly to gain connections as each other but after gaining a number of connections become much more attractive gaining more influence based on the number of connections they have.
-The second distinct shape appears provide some clear PA behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.
+@fig-pa shows the estimated preference function $b(k)$ alongside the 95% credible interval on a log-log plot.
+Although the credible interval becomes very large for the largest degrees, this is expected as not all of these networks had data in that region, and for those that do the credible interval is much narrower, as is the case for `sx-mathoverflow`.
+Looking at the shape of the preference function, there appears to be two distinct shapes of preference function.
+The first appears mostly flat (similar to uniform attachment) for the smallest degrees and then after a threshold preferential attachment kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`.
+The second distinct shape appears to provide some clear preferential attachment behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.
 This slowing down could be viewed as a kind of diminishing returns on the degree of a vertex i.e.
@@
 
-![: Posterior estimates of paramters for real data.](images/pars_plot.png)
+\newpage
 
-# Conclusion and Discussion
+# Conclusion and Discussion {#sec-conc}
 
-In this paper we introduced a flexible class of preference function that when used under the GPA scheme (in the tree setting) is guaranteed to generate a network with a heavy tailed degree distribution whilst remaining flexible in the body.
-Using simulations from networks using this class of preference function we showed that the model parameters are quite easily estimated from the degrees alone at a snapshot in time.
-Seeing this we applied this method to the degree distributions of real networks, estimating their model parameters assuming they evolved in the same way.
-Not only did this yield fairly good fits for the degree distribution, similar to that of the Zipf-IGP, it came with the added benefit of giving a posterior estimate for a preference function.
+In this paper we introduced a class of preference functions that, under the GPA scheme, generate a network with a flexible yet heavy-tailed degree distribution.
+From the simulation study we showed that the parameters can be recovered from fitting the model to the degrees alone.
+We also applied this method to the degree distributions of real networks, estimating their model parameters assuming they evolved in the same way.
+Not only did this yield fairly good fits for the degree distribution, similar to that of the Zipf-IGP, it also came with the added benefit of giving a posterior estimate for a preference function.
 
@@
 Future work could apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.
+This could include fixing the out-degree of new nodes at a constant greater than one, or allowing the out-degree of new nodes to vary.
 
@@
 
+# Additional Plots {.appendix}
+
+![: Posterior estimates of parameters for real data.](images/pars_plot.png){width=80%}
+
 # Proofs and Derivations {#sec-proofs .appendix}
@@
 Taking the form of the GPA degree survival: $$
-\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
+\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda+b(i)}
 $$ and substituting into the formula for $\Omega(F,n)$:
@@
 \begin{align*}
-\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
-&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
-&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
+\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
+&=\left(\log\frac{\lambda+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda+b(k+1)}{b(k+1)}\right)^{-1}\\
+&=\left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}.
 \end{align*}
 
-Clearly if $b(n)=c$ or $\lim_{n\rightarrow\infty}b(n)=c$ for some $c>0$ then $\Omega(F,n)=0$. Now consider a non-constant $b(n)$ and re-write $\Omega(F,n)$ as:
+Clearly if $b(k)=c$ or $\lim_{k\rightarrow\infty}b(k)=c$ for some $c>0$ then $\Omega(F,k)=0$. Now consider a non-constant $b(k)$ and re-write $\Omega(F,k)$ as:
 
 \begin{align*}
-\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}\\&+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
-&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\&+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
+\Omega(F,k) &= \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}+\frac{b(k+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}\\&+\frac{b(k+1)}{\lambda}  -\frac{b(k+1)}{\lambda}\\
+&=\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\&+\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}.
 \end{align*}
 
-Then if $\lim_{n\rightarrow\infty}b(n)=\infty$ it follows that:
+Then if $\lim_{k\rightarrow\infty}b(k)=\infty$ it follows that:
 
 \begin{align*}
-\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} \\&- \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
-&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
-&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
-&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
+\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} \\&- \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\
+&\qquad+\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
+&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
+&=\frac{1}{\lambda}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad \square
 \end{align*}
@@
 b(k) = \begin{cases}
-g(k),&k<k_0\\
-g(k_0) + \beta(k-k_0), &k\ge k_0
+g(k),&k<k_0,\\
+g(k_0) + \beta(k-k_0), &k\ge k_0,
 \end{cases}
@@
 \begin{align*}
-\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
+\hat\rho(\lambda) &= \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)}\\ &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
+&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}.
 \end{align*}
@@
 \prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
-$$ and reindexing the product in the second sum
+$$ and reindexing the product in the second sum,
 
@@
 &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
+&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}.
 \end{align*}
@@
 &=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
-&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
+&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}.
 \end{align*}
 
-this infinite sum does not converge when $x\le1$ as each term is $O(n^{-x})$. We can now use this in $\hat\rho(\lambda)$:
+This infinite sum does not converge when $x\le1$ as each term is $O(n^{-x})$. We can now use this in $\hat\rho(\lambda)$:
 
@@
 &=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
+&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}.\qquad \square
 \end{align*}
-
-\newpage
-
-# Additional Tables and Figures {.appendix}
 

### 2025-06-23 paper.tex
Updating paper draft

@@
 guaranteeing that the limiting degree distribution has a heavy tail. We
-relate the distributions tail heaviness directly to the model
-parameters, allowing direct inference of the parameters from the degree
-distribution alone.}
+relate the distributions tail index directly to the model parameters,
+allowing direct inference of the parameters from the degree distribution
+alone.}
 
@@
 Networks appear in many fields such as sociology, politics,
-epidemiology, and economics. Statistical methods have been and continue
-to be used to study networks that appear in these fields and beyond,
-from the use of stochastic block models to detect communities within the
-French political blogosphere \citep{Latouche11}, to using Exponential
-Random Graph Models (ERGMs) to analyse the global trade network
+epidemiology, and economics. Statistical methods have been used to study
+networks that appear in these fields and beyond, from the use of
+stochastic block models to detect communities within the French
+political blogosphere \citep{Latouche11}, to using Exponential Random
+Graph Models (ERGMs) to analyse the global trade network
 \citep{Setayesh22}, and the use of mechanistic models in neuroscience to
@@
 
-One example of a mechanistic model that may be considered is the
-preferential attachment (PA) model where as a network grows and new
-vertices join they connect to those already present in the network with
-weights proportional to some preference function \(b(\cdot)\) of their
-in-degree. The well-known Barabási-Albert (BA) model is a special case
-of this when \(b(k) = k+1\) \citep{Barabasi99}.
+One example of mechanistic models is the preferential attachment (PA)
+model, where as a network grows and new vertices join, they connect to
+those already present in the network with probabilities proportional to
+some preference function of their in-degrees. The well-known
+Barabási-Albert (BA) model is a special case of this when \(b(k) = k+1\)
+\citep{Barabasi99}.
 
@@
 about the growth are difficult to make. However, one statistic that is
-usually available for snapshots of networks is the degree distribution
-making it valuable to study the degrees in a network. Often, when
-looking at the degree distributions of real networks it can be
-attractive to use the power law \footnote{Power law:
-  \(\bar F(k) = L(k)\cdot k^{-1/\xi}\), where \(\xi>0\) and \(L(k)\) is
-  slowly varying in \(k\).} when modelling as it is seemingly observed
-in much of the data. It also comes with the added benefit of offering a
-connection to preferential attachment since the BA model results in a
-power law degree distribution \citep{Barabasi99}, perhaps suggesting
-that if the power law fits well then the network may have grown
-according to similar simple rules. However, attempting to explain the
-networks growth using preferential attachment when the degrees seemingly
-follow a power law may not be the best course of action as
-\citep{krapivsky01} shows that a general preference function does not
-necessarily lead to power law degrees, showing that a preference
-function of form \(b(k) = k^\alpha\) leads to light tails when
-\(\alpha<1\), a degenerate network (a finite number of vertices end up
-with all of the added edges) when \(\alpha>1\), and only yielding heavy
-tails when \(\alpha=1\). Additionally, it is still being argued whether
-or not a lot of real data actually follows a power law.
-\citet{Broido_2019} provides some evidence that suggests that real world
-networks rarely have degrees that follow a power law through the use of
+usually available for snapshots of networks is the degree distribution,
+making it valuable to study the degrees in a network. Often, it is
+appealing to model the degree distributions of real networks using the
+power law \footnote{Power law: \(\bar F(k) = L(k)\times k^{-1/\xi}\),
+  where \(\xi>0\) and \(L(k)\) is slowly varying in \(k\).}, as it is
+seemingly observed in much of the data. Furthermore, as the BA model
+results in a power law degree distribution \citep{Barabasi99}, that the
+power law fits well offers the interpretation that the network may have
+grown according to PA or similar simple rules. However, attempting to
+explain the networks growth using preferential attachment when the
+degrees seemingly follow a power law may not be the best course of
+action as \citet{krapivsky01} show that a general preference function
+does not necessarily lead to power law degrees. Specifically, showing
+that a preference function of form \(b(k) = k^\alpha\) leads to light
+tails when \(\alpha<1\), a degenerate network (a finite number of
+vertices end up with all of the added edges) when \(\alpha>1\), and only
+yields heavy tails when \(\alpha=1\). \citet{rudas07} derives the
+limiting degree distribution in terms of the preference function
+\(b(\cdot)\) and its parameters, we build upon this by finding a class
+of preference functions that result in realistic tails i.e.~heavy but
+not as heavy as implied by the power law. On the other hand, it is still
+being argued whether or not many real data actually follow the power
+law. \citet{Broido_2019} provides some evidence that real world networks
+rarely have degrees that follow a power law, through the use of
 statistical analysis of almost 1000 networks across a wide variety of
@@
 as a result of an unrealistic expectation of power law without
-deviations or noise. \citet{rudas07} derives the limiting degree
-distribution in terms of the preference function \(b(\cdot)\) and its
-parameters, we build upon this by finding a class of preference
-functions that result in realistic tails i.e.~heavy but not as heavy as
-implied by the power law.
+deviations or noise.
 
@@
 of degree distributions and whether they follow the power law. The
-absence of these methods essentially down weights the deviation of
+absence of these methods essentially downweights the deviation of
 vertices with the largest degrees, which are usually the most
-influential, from the power law. Somewhat recent publications
-\citep{Voitalov_2019, Lee24, wang2022random} have addressed this gap in
-the literature. \citet{Voitalov_2019} proposes estimating the power law
-exponent of a degree distribution using three estimators based in
-extreme value theory, in contrast to the estimators generally used in
-network science, the degree distributions are then split into one of
-three categories (not power-law , hardly power-law or power-law)
-depending on the values of the estimates obtained from each of the
-estimators. \citet{Lee24} proposes a framework for modelling entire
-degree distributions using a combination of two generalisations of the
-power law including a model selection step that is capable of suggesting
-whether or not the data is adequately modelled by a power law, with one
-conclusion being that most data seems to have a tail lighter than what
-is implied by the body whilst still being heavy tailed.
+influential, from the power law. \citet{Voitalov_2019} proposes
+estimating the power law exponent of a degree distribution using three
+estimators based in extreme value theory, in contrast to the estimators
+generally used in network science, the degree distributions are then
+split into one of three categories (not power-law, hardly power-law or
+power-law) depending on the values of the estimates obtained from each
+of the estimators. \citet{Lee24} proposes a framework for modelling
+entire degree distributions using a combination of two generalisations
+of the power law with a model selection step that is capable of
+suggesting whether or not the data is adequately modelled by a power
+law, with one conclusion being that most data seems to have a tail
+lighter than what is implied by the power law in body whilst still being
+heavy tailed. One of the models used by \citet{Lee24} is the Zipf-IGP
+model where up until a certain threshold the data is modelled using the
+Zipf disribution (a discrete power law) and beyond it a discrete version
+of the Generalised Pareto distribution is used, we will later compare
+our model to this as they share some similarities.
 \citet{wang2022random} considers a preferential attachment model with an
-affine preference function and reciprocal edges, they study the limiting
-degree distribution of this model for both the in-degrees and
-out-degrees - finding that the joint limiting distribution is
-multivariate regularly varying and has the property of hidden regular
-variation.
+affine preference function and reciprocal edges, finding that the joint
+limiting distribution of the in-degrees and out-degrees is multivariate
+regularly varying and has the property of hidden regular variation.
 
@@
 preference function is revealed even if preferential attachment was the
-underlying mechanism. This paper aims to address this gap - asking if
-given the degree distribution of a network assumed to come from the
+underlying mechanism. This paper aims to address this gap, asking if the
+degree distribution of a network is assumed to come from the
 preferential attachment model, can we directly infer the model
@@
 
-The remainder of this paper begins with a more detailed description of
-the PA model alongside the theoretical results for the limiting degree
-distribution using results from \citet{rudas07},
-(Section~\ref{sec-tail}) then goes on to analyse the tail behaviour of
-this limiting distribution in terms of the preference function
-\(b(\cdot)\) arguing that a heavy tailed degree distribution can only be
-the result of an asymptotically linear preference function shown by
-Proposition~\ref{prp-omega} - ending with a suggestion for a preference
-function that can guarantee heavy tails while remaining flexible.
-Section~\ref{sec-model} builds on the results from
-Section~\ref{sec-tail} and considers the limiting distribution obtained
-when using a preference function of the form stated at the end of
-Section~\ref{sec-tail} providing examples of how the shape of the
-distribution varies with respect to the model parameters as well as
-visualisations for how the tail heaviness changes when these parameters
-are varied - continued study of this distribution then also reveals a
-connection to the general Pareto distribution (GPD).
-Section~\ref{sec-sim} consists of a simulation study where networks are
-simulated from the PA model for various parameter combinations and then
-the model specified in Section~\ref{sec-model} is fitted to the
-resulting degree distributions using Bayesian inference, demonstrating
-that (for simulated data) the model parameters can be recovered from
-only the degree distribution. Finally, Section~\ref{sec-real} fits the
-model to some real data and provides posterior estimates for the
-preference function assuming the networks evolved according to the PA
-model.
+The remainder of this paper is as follows: Section 2 gives a detailed
+description of the PA model alongside the theoretical results for the
+limiting degree distribution with a focus on its tail behaviour in terms
+of the preference function \(b(\cdot)\). It is shown that a heavy tailed
+degree distribution can only be the result of an asymptotically linear
+preference function ending with a suggestion for a preference function
+that can guarantee heavy tails while remaining flexible.
+Section~\ref{sec-model} utilises the preference function proposed at the
+end of Section 2, and illustrates how the tail index of the degree
+distribution varies with the model parameters. Section~\ref{sec-sim}
+consists of a simulation study where networks are simulated from the
+proposed model for various parameter combinations, demonstrating that
+the parameters can be recovered from fitting the model to only the
+degree distribution. Section~\ref{sec-real} fits the model to some real
+data and provides posterior estimates for the preference function.
+Section~\ref{sec-conc} concludes the article.
 
@@
 
-The network generative model that we will be focussing on in this paper
-is dubbed General Preferential Attachment in \citep{rudas07} and is
-defined as follows:
+The network generative model that we will focus on in this paper is the
+General Preferential Attachment model in \citet{rudas07} and is defined
+as follows:
 
@@
 vertices already in the network with weights proportional to some
-function \(b\) of their degree where the preference function \(b\) is
-chosen such that:
-
-\begin{equation}\phantomsection\label{eq-condb1}{
-b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
-}\end{equation}
+preference function \(b(\cdot)\) of their degree, where
+\(b: N \mapsto \mathbb R^+\setminus\{0\}\) is such that:
 
@@
 \(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
-with \(\xi=1/2\) and the Uniform Attachment (UA) model where \(b(k)=c\)
+with \(\xi=1/2\), and the Uniform Attachment (UA) model where \(b(k)=c\)
 leading to a degree distribution that is light tailed.
 
-Given conditions \ref{eq-condb1} and \ref{eq-condb2} an expression for
-the survival of the limiting degree distribution can be found in the
-case that \(m=1\); obtained by considering a branching process that is
-equivalent to the growth of the network, as in \citep{rudas07}. Theorem
-1 from \citep{rudas07} states that for the tree \(\Upsilon(t)\) at time
-\(t\):
+Given condition \ref{eq-condb2} the survival function of the limiting
+degree distribution, called the limiting survival hereafter, can be
+analytically derived in the case where \(m=1\), obtained by considering
+a branching process that is equivalent to the growth of the network, as
+in \citet{rudas07}. Theorem 1 from \citet{rudas07} states that for the
+tree \(\Upsilon(t)\) at time \(t\):
 
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-}\end{equation} where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).
+}\end{equation} where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\)
+and \(\hat\rho\) is the Laplace transform of the density of the point
+process associated with the pure birth process that corresponds to a
+nodes individual growth, that is:
+
+\[
+\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t.
+\]
 
 The limiting survival can be viewed as the limit of the empirical
-proportion of vertices with degree over a threshold \(k\), that is:
+proportion of vertices with degree over a threshold \(k\in\mathbb N\),
+that is:
 
@@
 \[
-\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
-the limiting degree distribution can be shown to be:
+\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
+\] Therefore, the corresponding probability mass function of the degree
+distribution \(f(k) = \bar F(k-1) - \bar F(k)\) is
 
 \[
-f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
+f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)}.
 \]
 
-We are interested in how the tail behaviour of the limiting degrees is
-affected by the preference function \(b\).
-
-\citep{shimura12} introduces a quantity that will help in determining
-this for this discrete distribution, that is:
+We are interested in how the tail behaviour of the discrete limiting
+degree distribution is affected by the preference function \(b\), which
+can be determined by the results of \citet{shimura12}.
 
 For a distribution \(F\) with survival function \(\bar F\) and some
-\(k\in\mathbb Z^+\) let:
+\(k\in\mathbb Z^+\) let
 
 \[
-\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
+\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}.
 \]
 
-\citep{shimura12} then states that if
-\(\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is heavy tailed with \(\bar F(k) \sim k^{-\alpha}\).
-Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,k) = 0\) then the
-distribution is light tailed. This allows us to show the following:
+\citet{shimura12} states that if
+\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
+then \(F\) is heavy-tailed with \(\bar F(k) \sim k^{-\alpha}\). On the
+other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then the
+distribution is light-tailed. This allows us to show the following:
 
@@
 
-If \(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\) then \[
-\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
-\] and
-
-\[
+If \(\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}\) and
+\(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\), then \[
 \lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
@@
 
-See appendix \ref{sec-proofs} for the details of the proof.
+See Appendix \ref{sec-proofs} for the details of the proof.
 
 Proposition~\ref{prp-omega} aligns with the result from
-\citep{krapivsky01} demonstrating that a sub-linear preference function
-will lead to a light tailed distribution. So, in order for the degree
-distribution to be heavy tailed we need that the limit
-\(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is positive. We can
-in fact show that BA model produces a heavy tailed degree distribution
-with index \(\xi=0.5\) by considering the preference function
-\(b(k) = k + \varepsilon\), noting that the limit
-\(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) is simply
-\(\lim_{k\rightarrow\infty} 1 = 1\) leaving the tail heaviness to be
-\(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\).
+\citet{krapivsky01} demonstrating that a sub-linear preference function
+will lead to a light-tailed distribution, as
+\(\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0\) if \(b(k)=k^\alpha\) where
+\(\alpha < 1\). Proposition 2.1 also aligns with the fact that BA model
+produces a heavy-tailed degree distribution with index \(\xi=0.5\) by
+considering the preference function \(b(k) = k + \varepsilon\), as
+\(\lim_{k\rightarrow\infty}b(k+1)-b(k)=1\) leaving the tail index to be
+\(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\). So,
+in order for the degree distribution to be heavy-tailed we need that the
+limit \(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is positive.
+To determine the class of functions that will result in heavy-tailed
+limiting degree distributions, we use the following result:
 
@@
 
-Let \((a_n)_{n\ge1}\) and \((b_n)_{n\ge1}\) be two sequence of real
-numbers. Assume \((a_n)_{n\ge1}\) is a strictly monotone and divergent
+Let \((a_k)_{k\ge1}\) and \((b_k)_{k\ge1}\) be two sequence of real
+numbers. Assume \((a_k)_{k\ge1}\) is a strictly monotone and divergent
 sequence and the following limit exists:
@@
 Consider a GPA model with preference function \(b(\cdot)\) satisfying
-Proposition~\ref{prp-omega}. If
-\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} = c\), then the limiting
-degree distribution has a heavy tail with index \(c\over\lambda^*\) and
-this heavy tail may only be the result of a preference function of this
-form (asymptotically linear).
+Proposition~\ref{prp-omega}. Then the limiting degree distribution has a
+heavy tail with index \(c/\lambda^{*}\) if and only if
+\(\lim_{k\rightarrow\infty}b(k)/k = c > 0\).
 
@@
 
-\[
+\begin{equation}\phantomsection\label{eq-pref}{
 b(k) = \begin{cases}
-k^\alpha + \varepsilon,&k<k_0\\
+k^\alpha + \varepsilon,&k<k_0,\\
 k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
 \end{cases}
-\] for \(\alpha,\beta, \varepsilon>0\) and \(k_0\in\mathbb N\).
+}\end{equation} for \(\alpha,\beta, \varepsilon>0\) and
+\(k_0\in\mathbb N\).
 
 This preference function, as per Proposition~\ref{prp-omega}, will
-produce a degree distribution with tail heaviness \(\beta/\lambda^*\)
+produce a degree distribution with tail index \(\beta/\lambda^*\)
 guaranteeing a heavy tail. We study this preference function further in
@@
 
-Using a preference function with guaranteed linear behaviour in the
-limit, allows for the inclusion of sub/super linear behaviour without
-losing the heavy tails or ending up with a degenerate degree
-distribution.
-
-The limiting degree distribution resulting from using a preference
-function of this form can be found to have survival:
+In the previous section, we found that using a preference function with
+linearity in the limit allows for the inclusion of sub/super-linear
+behaviour below the threshold, while simultaneously guaranteeing a heavy
+tail. In this section, we study how the limiting degree distribution
+varies with the preference function according to Equation~\ref{eq-pref}.
+Its survival is
 
@@
 
-with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where:
+with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where
 
@@
 
-Some examples of what the degree distribution looks like for various
-parameters are shown below in Figure~\ref{fig-polylinsurv}:
+Some examples of the limiting degree distribution for various parameter
+combinations are shown below on log-log scale in
+Figure~\ref{fig-polylinsurv}:
+
+\begin{verbatim}
+     a   b  eps k0 x   lambda        surv
+10 0.5 1.5 0.01 20 0 1.500094 0.006622105
+\end{verbatim}
+
+\begin{verbatim}
+[1] 0.9999374
+\end{verbatim}
 
@@
 
-The survival function (\ref{eq-polysurv}) can be connected to a
-discretised variation of the generalised Pareto distribution (GPD)
-usually dubbed the Integer GPD (IGPD) with survival: \[
+Figure~\ref{fig-polylinsurv} demonstrates that this model can capture a
+wide variety of shapes for the survival function, including a large
+range of possible tail indices ranging from 0.035
+(\(\alpha=1.5, \beta=0.1, \varepsilon=1\)) to 0.999
+(\(\alpha=0.5, \beta=1.5, \varepsilon=0.01\)).
+
+The survival function (\ref{eq-polysurv}) can be connected to the
+discrete version of the generalised Pareto distribution (GP), called the
+Integer GP (IGP) \citep{Rohrbeck_2018} with conditional survival: \[
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
 \] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\), denoted as
-\(X|X>u \sim  \mathrm {IGPD}(\xi, \sigma, u)\) where \(\xi\) is the
-shape parameter controlling th tail heaviness.
+\(X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)\) where \(\xi\) is the shape
+parameter controlling the tail index
 
@@
 
-\[
+Therefore, \[
 \bar F(k) 
@@
 large \(k_0^\alpha\)) is approximated by
-\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).
+\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).
 
@@
 conditional survivals are shown in Figure~\ref{fig-approx_surv} in
-colour and their IGPD approximations are shown in grey. The
-approximation seems to hold up fairly well even for large degrees.
+colour and their IGP approximations are shown in grey. The approximation
+seems to hold up fairly well even for large degrees and more so when
+\(\alpha\) is larger.
 
@@
 \caption{\label{fig-approx_surv}Theoretical conditional survivals (grey)
-alongside their IGPD approximations (coloured).}
+alongside their IGP approximations (coloured).}
 
@@
 
-In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\), the shape
-parameter of the IGPD is positive and thus the distribution is heavy
-tailed. Additionally the value of the shape parameter \(\xi\) is shown
-in Figure~\ref{fig-polyheat} for various parameter choices. The darker
-regions on the heat maps correspond to a heavier tail and the lighter to
-a lighter tail, the red dashed line shows combinations of \(\alpha\) and
-\(\beta\) that produce a limiting degree distribution with the same tail
-heaviness as the Barabási-Albert model, \(\xi=0.5\).
+In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\) ensures that
+the shape parameter of the IGP is positive and thus the distribution is
+heavy-tailed. Additionally the value of the shape parameter \(\xi\) is
+shown in Figure~\ref{fig-polyheat} for various parameter choices. The
+darker and lighter regions on the heat maps correspond to a heavier and
+a lighter tail, respectively, the red dashed line shows combinations of
+\(\alpha\) and \(\beta\) that produce a limiting degree distribution
+with the same tail index as the Barabási-Albert model, that is,
+\(\xi=0.5\).
 
@@
 
-Finally to be able to perform inference of the model parameters, we
-consider a network with degree count vector
-\(\pmb n = (n_0, n_1, \ldots, n_M)\) and maximum degree \(M\). We can
-then write the likelihood of this model for this network as:
+\newpage
+
+To perform inference of the model parameters, we consider a network with
+degree count vector \(\pmb n = (n_0, n_1, \ldots, n_M)\), where \(M\) is
+the maximum degree. We can then write the likelihood as:
 
 \begin{align*}
-L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
-\end{align*}
+L(\pmb x,\pmb n | \pmb \theta) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
+\end{align*}\label{eq-lh}
 
-where \(B(y,z)\) is the the beta function and \(l\ge0\) is a variable
-that allows truncating the data such that the minimum degree is \(l\),
-which will allow the model to be fitted whilst ignoring the influence of
+where \(B(y,z)\) is the the beta function, and \(l\ge0\) is a variable
+that allows truncating the data such that the minimum degree is \(l\).
+This will allow the model to be fitted whilst ignoring the influence of
 the lower degrees (those less than \(l\)) as the model does not capture
-the behaviour at the lower degrees since \citet{rudas07} only provides
-results for the case of a preferential attachment tree.
+the behaviour well at the lower degrees since \citet{rudas07} only
+provides results for the case of a preferential attachment tree.
 
-\section{Simulation Study}\label{sec-sim}
+\section{Applications}\label{applications}
 
-This section aims to show that the parameters of the model in
-Section~\ref{sec-model} can be recovered from the degree distribution of
-a network simulated from it.
+\subsection{Simulated Data}\label{sec-sim}
+
+This first subsection aims to show that the parameters of the model in
+Section~\ref{sec-model} can be recovered from simulating a network from
+the model, and fitting it to the observed degree distribution, using the
+likelihood in \eqref{eq-lh}.
 
@@
 
-to obtain a posterior distribution that can then be used in an adaptive
-Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm to obtain
-posterior samples. The results of this inference are shown in
-Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
-networks \(l=0\).
+where Ga(a, b) is the Gamma distribution with shape a and rate b, and
+U(a, b) is uniform distribution with lower and upper bounds a and b, to
+obtain a posterior distribution, up to the proportionality constant.
+Posterior samples can then be obtained by an adaptive
+Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm. For these
+simulated networks \(l=0\).
 
@@
 
-\pandocbounded{\includegraphics[keepaspectratio]{images/mcmc_plot.jpg}}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/mcmc_plot.jpg}
 
@@
 Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} shows the estimations
-demonstrates that using this methodology it is possible to recover the
-model parameters fairly well from only the final degree distribution of
-a simulated network. This indicates that the method may also be able to
-be applied to the degree distributions of real networks, estimating the
-model parameters assuming they evolved according to the GPA scheme.
+demonstrates the usefulness of the model, as we can recover the model
+parameters fairly well from only the final degree distribution of the
+simulated network. This indicates that the method may also be applied to
+the degree distributions of real networks, estimating the model
+parameters assuming they evolved according to the GPA scheme.
 
-\section{Application to Real Data}\label{sec-real}
+\subsection{Real Data}\label{sec-real}
 
-Turning now to real data, the goal is to fit the model to the degree
-distributions of real networks from various sources and learn about the
-mechanics of their growth. Alongside fitting this model to the degree
-distributions, we then compare the fit to that of an mixture
-distribution that was used in \citet{Lee24}, but this method has the
-additional benefit of learning about a network's growth. The data
-consists of 12 networks sourced from \href{konect.cc}{KONECT} and the
+In this section, we fit the proposed model to the degree distributions
+of various real networks and learn about the mechanics of their growth.
+While we also compare the fit to that of the mixture distribution by
+\citet{Lee24} we note that the proposed method has the additional
+benefit of learning directly about the growth of a network from the
+inference results. The data consists of 12 networks sourced from
+\href{konect.cc}{KONECT} and the
 \href{https://networkrepository.com}{Network Data Repository}\citep{nr}:
@@
 for several real data sets and their 95\% credible intervals (dotted
-red)}
+red).}
 
@@
 function for various data sets, obtained from fitting the GPA model and
-the Zipf-IGP mixture model. In most cases, the GPA model does not
-necessarily provide an improvement in fit when compared to the Zipf-IGP
-model but where the GPA model fits well we gain additional information
-about the preference function assuming that the network evolved
-according the the GPA scheme. Figure~\ref{fig-shapes} shows the
-posterior of the shape parameter \(\xi\) obtained from the Zipf-IGP
-model alongside the posterior of the equivalent shape parameter
-\(\beta/\lambda^*\) obtained from fitting the GPA model. Generally, the
-GPA model performs similarly to the Zipf-IGP when estimating the tail
-behaviour of the degree distribution and where it doesn't it appears to
-either be because it is fitting better at the tail than the Zipf-IGP
-model or because of the threshold being estimated as too low forcing
-almost all of the data to be modelled by the linear part of the GPA.
-This again shows the effects that small degrees have on this model,
-which is somewhat expected as the theory used for this model is for
-trees and none of these real networks (nor many real networks) are.
+the Zipf-IGP mixture model from \citet{Lee24}. In most cases, the GPA
+model does not necessarily provide an improvement in fit when compared
+to the Zipf-IGP model but where the GPA model fits well we gain
+additional information about the preference function assuming that the
+network evolved according the the GPA scheme.
+
+Figure~\ref{fig-shapes} shows the posterior of the shape parameter
+\(\xi\) obtained from the Zipf-IGP model alongside the posterior of the
+equivalent shape parameter \(\beta/\lambda^*\) obtained from fitting the
+GPA model. Generally, the GPA model performs similarly to the Zipf-IGP
+when estimating the tail behaviour of the degree distribution. In the
+cases of substantial discrepancies, it is either because the GPA model
+fits the tail better than the Zipf-IGP model does, or because of the
+threshold being estimated as too low forcing almost all of the data to
+be modelled by the linear part of the GPA. This again shows the effects
+that small degrees have on this model, which is somewhat expected as the
+theory used for this model is for trees and none of these real networks
+(nor many real networks) are.
 
@@
 
-Figure~\ref{fig-pa} shows the estimated preference function \(b\)
+Figure~\ref{fig-pa} shows the estimated preference function \(b(k)\)
 alongside the 95\% credible interval on a log-log plot. Although the
 credible interval becomes very large for the largest degrees, this is
-expected as not all of these networks had data in that region, for those
-that do the credible interval is much narrower as is the case for
-\texttt{sx-mathoverflow}. The most insightful conclusion we can draw
-from these plots come from the shape of the preference function. There
-appears to be two distinct shapes of preference function. The first
-appears mostly flat (similar to UA) for the smallest degrees and then
-after a threshold PA kicks in, some with this shape are
-\texttt{pajek-erdos} and \texttt{sx-mathoverflow}, maybe suggesting that
-most vertices are just as liekly to gain connections as each other but
-after gaining a number of connections become much more attractive
-gaining more influence based on the number of connections they have. The
-second distinct shape appears provide some clear PA behaviour that then
-slows down after a certain point, examples of this are seen in the two
-infrastructure networks \texttt{opsahl-openflights} and
-\texttt{topology}. This slowing down could be viewed as a kind of
-diminishing returns on the degree of a vertex i.e.~as a vertex gets
-larger gaining more connections has less of an effect than it did before
-some threshold \(k_0\).
+expected as not all of these networks had data in that region, and for
+those that do the credible interval is much narrower, as is the case for
+\texttt{sx-mathoverflow}. Looking at the shape of the preference
+function, there appears to be two distinct shapes of preference
+function. The first appears mostly flat (similar to uniform attachment)
+for the smallest degrees and then after a threshold preferential
+attachment kicks in, some with this shape are \texttt{pajek-erdos} and
+\texttt{sx-mathoverflow}. The second distinct shape appears to provide
+some clear preferential attachment behaviour that then slows down after
+a certain point, examples of this are seen in the two infrastructure
+networks \texttt{opsahl-openflights} and \texttt{topology}. This slowing
+down could be viewed as a kind of diminishing returns on the degree of a
+vertex i.e.~as a vertex gets larger gaining more connections has less of
+an effect than it did before some threshold \(k_0\).
 
@@
 
-\begin{figure}[H]
+\newpage
 
-{\centering \pandocbounded{\includegraphics[keepaspectratio]{images/pars_plot.png}}
+\section{Conclusion and Discussion}\label{sec-conc}
 
-}
-
-\caption{: Posterior estimates of paramters for real data.}
-
-\end{figure}%
-
-\section{Conclusion and Discussion}\label{conclusion-and-discussion}
-
-In this paper we introduced a flexible class of preference function that
-when used under the GPA scheme (in the tree setting) is guaranteed to
-generate a network with a heavy tailed degree distribution whilst
-remaining flexible in the body. Using simulations from networks using
-this class of preference function we showed that the model parameters
-are quite easily estimated from the degrees alone at a snapshot in time.
-Seeing this we applied this method to the degree distributions of real
+In this paper we introduced a class of preference functions that, under
+the GPA scheme, generate a network with a flexible yet heavy-tailed
+degree distribution. From the simulation study we showed that the
+parameters can be recovered from fitting the model to the degrees alone.
+We also applied this method to the degree distributions of real
 networks, estimating their model parameters assuming they evolved in the
 same way. Not only did this yield fairly good fits for the degree
-distribution, similar to that of the Zipf-IGP, it came with the added
-benefit of giving a posterior estimate for a preference function.
+distribution, similar to that of the Zipf-IGP, it also came with the
+added benefit of giving a posterior estimate for a preference function.
 
@@
 using a similar method to this, allowing us to compare the results here
-something that is more accurate.
+something that is more accurate. This could include fixing the
+out-degree of new nodes at a constant greater than one, or allowing the
+out-degree of new nodes to vary.
 
@@
 
+\section{Additional Plots}\label{additional-plots}
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.png}
+
+}
+
+\caption{: Posterior estimates of parameters for real data.}
+
+\end{figure}%
+
 \section{Proofs and Derivations}\label{sec-proofs}
@@
 Taking the form of the GPA degree survival: \[
-\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
+\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda+b(i)}
 \] and substituting into the formula for \(\Omega(F,n)\):
@@
 \begin{align*}
-\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
-&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
-&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
+\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
+&=\left(\log\frac{\lambda+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda+b(k+1)}{b(k+1)}\right)^{-1}\\
+&=\left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}.
 \end{align*}
 
-Clearly if \(b(n)=c\) or \(\lim_{n\rightarrow\infty}b(n)=c\) for some
-\(c>0\) then \(\Omega(F,n)=0\). Now consider a non-constant \(b(n)\) and
-re-write \(\Omega(F,n)\) as:
+Clearly if \(b(k)=c\) or \(\lim_{k\rightarrow\infty}b(k)=c\) for some
+\(c>0\) then \(\Omega(F,k)=0\). Now consider a non-constant \(b(k)\) and
+re-write \(\Omega(F,k)\) as:
 
 \begin{align*}
-\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}\\&+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
-&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\&+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
+\Omega(F,k) &= \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}+\frac{b(k+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}\\&+\frac{b(k+1)}{\lambda}  -\frac{b(k+1)}{\lambda}\\
+&=\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\&+\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}.
 \end{align*}
 
-Then if \(\lim_{n\rightarrow\infty}b(n)=\infty\) it follows that:
+Then if \(\lim_{k\rightarrow\infty}b(k)=\infty\) it follows that:
 
 \begin{align*}
-\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} \\&- \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
-&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
-&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
-&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
+\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} \\&- \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\
+&\qquad+\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
+&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
+&=\frac{1}{\lambda}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad \square
 \end{align*}
@@
 \prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
-\] and reindexing the product in the second sum
+\] and reindexing the product in the second sum,
 
@@
 &=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
-&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
+&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}.
 \end{align*}
 
-this infinite sum does not converge when \(x\le1\) as each term is
+This infinite sum does not converge when \(x\le1\) as each term is
 \(O(n^{-x})\). We can now use this in \(\hat\rho(\lambda)\):
@@
 &=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
+&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}.\qquad \square
 \end{align*}
-
-\newpage
-
-\section{Additional Tables and
-Figures}\label{additional-tables-and-figures}
 

### 2025-06-23 scripts/data_fit_background.R

@@ -131,7 +131,7 @@ mixplt = readRDS('mix_survs.rds')
 comp_plot = ggpubr::ggarrange(mixplt, real_survs,labels=c('Zipf-IGP', 'GPA'), ncol=1)
 
 saveRDS(comp_plot, 'comp_plot.rds')
-fits = readRDS('modelfits.rds')
+fits = readRDS('../results/modelfits.rds')
 plots = list()
 for(i in 1:length(fits)){
   udf = as.data.frame(table(fits[[i]]$smps$k0))
@@ -206,9 +206,9 @@ saveRDS(PA_overlay, 'PA_overlay.rds')
 
 
 # -------------------------------------------------------------------------
-
+library(latex2exp)
 library(ggridges)
-
+nms = readRDS('../results/nms.rds')
 full_par_mat = data.frame(a=NA, eps=NA, k0=NA, b=NA, name=NA)
 
 for(i in 1:length(fits)){
@@ -219,10 +219,10 @@ for(i in 1:length(fits)){
 full_par_mat = full_par_mat[-1,]
 
 meanmat = full_par_mat |> aggregate(.~name, FUN=mean)
-aplot = ggplot() + geom_density_ridges2(data = full_par_mat, aes(x=a, y=name, fill=name))+theme(aspect.ratio = 1, axis.title.x = element_blank(), axis.title.y =element_blank()) + ggtitle('a')
-epsplot =ggplot() + geom_density_ridges2(data = full_par_mat, aes(x=eps, y=name, fill=name))+theme(aspect.ratio = 1, axis.title.x = element_blank(), axis.title.y =element_blank())+ ggtitle('eps')
-k0plot=ggplot() + geom_density_ridges2(data = full_par_mat, aes(x=k0, y=name, fill=name))+theme(aspect.ratio = 1, axis.title.x = element_blank(), axis.title.y =element_blank())+ ggtitle('k0')
-bplot = ggplot() + geom_density_ridges2(data = full_par_mat, aes(x=b, y=name, fill=name))+theme(aspect.ratio = 1, axis.title.x = element_blank(), axis.title.y =element_blank())+ ggtitle('b')
+aplot = ggplot() + geom_density_ridges2(data = full_par_mat, aes(x=a, y=name, fill=name))+theme(aspect.ratio = 1, axis.title.x = element_blank(), axis.title.y =element_blank()) + ggtitle(TeX('\\alpha'))
[…]

### 2025-06-23 scripts/funcs.R

@@ -78,8 +78,6 @@ sample_gpa_tree = function(n, g, b, k0, m=1,quiet=T){
   return(df)
 }
 
-
-
 rho = Vectorize(function(lambda, g, b, k0){
   n1 = 1:k0
   n2 = 0:(k0-1)

### 2025-06-23 scripts/real_data_fitting.R

@@ -200,8 +200,6 @@ nms = readRDS('nms.rds')
 
 plts = list()
 
-
-
 pref = function(pars, x){
   pars = unlist(pars)
   out = x

### 2025-06-23 scripts/testing.R

@@ -1,7 +1,93 @@
-source('funcs.R')
+source('scripts/funcs.R')
+
+rho_fitness_single = function(lambda, g, pars, b, k0){
+  n1 = 1:k0
+  n2 = 0:(k0-1)
+  out1 = sum(exp(cumsum(log(1-lambda/(lambda+g(n1-1,pars))))))
+  out2 = exp(sum(log(1-lambda/(lambda+g(n2,pars)))))
+  return(out1 + g(k0,pars)*out2/(lambda-b))
+}
+rho_fitness_dist = Vectorize(function(lambda, g, fitsample, b, k0, n=10){
+  smps = fitsample(n)
+  rhos = apply(smps, 1, rho_fitness_single, lambda=lambda, g=g, b=b, k0=k0)
+  return(mean(rhos)-1)
+}, vectorize.args = 'lambda')
+find_lambda_fitness = function(g, fitsample, b, k0, n=1e4){
+  rt = uniroot(rho_fitness_dist, c(0.1,10), g=g, fitsample=fitsample, b=b, k0=k0, n=n, extendInt = 'yes')$root
+  return(rt)
+}
+
+S_fitness_single = Vectorize(function(x, g,pars, lambda, k0, b){
+  if(x<0){
+    return(1)
+  }
+  if(x>=k0){
+    k0s = 0:(k0-1)
+    
+    phi_g = lgamma((lambda+g(k0,pars))/b)  + sum(log(g(k0s,pars)/(lambda+g(k0s,pars))))-
+      (lgamma(lambda/b)+lgamma(g(k0,pars)/b))
+    
+    return(exp(phi_g+lbeta(g(k0,pars)/b + x-k0+1, lambda/b)))
+  }
+  if(x>0){
+    ks = 0:(x)
+    return(prod(g(ks,pars)/(lambda+g(ks,pars))))
+  }
+  if(x==0){
+    return(g(x,pars)/(lambda+g(x,pars)))
+  }
+  
+},vectorize.args = 'x')
+
+S_fitness = Vectorize(function(x, g, fitsample, lambda, k0, b, n=1000){
+  smps = fitsample(n)
+  return(mean(apply(smps, 1, S_fitness_single, x=x, g=g, lambda=lambda, k0=k0, b=b)))
+}, vectorize.args = 'x')
+
+sample_gpa_tree_fitness = function(n, g,fitsample, b, k0,quiet=T){
+  root_fitness = fitsample(1)
[…]

### 2025-08-17 paper.qmd
Making changes

@@
 ---
-title: Tail Flexibility in the Degrees of Preferential
- Attachment Networks
+title: "Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions"
 format:
@@
 bibliography: refs.bib  
+keep-tex: true
+pdf-engine-opt: -outdir=''
+output-temp-dir: 'temporary_files'
 ---
@@
 
-Networks appear in many fields such as sociology, politics, epidemiology, and economics.
-Statistical methods have been used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17].
+Networks have become very powerful tools for representing and analysing complex systems, with uses in a large array of fields.
+In network science and statistics, they have been studies by various families of models, from stochastic block models for detecting communities online [@Latouche11], to exponential random graph models (ERGMs) for analysing the global trade network [@Setayesh22],  and mechanistic models for investigating patterns in neural systems [@Betzel17].
 
-One example of mechanistic models is the preferential attachment (PA) model, where as a network grows and new vertices join, they connect to those already present in the network with probabilities proportional to some preference function of their in-degrees.
-The well-known Barabási-Albert (BA) model is a special case of this when $b(k) = k+1$ [@Barabasi99].
+Amid the recent rise of interest in networks, there has been a debate on whether most real networks are scale-free.
+Claiming a real network is scale-free is equivalent to saying that its degree distribution follows a power law, that is the fraction of nodes with degree $k$ is proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with extreme value index $1/\alpha$.
+On the side against the claim is  @Broido_2019who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority in real networks.
+They compare the fits of a power law model against that of several non-scale-free models only finding strong evidence for scale-freeness in four percent and weak evidence in fifty-two percent of the networks.
+On the other side of this debate is @Voitalov_2019 who disagrees and claims that these networks are not nearly as rare and only appear to so as a result of an unrealistic expectation of a power law without deviations or noise.
+Additional evidence of these deviations from a power law is shown by @Lee24 who demonstrate that a lot of networks are partially scale-free, in that the body of the degree distribution is often modelled well by a power law, while the tail is lighter than what is implied by the body, albeit still regularly varying..
 
-These mechanistic models are generally useful when simulating networks, but not for fitting to real data as usually only snapshots of a network are available and not the full evolution - without this conclusions about the growth are difficult to make.
-However, one statistic that is usually available for snapshots of networks is the degree distribution, making it valuable to study the degrees in a network.
-Often, it is appealing to model the degree distributions of real networks using the power law [^1], as it is seemingly observed in much of the data.
-Furthermore, as the BA model results in a power law degree distribution [@Barabasi99], that the power law fits well offers the interpretation that the network may have grown according to PA or similar simple rules.
-However, attempting to explain the networks growth using preferential attachment when the degrees seemingly follow a power law may not be the best course of action as @krapivsky01 show that a general preference function does not necessarily lead to power law degrees.
-Specifically, showing that a preference function of form $b(k) = k^\alpha$ leads to light tails when $\alpha<1$, a degenerate network (a finite number of vertices end up with all of the added edges) when $\alpha>1$, and only yields heavy tails when $\alpha=1$.
-@rudas07 derives the limiting degree distribution in terms of the preference function $b(\cdot)$ and its parameters, we build upon this by finding a class of preference functions that result in realistic tails i.e.
-heavy but not as heavy as implied by the power law.
-On the other hand, it is still being argued whether or not many real data actually follow the power law.
-@Broido_2019 provides some evidence that real world networks rarely have degrees that follow a power law, through the use of statistical analysis of almost 1000 networks across a wide variety of fields; @Voitalov_2019, on the other hand, disagrees and claims that these kinds of networks are not nearly as rare and only appear so as a result of an unrealistic expectation of power law without deviations or noise.
+Nevertheless, most studies into the appropriateness of a power law for the degrees of real networks, the aforementioned references included, have been largely descriptive in the sense that no information about the growth of the networks is revealed.
 
-[^1]: Power law: $\bar F(k) = L(k)\times k^{-1/\xi}$, where $\xi>0$ and $L(k)$ is slowly varying in $k$.
+The popularity of using the power law for network degrees can be traced back to the preferential attachment (PA) model popularised by @Barabasi99.
+In the general model, as new nodes join the network, an existing node with degree k gains edges at a rate proportional to $b(k)$, where b is a non-negative preference function.
+@Barabasi99 showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 3.
+Subsequently, if a real network is shown to be scale-free, one can loosely justify preferential attachment as the underlying mechanism of its growth.
 
-Extreme value methods have been scarcely used when it comes to analysis of degree distributions and whether they follow the power law.
-The absence of these methods essentially downweights the deviation of vertices with the largest degrees, which are usually the most influential, from the power law.
-@Voitalov_2019 proposes estimating the power law exponent of a degree distribution using three estimators based in extreme value theory, in contrast to the estimators generally used in network science, the degree distributions are then split into one of three categories (not power-law, hardly power-law or power-law) depending on the values of the estimates obtained from each of the estimators.
-@Lee24 proposes a framework for modelling entire degree distributions using a combination of two generalisations of the power law with a model selection step that is capable of suggesting whether or not the data is adequately modelled by a power law, with one conclusion being that most data seems to have a tail lighter than what is implied by the power law in body whilst still being heavy tailed.
-One of the models used by @Lee24 is the Zipf-IGP model where up until a certain threshold the data is modelled using the Zipf disribution (a discrete power law) and beyond it a discrete version of the Generalised Pareto distribution is used, we will later compare our model to this as they share some similarities.
-@wang2022random considers a preferential attachment model with an affine preference function and reciprocal edges, finding that the joint limiting distribution of the in-degrees and out-degrees is multivariate regularly varying and has the property of hidden regular variation.
+The model from @Barabasi99 provided the foundations for various generalisations.
 
-Evidently, the study of degree distributions is a very active but the vast majority of research in the field (whether using extremes or not) are purely descriptive in the sense that no information about the preference function is revealed even if preferential attachment was the underlying mechanism.
-This paper aims to address this gap, asking if the degree distribution of a network is assumed to come from the preferential attachment model, can we directly infer the model parameters?
+@krapivsky01 considered $b(k) = (k+1)^\alpha$, and showed that the degree distribution is not regularly varying (and therefore not following the power law) when $0<\alpha<1$, and when $\alpha>1$ a finite number of nodes end up with all edges after a certain point resulting in a degenerate degree distribution.
+@wang2022random returns to a linear preference function of the form $b(k) = k+\varepsilon$ but adds the possibility for reciprocal edges to be sent, resulting in the joint distribution of in-degree and out-degree being multivariate regularly varying and having the property of hidden regular variation.
+@rudas07 follows in the footsteps of @krapivsky01 , by considering a preferential attachment tree and using theory from continuous branching processes, derives a limiting degree distribution in terms of the preference function $b(\cdot)$.
+Nevertheless, research in this area tends to only focus on the theoretical asymptotic results of network growth models with little analysis of real networks.
 
-The remainder of this paper is as follows: Section 2 gives a detailed description of the PA model alongside the theoretical results for the limiting degree distribution with a focus on its tail behaviour in terms of the preference function $b(\cdot)$.
-It is shown that a heavy tailed degree distribution can only be the result of an asymptotically linear preference function ending with a suggestion for a preference function that can guarantee heavy tails while remaining flexible.
-@sec-model utilises the preference function proposed at the end of Section 2, and illustrates how the tail index of the degree distribution varies with the model parameters.
+This paper aims to address the gap between the applied and theoretical works, by asking if a network is assumed to come from a preferential attachment model, can we use the degree distribution alone to directly infer the parameters of the the preference function and learn about the growth mechanisms?
+Moreover, proper consideration is given to the tail of the degree distribution, because otherwise the effects of the largest degrees, which correspond to the most influential nodes, deviating from the power law will be discounted.
+
+It is important to think about how we intend to consider the tail of the degree distributions, because if not done correctly we may end up essentially discounting the effects of the largest degrees (often the most influential nodes) deviating from a power law.
+
+For the above reason, we will use methods from discrete extreme value theory, in particular those by @shimura12, who provided theoretical guarantees for a discrete distribution to be regularly varying or not.
+Using these results, we demonstrate how the tail of the degree distribution is affected by $b(\cdot)$, and subsequently propose a class of preference functions that is tail realistic for real networks.
+These analytical results enables the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to preferential attachment, to be inferred directly.
+
+The remainder of the paper is as follows: @sec-tail gives a detailed description of the preferential attachment model alongside the theoretical results for the survival of the limiting degrees, with a focus on the tail behaviour in terms of the preference function $b(\cdot)$.
+A class of asymptotically linear preference functions will be introduced and shown to guarantee regular variation in the degrees while remaining flexible up until a threshold.
+@sec-model utilises the proposed preference function and illustrates numerically how the extreme value index (EVI) of the degree distribution varies with the model parameters.
 @sec-sim consists of a simulation study where networks are simulated from the proposed model for various parameter combinations, demonstrating that the parameters can be recovered from fitting the model to only the degree distribution.
@@
 
-The network generative model that we will focus on in this paper is the General Preferential Attachment model in @rudas07 and is defined as follows:
+The model that we will focus on in this paper is the General Preferential Attachment model in @rudas07 and is defined as follows:
 
-Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some preference function $b(\cdot)$ of their degree, where $b: N \mapsto \mathbb R^+\setminus\{0\}$ is such that:
+Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges from the new vertex; the target for each of these edges are selected from the vertices already in the network with weights proportional to some preference function $b(\cdot)$ of their degree, where $b: \mathbb N \mapsto \mathbb R^+\setminus\{0\}$ is such that:
 
@@
 
-Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with $\xi=1/2$, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is light tailed.
+Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with  EVI $\xi=1/2$, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
 
-Given condition [-@eq-condb2] the survival function of the limiting degree distribution, called the limiting survival hereafter, can be analytically derived in the case where $m=1$, obtained by considering a branching process that is equivalent to the growth of the network, as in @rudas07.
+The survival function of the limiting degree distribution, called the limiting survival hereafter, under condition [-@eq-condb2] can be analytically derived in the case where $m=1$, this is presented below.
+
+Consider a continuous time branching process $\zeta(t)$ driven by a Markovian pure birth process, with $\zeta(0)=0$ and birth rates depending on a non-negative function $b(\cdot)$:
+
+$$
+\Pr(\zeta(t+dt)=k+1|\zeta(t)=k) = b(k)dt + o(dt).
+$$
+Now let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$.
+
 Theorem 1 from @rudas07 states that for the tree $\Upsilon(t)$ at time $t$:
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-$$ {#eq-survlim} where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to a nodes individual growth, that is:
-
-$$
-\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t.
-$$
+$$ {#eq-survlim} 
+where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to the growth of an individual node, that is $\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t$
 
@@
 \bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
-$$ which can also be written using @eq-survlim as:
+$$
+which can also be written using @eq-survlim as:
 
@@
 \bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
-$$ Therefore, the corresponding probability mass function of the degree distribution $f(k) = \bar F(k-1) - \bar F(k)$ is
+$${#eq-surv-origin}
+
+Therefore, the corresponding probability mass function of the degree distribution $f(k) = \bar F(k-1) - \bar F(k)$ is
 
@@
 f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)}.
-$$
+$${#eq-pmf-origin}
 
-We are interested in how the tail behaviour of the discrete limiting degree distribution is affected by the preference function $b$, which can be determined by the results of @shimura12.
+We are interested in how the tail behaviour of the discrete limiting degree distribution is affected by the preference function $b$.
 
-For a distribution $F$ with survival function $\bar F$ and some $k\in\mathbb Z^+$ let
+For a distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ let
 
@@
 
-@shimura12 states that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy-tailed with $\bar F(k) \sim k^{-\alpha}$.
-On the other hand, if $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is light-tailed.
 This allows us to show the following:
@@
 
+@shimura12 states that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$.
+On the other hand, if $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then we will refer to the distribution as light-tailed.
+
 @prp-omega aligns with the result from @krapivsky01 demonstrating that a sub-linear preference function will lead to a light-tailed distribution, as $\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0$ if $b(k)=k^\alpha$ where $\alpha < 1$.
-Proposition 2.1 also aligns with the fact that BA model produces a heavy-tailed degree distribution with index $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$ leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
-So, in order for the degree distribution to be heavy-tailed we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
-To determine the class of functions that will result in heavy-tailed limiting degree distributions, we use the following result:
+Proposition 2.1 also aligns with the fact that BA model produces a regularly varying degree distribution with EVI $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$ leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
+So, in order for the degree distribution to be regularly varying  we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
+To determine the class of functions that will result in regularly varying limiting degree distributions, we use the following result:
 
@@
 
-::: {#thm-stolz}
-## Stolz-Cesàro Theorem [@cesaro]
-
-Let $(a_k)_{k\ge1}$ and $(b_k)_{k\ge1}$ be two sequence of real numbers. Assume $(a_k)_{k\ge1}$ is a strictly monotone and divergent sequence and the following limit exists:
-
-$$
-\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
-$$ 
-Then,
-
-$$
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
-$$
-:::
-
 :::{#prp-omega2}
@@
 Consider a GPA model with preference function $b(\cdot)$ satisfying @prp-omega.
-Then the limiting degree distribution has a heavy tail with index $c/\lambda^{*}$ if and only if $\lim_{k\rightarrow\infty}b(k)/k = c > 0$.
+Then the limiting degree distribution is regularly varying with EVI $c/\lambda^{*}$ if and only if $\lim_{k\rightarrow\infty}b(k)/k = c > 0$.
 
@@
 \lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
-$$ meaning by @thm-stolz:
+$$ Using the Stolz-Cesaro Theorem [@thm-stolz](Appendix -@sec-sup),
 
@@
 
-Considering @prp-omega2 we propose the following preference function:
+Using this result we can understand how the preference function is directly connected with the tail behaviour of the degree distribution.
+Specifically, regular variation is achieved if and only if $b(k)$ is asymptotically linear with $k$.
+We use this result to create a preference function that guarantees regular variation in the tail of the degree distribution, aligning with analysis of real networks, whilst allowing for the tail to deviate from the shape implied by the body.
+This gives the model the capability to produce realistic behaviour in the degrees like what was in @Lee24 by using a piecewise function inspired by the observed deviation from the power law after a certain threshold:
 
@@
 
-This preference function, as per @prp-omega, will produce a degree distribution with tail index $\beta/\lambda^*$ guaranteeing a heavy tail.
-We study this preference function further in the next section.
+Using @prp-omega, we can show that $\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta$, meaing the degree distribution obtained using this preference funnction is regularly varying with tail index $\lambda^*/\beta$.
 
@@
 
-# Preferential Attachment with Flexible Heavy Tail {#sec-model}
+# A Preferential Attachment Model with Flexible Regular Variation {#sec-model}
 
-In the previous section, we found that using a preference function with linearity in the limit allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing a heavy tail.
-In this section, we study how the limiting degree distribution varies with the preference function according to @eq-pref.
-Its survival is
+In the previous section, we found that using an asymptotically linear preference function allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing regular varation of the degrees.
+In this section, we demonstrate the flexibility of the preference function in @eq-pref, with regard to the tail behaviour of the limiting degree distribution.
+Using @eq-surv-origin, the limiting survival is
 
@@
 
-which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$.
+which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$.
 
-Some examples of the limiting degree distribution for various parameter combinations are shown below on log-log scale in @fig-polylinsurv:
+For some parameter combinations, the limiting survival $\bar F(k)$ is shown on log-log scale in @fig-polylinsurv:
 
@@
 #| fig-height: 4
-#| fig-cap: Theoretical survival distributions of the limiting degree distributions, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed preferential attachment model.
+#| fig-cap: The limiting survival, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed preferential attachment model.
 #| label: fig-polylinsurv
@@
 
-pars[which.max(pars$b/lambdas),]
-max(pars$b/lambdas)
+# pars[which.max(pars$b/lambdas),]
+# max(pars$b/lambdas)
 
@@
 ```
-@fig-polylinsurv demonstrates that this model can capture a wide variety of shapes for the survival function, including a large range of possible tail indices ranging from  0.035 ($\alpha=1.5, \beta=0.1, \varepsilon=1$) to 0.999  ($\alpha=0.5, \beta=1.5, \varepsilon=0.01$).
+@fig-polylinsurv demonstrates that this model can capture a range of tail behaviour, including a large range of possible tail indices ranging from  0.035 ($\alpha=1.5, \beta=0.1, \varepsilon=1$) to 0.999  ($\alpha=0.5, \beta=1.5, \varepsilon=0.01$).
 
-The survival function  (-@eq-polysurv) can be connected to the discrete version of the generalised Pareto distribution (GP), called the Integer GP (IGP) [@Rohrbeck_2018] with conditional survival:
+The analytic form of the survival function  in (-@eq-polysurv), offers a natural connection to the discrete version of the generalised Pareto distribution (GP), providing a link to a well established component of discrete extremes in the literature.
+Specifically, connecting to the Integer GP (IGP) seen in @Rohrbeck_2018 with conditional survival:
 $$
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$, denoted as $X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)$ where $\xi$ is the shape parameter controlling the tail index
+$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$, denoted as $X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)$ where $\xi$ is the shape parameter controlling the tail index.
 
@@
 &=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
-&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
+&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}.
 \end{align*}
@@
 Therefore, 
-$$
+$${#eq-igp-est}
 \bar F(k) 
@@
 \end{cases}
-$$ meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
+$$ meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
 
-To assess how close of an approximation this is, the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGP approximations are shown in grey.
+To assess how close of an approximation this is, the theoretical conditional survival [@eq-polysurv] are shown in @fig-approx_surv in colour and their IGP approximations [@eq-igp-est] are shown in grey.
 The approximation seems to hold up fairly well even for large degrees and more so when $\alpha$ is larger.
@@
 
-In agreement with @prp-omega2, $\beta>0$ ensures that the shape parameter of the IGP is positive and thus the distribution is heavy-tailed.
+In agreement with @prp-omega2, $\beta>0$ ensures that the shape parameter of the IGP is positive and thus the distribution is regularly varying.
 Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices.
 The darker and lighter regions on the heat maps correspond to a heavier and a lighter tail, respectively, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail index as the Barabási-Albert model, that is, $\xi=0.5$.
+
+Through the connection to the IGP, fitting this model is almost equivalent to fitting the IGP to the degrees and estimating its parameters.
+However, instead of only describing the shape of the degree distribution, we would also gain estimates for the shape of the preference function, thus gaining a direct understanding the mechanisms underlying the growth of the network.
 
@@
 
-To perform inference of the model parameters, we consider a network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$, where $M$ is the maximum degree. We can then write the likelihood as:
+To perform inference of the model parameters, we consider a network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$, where $M$ is the maximum degree.
+Using @eq-pmf-origin, the likelihood is:
 
 \begin{align*}
-L(\pmb x,\pmb n | \pmb \theta) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
+L(\pmb n | \pmb \theta,l) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
 \end{align*}\label{eq-lh}
 
-where $B(y,z)$ is the the beta function, and $l\ge0$ is a variable that allows truncating the data such that the minimum degree is $l$.
+where $\text{B}(\cdot,\cdot)$ is the the beta function, and $l\ge0$ is a quantity that allows truncating the data such that the minimum degree is $l$.
 This will allow the model to be fitted whilst ignoring the influence of the lower degrees (those less than $l$) as the model does not capture the behaviour well at the lower degrees since @rudas07 only provides results for the case of a preferential attachment tree.
@@
 
-This first subsection aims to show that the parameters of the model in @sec-model can be recovered from simulating a network from the model, and fitting it to the observed degree distribution, using the likelihood in \eqref{eq-lh}.
+This subsection aims to show that the parameters of the model (and therefore the preference function) in @sec-model can be recovered from simulating a network from the model, and fitting it to the observed degree distribution, using the likelihood in \eqref{eq-lh}.
 
@@
 \begin{align*}
-\alpha&\sim \text{Ga}(1,0.01),\\
-\beta &\sim  \text{Ga}(1,0.01),\\
+\alpha&\sim \text{Gamma}(1,0.01),\\
+\beta &\sim  \text{Gamma}(1,0.01),\\
 k_0 &\sim \text{U}(1,10,000),\\
-\varepsilon &\sim \text{Ga(1,0.01)},
+\varepsilon &\sim \text{Gamma(1,0.01)},
 \end{align*}
 
-where Ga(a, b) is the Gamma distribution with shape a and rate b, and U(a, b) is uniform distribution with lower and upper bounds a and b, to obtain a posterior distribution, up to the proportionality constant.
+where Gamma($a$, $b$) is the Gamma distribution with shape $a$ and rate $b$, and U($a$, $b$) is uniform distribution with lower and upper bounds a and b, to obtain a posterior distribution, up to the proportionality constant.
 Posterior samples can then be obtained by an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm.
@@
 #| fig-height: 8
-#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
+#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of $(\alpha,\beta,\varepsilon)$ and $k_0=20$.
 #| label: fig-rec1
@@
 
-![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2 width=80%}
+![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2 width=100%}
 
-@fig-rec1 and @fig-rec2 shows the estimations demonstrates the usefulness of the model, as we can recover the model parameters fairly well from only the final degree distribution of the simulated network.
-This indicates that the method may also be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.
+Figures [-@fig-rec1] and [-@fig-rec2] demonstrate the usefulness of the model, as we can recover the model parameters well from only the final degree distribution of the simulated network.
+This indicates that the method may also be applied to real networks, with the assumption that they evolved according to the GPA scheme.
 
@@
 - `ego-twitter`: network of twitter followers
-- `facebook-wosn-wall`: subset of network of Facebook wall pasts
+- `facebook-wosn-wall`: subset of network of Facebook wall posts
 - `maayan-faa`: USA FAA (Federal Aviation Administration) preferred routes as recommended by the NFDC (National Flight Data Center)
-- `maayan-Stelzl`: network representing interacting pairs of proteins in Humans
+- `maayan-Stelzl`: network representing interacting pairs of proteins in humans
 - `moreno-blogs-blogs`: network of URLs found on the
@@
 - `pajek-erdos`: co-authorship network around Paul Erdős
-- `reactome`: network of protein–protein interactions in Humans
+- `reactome`: network of protein–protein interactions in humans
 - `sx-mathoverflow`: interactions from the StackExchange site [MathOverflow](https://mathoverflow.net/)
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="100%"}
 #| fig-width: 9
@@
 @fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model from @Lee24.
-In most cases, the GPA model does not necessarily provide an improvement in fit when compared to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme.
+In most cases, the GPA model gives a similar fit to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme.
 
@@
 In the cases of substantial discrepancies, it is either because the GPA model fits the tail better than the Zipf-IGP model does, or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA.
-This again shows the effects that small degrees have on this model, which is somewhat expected as the theory used for this model is for trees and none of these real networks (nor many real networks) are.
+This again shows the effects that small degrees have on this model.
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="100%"}
 #| fig-width: 9
@@
 
-@fig-pa shows the estimated preference function $b(k)$ alongside the 95% credible interval on a log-log plot.
+@fig-pa shows the estimated preference function $b(k)$ alongside the 95% credible interval on a log-log scale.
 Although the credible interval becomes very large for the largest degrees, this is expected as not all of these networks had data in that region, and for those that do the credible interval is much narrower, as is the case for `sx-mathoverflow`.
@@
 
-In this paper we introduced a class of preference functions that, under the GPA scheme, generate a network with a flexible yet heavy-tailed degree distribution.
+In this paper we introduced a class of preference functions that, under the GPA scheme, generate a network with a flexible yet regularly varying degree distribution.
 From the simulation study we showed that the parameters can be recovered from fitting the model to the degrees alone.
@@
 
-# Additional Plots {.appendix}
+# Supplementary Results{#sec-sup .appendix}
+
+::: {#thm-stolz}
+## Stolz-Cesàro Theorem [@cesaro]
+
+Let $(a_k)_{k\ge1}$ and $(b_k)_{k\ge1}$ be two sequence of real numbers. Assume $(a_k)_{k\ge1}$ is a strictly monotone and divergent sequence and the following limit exists:
+
+$$
+\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
+$$ 
+Then,
+
+$$
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
+$$
+:::
+
+# Additional Plots{.appendix}
 

### 2025-08-17 paper.tex
Making changes

@@
 \makeatother
-\usepackage{xcolor}
-\setlength{\emergencystretch}{3em} % prevent overfull lines
-\setcounter{secnumdepth}{5}
 % Make \paragraph and \subparagraph free-standing
@@
 
-\providecommand{\tightlist}{%
-  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}\usepackage{longtable,booktabs,array}
+\usepackage{longtable,booktabs,array}
 \usepackage{calc} % for calculating minipage widths
@@
 \makeatother
+
+\setlength{\emergencystretch}{3em} % prevent overfull lines
+
+\providecommand{\tightlist}{%
+  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
 
@@
 \makeatother
-
 \usepackage{bookmark}
-
 \IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
-\urlstyle{same} % disable monospaced font for URLs
+\urlstyle{same}
 \hypersetup{
-  pdftitle={Tail Flexibility in the Degrees of Preferential Attachment Networks},
+  pdftitle={Learning Network Growth Mechanisms from Degree Distributions in Preferential Attachment Networks},
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
@@
 
-\title[Tail Flexibility in the Degrees of Preferential Attachment
-Networks]{Tail Flexibility in the Degrees of Preferential Attachment
-Networks}
+\title[Learning Network Growth Mechanisms from Degree Distributions in
+Preferential Attachment Networks]{Learning Network Growth Mechanisms
+from Degree Distributions in Preferential Attachment Networks}
 
@@
 
-Networks appear in many fields such as sociology, politics,
-epidemiology, and economics. Statistical methods have been used to study
-networks that appear in these fields and beyond, from the use of
-stochastic block models to detect communities within the French
-political blogosphere \citep{Latouche11}, to using Exponential Random
-Graph Models (ERGMs) to analyse the global trade network
-\citep{Setayesh22}, and the use of mechanistic models in neuroscience to
-investigate the patterns of connections in neural systems
-\citep{Betzel17}.
+Networks have become very powerful tools for representing and analysing
+complex systems, with uses in a large array of fields, from using
+stochastic block models to detect communities online \citep{Latouche11},
+to using Exponential Random Graph Models (ERGMs) to analyse the global
+trade network \citep{Setayesh22}, and even using mechanistic models to
+investigate the patterns in neural systems \citep{Betzel17}.
 
-One example of mechanistic models is the preferential attachment (PA)
-model, where as a network grows and new vertices join, they connect to
-those already present in the network with probabilities proportional to
-some preference function of their in-degrees. The well-known
-Barabási-Albert (BA) model is a special case of this when \(b(k) = k+1\)
-\citep{Barabasi99}.
+Along with the recent rise of interest in networks, has been a debate
+amongst those that study them regarding the claims that most real
+networks are scale-free. This claim is most often equivalent to claiming
+that the degree distributions of real networks follow a power law, that
+is the fraction of nodes with degree \(k\) is proportional to
+\(k^{-\alpha}\), and therefore has a regularly varying tail with extreme
+value index \(1/\alpha\). \citet{Broido_2019} provide empirical
+evidence, using almost one thousand networks, that these scale-free
+networks do not make up a large majority of real networks. They compare
+the fits of a power law model against that of several non-scale-free
+models only finding strong evidence for scale-freeness in four percent
+and weak evidence in fifty-two percent of the networks. On the other
+side of this debate is \citet{Voitalov_2019} who disagrees and claims
+that these networks are not nearly as rare and only appear to so as a
+result of an unrealistic expectation of a power law without deviations
+or noise. Evidence of these deviations from a power law is shown by
+\citet{Lee24} who demonstrate that a lot of networks are partially
+scale-free, in that the body of the degree distribution is modelled well
+by a power law while the tail is lighter than what is implied by the
+body (although still regularly varying).
 
-These mechanistic models are generally useful when simulating networks,
-but not for fitting to real data as usually only snapshots of a network
-are available and not the full evolution - without this conclusions
-about the growth are difficult to make. However, one statistic that is
-usually available for snapshots of networks is the degree distribution,
-making it valuable to study the degrees in a network. Often, it is
-appealing to model the degree distributions of real networks using the
-power law \footnote{Power law: \(\bar F(k) = L(k)\times k^{-1/\xi}\),
-  where \(\xi>0\) and \(L(k)\) is slowly varying in \(k\).}, as it is
-seemingly observed in much of the data. Furthermore, as the BA model
-results in a power law degree distribution \citep{Barabasi99}, that the
-power law fits well offers the interpretation that the network may have
-grown according to PA or similar simple rules. However, attempting to
-explain the networks growth using preferential attachment when the
-degrees seemingly follow a power law may not be the best course of
-action as \citet{krapivsky01} show that a general preference function
-does not necessarily lead to power law degrees. Specifically, showing
-that a preference function of form \(b(k) = k^\alpha\) leads to light
-tails when \(\alpha<1\), a degenerate network (a finite number of
-vertices end up with all of the added edges) when \(\alpha>1\), and only
-yields heavy tails when \(\alpha=1\). \citet{rudas07} derives the
-limiting degree distribution in terms of the preference function
-\(b(\cdot)\) and its parameters, we build upon this by finding a class
-of preference functions that result in realistic tails i.e.~heavy but
-not as heavy as implied by the power law. On the other hand, it is still
-being argued whether or not many real data actually follow the power
-law. \citet{Broido_2019} provides some evidence that real world networks
-rarely have degrees that follow a power law, through the use of
-statistical analysis of almost 1000 networks across a wide variety of
-fields; \citet{Voitalov_2019}, on the other hand, disagrees and claims
-that these kinds of networks are not nearly as rare and only appear so
-as a result of an unrealistic expectation of power law without
-deviations or noise.
+Most of the study into the degrees of real networks, including those
+mentioned here, has been purely descriptive in the sense that no
+information about the networks growth is revealed (aside from relating a
+power law to preferential attachment).
 
-Extreme value methods have been scarcely used when it comes to analysis
-of degree distributions and whether they follow the power law. The
-absence of these methods essentially downweights the deviation of
-vertices with the largest degrees, which are usually the most
-influential, from the power law. \citet{Voitalov_2019} proposes
-estimating the power law exponent of a degree distribution using three
-estimators based in extreme value theory, in contrast to the estimators
-generally used in network science, the degree distributions are then
-split into one of three categories (not power-law, hardly power-law or
-power-law) depending on the values of the estimates obtained from each
-of the estimators. \citet{Lee24} proposes a framework for modelling
-entire degree distributions using a combination of two generalisations
-of the power law with a model selection step that is capable of
-suggesting whether or not the data is adequately modelled by a power
-law, with one conclusion being that most data seems to have a tail
-lighter than what is implied by the power law in body whilst still being
-heavy tailed. One of the models used by \citet{Lee24} is the Zipf-IGP
-model where up until a certain threshold the data is modelled using the
-Zipf disribution (a discrete power law) and beyond it a discrete version
-of the Generalised Pareto distribution is used, we will later compare
-our model to this as they share some similarities.
-\citet{wang2022random} considers a preferential attachment model with an
-affine preference function and reciprocal edges, finding that the joint
-limiting distribution of the in-degrees and out-degrees is multivariate
-regularly varying and has the property of hidden regular variation.
+The study of network growth began to gain popularity when
+\citet{Barabasi99} showed that a network that has grown in such a way
+that nodes gain edges at a rate proportional to their degree results in
+a network that is scale-free. This is why the presence of real
+scale-free networks is such a hot topic, since if it is shown that they
+are often scale-free then a model such as preferential attachment gains
+a lot of justification for being used to model network growth. The
+preferential attachment model they introduced uses a preference
+function, \(b(k)=k+1\) where \(k\) is a node in-degree, that is fairly
+simple and provided the foundations for many modifications and
+extensions to the model. \citet{krapivsky01} considered a more general
+preference function of the form \(b(k) = (k+1)^\alpha\) and showed that
+the degree distribution only has regular variation in the tail when
+\(\alpha=1\), no regular variation when \(0<\alpha<1\) (and therefore
+not power law), and when \(\alpha>1\) a finite number of nodes end up
+with all edges after a certain point resulting in a degenerate degree
+distribution. \citet{wang2022random} returns to a linear preference
+function of the form \(b(k) = k+\varepsilon\) but adds the possibility
+for reciprocal edges to be sent, this results in joint in-degree
+out-degree distribution that is multivariate regular varying and has the
+property of hidden regular variation. \citet{rudas07} follows in the
+footsteps of \citet{krapivsky01} and considers a preferential attachment
+tree with a preference function \(b(\cdot)\) and uses theory from
+continuous time branching processes to derive a limiting degree
+distribution in terms of the preference function \(b(\cdot)\). Research
+in this area tends to only focus on the theoretical asymptotic results
+of network growth models with little analysis of real networks. This
+paper aims to address this gap, asking if a network is assumed to come
+from a preferential attachment model can we use the degree distribution
+alone to directly infer the model parameters?
 
-Evidently, the study of degree distributions is a very active but the
-vast majority of research in the field (whether using extremes or not)
-are purely descriptive in the sense that no information about the
-preference function is revealed even if preferential attachment was the
-underlying mechanism. This paper aims to address this gap, asking if the
-degree distribution of a network is assumed to come from the
-preferential attachment model, can we directly infer the model
-parameters?
+It is important to think about how we intend to consider the tail of the
+degree distributions, because if not done correctly we may end up
+essentially discounting the effects of the largest degrees (often the
+most influential nodes) deviating from a power law. For this reason we
+will use methods from discrete extreme value theory, as they best
+capture the nature of the data involved, more specifically we will use
+results from \citet{shimura12} to study how the tail of the degree
+distribution is affected by the preference function since we wish to
+align what we do with the recent findings that the a lot of real degree
+distributions are regularly varying.
 
-The remainder of this paper is as follows: Section 2 gives a detailed
-description of the PA model alongside the theoretical results for the
-limiting degree distribution with a focus on its tail behaviour in terms
-of the preference function \(b(\cdot)\). It is shown that a heavy tailed
-degree distribution can only be the result of an asymptotically linear
-preference function ending with a suggestion for a preference function
-that can guarantee heavy tails while remaining flexible.
-Section~\ref{sec-model} utilises the preference function proposed at the
-end of Section 2, and illustrates how the tail index of the degree
-distribution varies with the model parameters. Section~\ref{sec-sim}
-consists of a simulation study where networks are simulated from the
-proposed model for various parameter combinations, demonstrating that
-the parameters can be recovered from fitting the model to only the
-degree distribution. Section~\ref{sec-real} fits the model to some real
-data and provides posterior estimates for the preference function.
-Section~\ref{sec-conc} concludes the article.
+The remainder of the paper is as follows: Section~\ref{sec-tail} gives a
+detailed description of the preferential attachment model alongside the
+theoretical results for the survival of the limiting degrees with a
+focus on the tail behaviour in terms of the preference function
+\(b(\cdot)\). It is shown that a regularly varying degree distribution
+can only be the result of an asymptotically linear preference function.
+This section also introduces a preference function that can guarantee
+regular variation in the degrees while remaining flexible via a set of
+parameters that can provide behaviour similar to that in
+\citet{krapivsky01} up until a threshold. Section~\ref{sec-model}
+utilises the preference function proposed at the end of
+Section~\ref{sec-tail}, and illustrates how the extreme value index
+(EVI) of the degree distribution varies with the model parameters.
+Section~\ref{sec-sim} consists of a simulation study where networks are
+simulated from the proposed model for various parameter combinations,
+demonstrating that the parameters can be recovered from fitting the
+model to only the degree distribution. Section~\ref{sec-real} fits the
+model to some real data and provides posterior estimates for the
+preference function. Section~\ref{sec-conc} concludes the article.
 
@@
 
-The network generative model that we will focus on in this paper is the
-General Preferential Attachment model in \citet{rudas07} and is defined
-as follows:
+The model that we will focus on in this paper is the General
+Preferential Attachment model in \citet{rudas07} and is defined as
+follows:
 
@@
 \(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
-with \(\xi=1/2\), and the Uniform Attachment (UA) model where \(b(k)=c\)
-leading to a degree distribution that is light tailed.
+with EVI \(\xi=1/2\), and the Uniform Attachment (UA) model where
+\(b(k)=c\) leading to a degree distribution that is not regularly
+varying.
 
-Given condition \ref{eq-condb2} the survival function of the limiting
-degree distribution, called the limiting survival hereafter, can be
-analytically derived in the case where \(m=1\), obtained by considering
-a branching process that is equivalent to the growth of the network, as
-in \citet{rudas07}. Theorem 1 from \citet{rudas07} states that for the
-tree \(\Upsilon(t)\) at time \(t\):
+The survival function of the limiting degree distribution, called the
+limiting survival hereafter, under condition \ref{eq-condb2} can be
+analytically derived in the case where \(m=1\).
+
+Consider a continuous time branching process \(\zeta(t)\) driven by a
+Markovian pure birth process, with \(\zeta(0)=0\) abd birth rates
+depending on a non-negative function \(b(\cdot)\):
+
+\[
+\Pr(\zeta(t+dt)=k+1|\zeta(t)=k) = b(k)dt + o(dt).
+\] Now let \(\Upsilon(t)\) be the tree determined by \(\zeta(t)\) as
+follows: \(\Upsilon(t)=\{\emptyset\}\) and \(\Upsilon(t)=G\) where each
+existing node \(x\) in \(\Upsilon(t)\) gives birth to a child with rate
+\(b(\mathrm{deg}(x, \Upsilon(t))\) independently of the other nodes
+where \(\mathrm{deg}(x, \Upsilon(t))\) is the degree of node \(x\) in
+the tree \(\Upsilon(t)\) at time \(t\).
+
+Theorem 1 from \citet{rudas07} states that for the tree \(\Upsilon(t)\)
+at time \(t\):
 
@@
 We are interested in how the tail behaviour of the discrete limiting
-degree distribution is affected by the preference function \(b\), which
-can be determined by the results of \citet{shimura12}.
+degree distribution is affected by the preference function \(b\).
 
@@
 \(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is heavy-tailed with \(\bar F(k) \sim k^{-\alpha}\). On the
-other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then the
-distribution is light-tailed. This allows us to show the following:
+then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\). On
+the other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then we
+will refer to the distribution as light-tailed. This allows us to show
+the following:
 
@@
 \(\alpha < 1\). Proposition 2.1 also aligns with the fact that BA model
-produces a heavy-tailed degree distribution with index \(\xi=0.5\) by
+produces a regularly varying degree distribution with EVI \(\xi=0.5\) by
 considering the preference function \(b(k) = k + \varepsilon\), as
@@
 \(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\). So,
-in order for the degree distribution to be heavy-tailed we need that the
-limit \(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is positive.
-To determine the class of functions that will result in heavy-tailed
-limiting degree distributions, we use the following result:
-
-\begin{theorem}[Stolz-Cesàro Theorem
-\citep{cesaro}]\protect\hypertarget{thm-stolz}{}\label{thm-stolz}
-
-Let \((a_k)_{k\ge1}\) and \((b_k)_{k\ge1}\) be two sequence of real
-numbers. Assume \((a_k)_{k\ge1}\) is a strictly monotone and divergent
-sequence and the following limit exists:
-
-\[
-\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
-\] Then,
-
-\[
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
-\]
-
-\end{theorem}
+in order for the degree distribution to be regularly varying we need
+that the limit \(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is
+positive. To determine the class of functions that will result in
+regularly varying limiting degree distributions, we use the following
+result:
 
@@
 Consider a GPA model with preference function \(b(\cdot)\) satisfying
-Proposition~\ref{prp-omega}. Then the limiting degree distribution has a
-heavy tail with index \(c/\lambda^{*}\) if and only if
+Proposition~\ref{prp-omega}. Then the limiting degree distribution is
+regularly varying with EVI \(c/\lambda^{*}\) if and only if
 \(\lim_{k\rightarrow\infty}b(k)/k = c > 0\).
@@
 \lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
-\] meaning by Theorem~\ref{thm-stolz}:
+\] meaning by Theorem~\ref{thm-stolz} in Appendix \ref{sec-sup}:
 
@@
 
-Considering Proposition~\ref{prp-omega2} we propose the following
-preference function:
+Using this result we can understand how the preference function is
+directly connected with the EVI and regular variation of the degree
+distribution. We use this result to create a preference function that
+guarantees regular variation in the tail of the degree distribution,
+aligning with analysis of real networks, whilst allowing for the tail to
+deviate from the shape implied by the body. This gives the model the
+capability to produce realistic behaviour in the degrees like what was
+in \citet{Lee24} by using a piecewise function inspired by the observed
+deviation from the power law after a certain threshold.
 
@@
 This preference function, as per Proposition~\ref{prp-omega}, will
-produce a degree distribution with tail index \(\beta/\lambda^*\)
-guaranteeing a heavy tail. We study this preference function further in
-the next section.
+produce a degree distribution with EVI \(\xi=\beta/\lambda^*\)
+guaranteeing regular variation. We study this preference function
+further in the next section.
 
@@
 
-\section{Preferential Attachment with Flexible Heavy
-Tail}\label{sec-model}
+\section{A Preferential Attachment Model with Flexible Regular
+Variation}\label{sec-model}
 
@@
 linearity in the limit allows for the inclusion of sub/super-linear
-behaviour below the threshold, while simultaneously guaranteeing a heavy
-tail. In this section, we study how the limiting degree distribution
-varies with the preference function according to Equation~\ref{eq-pref}.
-Its survival is
+behaviour below the threshold, while simultaneously guaranteeing regular
+varation of the degrees. In this section, we study how the limiting
+degree distribution varies with the preference function according to
+Equation~\ref{eq-pref}. Its survival is
 
@@
 Figure~\ref{fig-polylinsurv}:
-
-\begin{verbatim}
-     a   b  eps k0 x   lambda        surv
-10 0.5 1.5 0.01 20 0 1.500094 0.006622105
-\end{verbatim}
-
-\begin{verbatim}
-[1] 0.9999374
-\end{verbatim}
 
@@
 the shape parameter of the IGP is positive and thus the distribution is
-heavy-tailed. Additionally the value of the shape parameter \(\xi\) is
-shown in Figure~\ref{fig-polyheat} for various parameter choices. The
+regularly varying. Additionally the value of the shape parameter \(\xi\)
+is shown in Figure~\ref{fig-polyheat} for various parameter choices. The
 darker and lighter regions on the heat maps correspond to a heavier and
@@
 \(\xi=0.5\).
+
+Through the connection to the IGP, fitting this model is almost
+equivalent to fitting the IGP to the degree and estimating the
+parameters but instead of only describing the shape of the degree
+distribution we would also gain estimates for the shape of the
+preference function while can help understand the mechanisms underlying
+the growth of the network.
 
@@
 
-This first subsection aims to show that the parameters of the model in
-Section~\ref{sec-model} can be recovered from simulating a network from
-the model, and fitting it to the observed degree distribution, using the
-likelihood in \eqref{eq-lh}.
+This first subsection aims to show that the parameters of the model (and
+therefore the preference function) in Section~\ref{sec-model} can be
+recovered from simulating a network from the model, and fitting it to
+the observed degree distribution, using the likelihood in \eqref{eq-lh}.
 
@@
 \begin{align*}
-\alpha&\sim \text{Ga}(1,0.01),\\
-\beta &\sim  \text{Ga}(1,0.01),\\
+\alpha&\sim \text{Gamma}(1,0.01),\\
+\beta &\sim  \text{Gamma}(1,0.01),\\
 k_0 &\sim \text{U}(1,10,000),\\
-\varepsilon &\sim \text{Ga(1,0.01)},
+\varepsilon &\sim \text{Gamma(1,0.01)},
 \end{align*}
 
-where Ga(a, b) is the Gamma distribution with shape a and rate b, and
+where Gamma(a, b) is the Gamma distribution with shape a and rate b, and
 U(a, b) is uniform distribution with lower and upper bounds a and b, to
@@
 demonstrates the usefulness of the model, as we can recover the model
-parameters fairly well from only the final degree distribution of the
-simulated network. This indicates that the method may also be applied to
-the degree distributions of real networks, estimating the model
-parameters assuming they evolved according to the GPA scheme.
+parameters well from only the final degree distribution of the simulated
+network. This indicates that the method may also be applied to the
+degree distributions of real networks, estimating the model parameters
+assuming they evolved according to the GPA scheme.
 
@@
 \item
-  \texttt{facebook-wosn-wall}: subset of network of Facebook wall pasts
+  \texttt{facebook-wosn-wall}: subset of network of Facebook wall posts
 \item
@@
   \texttt{maayan-Stelzl}: network representing interacting pairs of
-  proteins in Humans
+  proteins in humans
 \item
@@
 \item
-  \texttt{pajek-erdos}: co-authorship network around Paul Erdős
+  \texttt{pajek-erdos}: co-authorship network around Paul Erd\o``s
 \item
-  \texttt{reactome}: network of protein--protein interactions in Humans
+  \texttt{reactome}: network of protein--protein interactions in humans
 \item
@@
 the Zipf-IGP mixture model from \citet{Lee24}. In most cases, the GPA
-model does not necessarily provide an improvement in fit when compared
-to the Zipf-IGP model but where the GPA model fits well we gain
-additional information about the preference function assuming that the
-network evolved according the the GPA scheme.
+model gives a similar fit to the Zipf-IGP model but where the GPA model
+fits well we gain additional information about the preference function
+assuming that the network evolved according the the GPA scheme.
 
@@
 \newpage
+
+\section{Supplementary Results}\label{sec-sup}
+
+\begin{theorem}[Stolz-Cesàro Theorem
+\citep{cesaro}]\protect\hypertarget{thm-stolz}{}\label{thm-stolz}
+
+Let \((a_k)_{k\ge1}\) and \((b_k)_{k\ge1}\) be two sequence of real
+numbers. Assume \((a_k)_{k\ge1}\) is a strictly monotone and divergent
+sequence and the following limit exists:
+
+\[
+\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
+\] Then,
+
+\[
+\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
+\]
+
+\end{theorem}
 

### 2025-08-28 paper.qmd
Making edits

@@
 Amid the recent rise of interest in networks, there has been a debate on whether most real networks are scale-free.
-Claiming a real network is scale-free is equivalent to saying that its degree distribution follows a power law, that is the fraction of nodes with degree $k$ is proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with extreme value index $1/\alpha$.
+Claiming a real network is scale-free is equivalent to saying that its degree distribution follows a power law, that is the fraction of nodes with degree $k$ is proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with tail index $\alpha$.
 On the side against the claim is @Broido_2019who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority in real networks.
@@
 A class of asymptotically linear preference functions will be introduced and shown to guarantee regular variation in the degrees while remaining flexible up until a threshold.
-@sec-model utilises the proposed preference function and illustrates numerically how the extreme value index (EVI) of the degree distribution varies with the model parameters.
+@sec-model utilises the proposed preference function and illustrates numerically how the tail index of the degree distribution varies with the model parameters.
 @sec-sim consists of a simulation study where networks are simulated from the proposed model for various parameter combinations, demonstrating that the parameters can be recovered from fitting the model to only the degree distribution.
@@
 
-Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges from the new vertex; the target for each of these edges are selected from the vertices already in the network with weights proportional to some preference function $b(\cdot)$ of their degree, where $b: \mathbb N \mapsto \mathbb R^+\setminus\{0\}$ is such that:
+Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges from the new vertex; the target for each of these edges are selected from the vertices already in the network with weights proportional to some non-decreasing preference function $b(\cdot)$ of their degree, where $b: \mathbb N \mapsto \mathbb R^+\setminus\{0\}$ is such that:
 
@@
 
-Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with  EVI $\xi=1/2$, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
+Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
 
@@
 \Pr(\zeta(t+dt)=k+1|\zeta(t)=k) = b(k)dt + o(dt).
-$$
-Now let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$.
+$$ Now let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$.
 
-Theorem 1 from @rudas07 states that for the tree $\Upsilon(t)$ at time $t$:
+Theorem 1 from @rudas07 states that for the tree $\Upsilon(t)$ at time $t$ and a characteristic function of the tree $\varphi(\cdot)$ :
 
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-$$ {#eq-survlim} 
-where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to the growth of an individual node, that is $\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t$
+$$ {#eq-survlim} where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to the growth of an individual node, that is $\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t$
 
@@
 \bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
-$$
-which can also be written using @eq-survlim as:
+$$ which can also be written using @eq-survlim as:
 
@@
 \bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
-$${#eq-surv-origin}
+$$ {#eq-surv-origin}
 
@@
 f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)}.
-$${#eq-pmf-origin}
+$$ {#eq-pmf-origin}
 
@@
 ::: {#prp-omega}
-If $\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}$ and  $b(k) \rightarrow \infty$ as $k\rightarrow \infty$, then
-$$
+If $\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}$ and $b(k) \rightarrow \infty$ as $k\rightarrow \infty$, then $$
 \lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
@@
 @prp-omega aligns with the result from @krapivsky01 demonstrating that a sub-linear preference function will lead to a light-tailed distribution, as $\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0$ if $b(k)=k^\alpha$ where $\alpha < 1$.
-Proposition 2.1 also aligns with the fact that BA model produces a regularly varying degree distribution with EVI $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$ leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
+Proposition 2.1 also aligns with the fact that BA model produces a regularly varying degree distribution, with tail index 2, by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$ leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
 So, in order for the degree distribution to be regularly varying we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
-To determine the class of functions that will result in regularly varying limiting degree distributions, we use the following result:
+The following proposition determines the class of functions that will result in regular varying degree distributions.
 
@@
 
-:::{#prp-omega2}
-
+::: {#prp-omega2}
 Consider a GPA model with preference function $b(\cdot)$ satisfying @prp-omega.
-Then the limiting degree distribution is regularly varying with EVI $c/\lambda^{*}$ if and only if $\lim_{k\rightarrow\infty}b(k)/k = c > 0$.
+Then the limiting degree distribution is regularly varying with tail index $\lambda^{*}/c$ if and only if $\lim_{k\rightarrow\infty}b(k)/k = c > 0$.
 
@@
 
-From @prp-omega, we have that:
-$$
+From @prp-omega, we have that: $$
 \lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
@@
 \lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
-$$ Using the Stolz-Cesaro Theorem [@thm-stolz](Appendix -@sec-sup),
+$$ Using the Stolz-Cesaro Theorem [\@thm-stolz](Appendix%20-@sec-sup),
 
@@
 $$
-
 :::
@@
 \end{cases}
-$${#eq-pref} for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.
+$$ {#eq-pref} for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.
 
-Using @prp-omega, we can show that $\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta$, meaing the degree distribution obtained using this preference funnction is regularly varying with tail index $\lambda^*/\beta$.
+Using @prp-omega, we can show that $\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta$, meaning the degree distribution obtained using this preference funnction is regularly varying with tail index $\lambda^*/\beta$.
 
@@
 ```
+
 @fig-polylinsurv demonstrates that this model can capture a range of tail behaviour, including a large range of possible tail indices ranging from 0.035 ($\alpha=1.5, \beta=0.1, \varepsilon=1$) to 0.999 ($\alpha=0.5, \beta=1.5, \varepsilon=0.01$).
 
-The analytic form of the survival function  in (-@eq-polysurv), offers a natural connection to the discrete version of the generalised Pareto distribution (GP), providing a link to a well established component of discrete extremes in the literature.
-Specifically, connecting to the Integer GP (IGP) seen in @Rohrbeck_2018 with conditional survival:
-$$
+The analytic form of the survival function in ([-@eq-polysurv]), offers a natural connection to the discrete version of the generalised Pareto distribution (GP), providing a link to a well established component of discrete extremes in the literature.
+Specifically, connecting to the Integer GP (IGP) seen in @Rohrbeck_2018 with conditional survival: $$
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
@@
 
-Therefore, 
-$${#eq-igp-est}
+Therefore, $$
 \bar F(k) 
@@
 \end{cases}
-$$ meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
+$${#eq-igp-est} meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
 
@@
 Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices.
-The darker and lighter regions on the heat maps correspond to a heavier and a lighter tail, respectively, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail index as the Barabási-Albert model, that is, $\xi=0.5$.
+The darker and lighter regions on the heat maps correspond to a heavier and a lighter tail, respectively, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail index as the Barabási-Albert model.
 
@@
 
-![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2 width=100%}
+![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2 width="100%"}
 
@@
 -   `maayan-Stelzl`: network representing interacting pairs of proteins in humans
-- `moreno-blogs-blogs`: network of URLs found on the
-first pages of individual blogs
+-   `moreno-blogs-blogs`: network of URLs found on the first pages of individual blogs
 -   `opsahl−openflights`: network containing flights between airports of the world.
@@
 
+Its worth noting the recent work by @banerjee25, that provides results for more general networks beyond trees utilising the same underlying branching process.
+They consider a model that grows in the exact same way as the model that we have considered, but then they collapse the nodes of the tree resulting in a more general network much more like those in reality.
+However, there is no expression for the probability mass function of the degrees and therefore no likelihood that can be used for modeling in the same way that we have.
+In spite of this, using Remark 2 from @banerjee25 it can be shown that the degree distribution is still heavy tailed although not necessarily regularly varying.
+
+Obtaining an expression for the degree distributions of more general cases of preferential attachment models are, for the moment, seemingly inaccessible and provide a real barrier to using more realistic models in a way similar to what we have in this paper.
+
 \setcounter{section}{0}
@@
 
-# Supplementary Results{#sec-sup .appendix}
+# Supplementary Results {#sec-sup .appendix}
 
@@
 \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
-$$ 
-Then,
+$$ Then,
 
@@
 
-# Additional Plots{.appendix}
+# Additional Plots {.appendix}
 
-![: Posterior estimates of parameters for real data.](images/pars_plot.png){width=80%}
+![: Posterior estimates of parameters for real data.](images/pars_plot.png){width="80%"}
 
@@
 
-Taking the form of the GPA degree survival: $$
-\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda+b(i)}
+Taking the form of the GPA degree survival from @eq-surv-origin : $$
+\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^*+b(i)}
 $$ and substituting into the formula for $\Omega(F,n)$:
@@
 
-Clearly if $b(k)=c$ or $\lim_{k\rightarrow\infty}b(k)=c$ for some $c>0$ then $\Omega(F,k)=0$. Now consider a non-constant $b(k)$ and re-write $\Omega(F,k)$ as:
+Clearly if $\lim_{k\rightarrow\infty}b(k)=c$ for some $c>0$ then $\Omega(F,k)= 0 (= \lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}} [b(k+2)-b(k+1)])$.
+Now consider a non-constant $b(k)$ and re-write $\Omega(F,k)$ as:
 
@@
 $$
-\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
+\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma\left(\frac{x}{y}+n+1\right)}{\Gamma\left(\frac{x}{y}\right)}
 $$ and reindexing the product in the second sum,

### 2025-08-28 paper.tex
Making edits

@@
 \theoremstyle{plain}
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{plain}
 \newtheorem{theorem}{Theorem}[section]
-\theoremstyle{plain}
-\newtheorem{proposition}{Proposition}[section]
 \theoremstyle{remark}
@@
 \hypersetup{
-  pdftitle={Learning Network Growth Mechanisms from Degree Distributions in Preferential Attachment Networks},
+  pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
@@
 
-\title[Learning Network Growth Mechanisms from Degree Distributions in
-Preferential Attachment Networks]{Learning Network Growth Mechanisms
-from Degree Distributions in Preferential Attachment Networks}
+\title[Learning growth mechanisms of tail realistic preferential
+attachment models from network degree distributions]{Learning growth
+mechanisms of tail realistic preferential attachment models from network
+degree distributions}
 
@@
 Networks have become very powerful tools for representing and analysing
-complex systems, with uses in a large array of fields, from using
-stochastic block models to detect communities online \citep{Latouche11},
-to using Exponential Random Graph Models (ERGMs) to analyse the global
-trade network \citep{Setayesh22}, and even using mechanistic models to
-investigate the patterns in neural systems \citep{Betzel17}.
+complex systems, with uses in a large array of fields. In network
+science and statistics, they have been studies by various families of
+models, from stochastic block models for detecting communities online
+\citep{Latouche11}, to exponential random graph models (ERGMs) for
+analysing the global trade network \citep{Setayesh22}, and mechanistic
+models for investigating patterns in neural systems \citep{Betzel17}.
 
-Along with the recent rise of interest in networks, has been a debate
-amongst those that study them regarding the claims that most real
-networks are scale-free. This claim is most often equivalent to claiming
-that the degree distributions of real networks follow a power law, that
-is the fraction of nodes with degree \(k\) is proportional to
-\(k^{-\alpha}\), and therefore has a regularly varying tail with extreme
-value index \(1/\alpha\). \citet{Broido_2019} provide empirical
-evidence, using almost one thousand networks, that these scale-free
-networks do not make up a large majority of real networks. They compare
-the fits of a power law model against that of several non-scale-free
-models only finding strong evidence for scale-freeness in four percent
-and weak evidence in fifty-two percent of the networks. On the other
-side of this debate is \citet{Voitalov_2019} who disagrees and claims
-that these networks are not nearly as rare and only appear to so as a
-result of an unrealistic expectation of a power law without deviations
-or noise. Evidence of these deviations from a power law is shown by
+Amid the recent rise of interest in networks, there has been a debate on
+whether most real networks are scale-free. Claiming a real network is
+scale-free is equivalent to saying that its degree distribution follows
+a power law, that is the fraction of nodes with degree \(k\) is
+proportional to \(k^{-\alpha}\), and therefore has a regularly varying
+tail with tail index \(\alpha\). On the side against the claim is
+\citet{Broido_2019who} compared the fits of a power law model against
+that of several non-scale-free models to nearly a thousand networks,
+only to find strong evidence for scale-freeness in four percent and weak
+evidence in over half of the networks, thus claiming that scale-free
+networks do not make up a majority in real networks. They compare the
+fits of a power law model against that of several non-scale-free models
+only finding strong evidence for scale-freeness in four percent and weak
+evidence in fifty-two percent of the networks. On the other side of this
+debate is \citet{Voitalov_2019} who disagrees and claims that these
+networks are not nearly as rare and only appear to so as a result of an
+unrealistic expectation of a power law without deviations or noise.
+Additional evidence of these deviations from a power law is shown by
 \citet{Lee24} who demonstrate that a lot of networks are partially
-scale-free, in that the body of the degree distribution is modelled well
-by a power law while the tail is lighter than what is implied by the
-body (although still regularly varying).
+scale-free, in that the body of the degree distribution is often
+modelled well by a power law, while the tail is lighter than what is
+implied by the body, albeit still regularly varying..
 
-Most of the study into the degrees of real networks, including those
-mentioned here, has been purely descriptive in the sense that no
-information about the networks growth is revealed (aside from relating a
-power law to preferential attachment).
+Nevertheless, most studies into the appropriateness of a power law for
+the degrees of real networks, the aforementioned references included,
+have been largely descriptive in the sense that no information about the
+growth of the networks is revealed.
 
-The study of network growth began to gain popularity when
-\citet{Barabasi99} showed that a network that has grown in such a way
-that nodes gain edges at a rate proportional to their degree results in
-a network that is scale-free. This is why the presence of real
-scale-free networks is such a hot topic, since if it is shown that they
-are often scale-free then a model such as preferential attachment gains
-a lot of justification for being used to model network growth. The
-preferential attachment model they introduced uses a preference
-function, \(b(k)=k+1\) where \(k\) is a node in-degree, that is fairly
-simple and provided the foundations for many modifications and
-extensions to the model. \citet{krapivsky01} considered a more general
-preference function of the form \(b(k) = (k+1)^\alpha\) and showed that
-the degree distribution only has regular variation in the tail when
-\(\alpha=1\), no regular variation when \(0<\alpha<1\) (and therefore
-not power law), and when \(\alpha>1\) a finite number of nodes end up
-with all edges after a certain point resulting in a degenerate degree
-distribution. \citet{wang2022random} returns to a linear preference
-function of the form \(b(k) = k+\varepsilon\) but adds the possibility
-for reciprocal edges to be sent, this results in joint in-degree
-out-degree distribution that is multivariate regular varying and has the
-property of hidden regular variation. \citet{rudas07} follows in the
-footsteps of \citet{krapivsky01} and considers a preferential attachment
-tree with a preference function \(b(\cdot)\) and uses theory from
-continuous time branching processes to derive a limiting degree
-distribution in terms of the preference function \(b(\cdot)\). Research
-in this area tends to only focus on the theoretical asymptotic results
-of network growth models with little analysis of real networks. This
-paper aims to address this gap, asking if a network is assumed to come
-from a preferential attachment model can we use the degree distribution
-alone to directly infer the model parameters?
+The popularity of using the power law for network degrees can be traced
+back to the preferential attachment (PA) model popularised by
+\citet{Barabasi99}. In the general model, as new nodes join the network,
+an existing node with degree k gains edges at a rate proportional to
+\(b(k)\), where b is a non-negative preference function.
+\citet{Barabasi99} showed that, when \(b(k) = k + 1\), in the limit the
+resulting degree distribution is regularly varying with index 3.
+Subsequently, if a real network is shown to be scale-free, one can
+loosely justify preferential attachment as the underlying mechanism of
+its growth.
+
+The model from \citet{Barabasi99} provided the foundations for various
+generalisations.
+
+\citet{krapivsky01} considered \(b(k) = (k+1)^\alpha\), and showed that
+the degree distribution is not regularly varying (and therefore not
+following the power law) when \(0<\alpha<1\), and when \(\alpha>1\) a
+finite number of nodes end up with all edges after a certain point
+resulting in a degenerate degree distribution. \citet{wang2022random}
+returns to a linear preference function of the form
+\(b(k) = k+\varepsilon\) but adds the possibility for reciprocal edges
+to be sent, resulting in the joint distribution of in-degree and
+out-degree being multivariate regularly varying and having the property
+of hidden regular variation. \citet{rudas07} follows in the footsteps of
+\citet{krapivsky01} , by considering a preferential attachment tree and
+using theory from continuous branching processes, derives a limiting
+degree distribution in terms of the preference function \(b(\cdot)\).
+Nevertheless, research in this area tends to only focus on the
+theoretical asymptotic results of network growth models with little
+analysis of real networks.
+
+This paper aims to address the gap between the applied and theoretical
+works, by asking if a network is assumed to come from a preferential
+attachment model, can we use the degree distribution alone to directly
+infer the parameters of the the preference function and learn about the
+growth mechanisms? Moreover, proper consideration is given to the tail
+of the degree distribution, because otherwise the effects of the largest
+degrees, which correspond to the most influential nodes, deviating from
+the power law will be discounted.
 
@@
 essentially discounting the effects of the largest degrees (often the
-most influential nodes) deviating from a power law. For this reason we
-will use methods from discrete extreme value theory, as they best
-capture the nature of the data involved, more specifically we will use
-results from \citet{shimura12} to study how the tail of the degree
-distribution is affected by the preference function since we wish to
-align what we do with the recent findings that the a lot of real degree
-distributions are regularly varying.
+most influential nodes) deviating from a power law.
+
+For the above reason, we will use methods from discrete extreme value
+theory, in particular those by \citet{shimura12}, who provided
+theoretical guarantees for a discrete distribution to be regularly
+varying or not. Using these results, we demonstrate how the tail of the
+degree distribution is affected by \(b(\cdot)\), and subsequently
+propose a class of preference functions that is tail realistic for real
+networks. These analytical results enables the likelihood of the degree
+distribution to be expressed in terms of the parameters of \(b(\cdot)\),
+which in turn allows the underlying mechanism of the network, assumed to
+grow according to preferential attachment, to be inferred directly.
 
@@
 detailed description of the preferential attachment model alongside the
-theoretical results for the survival of the limiting degrees with a
+theoretical results for the survival of the limiting degrees, with a
 focus on the tail behaviour in terms of the preference function
-\(b(\cdot)\). It is shown that a regularly varying degree distribution
-can only be the result of an asymptotically linear preference function.
-This section also introduces a preference function that can guarantee
-regular variation in the degrees while remaining flexible via a set of
-parameters that can provide behaviour similar to that in
-\citet{krapivsky01} up until a threshold. Section~\ref{sec-model}
-utilises the preference function proposed at the end of
-Section~\ref{sec-tail}, and illustrates how the extreme value index
-(EVI) of the degree distribution varies with the model parameters.
-Section~\ref{sec-sim} consists of a simulation study where networks are
-simulated from the proposed model for various parameter combinations,
-demonstrating that the parameters can be recovered from fitting the
-model to only the degree distribution. Section~\ref{sec-real} fits the
-model to some real data and provides posterior estimates for the
-preference function. Section~\ref{sec-conc} concludes the article.
+\(b(\cdot)\). A class of asymptotically linear preference functions will
+be introduced and shown to guarantee regular variation in the degrees
+while remaining flexible up until a threshold. Section~\ref{sec-model}
+utilises the proposed preference function and illustrates numerically
+how the tail index of the degree distribution varies with the model
+parameters. Section~\ref{sec-sim} consists of a simulation study where
+networks are simulated from the proposed model for various parameter
+combinations, demonstrating that the parameters can be recovered from
+fitting the model to only the degree distribution.
+Section~\ref{sec-real} fits the model to some real data and provides
+posterior estimates for the preference function. Section~\ref{sec-conc}
+concludes the article.
 
@@
 each have no edges, at times \(t=1,2,\ldots\) a new vertex is added to
-the network bringing with it \(m\) directed edges (with the new vertex
-as the source); the target for each of these edges are selected from the
-vertices already in the network with weights proportional to some
+the network bringing with it \(m\) directed edges from the new vertex;
+the target for each of these edges are selected from the vertices
+already in the network with weights proportional to some non-decreasing
 preference function \(b(\cdot)\) of their degree, where
-\(b: N \mapsto \mathbb R^+\setminus\{0\}\) is such that:
+\(b: \mathbb N \mapsto \mathbb R^+\setminus\{0\}\) is such that:
 
@@
 \(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
-with EVI \(\xi=1/2\), and the Uniform Attachment (UA) model where
+with tail index 2, and the Uniform Attachment (UA) model where
 \(b(k)=c\) leading to a degree distribution that is not regularly
@@
 limiting survival hereafter, under condition \ref{eq-condb2} can be
-analytically derived in the case where \(m=1\).
+analytically derived in the case where \(m=1\), this is presented below.
 
 Consider a continuous time branching process \(\zeta(t)\) driven by a
-Markovian pure birth process, with \(\zeta(0)=0\) abd birth rates
+Markovian pure birth process, with \(\zeta(0)=0\) and birth rates
 depending on a non-negative function \(b(\cdot)\):
@@
 Theorem 1 from \citet{rudas07} states that for the tree \(\Upsilon(t)\)
-at time \(t\):
+at time \(t\) and a characteristic function of the tree
+\(\varphi(\cdot)\) :
 
@@
 and \(\hat\rho\) is the Laplace transform of the density of the point
-process associated with the pure birth process that corresponds to a
-nodes individual growth, that is:
-
-\[
-\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t.
-\]
+process associated with the pure birth process that corresponds to the
+growth of an individual node, that is
+\(\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t\)
 
@@
 
-\[
+\begin{equation}\phantomsection\label{eq-surv-origin}{
 \bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
-\] Therefore, the corresponding probability mass function of the degree
+}\end{equation}
+
+Therefore, the corresponding probability mass function of the degree
 distribution \(f(k) = \bar F(k-1) - \bar F(k)\) is
 
-\[
+\begin{equation}\phantomsection\label{eq-pmf-origin}{
 f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)}.
-\]
+}\end{equation}
 
@@
 
-For a distribution \(F\) with survival function \(\bar F\) and some
+For a distribution \(F\) with survival function \(\bar F\) and
 \(k\in\mathbb Z^+\) let
@@
 
-\citet{shimura12} states that if
-\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\). On
-the other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then we
-will refer to the distribution as light-tailed. This allows us to show
-the following:
+This allows us to show the following:
 
@@
 
+\citet{shimura12} states that if
+\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
+then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\). On
+the other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then we
+will refer to the distribution as light-tailed.
+
 Proposition~\ref{prp-omega} aligns with the result from
@@
 \(\alpha < 1\). Proposition 2.1 also aligns with the fact that BA model
-produces a regularly varying degree distribution with EVI \(\xi=0.5\) by
+produces a regularly varying degree distribution, with tail index 2, by
 considering the preference function \(b(k) = k + \varepsilon\), as
@@
 that the limit \(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is
-positive. To determine the class of functions that will result in
-regularly varying limiting degree distributions, we use the following
-result:
+positive. The following proposition determines the class of functions
+that will result in regular varying degree distributions.
 
@@
 Proposition~\ref{prp-omega}. Then the limiting degree distribution is
-regularly varying with EVI \(c/\lambda^{*}\) if and only if
+regularly varying with tail index \(\lambda^{*}/c\) if and only if
 \(\lim_{k\rightarrow\infty}b(k)/k = c > 0\).
@@
 \lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
-\] meaning by Theorem~\ref{thm-stolz} in Appendix \ref{sec-sup}:
+\] Using the Stolz-Cesaro Theorem
+\href{Appendix\%20-@sec-sup}{@thm-stolz},
 
@@
 Using this result we can understand how the preference function is
-directly connected with the EVI and regular variation of the degree
-distribution. We use this result to create a preference function that
-guarantees regular variation in the tail of the degree distribution,
-aligning with analysis of real networks, whilst allowing for the tail to
-deviate from the shape implied by the body. This gives the model the
-capability to produce realistic behaviour in the degrees like what was
-in \citet{Lee24} by using a piecewise function inspired by the observed
-deviation from the power law after a certain threshold.
+directly connected with the tail behaviour of the degree distribution.
+Specifically, regular variation is achieved if and only if \(b(k)\) is
+asymptotically linear with \(k\). We use this result to create a
+preference function that guarantees regular variation in the tail of the
+degree distribution, aligning with analysis of real networks, whilst
+allowing for the tail to deviate from the shape implied by the body.
+This gives the model the capability to produce realistic behaviour in
+the degrees like what was in \citet{Lee24} by using a piecewise function
+inspired by the observed deviation from the power law after a certain
+threshold:
 
@@
 
-This preference function, as per Proposition~\ref{prp-omega}, will
-produce a degree distribution with EVI \(\xi=\beta/\lambda^*\)
-guaranteeing regular variation. We study this preference function
-further in the next section.
+Using Proposition~\ref{prp-omega}, we can show that
+\(\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta\), meaning the
+degree distribution obtained using this preference funnction is
+regularly varying with tail index \(\lambda^*/\beta\).
 
@@
 
-In the previous section, we found that using a preference function with
-linearity in the limit allows for the inclusion of sub/super-linear
+In the previous section, we found that using an asymptotically linear
+preference function allows for the inclusion of sub/super-linear
 behaviour below the threshold, while simultaneously guaranteeing regular
-varation of the degrees. In this section, we study how the limiting
-degree distribution varies with the preference function according to
-Equation~\ref{eq-pref}. Its survival is
+varation of the degrees. In this section, we demonstrate the flexibility
+of the preference function in Equation~\ref{eq-pref}, with regard to the
+tail behaviour of the limiting degree distribution. Using
+Equation~\ref{eq-surv-origin}, the limiting survival is
 
@@
 
-which must be solved numerically for most parameter choices. Also, note
-that \(\lambda>\beta\).
+which has to be solved numerically for most parameter choices. Also,
+note that \(\lambda>\beta\).
 
-Some examples of the limiting degree distribution for various parameter
-combinations are shown below on log-log scale in
-Figure~\ref{fig-polylinsurv}:
+For some parameter combinations, the limiting survival \(\bar F(k)\) is
+shown on log-log scale in Figure~\ref{fig-polylinsurv}:
 
@@
 
-\caption{\label{fig-polylinsurv}Theoretical survival distributions of
-the limiting degree distributions, according to various combinations of
-\((\alpha, \beta, \varepsilon)\) and \(k_0=20\) of the proposed
-preferential attachment model.}
+\caption{\label{fig-polylinsurv}The limiting survival, according to
+various combinations of \((\alpha, \beta, \varepsilon)\) and \(k_0=20\)
+of the proposed preferential attachment model.}
 
@@
 Figure~\ref{fig-polylinsurv} demonstrates that this model can capture a
-wide variety of shapes for the survival function, including a large
-range of possible tail indices ranging from 0.035
-(\(\alpha=1.5, \beta=0.1, \varepsilon=1\)) to 0.999
-(\(\alpha=0.5, \beta=1.5, \varepsilon=0.01\)).
+range of tail behaviour, including a large range of possible tail
+indices ranging from 0.035 (\(\alpha=1.5, \beta=0.1, \varepsilon=1\)) to
+0.999 (\(\alpha=0.5, \beta=1.5, \varepsilon=0.01\)).
 
-The survival function (\ref{eq-polysurv}) can be connected to the
-discrete version of the generalised Pareto distribution (GP), called the
-Integer GP (IGP) \citep{Rohrbeck_2018} with conditional survival: \[
+The analytic form of the survival function in (\ref{eq-polysurv}),
+offers a natural connection to the discrete version of the generalised
+Pareto distribution (GP), providing a link to a well established
+component of discrete extremes in the literature. Specifically,
+connecting to the Integer GP (IGP) seen in \citet{Rohrbeck_2018} with
+conditional survival: \[
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
@@
 \(X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)\) where \(\xi\) is the shape
-parameter controlling the tail index
+parameter controlling the tail index.
 
@@
 &=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
-&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
+&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}.
 \end{align*}
 
-Therefore, \[
+Therefore, \[{#eq-igp-est}
 \bar F(k) 
@@
 \] meaning that for \(k\ge k_0\) the limiting degree distribution (for
-large \(k_0^\alpha\)) is approximated by
+large \(k_0^\alpha\)) is approximated by the
 \(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).
@@
 To assess how close of an approximation this is, the theoretical
-conditional survivals are shown in Figure~\ref{fig-approx_surv} in
-colour and their IGP approximations are shown in grey. The approximation
-seems to hold up fairly well even for large degrees and more so when
-\(\alpha\) is larger.
+conditional survival Equation~\ref{eq-polysurv} are shown in
+Figure~\ref{fig-approx_surv} in colour and their IGP approximations
+\textbf{?@eq-igp-est} are shown in grey. The approximation seems to hold
+up fairly well even for large degrees and more so when \(\alpha\) is
+larger.
 
@@
 \(\alpha\) and \(\beta\) that produce a limiting degree distribution
-with the same tail index as the Barabási-Albert model, that is,
-\(\xi=0.5\).
+with the same tail index as the Barabási-Albert model.
 
 Through the connection to the IGP, fitting this model is almost
-equivalent to fitting the IGP to the degree and estimating the
-parameters but instead of only describing the shape of the degree
-distribution we would also gain estimates for the shape of the
-preference function while can help understand the mechanisms underlying
-the growth of the network.
+equivalent to fitting the IGP to the degrees and estimating its
+parameters. However, instead of only describing the shape of the degree
+distribution, we would also gain estimates for the shape of the
+preference function, thus gaining a direct understanding the mechanisms
+underlying the growth of the network.
 
@@
 degree count vector \(\pmb n = (n_0, n_1, \ldots, n_M)\), where \(M\) is
-the maximum degree. We can then write the likelihood as:
+the maximum degree. Using Equation~\ref{eq-pmf-origin}, the likelihood
+is:
 
 \begin{align*}
-L(\pmb x,\pmb n | \pmb \theta) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
+L(\pmb n | \pmb \theta,l) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
 \end{align*}\label{eq-lh}
 
-where \(B(y,z)\) is the the beta function, and \(l\ge0\) is a variable
-that allows truncating the data such that the minimum degree is \(l\).
-This will allow the model to be fitted whilst ignoring the influence of
-the lower degrees (those less than \(l\)) as the model does not capture
-the behaviour well at the lower degrees since \citet{rudas07} only
-provides results for the case of a preferential attachment tree.
+where \(\text{B}(\cdot,\cdot)\) is the the beta function, and \(l\ge0\)
+is a quantity that allows truncating the data such that the minimum
+degree is \(l\). This will allow the model to be fitted whilst ignoring
+the influence of the lower degrees (those less than \(l\)) as the model
+does not capture the behaviour well at the lower degrees since
+\citet{rudas07} only provides results for the case of a preferential
+attachment tree.
 
@@
 
-This first subsection aims to show that the parameters of the model (and
+This subsection aims to show that the parameters of the model (and
 therefore the preference function) in Section~\ref{sec-model} can be
@@
 
-where Gamma(a, b) is the Gamma distribution with shape a and rate b, and
-U(a, b) is uniform distribution with lower and upper bounds a and b, to
-obtain a posterior distribution, up to the proportionality constant.
-Posterior samples can then be obtained by an adaptive
-Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm. For these
-simulated networks \(l=0\).
+where Gamma(\(a\), \(b\)) is the Gamma distribution with shape \(a\) and
+rate \(b\), and U(\(a\), \(b\)) is uniform distribution with lower and
+upper bounds a and b, to obtain a posterior distribution, up to the
+proportionality constant. Posterior samples can then be obtained by an
+adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm.
+For these simulated networks \(l=0\).
 
@@
 data simulated from the proposed model with various combinations of
-(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
+\((\alpha,\beta,\varepsilon)\) and \(k_0=20\).}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/mcmc_plot.jpg}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{images/mcmc_plot.jpg}
 
@@
 
-Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} shows the estimations
-demonstrates the usefulness of the model, as we can recover the model
-parameters well from only the final degree distribution of the simulated
-network. This indicates that the method may also be applied to the
-degree distributions of real networks, estimating the model parameters
-assuming they evolved according to the GPA scheme.
+Figures \ref{fig-rec1} and \ref{fig-rec2} demonstrate the usefulness of
+the model, as we can recover the model parameters well from only the
+final degree distribution of the simulated network. This indicates that
+the method may also be applied to real networks, with the assumption
+that they evolved according to the GPA scheme.
 
@@
 \item
-  \texttt{pajek-erdos}: co-authorship network around Paul Erd\o``s
+  \texttt{pajek-erdos}: co-authorship network around Paul Erdős
 \item
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
 
@@
 be modelled by the linear part of the GPA. This again shows the effects
-that small degrees have on this model, which is somewhat expected as the
-theory used for this model is for trees and none of these real networks
-(nor many real networks) are.
+that small degrees have on this model.
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
+\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
 
@@
 Figure~\ref{fig-pa} shows the estimated preference function \(b(k)\)
-alongside the 95\% credible interval on a log-log plot. Although the
+alongside the 95\% credible interval on a log-log scale. Although the
 credible interval becomes very large for the largest degrees, this is
@@
 In this paper we introduced a class of preference functions that, under
-the GPA scheme, generate a network with a flexible yet heavy-tailed
+the GPA scheme, generate a network with a flexible yet regularly varying
 degree distribution. From the simulation study we showed that the
@@
 out-degree of new nodes to vary.
+
+Its worth noting the recent work by \citet{banerjee25}, that provides
+results for more general networks beyond trees utilising the same
+underlying branching process. They consider a model that grows in the
+exact same way as the model that we have considered, but then they
+collapse the nodes of the tree resulting in a more general network much
+more like those in reality. However, there is no expression for the
+probability mass function of the degrees and therefore no likelihood
+that can be used for modeling in the same way that we have. In spite of
+this, using Remark 2 from \citet{banerjee25} it can be shown that the
+degree distribution is still heavy tailed although not necessarily
+regularly varying.
+
+Obtaining an expression for the degree distributions of more general
+cases of preferential attachment models are, for the moment, seemingly
+inaccessible and provide a real barrier to using more realistic models
+in a way similar to what we have in this paper.
 
@@
 
-Taking the form of the GPA degree survival: \[
-\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda+b(i)}
+Taking the form of the GPA degree survival from
+Equation~\ref{eq-surv-origin} : \[
+\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^*+b(i)}
 \] and substituting into the formula for \(\Omega(F,n)\):
@@
 
-Clearly if \(b(k)=c\) or \(\lim_{k\rightarrow\infty}b(k)=c\) for some
-\(c>0\) then \(\Omega(F,k)=0\). Now consider a non-constant \(b(k)\) and
-re-write \(\Omega(F,k)\) as:
+Clearly if \(\lim_{k\rightarrow\infty}b(k)=c\) for some \(c>0\) then
+\(\Omega(F,k)= 0 (= \lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}} [b(k+2)-b(k+1)])\).
+Now consider a non-constant \(b(k)\) and re-write \(\Omega(F,k)\) as:
 
@@
 \[
-\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
+\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma\left(\frac{x}{y}+n+1\right)}{\Gamma\left(\frac{x}{y}\right)}
 \] and reindexing the product in the second sum,

### 2025-09-15 paper.qmd
Implementing latest comments + change to prop 2.2

@@
 
-Networks have become very powerful tools for representing and analysing complex systems, with uses in a large array of fields.
-In network science and statistics, they have been studies by various families of models, from stochastic block models for detecting communities online [@Latouche11], to exponential random graph models (ERGMs) for analysing the global trade network [@Setayesh22], and mechanistic models for investigating patterns in neural systems [@Betzel17].
+Networks have become powerful tools for representing and analysing complex systems, with uses in a large array of fields.
+In network science and statistics, they have been studied by various families of models, from stochastic block models for detecting communities online [@Latouche11], to exponential random graph models (ERGMs) for analysing the global trade network [@Setayesh22], and mechanistic models for investigating patterns in neural systems [@Betzel17].
 
@@
 Claiming a real network is scale-free is equivalent to saying that its degree distribution follows a power law, that is the fraction of nodes with degree $k$ is proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with tail index $\alpha$.
-On the side against the claim is @Broido_2019who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority in real networks.
-They compare the fits of a power law model against that of several non-scale-free models only finding strong evidence for scale-freeness in four percent and weak evidence in fifty-two percent of the networks.
+On the side against the claim is @Broido_2019 who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority in real networks.
 On the other side of this debate is @Voitalov_2019 who disagrees and claims that these networks are not nearly as rare and only appear to so as a result of an unrealistic expectation of a power law without deviations or noise.
-Additional evidence of these deviations from a power law is shown by @Lee24 who demonstrate that a lot of networks are partially scale-free, in that the body of the degree distribution is often modelled well by a power law, while the tail is lighter than what is implied by the body, albeit still regularly varying..
-
+Additional evidence of these deviations from a power law is shown by @Lee24 who demonstrate that a lot of networks are partially scale-free, in that the body of the degree distribution is often modelled well by a power law, while the tail is lighter than what is implied by the body, albeit still regularly varying.
 Nevertheless, most studies into the appropriateness of a power law for the degrees of real networks, the aforementioned references included, have been largely descriptive in the sense that no information about the growth of the networks is revealed.
@@
 The popularity of using the power law for network degrees can be traced back to the preferential attachment (PA) model popularised by @Barabasi99.
-In the general model, as new nodes join the network, an existing node with degree k gains edges at a rate proportional to $b(k)$, where b is a non-negative preference function.
-@Barabasi99 showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 3.
-Subsequently, if a real network is shown to be scale-free, one can loosely justify preferential attachment as the underlying mechanism of its growth.
+In the general model, as new nodes join the network, an existing node with degree $k$ gains edges at a rate proportional to $b(k)$, where $b(\cdot)$ is a non-negative preference function.
+@Barabasi99 showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 2.
+Subsequently, if a real network is shown to be scale-free, one can loosely justify PA as the underlying mechanism of its growth.
 
-The model from @Barabasi99 provided the foundations for various generalisations.
-
-@krapivsky01 considered $b(k) = (k+1)^\alpha$, and showed that the degree distribution is not regularly varying (and therefore not following the power law) when $0<\alpha<1$, and when $\alpha>1$ a finite number of nodes end up with all edges after a certain point resulting in a degenerate degree distribution.
+The model from @Barabasi99 provided the foundations for various generalisations — @krapivsky01 considered $b(k) = (k+1)^\alpha$, and showed that the degree distribution is not regularly varying (and therefore not following the power law) when $0<\alpha<1$, and when $\alpha>1$ a finite number of nodes end up with all edges after a certain point resulting in a degenerate degree distribution.
 @wang2022random returns to a linear preference function of the form $b(k) = k+\varepsilon$ but adds the possibility for reciprocal edges to be sent, resulting in the joint distribution of in-degree and out-degree being multivariate regularly varying and having the property of hidden regular variation.
-@rudas07 follows in the footsteps of @krapivsky01 , by considering a preferential attachment tree and using theory from continuous branching processes, derives a limiting degree distribution in terms of the preference function $b(\cdot)$.
+@rudas07 follows in the footsteps of @krapivsky01, by considering a PA tree and using theory from continuous branching processes, derives a limiting degree distribution in terms of the preference function $b(\cdot)$.
 Nevertheless, research in this area tends to only focus on the theoretical asymptotic results of network growth models with little analysis of real networks.
 
-This paper aims to address the gap between the applied and theoretical works, by asking if a network is assumed to come from a preferential attachment model, can we use the degree distribution alone to directly infer the parameters of the the preference function and learn about the growth mechanisms?
+This paper aims to address the gap between the applied and theoretical works, by asking if a network is assumed to come from a PA model, can we use the degree distribution alone to directly infer the parameters of the the preference function and learn about the growth mechanisms?
 Moreover, proper consideration is given to the tail of the degree distribution, because otherwise the effects of the largest degrees, which correspond to the most influential nodes, deviating from the power law will be discounted.
 
-It is important to think about how we intend to consider the tail of the degree distributions, because if not done correctly we may end up essentially discounting the effects of the largest degrees (often the most influential nodes) deviating from a power law.
+As the degrees of networks are discrete values, we will use methods from discrete extreme value theory, in particular those by @shimura12, who provided theoretical guarantees for a discrete distribution to be regularly varying or not.
+Using these results, we demonstrate how the tail of the degree distribution is affected by $b(\cdot)$, and subsequently propose a class of preference functions that is tail realistic for real networks.
+These analytical results enables the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to PA, to be inferred directly.
 
-For the above reason, we will use methods from discrete extreme value theory, in particular those by @shimura12, who provided theoretical guarantees for a discrete distribution to be regularly varying or not.
-Using these results, we demonstrate how the tail of the degree distribution is affected by $b(\cdot)$, and subsequently propose a class of preference functions that is tail realistic for real networks.
-These analytical results enables the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to preferential attachment, to be inferred directly.
-
-The remainder of the paper is as follows: @sec-tail gives a detailed description of the preferential attachment model alongside the theoretical results for the survival of the limiting degrees, with a focus on the tail behaviour in terms of the preference function $b(\cdot)$.
-A class of asymptotically linear preference functions will be introduced and shown to guarantee regular variation in the degrees while remaining flexible up until a threshold.
+The remainder of the paper is as follows: @sec-tail gives a detailed description of the PA model alongside the theoretical results for the survival function of the limiting degree distribution, with a focus on the tail behaviour in terms of the preference function $b(\cdot)$.
+A class of asymptotically linear preference functions will be introduced and shown to guarantee regular variation in the degree distribution while remaining flexible up until a threshold.
 @sec-model utilises the proposed preference function and illustrates numerically how the tail index of the degree distribution varies with the model parameters.
-@sec-sim consists of a simulation study where networks are simulated from the proposed model for various parameter combinations, demonstrating that the parameters can be recovered from fitting the model to only the degree distribution.
+The simulation study in @sec-sim demonstrates that the parameters can be recovered from fitting the model to only the degree distribution.
 @sec-real fits the model to some real data and provides posterior estimates for the preference function.
-@sec-conc concludes the article.
+@sec-conc provides a discussion of this paper and possible avenues for future work.
 
@@
 
-The model that we will focus on in this paper is the General Preferential Attachment model in @rudas07 and is defined as follows:
+The model that we will focus on in this paper is the General Preferential Attachment (GPA) model in @rudas07 and is defined as follows:
 
@@
 
-Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
+Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
 
-The survival function of the limiting degree distribution, called the limiting survival hereafter, under condition [-@eq-condb2] can be analytically derived in the case where $m=1$, this is presented below.
+The survival function of the limiting degree distribution, called the limiting survival hereafter, under condition [-@eq-condb2] can be analytically derived in the case where $m=1$, which is presented below.
 
@@
 $$
-\Pr(\zeta(t+dt)=k+1|\zeta(t)=k) = b(k)dt + o(dt).
-$$ Now let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$.
+\Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
+$$ Now let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
 
@@
 $$
-\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-$$ {#eq-survlim} where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to the growth of an individual node, that is $\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t$
+\lim*{t*\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum{x\in\Upsilon(t)}\varphi(\Upsilon(t)\_{\downarrow x}) = \lambda\^\* \int\_0\infty e{-\lambda\^\* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t
+$$ {#eq-survlim}
+
+where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to the growth of an individual node, that is $\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t$.
 
@@
 $$
-\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
+\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]\text{d}t}{\int_0^\infty e^{-\lambda^* t}\text{d}t} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
 $$ {#eq-surv-origin}
@@
 We are interested in how the tail behaviour of the discrete limiting degree distribution is affected by the preference function $b$.
+In order to study this, we use the quantity $\Omega(F, k)$ as stated in @shimura12 and is repeated below:
 
-For a distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ let
+For a discrete distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ let
 
@@
 
-This allows us to show the following:
+The following proposition gives the limiting behaviour of $\Omega(F,k)$ when \$F\$ is a limiting degree distribution resulting from the GPA model with preference function $b(\cdot)$.
 
@@
 @prp-omega aligns with the result from @krapivsky01 demonstrating that a sub-linear preference function will lead to a light-tailed distribution, as $\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0$ if $b(k)=k^\alpha$ where $\alpha < 1$.
-Proposition 2.1 also aligns with the fact that BA model produces a regularly varying degree distribution, with tail index 2, by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$ leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
+@prp-omega also aligns with the fact that BA model produces a regularly varying degree distribution, with tail index 2, by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$, leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
 So, in order for the degree distribution to be regularly varying we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
@@
 ::: {#prp-omega2}
-Consider a GPA model with preference function $b(\cdot)$ satisfying @prp-omega.
-Then the limiting degree distribution is regularly varying with tail index $\lambda^{*}/c$ if and only if $\lim_{k\rightarrow\infty}b(k)/k = c > 0$.
+The limiting survival of a GPA model is regularly varying if and only if the preference function is *eventually* of the form:
+
+\begin{align*}
+&b(k) = \beta k + r(k)\\
+&\text{where}\\
+&r(k+1)-r(k)\rightarrow 0,
+\end{align*}
+
+and if the preference function is of this form then the tail index is exactly $\lambda^*/\beta$ where $\hat\rho(\lambda\^\*) = 1$.
 
@@
 
-From @prp-omega, we have that: $$
-\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
-$$ Now, setting $b_k = b(k)$ and $a_k = k$:
+Letting $b(k) = \beta k  + o(k)$ and using @prp-omega we have that:
 
 $$
-\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
-$$ Using the Stolz-Cesaro Theorem [\@thm-stolz](Appendix%20-@sec-sup),
+\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \lim_{k\rightarrow\infty} \frac{\beta + r(k+1) - r(k)}{\lambda^*} = \frac{\beta}{\lambda^*},
+$$
+
+and thus the tail index of the limiting survival must be $\lambda^*/\beta$.
+
+\
+If we now assume that the limiting survival is regularly varying with tail index $\delta$, we must have that:
 
 $$
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0\qquad \square.
+\lim_{k\rightarrow\infty}\Omega(F,k) = \delta
 $$
+
+and again by @prp-omega
+
+$$
+\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \delta
+$$
+
+Letting $b_k=b(k)$ and $a_k = k$ we can use the Stolz-Cesàro theorem (@thm-stolz) to obtain:
+
+$$
+\lim_{k\rightarrow\infty}\frac{b_{k+1}-b_k}{a_{k+1}-a_k} = \delta \implies \lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \delta
+$$
+
+and so
+
+$$
+\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \delta
+$$
+
+which $b(k) = \beta k + r(k)$ fulfills.$\square$
 :::
@@
 
-Using @prp-omega, we can show that $\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta$, meaning the degree distribution obtained using this preference funnction is regularly varying with tail index $\lambda^*/\beta$.
+Using @prp-omega, we can show that $\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta$, meaning the degree distribution obtained using this preference function is regularly varying with tail index $\lambda^*/\beta$.
 
@@
 
-# A Preferential Attachment Model with Flexible Regular Variation {#sec-model}
+# A PA Model with Flexible Regular Variation {#sec-model}
 
-In the previous section, we found that using an asymptotically linear preference function allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing regular varation of the degrees.
+In the previous section, we found that using an asymptotically linear preference function allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing regular variation of the degrees.
 In this section, we demonstrate the flexibility of the preference function in @eq-pref, with regard to the tail behaviour of the limiting degree distribution.
@@
 #| fig-height: 4
-#| fig-cap: The limiting survival, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed preferential attachment model.
+#| fig-cap: The limiting survival, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed PA model.
 #| label: fig-polylinsurv
@@
 
-The analytic form of the survival function in ([-@eq-polysurv]), offers a natural connection to the discrete version of the generalised Pareto distribution (GP), providing a link to a well established component of discrete extremes in the literature.
-Specifically, connecting to the Integer GP (IGP) seen in @Rohrbeck_2018 with conditional survival: $$
+The analytic form of the survival function in ([-@eq-polysurv]), offers a natural connection to the discrete version of the generalised Pareto (GP) distribution , providing a link to a well established component of discrete extremes in the literature.
+Specifically, it is connected to the Integer GP (IGP) distribution seen in @Rohrbeck_2018 with conditional survival: $$
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
-$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$, denoted as $X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)$ where $\xi$ is the shape parameter controlling the tail index.
+$$ for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$, denoted as $X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)$ where $\xi$ is the shape parameter and reciprocal of the tail index.
 
@@
 \end{cases}
-$${#eq-igp-est} meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.
+$$ {#eq-igp-est} meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$ distribution.
 
 To assess how close of an approximation this is, the theoretical conditional survival [@eq-polysurv] are shown in @fig-approx_surv in colour and their IGP approximations [@eq-igp-est] are shown in grey.
-The approximation seems to hold up fairly well even for large degrees and more so when $\alpha$ is larger.
+The approximation holds up fairly well even for large degrees and more so when $\alpha$ is larger.
 
@@
 #| fig-height: 4
-#| fig-cap: Theoretical conditional survivals (grey) alongside their IGP approximations (coloured).
+#| fig-cap: Theoretical conditional survivals (grey) of the proposed model alongside their IGP approximations (coloured).
 #| label: fig-approx_surv
@@
 
-In agreement with @prp-omega2, $\beta>0$ ensures that the shape parameter of the IGP is positive and thus the distribution is regularly varying.
-Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices.
-The darker and lighter regions on the heat maps correspond to a heavier and a lighter tail, respectively, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail index as the Barabási-Albert model.
+In agreement with @prp-omega2, $\beta>0$ ensures that the shape parameter of the IGP distribution is positive and thus the distribution is regularly varying.
+Additionally the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices.
+The darker and lighter regions on the heat maps correspond to a heavier and a lighter tail, respectively, and the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail index as the BA model.
 
-Through the connection to the IGP, fitting this model is almost equivalent to fitting the IGP to the degrees and estimating its parameters.
+Through the connection implied by @eq-igp-est, fitting the proposed model is almost equivalent to fitting the IGP distribution to the degrees and estimating its parameters.
 However, instead of only describing the shape of the degree distribution, we would also gain estimates for the shape of the preference function, thus gaining a direct understanding the mechanisms underlying the growth of the network.
@@
 
-where $\text{B}(\cdot,\cdot)$ is the the beta function, and $l\ge0$ is a quantity that allows truncating the data such that the minimum degree is $l$.
-This will allow the model to be fitted whilst ignoring the influence of the lower degrees (those less than $l$) as the model does not capture the behaviour well at the lower degrees since @rudas07 only provides results for the case of a preferential attachment tree.
+where $\text{B}(\cdot,\cdot)$ is the the beta function, $\pmb \theta = (\alpha, \varepsilon, k_0,\beta)$, and $l\ge0$ is a quantity that allows truncating the data such that the minimum degree is $l$.
+This will allow the model to be fitted whilst ignoring the influence of the lower degrees (those less than $l$) as the model does not capture the behaviour well at the lower degrees, since @rudas07 only provides results for the case of a preferential attachment tree.
 
@@
 
-The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongisde the priors:
+The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongside the priors:
 
@@
 
-where Gamma($a$, $b$) is the Gamma distribution with shape $a$ and rate $b$, and U($a$, $b$) is uniform distribution with lower and upper bounds a and b, to obtain a posterior distribution, up to the proportionality constant.
+where Gamma($a$, $b$) is the gamma distribution with shape $a$ and rate $b$, and U($a$, $b$) is uniform distribution with lower and upper bounds $a$ and $b$, to obtain a posterior distribution, up to the proportionality constant.
 Posterior samples can then be obtained by an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm.
@@
 #| fig-height: 8
-#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of $(\alpha,\beta,\varepsilon)$ and $k_0=20$.
+#| fig-cap: Empirical (black dots) and fitted (red line) survival functions for data simulated from the proposed model with various combinations of $(\alpha,\beta,\epsilon)$ and $k_0=20$.
+The 95% credible (dashed lines) are included but too narrow to be seen clearly.
 #| label: fig-rec1
@@
 
-![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2 width="100%"}
+![: Posterior densities of parameters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
+True parameter values shown with black dots.](images/mcmc_plot.jpg){#fig-rec2 width="100%"}
 
@@
 
-In this section, we fit the proposed model to the degree distributions of various real networks and learn about the mechanics of their growth.
+In this subsection, we fit the proposed model to the degree distributions of various real networks and learn about the mechanics of their growth.
 While we also compare the fit to that of the mixture distribution by @Lee24 we note that the proposed method has the additional benefit of learning directly about the growth of a network from the inference results.
@@
 #| fig-height: 9
-#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red).
+#| fig-cap: Empirical (black dots) and posterior medians (solid red) of the fitted survival function for several real data sets and their 95% credible intervals (dotted red).
 #| label: fig-real1
@@
 @fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model from @Lee24.
-In most cases, the GPA model gives a similar fit to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme.
+In most cases, the GPA model gives a similar fit to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function, assuming that the network evolved according the the GPA scheme.
+
+![ Posterior densities of the parameters of the proposed model fitted to real data.](images/pars_plot.png){width="80%"}
 
@@
 Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution.
-In the cases of substantial discrepancies, it is either because the GPA model fits the tail better than the Zipf-IGP model does, or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA.
+In the cases of substantial discrepancies, it is either because the GPA model fits the tail better than the Zipf-IGP model does, or because of the threshold being too low, forcing almost all of the data to be modelled by the linear part of the GPA.
 This again shows the effects that small degrees have on this model.
@@
 #| fig-height: 9
-#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red).
+#| fig-cap: Posterior estimates of shape parameter of Zipf-IGP distribution (right) and the analagous quantity obtained from fitting the proposed model (left).
 #| label: fig-shapes
@@
 Looking at the shape of the preference function, there appears to be two distinct shapes of preference function.
-The first appears mostly flat (similar to uniform attachment) for the smallest degrees and then after a threshold preferential attachment kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`.
-The second distinct shape appears to provide some clear preferential attachment behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.
+The first appears mostly flat (similar to uniform attachment) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`.
+The second distinct shape appears to provide some clear PA behaviour that then slows down after a certain point, and examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.
 This slowing down could be viewed as a kind of diminishing returns on the degree of a vertex i.e.
@@
 #| fig-height: 9
-#| fig-cap: Posterior estimate for preference function (solid) with 95% credible interval (dashed) on log-log scale.
+#| fig-cap: Posterior median for the preference function (solid) with 95% credible interval (dashed) on log-log scale.
 #| label: fig-pa
@@
 
-One limitation of this method is that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model as a result of using theory developed for trees and applying it to general networks.
+This paper contributes to the understanding of the relationship between the mechanism underlying a networks growth and the resulting degree distribution.
+As well as, demonstrating that under certain conditions information about this mechanism can be garnered from the degrees alone.
+Perhaps in future, something similar could be done by looking at another statistic for the network (e.g.
+the triangle distribution) and using it in combination with the degrees to gain  further insights into the growth mechanism.
+
+One limitation of this method is that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model, as a result of using theory developed for trees and applying it to general networks.
 Future work could apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.
@@
 However, there is no expression for the probability mass function of the degrees and therefore no likelihood that can be used for modeling in the same way that we have.
-In spite of this, using Remark 2 from @banerjee25 it can be shown that the degree distribution is still heavy tailed although not necessarily regularly varying.
+In spite of this, using Remark 2 from @banerjee25 it can be shown that the survival function of the limiting degree distribution (when using a preference function of the form from @prp-omega2) can be bounded by two regularly varying functions showing that the degree distribution is still heavy tailed, although not necessarily regularly varying.
 
-Obtaining an expression for the degree distributions of more general cases of preferential attachment models are, for the moment, seemingly inaccessible and provide a real barrier to using more realistic models in a way similar to what we have in this paper.
+Obtaining an expression for the degree distributions of more general cases of preferential attachment models are, for the moment, seemingly inaccessible and provide a real barrier to using more realistic models in a way similar to what we have in this paper, we leave these open problems for  future analysis.
 
@@
 :::
-
-# Additional Plots {.appendix}
-
-![: Posterior estimates of parameters for real data.](images/pars_plot.png){width="80%"}
 

### 2025-09-15 paper.tex
Implementing latest comments + change to prop 2.2

@@
 
-Networks have become very powerful tools for representing and analysing
+Networks have become powerful tools for representing and analysing
 complex systems, with uses in a large array of fields. In network
-science and statistics, they have been studies by various families of
+science and statistics, they have been studied by various families of
 models, from stochastic block models for detecting communities online
@@
 tail with tail index \(\alpha\). On the side against the claim is
-\citet{Broido_2019who} compared the fits of a power law model against
+\citet{Broido_2019} who compared the fits of a power law model against
 that of several non-scale-free models to nearly a thousand networks,
@@
 evidence in over half of the networks, thus claiming that scale-free
-networks do not make up a majority in real networks. They compare the
-fits of a power law model against that of several non-scale-free models
-only finding strong evidence for scale-freeness in four percent and weak
-evidence in fifty-two percent of the networks. On the other side of this
-debate is \citet{Voitalov_2019} who disagrees and claims that these
-networks are not nearly as rare and only appear to so as a result of an
-unrealistic expectation of a power law without deviations or noise.
-Additional evidence of these deviations from a power law is shown by
-\citet{Lee24} who demonstrate that a lot of networks are partially
+networks do not make up a majority in real networks. On the other side
+of this debate is \citet{Voitalov_2019} who disagrees and claims that
+these networks are not nearly as rare and only appear to so as a result
+of an unrealistic expectation of a power law without deviations or
+noise. Additional evidence of these deviations from a power law is shown
+by \citet{Lee24} who demonstrate that a lot of networks are partially
 scale-free, in that the body of the degree distribution is often
 modelled well by a power law, while the tail is lighter than what is
-implied by the body, albeit still regularly varying..
-
-Nevertheless, most studies into the appropriateness of a power law for
-the degrees of real networks, the aforementioned references included,
-have been largely descriptive in the sense that no information about the
-growth of the networks is revealed.
+implied by the body, albeit still regularly varying. Nevertheless, most
+studies into the appropriateness of a power law for the degrees of real
+networks, the aforementioned references included, have been largely
+descriptive in the sense that no information about the growth of the
+networks is revealed.
 
@@
 \citet{Barabasi99}. In the general model, as new nodes join the network,
-an existing node with degree k gains edges at a rate proportional to
-\(b(k)\), where b is a non-negative preference function.
+an existing node with degree \(k\) gains edges at a rate proportional to
+\(b(k)\), where \(b(\cdot)\) is a non-negative preference function.
 \citet{Barabasi99} showed that, when \(b(k) = k + 1\), in the limit the
-resulting degree distribution is regularly varying with index 3.
+resulting degree distribution is regularly varying with index 2.
 Subsequently, if a real network is shown to be scale-free, one can
-loosely justify preferential attachment as the underlying mechanism of
-its growth.
+loosely justify PA as the underlying mechanism of its growth.
 
 The model from \citet{Barabasi99} provided the foundations for various
-generalisations.
-
-\citet{krapivsky01} considered \(b(k) = (k+1)^\alpha\), and showed that
-the degree distribution is not regularly varying (and therefore not
-following the power law) when \(0<\alpha<1\), and when \(\alpha>1\) a
-finite number of nodes end up with all edges after a certain point
-resulting in a degenerate degree distribution. \citet{wang2022random}
-returns to a linear preference function of the form
-\(b(k) = k+\varepsilon\) but adds the possibility for reciprocal edges
-to be sent, resulting in the joint distribution of in-degree and
-out-degree being multivariate regularly varying and having the property
-of hidden regular variation. \citet{rudas07} follows in the footsteps of
-\citet{krapivsky01} , by considering a preferential attachment tree and
-using theory from continuous branching processes, derives a limiting
-degree distribution in terms of the preference function \(b(\cdot)\).
+generalisations --- \citet{krapivsky01} considered
+\(b(k) = (k+1)^\alpha\), and showed that the degree distribution is not
+regularly varying (and therefore not following the power law) when
+\(0<\alpha<1\), and when \(\alpha>1\) a finite number of nodes end up
+with all edges after a certain point resulting in a degenerate degree
+distribution. \citet{wang2022random} returns to a linear preference
+function of the form \(b(k) = k+\varepsilon\) but adds the possibility
+for reciprocal edges to be sent, resulting in the joint distribution of
+in-degree and out-degree being multivariate regularly varying and having
+the property of hidden regular variation. \citet{rudas07} follows in the
+footsteps of \citet{krapivsky01}, by considering a PA tree and using
+theory from continuous branching processes, derives a limiting degree
+distribution in terms of the preference function \(b(\cdot)\).
 Nevertheless, research in this area tends to only focus on the
@@
 This paper aims to address the gap between the applied and theoretical
-works, by asking if a network is assumed to come from a preferential
-attachment model, can we use the degree distribution alone to directly
-infer the parameters of the the preference function and learn about the
-growth mechanisms? Moreover, proper consideration is given to the tail
-of the degree distribution, because otherwise the effects of the largest
-degrees, which correspond to the most influential nodes, deviating from
-the power law will be discounted.
-
-It is important to think about how we intend to consider the tail of the
-degree distributions, because if not done correctly we may end up
-essentially discounting the effects of the largest degrees (often the
-most influential nodes) deviating from a power law.
-
-For the above reason, we will use methods from discrete extreme value
-theory, in particular those by \citet{shimura12}, who provided
-theoretical guarantees for a discrete distribution to be regularly
-varying or not. Using these results, we demonstrate how the tail of the
-degree distribution is affected by \(b(\cdot)\), and subsequently
-propose a class of preference functions that is tail realistic for real
-networks. These analytical results enables the likelihood of the degree
-distribution to be expressed in terms of the parameters of \(b(\cdot)\),
-which in turn allows the underlying mechanism of the network, assumed to
-grow according to preferential attachment, to be inferred directly.
+works, by asking if a network is assumed to come from a PA model, can we
+use the degree distribution alone to directly infer the parameters of
+the the preference function and learn about the growth mechanisms?
+Moreover, proper consideration is given to the tail of the degree
+distribution, because otherwise the effects of the largest degrees,
+which correspond to the most influential nodes, deviating from the power
+law will be discounted.
+
+As the degrees of networks are discrete values, we will use methods from
+discrete extreme value theory, in particular those by \citet{shimura12},
+who provided theoretical guarantees for a discrete distribution to be
+regularly varying or not. Using these results, we demonstrate how the
+tail of the degree distribution is affected by \(b(\cdot)\), and
+subsequently propose a class of preference functions that is tail
+realistic for real networks. These analytical results enables the
+likelihood of the degree distribution to be expressed in terms of the
+parameters of \(b(\cdot)\), which in turn allows the underlying
+mechanism of the network, assumed to grow according to PA, to be
+inferred directly.
 
 The remainder of the paper is as follows: Section~\ref{sec-tail} gives a
-detailed description of the preferential attachment model alongside the
-theoretical results for the survival of the limiting degrees, with a
+detailed description of the PA model alongside the theoretical results
+for the survival function of the limiting degree distribution, with a
 focus on the tail behaviour in terms of the preference function
 \(b(\cdot)\). A class of asymptotically linear preference functions will
-be introduced and shown to guarantee regular variation in the degrees
-while remaining flexible up until a threshold. Section~\ref{sec-model}
-utilises the proposed preference function and illustrates numerically
-how the tail index of the degree distribution varies with the model
-parameters. Section~\ref{sec-sim} consists of a simulation study where
-networks are simulated from the proposed model for various parameter
-combinations, demonstrating that the parameters can be recovered from
-fitting the model to only the degree distribution.
+be introduced and shown to guarantee regular variation in the degree
+distribution while remaining flexible up until a threshold.
+Section~\ref{sec-model} utilises the proposed preference function and
+illustrates numerically how the tail index of the degree distribution
+varies with the model parameters. The simulation study in
+Section~\ref{sec-sim} demonstrates that the parameters can be recovered
+from fitting the model to only the degree distribution.
 Section~\ref{sec-real} fits the model to some real data and provides
 posterior estimates for the preference function. Section~\ref{sec-conc}
-concludes the article.
+provides a discussion of this paper and possible avenues for future
+work.
 
@@
 The model that we will focus on in this paper is the General
-Preferential Attachment model in \citet{rudas07} and is defined as
+Preferential Attachment (GPA) model in \citet{rudas07} and is defined as
 follows:
@@
 Special cases of this model include the Barabási-Albert (BA) model when
-\(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
-with tail index 2, and the Uniform Attachment (UA) model where
-\(b(k)=c\) leading to a degree distribution that is not regularly
-varying.
+\(b(k) = k+1\), which in the limit of \(t\rightarrow \infty\) leads to a
+power-law degree distribution with tail index 2, and the Uniform
+Attachment (UA) model where \(b(k)=c\) leading to a degree distribution
+that is not regularly varying.
 
@@
 limiting survival hereafter, under condition \ref{eq-condb2} can be
-analytically derived in the case where \(m=1\), this is presented below.
+analytically derived in the case where \(m=1\), which is presented
+below.
 
@@
 \[
-\Pr(\zeta(t+dt)=k+1|\zeta(t)=k) = b(k)dt + o(dt).
+\Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
 \] Now let \(\Upsilon(t)\) be the tree determined by \(\zeta(t)\) as
@@
 where \(\mathrm{deg}(x, \Upsilon(t))\) is the degree of node \(x\) in
-the tree \(\Upsilon(t)\) at time \(t\).
+the tree \(\Upsilon(t)\) at time \(t\), denote by
+\(\Upsilon(t)_{\downarrow x}\) the tree when treating node \(x\) as the
+root.
 
@@
 \begin{equation}\phantomsection\label{eq-survlim}{
-\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
-}\end{equation} where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\)
-and \(\hat\rho\) is the Laplace transform of the density of the point
-process associated with the pure birth process that corresponds to the
-growth of an individual node, that is
-\(\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t\)
+\lim*{t*\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum{x\in\Upsilon(t)}\varphi(\Upsilon(t)\_{\downarrow x}) = \lambda\^\* \int\_0\infty e{-\lambda\^\* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t
+}\end{equation}
+
+where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\) and \(\hat\rho\)
+is the Laplace transform of the density of the point process associated
+with the pure birth process that corresponds to the growth of an
+individual node, that is
+\(\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t\).
 
@@
 \begin{equation}\phantomsection\label{eq-surv-origin}{
-\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
+\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]\text{d}t}{\int_0^\infty e^{-\lambda^* t}\text{d}t} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}.
 }\end{equation}
@@
 We are interested in how the tail behaviour of the discrete limiting
-degree distribution is affected by the preference function \(b\).
-
-For a distribution \(F\) with survival function \(\bar F\) and
+degree distribution is affected by the preference function \(b\). In
+order to study this, we use the quantity \(\Omega(F, k)\) as stated in
+\citet{shimura12} and is repeated below:
+
+For a discrete distribution \(F\) with survival function \(\bar F\) and
 \(k\in\mathbb Z^+\) let
@@
 
-This allows us to show the following:
+The following proposition gives the limiting behaviour of
+\(\Omega(F,k)\) when \$F\$ is a limiting degree distribution resulting
+from the GPA model with preference function \(b(\cdot)\).
 
@@
 \(\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0\) if \(b(k)=k^\alpha\) where
-\(\alpha < 1\). Proposition 2.1 also aligns with the fact that BA model
-produces a regularly varying degree distribution, with tail index 2, by
-considering the preference function \(b(k) = k + \varepsilon\), as
-\(\lim_{k\rightarrow\infty}b(k+1)-b(k)=1\) leaving the tail index to be
+\(\alpha < 1\). Proposition~\ref{prp-omega} also aligns with the fact
+that BA model produces a regularly varying degree distribution, with
+tail index 2, by considering the preference function
+\(b(k) = k + \varepsilon\), as
+\(\lim_{k\rightarrow\infty}b(k+1)-b(k)=1\), leaving the tail index to be
 \(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\). So,
@@
 
-Consider a GPA model with preference function \(b(\cdot)\) satisfying
-Proposition~\ref{prp-omega}. Then the limiting degree distribution is
-regularly varying with tail index \(\lambda^{*}/c\) if and only if
-\(\lim_{k\rightarrow\infty}b(k)/k = c > 0\).
+The limiting survival of a GPA model is regularly varying if and only if
+the preference function is \emph{eventually} of the form:
+
+\begin{align*}
+b(k) &= \beta k + r(k),\\
+&\text{and}\\
+r(k+1)-r(k)&\rightarrow 0,
+\end{align*}
+
+and if the preference function is of this form then the tail index is
+exactly \(\lambda^*/\beta\) where \(\hat\rho(\lambda\^\*) = 1\).
 
@@
 
-From Proposition~\ref{prp-omega}, we have that: \[
-\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = c>0.
-\] Now, setting \(b_k = b(k)\) and \(a_k = k\):
-
-\[
-\lim_{k\rightarrow\infty}[b(k+1)-b(k)] = \lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = c>0,
-\] Using the Stolz-Cesaro Theorem
-\href{Appendix\%20-@sec-sup}{@thm-stolz},
-
-\[
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \lim_{k\rightarrow\infty}\frac{b(k)}{k} = c>0\qquad \square.
+Letting \(b(k) = \beta k  + o(k)\) and using Proposition~\ref{prp-omega}
+we have that:
+
+\[
+\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \lim_{k\rightarrow\infty} \frac{\beta + o(k+1) - o(k)}{\lambda^*} = \frac{\beta}{\lambda^*},
 \]
+
+and thus the tail index of the limiting survival must be
+\(\lambda^*/\beta\).
+
+\hfill\break
+If we now assume that the limiting survival is regularly varying with
+tail index \(\delta\), we must have that:
+
+\[
+\lim_{k\rightarrow\infty}\Omega(F,k) = \delta
+\]
+
+and again by Proposition~\ref{prp-omega}
+
+\[
+\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \delta
+\]
+
+Letting \(b_k=b(k)\) and \(a_k = k\) we can use the Stolz-Cesàro theorem
+(Theorem~\ref{thm-stolz}) to obtain:
+
+\[
+\lim_{k\rightarrow\infty}\frac{b_{k+1}-b_k}{a_{k+1}-a_k} = \delta \implies \lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \delta
+\]
+
+and so
+
+\[
+\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \delta
+\]
+
+which \(b(k) = \beta k + o(k)\) fulfills.\(\square\)
 
@@
 \(\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta\), meaning the
-degree distribution obtained using this preference funnction is
-regularly varying with tail index \(\lambda^*/\beta\).
+degree distribution obtained using this preference function is regularly
+varying with tail index \(\lambda^*/\beta\).
 
@@
 
-\section{A Preferential Attachment Model with Flexible Regular
-Variation}\label{sec-model}
+\section{A PA Model with Flexible Regular Variation}\label{sec-model}
 
@@
 behaviour below the threshold, while simultaneously guaranteeing regular
-varation of the degrees. In this section, we demonstrate the flexibility
-of the preference function in Equation~\ref{eq-pref}, with regard to the
-tail behaviour of the limiting degree distribution. Using
+variation of the degrees. In this section, we demonstrate the
+flexibility of the preference function in Equation~\ref{eq-pref}, with
+regard to the tail behaviour of the limiting degree distribution. Using
 Equation~\ref{eq-surv-origin}, the limiting survival is
@@
 various combinations of \((\alpha, \beta, \varepsilon)\) and \(k_0=20\)
-of the proposed preferential attachment model.}
+of the proposed PA model.}
 
@@
 offers a natural connection to the discrete version of the generalised
-Pareto distribution (GP), providing a link to a well established
-component of discrete extremes in the literature. Specifically,
-connecting to the Integer GP (IGP) seen in \citet{Rohrbeck_2018} with
-conditional survival: \[
+Pareto (GP) distribution , providing a link to a well established
+component of discrete extremes in the literature. Specifically, it is
+connected to the Integer GP (IGP) distribution seen in
+\citet{Rohrbeck_2018} with conditional survival: \[
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
@@
 \(X|X>u \sim  \mathrm {IGP}(\xi, \sigma, u)\) where \(\xi\) is the shape
-parameter controlling the tail index.
+parameter and reciprocal of the tail index.
 
@@
 
-Therefore, \[{#eq-igp-est}
+Therefore, \begin{equation}\phantomsection\label{eq-igp-est}{
 \bar F(k) 
@@
 \end{cases}
-\] meaning that for \(k\ge k_0\) the limiting degree distribution (for
-large \(k_0^\alpha\)) is approximated by the
-\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).
+}\end{equation} meaning that for \(k\ge k_0\) the limiting degree
+distribution (for large \(k_0^\alpha\)) is approximated by the
+\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\)
+distribution.
 
@@
 Figure~\ref{fig-approx_surv} in colour and their IGP approximations
-\textbf{?@eq-igp-est} are shown in grey. The approximation seems to hold
-up fairly well even for large degrees and more so when \(\alpha\) is
+Equation~\ref{eq-igp-est} are shown in grey. The approximation holds up
+fairly well even for large degrees and more so when \(\alpha\) is
 larger.
@@
 \caption{\label{fig-approx_surv}Theoretical conditional survivals (grey)
-alongside their IGP approximations (coloured).}
+of the proposed model alongside their IGP approximations (coloured).}
 
@@
 In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\) ensures that
-the shape parameter of the IGP is positive and thus the distribution is
-regularly varying. Additionally the value of the shape parameter \(\xi\)
-is shown in Figure~\ref{fig-polyheat} for various parameter choices. The
-darker and lighter regions on the heat maps correspond to a heavier and
-a lighter tail, respectively, the red dashed line shows combinations of
-\(\alpha\) and \(\beta\) that produce a limiting degree distribution
-with the same tail index as the Barabási-Albert model.
-
-Through the connection to the IGP, fitting this model is almost
-equivalent to fitting the IGP to the degrees and estimating its
-parameters. However, instead of only describing the shape of the degree
-distribution, we would also gain estimates for the shape of the
-preference function, thus gaining a direct understanding the mechanisms
-underlying the growth of the network.
+the shape parameter of the IGP distribution is positive and thus the
+distribution is regularly varying. Additionally the shape parameter
+\(\xi\) is shown in Figure~\ref{fig-polyheat} for various parameter
+choices. The darker and lighter regions on the heat maps correspond to a
+heavier and a lighter tail, respectively, and the red dashed line shows
+combinations of \(\alpha\) and \(\beta\) that produce a limiting degree
+distribution with the same tail index as the BA model.
+
+Through the connection implied by Equation~\ref{eq-igp-est}, fitting the
+proposed model is almost equivalent to fitting the IGP distribution to
+the degrees and estimating its parameters. However, instead of only
+describing the shape of the degree distribution, we would also gain
+estimates for the shape of the preference function, thus gaining a
+direct understanding the mechanisms underlying the growth of the
+network.
 
@@
 
-where \(\text{B}(\cdot,\cdot)\) is the the beta function, and \(l\ge0\)
-is a quantity that allows truncating the data such that the minimum
-degree is \(l\). This will allow the model to be fitted whilst ignoring
-the influence of the lower degrees (those less than \(l\)) as the model
-does not capture the behaviour well at the lower degrees since
+where \(\text{B}(\cdot,\cdot)\) is the the beta function,
+\(\pmb \theta = (\alpha, \varepsilon, k_0,\beta)\), and \(l\ge0\) is a
+quantity that allows truncating the data such that the minimum degree is
+\(l\). This will allow the model to be fitted whilst ignoring the
+influence of the lower degrees (those less than \(l\)) as the model does
+not capture the behaviour well at the lower degrees, since
 \citet{rudas07} only provides results for the case of a preferential
@@
 \(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
-counts and using the likelihood from the previous section alongisde the
+counts and using the likelihood from the previous section alongside the
 priors:
@@
 
-where Gamma(\(a\), \(b\)) is the Gamma distribution with shape \(a\) and
+where Gamma(\(a\), \(b\)) is the gamma distribution with shape \(a\) and
 rate \(b\), and U(\(a\), \(b\)) is uniform distribution with lower and
-upper bounds a and b, to obtain a posterior distribution, up to the
-proportionality constant. Posterior samples can then be obtained by an
-adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm.
-For these simulated networks \(l=0\).
+upper bounds \(a\) and \(b\), to obtain a posterior distribution, up to
+the proportionality constant. Posterior samples can then be obtained by
+an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC)
+algorithm. For these simulated networks \(l=0\).
 
@@
 
-\caption{\label{fig-rec1}Posterior estimates of survival function for
-data simulated from the proposed model with various combinations of
-\((\alpha,\beta,\varepsilon)\) and \(k_0=20\).}
+\caption{\label{fig-rec1}Empirical (black dots) and fitted (red line)
+survival functions for data simulated from the proposed model with
+various combinations of \((\alpha,\beta,\epsilon)\) and \(k_0=20\). The
+95\% credible (dashed lines) are included but too narrow to be seen
+clearly.}
 
@@
 
-\caption{\label{fig-rec2}: Posterior estimates of paramters for data
+\caption{\label{fig-rec2}: Posterior densities of parameters for data
 simulated from the proposed model with various combinations of
-(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}
+(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\). True parameter
+values shown with black dots.}
 
@@
 
-In this section, we fit the proposed model to the degree distributions
-of various real networks and learn about the mechanics of their growth.
-While we also compare the fit to that of the mixture distribution by
-\citet{Lee24} we note that the proposed method has the additional
-benefit of learning directly about the growth of a network from the
-inference results. The data consists of 12 networks sourced from
-\href{konect.cc}{KONECT} and the
+In this subsection, we fit the proposed model to the degree
+distributions of various real networks and learn about the mechanics of
+their growth. While we also compare the fit to that of the mixture
+distribution by \citet{Lee24} we note that the proposed method has the
+additional benefit of learning directly about the growth of a network
+from the inference results. The data consists of 12 networks sourced
+from \href{konect.cc}{KONECT} and the
 \href{https://networkrepository.com}{Network Data Repository}\citep{nr}:
@@
 
-\caption{\label{fig-real1}Posterior estimates (solid red) of survival
-for several real data sets and their 95\% credible intervals (dotted
-red).}
+\caption{\label{fig-real1}Empirical (black dots) and posterior medians
+(solid red) of the fitted survival function for several real data sets
+and their 95\% credible intervals (dotted red).}
 
@@
 model gives a similar fit to the Zipf-IGP model but where the GPA model
-fits well we gain additional information about the preference function
+fits well we gain additional information about the preference function,
 assuming that the network evolved according the the GPA scheme.
+
+\begin{figure}[H]
+
+{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.png}
+
+}
+
+\caption{Posterior densities of the parameters of the proposed model
+fitted to real data.}
+
+\end{figure}%
 
@@
 fits the tail better than the Zipf-IGP model does, or because of the
-threshold being estimated as too low forcing almost all of the data to
-be modelled by the linear part of the GPA. This again shows the effects
-that small degrees have on this model.
+threshold being too low, forcing almost all of the data to be modelled
+by the linear part of the GPA. This again shows the effects that small
+degrees have on this model.
 
@@
 
-\caption{\label{fig-shapes}Posterior estimates (solid red) of survival
-for several real data sets and their 95\% credible intervals (dotted
-red).}
+\caption{\label{fig-shapes}Posterior estimates of shape parameter of
+Zipf-IGP distribution (right) and the analagous quantity obtained from
+fitting the proposed model (left).}
 
@@
 function. The first appears mostly flat (similar to uniform attachment)
-for the smallest degrees and then after a threshold preferential
-attachment kicks in, some with this shape are \texttt{pajek-erdos} and
-\texttt{sx-mathoverflow}. The second distinct shape appears to provide
-some clear preferential attachment behaviour that then slows down after
-a certain point, examples of this are seen in the two infrastructure
-networks \texttt{opsahl-openflights} and \texttt{topology}. This slowing
-down could be viewed as a kind of diminishing returns on the degree of a
-vertex i.e.~as a vertex gets larger gaining more connections has less of
-an effect than it did before some threshold \(k_0\).
+for the smallest degrees and then after a threshold PA kicks in, some
+with this shape are \texttt{pajek-erdos} and \texttt{sx-mathoverflow}.
+The second distinct shape appears to provide some clear PA behaviour
+that then slows down after a certain point, and examples of this are
+seen in the two infrastructure networks \texttt{opsahl-openflights} and
+\texttt{topology}. This slowing down could be viewed as a kind of
+diminishing returns on the degree of a vertex i.e.~as a vertex gets
+larger gaining more connections has less of an effect than it did before
+some threshold \(k_0\).
 
@@
 
-\caption{\label{fig-pa}Posterior estimate for preference function
+\caption{\label{fig-pa}Posterior median for the preference function
 (solid) with 95\% credible interval (dashed) on log-log scale.}
@@
 
+This paper contributes to the understanding of the relationship between
+the mechanism underlying a networks growth and the resulting degree
+distribution. As well as, demonstrating that under certain conditions
+information about this mechanism can be garnered from the degrees alone.
+Perhaps in future, something similar could be done by looking at another
+statistic for the network (e.g.~the triangle distribution) and using it
+in combination with the degrees to gain further insights into the growth
+mechanism.
+
 One limitation of this method is that the lowest degrees needed to be
-truncated as they had a very large effect on the fit of the model as a
+truncated as they had a very large effect on the fit of the model, as a
 result of using theory developed for trees and applying it to general
@@
 this, using Remark 2 from \citet{banerjee25} it can be shown that the
-degree distribution is still heavy tailed although not necessarily
-regularly varying.
+survival function of the limiting degree distribution (when using a
+preference function of the form from Proposition~\ref{prp-omega2}) can
+be bounded by two regularly varying functions showing that the degree
+distribution is still heavy tailed, although not necessarily regularly
+varying.
 
@@
 inaccessible and provide a real barrier to using more realistic models
-in a way similar to what we have in this paper.
+in a way similar to what we have in this paper, we leave these open
+problems for future analysis.
 
@@
 \end{theorem}
-
-\section{Additional Plots}\label{additional-plots}
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.png}
-
-}
-
-\caption{: Posterior estimates of parameters for real data.}
-
-\end{figure}%
 

### 2025-09-22 paper.qmd
Fixing figures + Section 2

@@
 title: "Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions"
+crossref:
+  fig-title: '**Fig.**'
 format:
@@
     keep-tex: true
-    equal-margins: false
+    equal-margins: true
     number-sections: true
@@
   - name: Thomas William Boughen
+    orcid: 0009-0005-1979-2479
+    email: t.w.boughen1@newcastle.ac.uk
+    corresponding: true
     affiliations:
@@
 One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law.
-Consequently, the growth of real-life networks that roughly display such power-law behaviour is commonly modelled by preferential attachment.
+Consequently, the growth of real-life networks that display such power-law behaviour is commonly modelled by preferential attachment.
 However, the validity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the right tail of the degree distribution is often lighter than implied by the body, whilst still being heavy.
@@
 
-Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
-
+```{=tex}
+Special cases of this model include the Barab\'asi-Albert (BA) model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
+```
 The survival function of the limiting degree distribution, called the limiting survival hereafter, under condition [-@eq-condb2] can be analytically derived in the case where $m=1$, which is presented below.
@@
 \Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
-$$ Now let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
+$$
+
+Denote the density of the point process associated with the pure birth process corresponding to the growth of an individual node by $\rho(t)$, and its Laplace transform by $\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t$.
+Next, denote the Malthusian parameter of this process by $\lambda^*$, that is $\lambda^*$ satisfies $\hat\rho(\lambda^*) = 1$.
+
+Now, let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
 
@@
 $$
-\lim*{t*\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum{x\in\Upsilon(t)}\varphi(\Upsilon(t)\_{\downarrow x}) = \lambda\^\* \int\_0\infty e{-\lambda\^\* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t
+\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda \int_0^\infty e^{-\lambda t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
 $$ {#eq-survlim}
-
-where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$ and $\hat\rho$ is the Laplace transform of the density of the point process associated with the pure birth process that corresponds to the growth of an individual node, that is $\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t$.
 
@@
 $$
-\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
+\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1},
 $$ which can also be written using @eq-survlim as:
@@
 
-We are interested in how the tail behaviour of the discrete limiting degree distribution is affected by the preference function $b$.
-In order to study this, we use the quantity $\Omega(F, k)$ as stated in @shimura12 and is repeated below:
+We investigate how the tail behaviour of the discrete limiting degree distribution is influenced by the preference function $b(·)$.
+Since the limiting degree distribution is discrete, standard tools for analysing extremes of continuous distributions are not directly applicable.
+A key difficulty arises because many standard discrete distributions, such as the Poisson, Geometric, or Negative Binomial, do not satisfy the conditions required to belong to an MDA.
+@shimura12 showed that for a discrete random variable X to lie in an MDA, its distribution must be long-tailed and introduced the notion of recoverability to the MDA, sometimes referred to as the discrete MDA (see e.g.
+@hitz24).
+In particular, if a distribution is both long-tailed and recoverable, then it can be placed in the appropriate MDA.
 
-For a discrete distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ let
+A central tool in this characterisation is the quantity $\Omega(F,k)$. For a discrete distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ define:
 
@@
 
-The following proposition gives the limiting behaviour of $\Omega(F,k)$ when \$F\$ is a limiting degree distribution resulting from the GPA model with preference function $b(\cdot)$.
+@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Frechet MDA.
+If instead, $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is recoverable to the Gumbel MDA; in this cases we say that the distribution is light-tailed.
+Moreover, if the distribution is also long-tailed, we can conclude that $F$ itself lies in the Gumbel MDA.
+
+This framework offers a natural approach for analysing the tail behaviour of the limiting degree distribution.
+The following proposition gives the limiting behaviour of $\Omega(F,k)$ when $F$ is a limiting degree distribution resulting from the GPA model with preference function $b(\cdot)$.
 
 ::: {#prp-omega}
-If $\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}$ and $b(k) \rightarrow \infty$ as $k\rightarrow \infty$, then $$
+If $\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}$ with $b(k) \rightarrow \infty$ as $k\rightarrow \infty$, then $$
 \lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
 $$
+
+Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process. See Appendix [-@sec-proofs] for the details of the proof.
 :::
 
-See Appendix [-@sec-proofs] for the details of the proof.
-
-@shimura12 states that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$.
-On the other hand, if $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then we will refer to the distribution as light-tailed.
-
-@prp-omega aligns with the result from @krapivsky01 demonstrating that a sub-linear preference function will lead to a light-tailed distribution, as $\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0$ if $b(k)=k^\alpha$ where $\alpha < 1$.
-@prp-omega also aligns with the fact that BA model produces a regularly varying degree distribution, with tail index 2, by considering the preference function $b(k) = k + \varepsilon$, as $\lim_{k\rightarrow\infty}b(k+1)-b(k)=1$, leaving the tail index to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.
-So, in order for the degree distribution to be regularly varying we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive.
-The following proposition determines the class of functions that will result in regular varying degree distributions.
-
-<!-- We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e.
-$\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$.
-First we must consider the following theorem: -->
-
-::: {#prp-omega2}
-The limiting survival of a GPA model is regularly varying if and only if the preference function is *eventually* of the form:
-
-\begin{align*}
-&b(k) = \beta k + r(k)\\
-&\text{where}\\
-&r(k+1)-r(k)\rightarrow 0,
-\end{align*}
-
-and if the preference function is of this form then the tail index is exactly $\lambda^*/\beta$ where $\hat\rho(\lambda\^\*) = 1$.
-
-*Proof*
-
-Letting $b(k) = \beta k  + o(k)$ and using @prp-omega we have that:
-
-$$
-\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \lim_{k\rightarrow\infty} \frac{\beta + r(k+1) - r(k)}{\lambda^*} = \frac{\beta}{\lambda^*},
-$$
-
-and thus the tail index of the limiting survival must be $\lambda^*/\beta$.
-
-\
-If we now assume that the limiting survival is regularly varying with tail index $\delta$, we must have that:
-
-$$
-\lim_{k\rightarrow\infty}\Omega(F,k) = \delta
-$$
-
-and again by @prp-omega
-
-$$
-\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \delta
-$$
-
-Letting $b_k=b(k)$ and $a_k = k$ we can use the Stolz-Cesàro theorem (@thm-stolz) to obtain:
-
-$$
-\lim_{k\rightarrow\infty}\frac{b_{k+1}-b_k}{a_{k+1}-a_k} = \delta \implies \lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \delta
-$$
-
-and so
-
-$$
-\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \delta
-$$
-
-which $b(k) = \beta k + r(k)$ fulfills.$\square$
+::: {#rem-omega}
+If $b(k)\rightarrow <\infty$, then $\Omega(F, k) = 0$.
 :::
 
-Using this result we can understand how the preference function is directly connected with the tail behaviour of the degree distribution.
-Specifically, regular variation is achieved if and only if $b(k)$ is asymptotically linear with $k$.
-We use this result to create a preference function that guarantees regular variation in the tail of the degree distribution, aligning with analysis of real networks, whilst allowing for the tail to deviate from the shape implied by the body.
-This gives the model the capability to produce realistic behaviour in the degrees like what was in @Lee24 by using a piecewise function inspired by the observed deviation from the power law after a certain threshold:
+::: {#cor-omega2}
+Let $\bar F (k)$ bet he limiting survival function of degree in a GPA network with preference function $b$ such that $b(k)\rightarrow\infty$. Then:
+
+-   $\bar F(k)$ is regularly varying if and only if $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=c>0$, in which case the tail index is $\lambda^*/c$, where $\lambda^*$ is the Malthusian parameter of the corresponding branching process.
+
+-   $\bar F(k)$ is light-tailed if and only if $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$.
+:::
+
+@cor-omega2 is a direct consequence of @prp-omega and the results of @shimura12, and it aligns with previous findings in the literature.
+In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow}[b(k+1)-b(k)]=0$.
+Conversely, it also agrees with the limiting behavior of BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+\varepsilon$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $1/\lambda^*$, in which cases the Malthusian parameter $\lambda^*$ is 2.
+
+Using @cor-omega2 we can directly connect the preference function to the tail behaviour of the degree distribution.
+By choosing an appropriate preference function, we can ensure that the tail of the degree distribution is regularly varying while allowing the body to deviate from the power-law shape, as observed in real networks.
+Inspired by @Lee24, we consider a piecewise function:
 
@@
 \end{cases}
-$$ {#eq-pref} for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.
+$$ {#eq-pref} for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$. By @cor-omega2, the resulting degree distribution is regularly varying with tail index $\lambda^*/\beta$, since
 
-Using @prp-omega, we can show that $\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta$, meaning the degree distribution obtained using this preference function is regularly varying with tail index $\lambda^*/\beta$.
+$$
+\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda}.
+$$
+
+```{r, echo=FALSE, out.width="60%", warning=FALSE, message=FALSE, cache=TRUE}
+#| fig-width: 6
+#| fig-height: 6
+#| fig-cap: Example construction of the preference function in @eq-pref
+#| label: fig-ex
+x = seq(0,10,l=100)
+k0 = 5
+a = 0.5
+eps = 0.5
+y2 = ifelse(x<=k0, (eps + x^1) , (x-k0) + k0^1 + eps) +  (k0^a - k0)
+y3 = ifelse(x<=k0, (eps + x^a), (x-k0) + k0^a + eps)
+y4 = eps + x^a
+
+textsize = 5
+lwd = 1
+ggplot() +
+  geom_line(aes(x=x,y= y3), lwd=lwd) + 
+  geom_segment(aes(x=k0, y=0.3,yend=eps + k0^a), lty=2, lwd=lwd)+
+  geom_text(aes(x=k0, y=0, label='k[0]'), parse = TRUE, size=textsize)+
+  geom_text(aes(x=0.25, y=eps, label='epsilon'), parse=TRUE, size=textsize)+
+  geom_text(aes(x=6.3, y=4.3, label='beta'), parse=TRUE, size=textsize)+
+  theme_bw() + xlab('Degree')+ylab('Preference') + ylim(0,6)+
+  theme(axis.ticks = element_blank(),
+        axis.text = element_blank(), legend.position='bottom')
+
+```
 
@@
 To assess how close of an approximation this is, the theoretical conditional survival [@eq-polysurv] are shown in @fig-approx_surv in colour and their IGP approximations [@eq-igp-est] are shown in grey.
-The approximation holds up fairly well even for large degrees and more so when $\alpha$ is larger.
+The approximation holds up well even for large degrees and more so when $\alpha$ is larger.
 
@@
 
-In agreement with @prp-omega2, $\beta>0$ ensures that the shape parameter of the IGP distribution is positive and thus the distribution is regularly varying.
+In agreement with @cor-omega2, $\beta>0$ ensures that the shape parameter of the IGP distribution is positive and thus the distribution is regularly varying.
 Additionally the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices.
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
-#| fig-width: 12
-#| fig-height: 6
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+#| fig-width: 8
+#| fig-height: 4
 #| fig-cap: Heat maps of $\xi$ for various combinations of the parameters of the proposed model.
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="90%"}
 #| fig-width: 8
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="100%"}
-#| fig-width: 9
-#| fig-height: 9
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+#| fig-width: 10
+#| fig-height: 10
 #| fig-cap: Empirical (black dots) and posterior medians (solid red) of the fitted survival function for several real data sets and their 95% credible intervals (dotted red).
@@
 
-![ Posterior densities of the parameters of the proposed model fitted to real data.](images/pars_plot.png){width="80%"}
-
 @fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model.
@@
 
+![Posterior densities of the parameters of the proposed model fitted to real data.](images/pars_plot.jpg){#fig-parplot width="80%"}
+
 ```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="100%"}
-#| fig-width: 9
-#| fig-height: 9
+#| fig-width: 10
+#| fig-height: 10
 #| fig-cap: Posterior estimates of shape parameter of Zipf-IGP distribution (right) and the analagous quantity obtained from fitting the proposed model (left).
@@
 
-In this paper we introduced a class of preference functions that, under the GPA scheme, generate a network with a flexible yet regularly varying degree distribution.
+We introduced a class of preference functions that, under the GPA scheme, generate a network with a flexible yet regularly varying degree distribution.
 From the simulation study we showed that the parameters can be recovered from fitting the model to the degrees alone.
 We also applied this method to the degree distributions of real networks, estimating their model parameters assuming they evolved in the same way.
-Not only did this yield fairly good fits for the degree distribution, similar to that of the Zipf-IGP, it also came with the added benefit of giving a posterior estimate for a preference function.
+Not only did this yield good fits for the degree distribution, similar to that of the Zipf-IGP, it also came with the added benefit of giving a posterior estimate for a preference function.
 
 This paper contributes to the understanding of the relationship between the mechanism underlying a networks growth and the resulting degree distribution.
-As well as, demonstrating that under certain conditions information about this mechanism can be garnered from the degrees alone.
-Perhaps in future, something similar could be done by looking at another statistic for the network (e.g.
-the triangle distribution) and using it in combination with the degrees to gain  further insights into the growth mechanism.
+As well as demonstrating that under certain conditions information about this mechanism can be garnered from the degrees alone.
+In future, something similar could be done by looking at another statistic for the network (e.g.
+the triangle distribution) and using it in combination with the degrees to gain further insights into the growth mechanism when the networks full evolution is not available.
+The triangle distribution would be a good avenue for future work within the realm of extremes as it has also been shown by @kang11 that many real networks seem to have power law triangle distributions.
 
@@
 However, there is no expression for the probability mass function of the degrees and therefore no likelihood that can be used for modeling in the same way that we have.
-In spite of this, using Remark 2 from @banerjee25 it can be shown that the survival function of the limiting degree distribution (when using a preference function of the form from @prp-omega2) can be bounded by two regularly varying functions showing that the degree distribution is still heavy tailed, although not necessarily regularly varying.
+In spite of this, using Remark 2 from @banerjee25 it can be shown that the survival function of the limiting degree distribution (when using a preference function of the form from @cor-omega2) can be bounded by two regularly varying functions showing that the degree distribution is still heavy tailed, although not necessarily regularly varying.
 
@@
 \newpage
-
-# Supplementary Results {#sec-sup .appendix}
-
-::: {#thm-stolz}
-## Stolz-Cesàro Theorem [@cesaro]
-
-Let $(a_k)_{k\ge1}$ and $(b_k)_{k\ge1}$ be two sequence of real numbers. Assume $(a_k)_{k\ge1}$ is a strictly monotone and divergent sequence and the following limit exists:
-
-$$
-\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
-$$ Then,
-
-$$
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
-$$
-:::
 

### 2025-09-22 paper.tex
Fixing figures + Section 2

@@
 ]{sn-jnl}
+
+\setlength\evensidemargin{\oddsidemargin}
 
@@
 \ifdefined\figurename
-  \renewcommand*\figurename{Figure}
+  \renewcommand*\figurename{\textbf{Fig.}}
 \else
-  \newcommand\figurename{Figure}
+  \newcommand\figurename{\textbf{Fig.}}
 \fi
@@
 \theoremstyle{plain}
-\newtheorem{theorem}{Theorem}[section]
+\newtheorem{corollary}{Corollary}[section]
 \theoremstyle{remark}
@@
 % author setup
-\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
+\author*[1]{\fnm{Thomas William} \sur{Boughen}}\email{t.w.boughen1@newcastle.ac.uk}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
 % affil setup
@@
 results in a degree distribution that follows the power law.
-Consequently, the growth of real-life networks that roughly display such
+Consequently, the growth of real-life networks that display such
 power-law behaviour is commonly modelled by preferential attachment.
@@
 
-Special cases of this model include the Barabási-Albert (BA) model when
-\(b(k) = k+1\), which in the limit of \(t\rightarrow \infty\) leads to a
-power-law degree distribution with tail index 2, and the Uniform
-Attachment (UA) model where \(b(k)=c\) leading to a degree distribution
-that is not regularly varying.
+Special cases of this model include the Barab\'asi-Albert (BA) model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
 
@@
 \Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
-\] Now let \(\Upsilon(t)\) be the tree determined by \(\zeta(t)\) as
+\]
+
+Denote the density of the point process associated with the pure birth
+process corresponding to the growth of an individual node by
+\(\rho(t)\), and its Laplace transform by
+\(\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t\).
+Next, denote the Malthusian parameter of this process by \(\lambda^*\),
+that is \(\lambda^*\) satisfies \(\hat\rho(\lambda^*) = 1\).
+
+Now, let \(\Upsilon(t)\) be the tree determined by \(\zeta(t)\) as
 follows: \(\Upsilon(t)=\{\emptyset\}\) and \(\Upsilon(t)=G\) where each
@@
 \begin{equation}\phantomsection\label{eq-survlim}{
-\lim*{t*\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum{x\in\Upsilon(t)}\varphi(\Upsilon(t)\_{\downarrow x}) = \lambda\^\* \int\_0\infty e{-\lambda\^\* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t
+\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda \int_0^\infty e^{-\lambda t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
 }\end{equation}
-
-where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\) and \(\hat\rho\)
-is the Laplace transform of the density of the point process associated
-with the pure birth process that corresponds to the growth of an
-individual node, that is
-\(\hat\rho(\lambda) \coloneq \int_0^\infty e^{-\lambda t}\rho(t)\mathrm{d}t\).
 
@@
 \[
-\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
+\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1},
 \] which can also be written using Equation~\ref{eq-survlim} as:
@@
 
-We are interested in how the tail behaviour of the discrete limiting
-degree distribution is affected by the preference function \(b\). In
-order to study this, we use the quantity \(\Omega(F, k)\) as stated in
-\citet{shimura12} and is repeated below:
-
+We investigate how the tail behaviour of the discrete limiting degree
+distribution is influenced by the preference function \(b(·)\). Since
+the limiting degree distribution is discrete, standard tools for
+analysing extremes of continuous distributions are not directly
+applicable. A key difficulty arises because many standard discrete
+distributions, such as the Poisson, Geometric, or Negative Binomial, do
+not satisfy the conditions required to belong to an MDA.
+\citet{shimura12} showed that for a discrete random variable X to lie in
+an MDA, its distribution must be long-tailed and introduced the notion
+of recoverability to the MDA, sometimes referred to as the discrete MDA
+(see e.g. \citet{hitz24}). In particular, if a distribution is both
+long-tailed and recoverable, then it can be placed in the appropriate
+MDA.
+
+A central tool in this characterisation is the quantity \(\Omega(F,k)\).
 For a discrete distribution \(F\) with survival function \(\bar F\) and
-\(k\in\mathbb Z^+\) let
+\(k\in\mathbb Z^+\) define:
 
@@
 
-The following proposition gives the limiting behaviour of
-\(\Omega(F,k)\) when \$F\$ is a limiting degree distribution resulting
-from the GPA model with preference function \(b(\cdot)\).
+\citet{shimura12} established that if
+\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
+then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\), and
+hence belongs to the Frechet MDA. If instead,
+\(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then the distribution is
+recoverable to the Gumbel MDA; in this cases we say that the
+distribution is light-tailed. Moreover, if the distribution is also
+long-tailed, we can conclude that \(F\) itself lies in the Gumbel MDA.
+
+This framework offers a natural approach for analysing the tail
+behaviour of the limiting degree distribution. The following proposition
+gives the limiting behaviour of \(\Omega(F,k)\) when \(F\) is a limiting
+degree distribution resulting from the GPA model with preference
+function \(b(\cdot)\).
 
@@
 
-If \(\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}\) and
+If \(\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}\) with
 \(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\), then \[
@@
 
+Here, \(\lambda^*\) is the Malthusian parameter of the corresponding
+branching process. See Appendix \ref{sec-proofs} for the details of the
+proof.
+
 \end{proposition}
 
-See Appendix \ref{sec-proofs} for the details of the proof.
-
-\citet{shimura12} states that if
-\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\). On
-the other hand, if \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then we
-will refer to the distribution as light-tailed.
-
-Proposition~\ref{prp-omega} aligns with the result from
-\citet{krapivsky01} demonstrating that a sub-linear preference function
-will lead to a light-tailed distribution, as
-\(\lim_{k\rightarrow\infty} b(k+1)-b(k) = 0\) if \(b(k)=k^\alpha\) where
-\(\alpha < 1\). Proposition~\ref{prp-omega} also aligns with the fact
-that BA model produces a regularly varying degree distribution, with
-tail index 2, by considering the preference function
-\(b(k) = k + \varepsilon\), as
-\(\lim_{k\rightarrow\infty}b(k+1)-b(k)=1\), leaving the tail index to be
-\(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\). So,
-in order for the degree distribution to be regularly varying we need
-that the limit \(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is
-positive. The following proposition determines the class of functions
-that will result in regular varying degree distributions.
-
-\begin{proposition}[]\protect\hypertarget{prp-omega2}{}\label{prp-omega2}
-
-The limiting survival of a GPA model is regularly varying if and only if
-the preference function is \emph{eventually} of the form:
-
-\begin{align*}
-b(k) &= \beta k + r(k),\\
-&\text{and}\\
-r(k+1)-r(k)&\rightarrow 0,
-\end{align*}
-
-and if the preference function is of this form then the tail index is
-exactly \(\lambda^*/\beta\) where \(\hat\rho(\lambda\^\*) = 1\).
-
-\emph{Proof}
-
-Letting \(b(k) = \beta k  + o(k)\) and using Proposition~\ref{prp-omega}
-we have that:
-
-\[
-\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \lim_{k\rightarrow\infty} \frac{\beta + o(k+1) - o(k)}{\lambda^*} = \frac{\beta}{\lambda^*},
-\]
-
-and thus the tail index of the limiting survival must be
-\(\lambda^*/\beta\).
-
-\hfill\break
-If we now assume that the limiting survival is regularly varying with
-tail index \(\delta\), we must have that:
-
-\[
-\lim_{k\rightarrow\infty}\Omega(F,k) = \delta
-\]
-
-and again by Proposition~\ref{prp-omega}
-
-\[
-\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} = \delta
-\]
-
-Letting \(b_k=b(k)\) and \(a_k = k\) we can use the Stolz-Cesàro theorem
-(Theorem~\ref{thm-stolz}) to obtain:
-
-\[
-\lim_{k\rightarrow\infty}\frac{b_{k+1}-b_k}{a_{k+1}-a_k} = \delta \implies \lim_{k\rightarrow\infty}\frac{b_k}{a_k} = \delta
-\]
-
-and so
-
-\[
-\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \delta
-\]
-
-which \(b(k) = \beta k + o(k)\) fulfills.\(\square\)
-
-\end{proposition}
-
-Using this result we can understand how the preference function is
-directly connected with the tail behaviour of the degree distribution.
-Specifically, regular variation is achieved if and only if \(b(k)\) is
-asymptotically linear with \(k\). We use this result to create a
-preference function that guarantees regular variation in the tail of the
-degree distribution, aligning with analysis of real networks, whilst
-allowing for the tail to deviate from the shape implied by the body.
-This gives the model the capability to produce realistic behaviour in
-the degrees like what was in \citet{Lee24} by using a piecewise function
-inspired by the observed deviation from the power law after a certain
-threshold:
+\begin{refremark}
+If \(b(k)\rightarrow <\infty\), then \(\Omega(F, k) = 0\).
+
+\label{rem-omega}
+
+\end{refremark}
+
+\begin{corollary}[]\protect\hypertarget{cor-omega2}{}\label{cor-omega2}
+
+Let \(\bar F (k)\) bet he limiting survival function of degree in a GPA
+network with preference function \(b\) such that
+\(b(k)\rightarrow\infty\). Then:
+
+\begin{itemize}
+\item
+  \(\bar F(k)\) is regularly varying if and only if
+  \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=c>0\), in which case the tail
+  index is \(\lambda^*/c\), where \(\lambda^*\) is the Malthusian
+  parameter of the corresponding branching process.
+\item
+  \(\bar F(k)\) is light-tailed if and only if
+  \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\).
+\end{itemize}
+
+\end{corollary}
+
+Corollary~\ref{cor-omega2} is a direct consequence of
+Proposition~\ref{prp-omega} and the results of \citet{shimura12}, and it
+aligns with previous findings in the literature. In particular it is
+consistent with \citet{krapivsky01}, who showed that a sub-linear
+preference function leads to a light-tailed distributions, since for
+\(b(k)=k^\alpha\) with \(\alpha<1\), we have
+\(\lim_{k\rightarrow}[b(k+1)-b(k)]=0\). Conversely, it also agrees with
+the limiting behavior of BA model, which produces a regularly degree
+distribution with tail index 2; for the preference function
+\(b(k)=k+\varepsilon\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\),
+giving a tail index of \(1/\lambda^*\), in which cases the Malthusian
+parameter \(\lambda^*\) is 2.
+
+Using Corollary~\ref{cor-omega2} we can directly connect the preference
+function to the tail behaviour of the degree distribution. By choosing
+an appropriate preference function, we can ensure that the tail of the
+degree distribution is regularly varying while allowing the body to
+deviate from the power-law shape, as observed in real networks. Inspired
+by \citet{Lee24}, we consider a piecewise function:
 
@@
 }\end{equation} for \(\alpha,\beta, \varepsilon>0\) and
-\(k_0\in\mathbb N\).
-
-Using Proposition~\ref{prp-omega}, we can show that
-\(\lim_{n\rightarrow\infty}\Omega(F,k)=\lambda^{*}/\beta\), meaning the
-degree distribution obtained using this preference function is regularly
-varying with tail index \(\lambda^*/\beta\).
+\(k_0\in\mathbb N\). By Corollary~\ref{cor-omega2}, the resulting degree
+distribution is regularly varying with tail index \(\lambda^*/\beta\),
+since
+
+\[
+\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda}.
+\]
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-ex-1.pdf}
+
+}
+
+\caption{\label{fig-ex}Example construction of the preference function
+in Equation~\ref{eq-pref}}
+
+\end{figure}%
 
@@
 Equation~\ref{eq-igp-est} are shown in grey. The approximation holds up
-fairly well even for large degrees and more so when \(\alpha\) is
-larger.
+well even for large degrees and more so when \(\alpha\) is larger.
 
@@
 
-In agreement with Proposition~\ref{prp-omega2}, \(\beta>0\) ensures that
+In agreement with Corollary~\ref{cor-omega2}, \(\beta>0\) ensures that
 the shape parameter of the IGP distribution is positive and thus the
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
+\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
 
@@
 
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
 
@@
 assuming that the network evolved according the the GPA scheme.
-
-\begin{figure}[H]
-
-{\centering \includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.png}
-
-}
-
-\caption{Posterior densities of the parameters of the proposed model
-fitted to real data.}
-
-\end{figure}%
 
@@
 degrees have on this model.
+
+\begin{figure}
+
+\centering{
+
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.jpg}
+
+}
+
+\caption{\label{fig-parplot}Posterior densities of the parameters of the
+proposed model fitted to real data.}
+
+\end{figure}%
 
@@
 
-In this paper we introduced a class of preference functions that, under
-the GPA scheme, generate a network with a flexible yet regularly varying
-degree distribution. From the simulation study we showed that the
-parameters can be recovered from fitting the model to the degrees alone.
-We also applied this method to the degree distributions of real
-networks, estimating their model parameters assuming they evolved in the
-same way. Not only did this yield fairly good fits for the degree
-distribution, similar to that of the Zipf-IGP, it also came with the
-added benefit of giving a posterior estimate for a preference function.
+We introduced a class of preference functions that, under the GPA
+scheme, generate a network with a flexible yet regularly varying degree
+distribution. From the simulation study we showed that the parameters
+can be recovered from fitting the model to the degrees alone. We also
+applied this method to the degree distributions of real networks,
+estimating their model parameters assuming they evolved in the same way.
+Not only did this yield good fits for the degree distribution, similar
+to that of the Zipf-IGP, it also came with the added benefit of giving a
+posterior estimate for a preference function.
 
@@
 the mechanism underlying a networks growth and the resulting degree
-distribution. As well as, demonstrating that under certain conditions
+distribution. As well as demonstrating that under certain conditions
 information about this mechanism can be garnered from the degrees alone.
-Perhaps in future, something similar could be done by looking at another
+In future, something similar could be done by looking at another
 statistic for the network (e.g.~the triangle distribution) and using it
 in combination with the degrees to gain further insights into the growth
-mechanism.
+mechanism when the networks full evolution is not available. The
+triangle distribution would be a good avenue for future work within the
+realm of extremes as it has also been shown by \citet{kang11} that many
+real networks seem to have power law triangle distributions.
 
@@
 survival function of the limiting degree distribution (when using a
-preference function of the form from Proposition~\ref{prp-omega2}) can
-be bounded by two regularly varying functions showing that the degree
+preference function of the form from Corollary~\ref{cor-omega2}) can be
+bounded by two regularly varying functions showing that the degree
 distribution is still heavy tailed, although not necessarily regularly
@@
 \newpage
-
-\section{Supplementary Results}\label{sec-sup}
-
-\begin{theorem}[Stolz-Cesàro Theorem
-\citep{cesaro}]\protect\hypertarget{thm-stolz}{}\label{thm-stolz}
-
-Let \((a_k)_{k\ge1}\) and \((b_k)_{k\ge1}\) be two sequence of real
-numbers. Assume \((a_k)_{k\ge1}\) is a strictly monotone and divergent
-sequence and the following limit exists:
-
-\[
-\lim_{k\rightarrow\infty}\frac{b_{k+1} - b_k}{a_{k+1} - a_k} = l.
-\] Then,
-
-\[
-\lim_{k\rightarrow\infty}\frac{b_k}{a_k} = l.
-\]
-
-\end{theorem}
 

### 2025-09-22 scripts/data_fit_background.R

@@ -13,9 +13,9 @@ trunc.at = 3
 
 
 
-data.files = list.files('./data') 
+data.files = list.files('../data') 
 nms = tstrsplit(data.files,'out.',keep=2)[[1]]
-
+nms=nms[!is.na(nms)]
 saveRDS(nms, 'nms.rds')
 
 
@@ -62,7 +62,7 @@ for(i in 1:length(fits)){
 tmp = fits[[4]]
 
 
-fits = readRDS('modelfits.rds')
+fits = readRDS('../results/modelfits.rds')
 
 fits[[4]] = tmp
 
@@ -160,14 +160,14 @@ comp_plot
 
 # -------------------------------------------------------------------------
 
-mixfits = readRDS('mixfits.rds')
+mixfits = readRDS('../results/mixfits.rds')
 plots = list()
 for(i in 1:length(fits)){
   plots[[i]] = ggplot() + geom_boxplot(aes(y=!!fits[[i]]$mcmc$pars[-(1:2e4),4]/!!fits[[i]]$mcmc$lambdas[-(1:2e4)], x='GPA', fill='GPA')) +
     geom_boxplot(aes(y = !!mixfits[[i]]$pars$shape, x='Zipf-IGP', fill = 'Zipf-IGP')) + ylim(-1,1) + ggtitle(nms[i]) + xlab('')+ylab('')
 }
 
-shape_plot = ggpubr::ggarrange(plotlist = plots, common.legend = T,label.x='', label.y='\\xi', legend='right')
+shape_plot = ggpubr::ggarrange(plotlist = plots, common.legend = T,label.x='', label.y='\\xi', legend='right', nrow=4, ncol=3)
 shape_plot
 saveRDS(shape_plot, 'shape_plot.rds')
 
@@ -206,6 +206,9 @@ saveRDS(PA_overlay, 'PA_overlay.rds')
 
 
 # -------------------------------------------------------------------------
+
+
+
 library(latex2exp)
 library(ggridges)
 nms = readRDS('../results/nms.rds')

### 2025-09-22 Files/archive/paper.qmd
Restructuring

---
title: Tail Flexibility in the Degrees of Preferential
 Attachment Networks
format:
  nature-pdf:
    # journal.cite-style is included in the tex file but ignored by pandoc if 
    # cite-method is not `natbib`.
    journal:
      cite-style: sn-basic
    # `citeproc` is the pandoc default. Set `cite-method: natbib` if required
    # to use the bst styles from the upstream template.
    cite-method: natbib
    keep-tex: true
    equal-margins: false
    number-sections: true
author:
  - name: Thomas William Boughen
    affiliations:
      - name: Newcastle University
        id: 1
        department: School of Mathematics, Statistics and Physics
  - name: Clement Lee
    affiliations:
      - ref: 1
  - name: Vianey Palacios Ramirez
    affiliations:
      - ref: 1
abstract: |
  Devising the underlying generating mechanism of a real-life network is difficult as, more often than not, only its snapshots are available, but not its full evolution. One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law. Consequently, the growth of real-life networks that roughly display such power-law behaviour is commonly modelled by preferential attachment. However, the validity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the right tail of the degree distribution is often lighter than implied by the body, whilst still being heavy. In this paper, we study a modified version of the model with a flexible preference function that allows super/sub-linear behaviour whilst also guaranteeing that the limiting degree distribution has a heavy tail. We relate the distributions tail heaviness directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
keywords: [networks, extremes]
bibliography: refs.bib

---
```{r,echo=FALSE, warning=FALSE, message=FALSE}
library(ggplot2)
library(paletteer)
library(gridExtra)
library(latex2exp)
library(ggpubr)
library(Rcpp)
library(RcppArmadillo)
```

# Introduction {#sec-intro}

Networks appear in many fields such as sociology, politics, epidemiology, and economics. Statistical methods have been and continue to be used to study networks that appear in these fields and beyond, from the use of stochastic block models to detect communities within the French political blogosphere [@Latouche11], to using Exponential Random Graph Models (ERGMs) to analyse the global trade network [@Setayesh22], and the use of mechanistic models in neuroscience to investigate the patterns of connections in neural systems [@Betzel17]. One such mechanistic model is the general preferential attachment (PA) model where as a network grows and new vertices join they connect to those already present in the network with weights proportional to some preference function $b$ of their degree. The well-known Barabási-Albert (BA) model is a special case of this when $b\sim k$ [BA99].

Mechanistic models, like the PA model, are in general only particularly useful when simulating networks and not fitting to the real data as usually only snapshots of real networks are available and not the entire evolution. As such, one of the most informative statistics when regarding a network is the degree distribution. When looking at the degree distributions of real networks it is attractive to many to use the power law to model the degree distribution, as the BA model generates degrees that follow a power law, and therefore claim that since the degree distribution is roughly power law then the underlying growth mechanism may be preferential attachment such as the BA model.  However, it has been shown that a general preference function will not always lead to a power law degree distribution [@krapivsky01] and additionally it has been debated for decades whether most real data actually follows a power law [@Voitalov_2019] or not[@Broido_2019].

Throughout this debate the use of methods from extremes has been absent until relatively recently with the papers such as [@Voitalov_2019] and [Lee]. Forgoing the use of extreme value methods essentially downplays how deviations from the power law in the largest degrees may affect the network given that they are often much more influential than the vertices with smaller degrees.

Noting this, many have now started to use these extreme value methods when analysing networks. [@wan19] showed that extreme value estimation methods prove more robust than existing parametric approaches especially when the data are corrupted or the model has been misspecified, demonstrating that extreme value methods can be a valuable tool when studying networks. [@wang2022random] investigates a PA model with a affine preference function (similar to the BA model) with the added behaviour of heterogeneous reciprocal edges, they show that the distribution of in-degrees and out-degrees produced by such a model is multivariate regularly varying and they also have hidden regular variation. [Lee] using methods from discrete extremes modelled the right tail of network degrees with a  discretised variation of the generalised Pareto distribution (GPD) usually dubbed the Integer GPD (IGPD) where $X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)$ is defined by the survival:

$$
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
$$
for $v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R$.

The study of degree distributions, whether using extremes or not, usually does not reveal any information about the preference function even if PA was the underlying generating mechanism of the network. Here we address this gap in the literature, more specifically investigating if given the degree distribution of a network assumed to come from the PA model can we directly infer the model parameters? 
We build upon the foundation provided by [@rudas07] deriving the limiting degree distribution in terms of the preference function and suggest a class of preference functions that result in realistic tails that are heavy but are lighter than implied by the power law.

## Structure of this paper

@sec-gpa looks at the theoretical limiting behaviour of a GPA model when $m=1$ according to the results from [@rudas07], going on to introduce a flexible class of preference functions that can guarantee a heavy tailed degree distribution while still being flexible in the body. @sec-rec shows that the preference function parameters can be fairly well recovered from the degree distributions of networks simulated from the GPA model using various functions of the class introduced in @sec-gpa. Having shown that for simulated data the model parameters are recoverable from the degree distribution alone, @sec-real builds upon this and attempts to fit the model to degree distributions of real networks; modelling not only the degree distribution itself but also the estimates for the parameters of the preference function assuming the networks evolved according to the GPA scheme. Ultimately, @sec-conc discusses the main results of the paper, pitfalls of this method and future work.

# Preferential attachment model {#sec-gpa}

The network generative model that we will be focussing on in this paper is dubbed General Preferential Attachment in [@rudas07] and is defined as follows:

Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges (with the new vertex as the source); the target for each of these edges are selected from the vertices already in the network with weights proportional to some function $b$ of their degree where the preference function $b$ is chosen such that:

$$
b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
$$ {#eq-condb1}

$$
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
$$ {#eq-condb2}

Special cases of this model include the Barabási-Albert (BA) model when $b(k) = k+\varepsilon$, which leads to a power-law degree distribution with index 2 and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution in the Gumbel maximum domain of attraction.

Given conditions [@eq-condb1] and [@eq-condb2] an expression for the survival of the limiting degree distribution can be found in the case that $m=1$; obtained by considering a branching process that is equivalent to the growth of the network, as in [@rudas07]. Theorem 1 from [@rudas07] states that for the tree $\Upsilon(t)$ at time $t$:

$$
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
$$
where $\lambda^*$ satisfies $\hat\rho(\lambda^*)=1$.

The limiting survival can be viewed as the limit of the empirical proportion of vertices with degree over a threshold $k$, that is:

$$
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
$$
which by the previously stated theorem can also be written as:

$$
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$
Additionally, using the fact that $f(k) = \bar F(k-1) - \bar F(k)$ the probability mass function (p.m.f.) can be shown to be:

$$
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
$$
We are specifically interested in how the tail heaviness ($\xi$ in the IGPD) is affected by the preference function $b$, [@shimura12] introduces a quantity that will help in determining what domain of attraction this discrete distribution belongs to, that is:

For a distribution $F$ with survival function $\bar F$ and some $k\in\mathbb Z^+$ let:

$$
\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
$$

[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(k) \sim k^{-\alpha}$. Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is light tailed. This allows us to show the following:

:::{#prp-omega}
If $b(k) \rightarrow \infty$ as $k\rightarrow \infty$ then
$$
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
$$
and

$$
\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
$$
:::
See [appendix] for the details of the proof.

@prp-omega aligns with the result from [@krapivsky01] demonstrating that a sub-linear preference function will lead to a light tailed distribution. So, in order for the degree distribution to be heavy tailed we need that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ exists and is positive. We can in fact show that BA model produces a heavy tailed degree distribution with index $\xi=0.5$ by considering the preference function $b(k) = k + \varepsilon$,  noting that the limit $\lim_{k\rightarrow\infty} b(k+1)-b(k)$ is simply $\lim_{k\rightarrow\infty} 1 = 1$ leaving the tail heaviness to be $1/\lambda^*$ which using $\hat\rho$ can be found to be $1/2$.

We can show that in order for the degree distribution to be heavy tailed, the preference function must be asymptotically linear i.e. $\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0$.

Consider the definition of heavy tails:

\begin{align*}
\lim_{k\rightarrow\infty}\Omega(F,n) &= c>0\\
\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} &= c\\
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] &= \lambda^*c\\
\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{(k+1)-(k)} &= \lambda^*c\\
\end{align*}

by Stolz-Cesaro Theorem:

$$
\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \lambda^*c > 0.
$$

Since $b$ must be asymptotically linear in order to have positive tail heaviness we propose the following preference function:

$$
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.

This preference function, as per @prp-omega, will produce a degree distribution with tail heaviness $\beta/\lambda^*$ guaranteeing a heavy tail. We study this preference function further in the next section.

# Preferential Attachment with flexible heavy tail {#sec-model}

Using a preference function with guaranteed linear behaviour in the limit, allows for the inclusion of sub/super linear behaviour without losing the heavy tails or ending up with a degenerate degree distribution.

The limiting degree distribution resulting from using a preference function of this form can be found to have survival:

$$
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
\end{cases}
$${#eq-polysurv}

with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where:

$$
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
$$

which must be solved numerically for most parameter choices. Also, note that $\lambda>\beta$.

Some examples of what the degree distribution looks like are shown below in @fig-polylinsurv:

```{r, echo=FALSE}
source('scripts/funcs.R')
```

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical survival distributions of the limiting degree distributions, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed preferential attachment model.
#| label: fig-polylinsurv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = 0:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
}
pars = cbind(pars,surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

ggplot(data = pars) + geom_line(aes(x=(x+1),y=surv, linetype=as.character(b),colour =as.character(b))) +
  scale_x_log10() + scale_y_log10(limits=c(1e-5,1))+theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) + 
  facet_grid(eps~a,labeller = labeller, scales='free')
```

The survival (@eq-polysurv) can be connected to the IGPD mentioned in [@sec-intro] by using Stirling's approximation of the gamma function to obtain:

$$
\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
$$
and utilising the expression from @eq-polysurv:

\begin{align*}
\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
\end{align*}

$$
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
$$ meaning that for $k\ge k_0$ the limiting degree distribution is similar to $\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$.

To assess how close of an approximation this is the theoretical conditional survivals are shown in @fig-approx_surv in colour and their IGPD approximations are shown in grey. The approximation seems to hold up fairly well even for large degrees.

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 8
#| fig-height: 4
#| fig-cap: Theoretical conditional survivals (grey) alongside their IGPD approximations (coloured).
#| label: fig-approx_surv

as = c(0.5,1,1.5)
bs = c(0.1,0.5,1,1.5)
eps = c(.01,.1,.5,1)
k0s = 20
x = k0s:200
pars = expand.grid(as,bs,eps,k0s,x)
names(pars) = c('a','b','eps','k0','x')
lambdas = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
}
pars = cbind(pars,lambdas)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda')
surv =numeric(nrow(pars))
gp_surv =numeric(nrow(pars))
for(i in 1:nrow(pars)){
  surv[i] = S(pars$x[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])/
    S(pars$k0[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
  gp_surv[i] = mev::pgp(pars$x[i],pars$k0[i]-1,((pars$k0[i]+1)^pars$a[i] +pars$eps[i]-1 )/pars$lambda[i],pars$b[i]/pars$lambda[i],lower.tail = F)
}
pars = cbind(pars,surv,gp_surv)
names(pars) = c('a', 'b', 'eps', 'k0','x','lambda','surv','gp_surv')

library(latex2exp)
labeller = label_bquote(rows = `epsilon`==.(eps),cols = `alpha`==.(a))

library(ggh4x)

l1 = scale_y_log10(limits=c(1e-1,1))
l2 = scale_y_log10(limits=c(1e-2,1))
l3 = scale_y_log10(limits=c(1e-3,1))
l4 = scale_y_log10(limits=c(1e-4,1))

limlist = list(l1,l2,l3,l4)

ggplot(data = pars) + geom_line(aes(x=(x),y=surv, linetype=as.character(b),colour =as.character(b)), lwd=1) +
  geom_line(data = pars[pars$x>=pars$k0,],aes(x=(x+1),y=gp_surv, linetype=as.character(b)),colour ='grey')+
  scale_x_log10()+
  theme(aspect.ratio = 1/2) + theme_bw() + xlab('Total Degree')  +ylab('Survival') + labs(linetype=TeX('\\beta'), colour =TeX('\\beta')) +
  facet_grid(eps~a,labeller = labeller,scales='free') + facetted_pos_scales(y=limlist)

```

Since $\beta>0$, the shape parameter of the IGPD is positive and thus the distribution is heavy tailed. Additionally the value of the shape parameter $\xi$ is shown in @fig-polyheat for various parameter choices:

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 12
#| fig-height: 6
#| fig-cap: Heat maps of $\xi$ for various combinations of the parameters of the proposed model.
#| label: fig-polyheat
#| 
rho_optim_ba= Vectorize(function(a,eps, b, k0){
  return(abs(rho(2*b, polylin(a, eps), b, k0)-1))
},vectorize.args = 'a')
find_a_ba = Vectorize(function(b, eps, k0){
  out = optimise(rho_optim_ba, c(0.00001,3), b=b, eps=eps, k0=k0)$minimum
  return(out)
}, vectorize.args = 'b')
N =50
as = seq(0,2,l=N+1)[-1]
bs = seq(0,2,l=N+1)[-1]
eps = c(.01,.1,.5,1)
k0 = c(25,100)
pars = expand.grid(as,bs,eps,k0)
names(pars) = c('a', 'b', 'eps', 'k0')
lambdas = numeric(nrow(pars))
a_for_ba = numeric(nrow(pars))
for(i in 1:nrow(pars)){
  lambdas[i] = find_lambda2(polylin(pars$a[i], pars$eps[i]), pars$b[i], pars$k0[i])
  a_for_ba[i] = find_a_ba(pars$b[i], pars$eps[i], pars$k0[i])
}
pars = cbind(pars,lambdas,a_for_ba)
names(pars) = c('a', 'b', 'eps', 'k0','lambda','ba')
labeller = label_bquote(cols = `epsilon`==.(eps),rows = ~k[0]==.(k0))

ggplot(pars) + geom_raster(aes(x=a,y=b,fill=b/lambda)) +
  geom_line(aes(x=ba, y=b),linetype='dashed', colour='red',lwd=1)+
  scale_fill_paletteer_c(palette='grDevices::Blues',limits=c(0,1),direction=-1)+theme(aspect.ratio = 1)+ylim(min(bs), max(bs))+xlim(min(as), max(as))+
  facet_grid(k0~eps,labeller = labeller) + labs(fill=TeX('\\xi')) + xlab(TeX('\\alpha')) + ylab(TeX('\\beta')) + theme_minimal()

```

The darker regions on the heat maps correspond to a heavier tail and the lighter to a lighter tail, the red dashed line shows combinations of $\alpha$ and $\beta$ that produce a limiting degree distribution with the same tail heaviness as the Barabási-Albert model, $\xi=0.5$.

For a given network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$ and maximum degree $M$ the likelihood is given by:

\begin{align*}
L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}

where $B(y,z)$ is the the beta function and $l\ge0$ variable that allows truncating the data such that the minimum degree is $l$.

This likelihood allows for the parameters to be inferred using either a frequentist or a Bayesian approach.

# Recovery {#sec-rec}

The goal of was to provide a possible preference function that is able to explain the growth of real networks with varied tail behaviour in their degree distributions. This section aims to show that the parameters of the model in @sec-model can be recovered from the degree distribution of a network simulated from it.

The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongisde the priors:

\begin{align*}
\alpha&\sim \text{Ga}(1,0.01),\\
\beta &\sim  \text{Ga}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Ga(1,0.01)},
\end{align*}

to obtain a posterior distribution that can then be used in an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm  to obtain posterior samples. The results of this inference are shown in @fig-rec1 and @fig-rec2. For these simulated networks $l=0$.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 8
#| fig-height: 8
#| fig-cap: Posterior estimates of survival function for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.
#| label: fig-rec1

surfit_list = list()
pars = readRDS('results/recovery_pars.rds')
recovery_list = readRDS('results/recovery_dat.rds')
selected = c(1,27,14,35)
# selected = 1:36
for(k in 1:length(selected)){
  j=selected[k]
  x = recovery_list[[j]]$mcmc$dat[,1]
  ls = c()

  y_975 = recovery_list[[j]]$mcmc$surv$CI[2,][order(x)]
  y_025 = recovery_list[[j]]$mcmc$surv$CI[1,][order(x)]
  y_50 = recovery_list[[j]]$mcmc$surv$est[order(x)]
  x = sort(x)
  surfit_list[[k]] = ggplot() + geom_point(data=twbfn::deg_surv(recovery_list[[j]]$degs), aes(x=degree, y=surv)) +
    geom_line(data=NULL, aes(x=!!x, y=!!y_975), colour = 'red', linetype='dashed')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_50),colour='red')+
    geom_line(data=NULL, aes(x=!!x, y=!!y_025), colour = 'red', linetype='dashed')+
    scale_x_log10(limits=c(1,1e5))  +scale_y_log10(limits = c(1/length(recovery_list[[j]]$degs),1))+theme_bw() + theme(aspect.ratio = 0.66,axis.title.y=element_blank(),                                                                   axis.text.y=element_blank(),                                                                     axis.ticks.y=element_blank()) + xlab('')+ylab('') +
    ggtitle(TeX(paste0('$\\alpha = $',pars$a[j],
                       ', $\\epsilon = $',pars$eps[j],
                       ', $\\beta = $',pars$b[j])))
}

fig = ggpubr::ggarrange(plotlist = surfit_list,
          nrow=round(sqrt(length(surfit_list)),0),ncol=round(sqrt(length(surfit_list)),0),
          label.x='Degree', label.y = 'Survival')

ggpubr::annotate_figure(fig, bottom='Degree', left='Survival')

```

![: Posterior estimates of paramters for data simulated from the proposed model with various combinations of ($\alpha$,$\beta$,$\varepsilon$) and $k_0=20$.](images/mcmc_plot.jpg){#fig-rec2}

@fig-rec1 and @fig-rec2 demonstrate that using this methodology it is possible to recover the model parameters fairly well from only the final degree distribution of a simulated network. This indicates that the method may also be able to be applied to the degree distributions of real networks, estimating the model parameters assuming they evolved according to the GPA scheme.

# Application to Real Data {#sec-real}

Turning now to real data, the goal is to fit the model to the degree distributions of real networks from various sources. [describe networks being used]. Alongside fitting this model to the degree distributions, we then compare the fit to that of an mixture distribution that was used in [clement] and is similar to others that have been used to model degree distributions [others].

```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
#| label: fig-real1

plt1 = readRDS('results/comp_plot.rds')

plot(plt1)

```

@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model. In most cases, the GPA model does not necessarily provide an improvement in fit when compared to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function assuming that the network evolved according the the GPA scheme. @fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model. Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution and where it doesn't it appears to either be because it is fitting better at the tail than the Zipf-IGP model or because of the threshold being estimated as too low forcing almost all of the data to be modelled by the linear part of the GPA. This again shows the effects that small degrees have on this model, which is somewhat expected as the theory used for this model is for trees and none of these real networks (nor many real networks) are.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimates (solid red) of survival for several real data sets and their 95% credible intervals (dotted red)
#| label: fig-shapes

plt2 = readRDS('results/shape_plot.rds')

plot(plt2)

```

@fig-pa shows the estimated preference function $b$ alongside the 95% credible interval  on a log-log plot. Although the credible interval becomes very large for the largest degrees, this is expected as not all of these networks had data in that region, for those that do the credible interval is much narrower as is the case for `sx-mathoverflow`. The most insightful conclusion we can draw from these plots come from the shape of the preference function. There appears to be two distinct shapes of preference function. The first appears mostly flat (similar to UA) for the smallest degrees and then after a threshold PA kicks in, some with this shape are `pajek-erdos` and `sx-mathoverflow`. The second distinct shape appears provide some clear PA behaviour that then slows down after a certain point, examples of this are seen in the two infrastructure networks `opsahl-openflights` and `topology`.

```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
#| fig-width: 9
#| fig-height: 9
#| fig-cap: Posterior estimate for preference function (solid) with 95% credible interval (dashed) on log-log scale.
#| label: fig-pa

plt3 = readRDS('results/PA_plot.rds')
plot(plt3)

```

![: Posterior estimates of paramters for real data.](images/pars_plot.png)

# Conclusion and Discussion {#sec-conc}

In this paper we introduced a flexible class of preference function that when used under the GPA scheme (in the tree setting) is guaranteed to generate a network with a heavy tailed degree distribution whilst remaining flexible in the body. Using simulations from networks using this class of preference function we showed that the model parameters are quite easily estimated from the degrees alone at a snapshot in time. Seeing this we applied this method to the degree distributions of real networks, estimating their model parameters assuming they evolved in the same way. Not only did this yield fairly good fits for the degree distribution, similar to that of the Zipf-IGP, it came with the added benefit of giving a posterior estimate for a preference function.

Obviously this method had its flaws in that the lowest degrees needed to be truncated as they had a very large effect on the fit of the model as a result of using theory developed for trees and applying it to general networks. Hopefully, as the field progresses we will be able to apply theory developed for general networks using a similar method to this, allowing us to compare the results here something that is more accurate.

GPA is also a fairly simplistic one on its own despite having large flexibility in the preference function allowing for edges to only be added at times when a new node is introduced into the network and not allowing for anything like vertex/edge death, reciprocal edges and rewiring steps. [@deijfen2015] introduces some theory for GPA with vertex death based on the work by [@rudas07], but is unable to come to an expression for the degree distribution. However, adding something like fitness scores into the GPA model seems fairly simple to do even going so far as making each of the parameters used in the class of functions here node-wise instead of universal. For example having the preference weight for a vertex be given by $b(k_i, \underline{\gamma_i})$  where $\underline{\gamma_i}\sim G$ is the vertex's fitness and $k_i$ is the vertex's degree, it is reasonable to expect that the degree distribution for a model like this to be given by:

$$
\bar F(k) = \mathbb E_G\left[\prod_{i=0}^k \frac{b(i, \gamma)}{\lambda^* + b(i,\gamma)}\right],\qquad \gamma\sim G,\, k=0,1,2,\ldots
$$

If this were the case, the posterior distributions of parameters from @sec-real could be interpreted as the approximate distributions from which a vertex's fitness values are drawn instead of providing estimates of the universal parameter values. Additionally, it allows the ability to set some parameters as universal and some as vertex-wise for example one could let $\alpha, \beta, k_0$ be the same for all vertices but let each vertex in the network have its own "zero appeal" $\varepsilon_i\overset{\text{iid}}{\sim} \text{Ga}(\theta_1, \theta_2)$; moving the model to being more realistic and accounting for the fact that in many real networks the objects vertices represent have some value intrinsic to themselves other than the number of connections they have.

# Appendix {.appendix}

## Additional Results 
### Tail heaviness of general preferential attachment model

Recall the limiting survival function:

$$
\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
$$

The aim is to determine how this distribution behaves for different choices of $b$, specifically what maximum domain of attraction does this distribution belong to and is it affected by the choice of preference function $b$. @shimura12 introduces a quantity that will help in determining what domain of attraction a discrete distribution belongs to, that is:

For a distribution $F$ with survival function $\bar F$ and some $n\in\mathbb Z^+$ let:

$$
\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
$$

[@shimura12] then states that if $\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha$ ($\alpha>0$), then $F$ is heavy tailed with $\bar F(n) \sim n^{-\alpha}$. Additionally, if $\lim_{n\rightarrow\infty} \Omega(F,n) = 0$ then the distribution is light tailed.

Consider a preference function $b(\cdot)$ that fulfills:

$$
\lim_{n\rightarrow \infty} b(n) = \infty
$${#eq-condb1}

Substituting in the form of $\bar F(n)$ from @eq-surv and taking the limit, subject to @eq-condb1:

$$
\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
$$ {#eq-omegares} (see appendix for full proof).

Whilst there are many classes of functions that satisfy @eq-condb1 and @eq-condb2, the vast majority of these result in $\Omega(F,n) \rightarrow 0$, meaning that the limiting degree distribution is light tailed. In fact, in order for the limiting degree distribution to be heavy tailed, $b$ must be asymptotically linear ($b(k)\sim k$).

## Proofs

### Tail heaviness of GPA

Taking the form of the GPA degree survival: $$
\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
$$ and substituting into the formula for $\Omega(F,n)$:

```{=tex}
\begin{align*}
\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
\end{align*}
```
Clearly if $b(n)=c$ or $\lim_{n\rightarrow\infty}b(n)=c$ for some $c>0$ then $\Omega(F,n)=0$. Now consider a non-constant $b(n)$ and re-write $\Omega(F,n)$ as:

```{=tex}
\begin{align*}
\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
\end{align*}
```
Then if $\lim_{n\rightarrow\infty}b(n)=\infty$ it follows that:

```{=tex}
\begin{align*}
\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
\end{align*}
```

### Removing the infinite sum

For a preference function of the form:

$$
b(k) = \begin{cases}
g(k),&k<k_0\\
g(k_0) + \beta(k-k_0), &k\ge k_0
\end{cases}
$$ for $\beta>0, k_0\in\mathbb N$ we have that

```{=tex}
\begin{align*}
\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
\end{align*}
```
Now using the fact that:

$$
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
$$ and reindexing the product in the second sum

```{=tex}
\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
\end{align*}
```
In order to simplify the infinite sum, consider:

```{=tex}
\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
\end{align*}
```
this infinite sum does not converge when $x\le1$ as each term is $O(n^{-x})$. We can now use this in $\hat\rho(\lambda)$:

```{=tex}
\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
\end{align*}
```

# References
<!-- the pattern below controls the placement of the references -->
:::{#refs}
:::

### 2025-09-22 Files/archive/paper.tex
Restructuring

% Options for packages loaded elsewhere
% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  sn-basic,
]{sn-jnl}

\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{5}
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
% Make \paragraph and \subparagraph free-standing
\makeatletter
\ifx\paragraph\undefined\else
  \let\oldparagraph\paragraph
  \renewcommand{\paragraph}{
    \@ifstar
      \xxxParagraphStar
      \xxxParagraphNoStar
  }
  \newcommand{\xxxParagraphStar}[1]{\oldparagraph*{#1}\mbox{}}
  \newcommand{\xxxParagraphNoStar}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
  \let\oldsubparagraph\subparagraph
  \renewcommand{\subparagraph}{
    \@ifstar
      \xxxSubParagraphStar
      \xxxSubParagraphNoStar
  }
  \newcommand{\xxxSubParagraphStar}[1]{\oldsubparagraph*{#1}\mbox{}}
  \newcommand{\xxxSubParagraphNoStar}[1]{\oldsubparagraph{#1}\mbox{}}
\fi
\makeatother

\usepackage{longtable,booktabs,array}
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\usepackage{graphicx}
\makeatletter
\newsavebox\pandoc@box
\newcommand*\pandocbounded[1]{% scales image to fit in text height/width
  \sbox\pandoc@box{#1}%
  \Gscale@div\@tempa{\textheight}{\dimexpr\ht\pandoc@box+\dp\pandoc@box\relax}%
  \Gscale@div\@tempb{\linewidth}{\wd\pandoc@box}%
  \ifdim\@tempb\p@<\@tempa\p@\let\@tempa\@tempb\fi% select the smaller of both
  \ifdim\@tempa\p@<\p@\scalebox{\@tempa}{\usebox\pandoc@box}%
  \else\usebox{\pandoc@box}%
  \fi%
}
% Set default figure placement to htbp
\def\fps@figure{htbp}
\makeatother

\setlength{\emergencystretch}{3em} % prevent overfull lines

\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}

%%%% Standard Packages

\usepackage{graphicx}%
\usepackage{multirow}%
\usepackage{amsmath,amssymb,amsfonts}%
\usepackage{amsthm}%
\usepackage{mathrsfs}%
\usepackage[title]{appendix}%
\usepackage{xcolor}%
\usepackage{textcomp}%
\usepackage{manyfoot}%
\usepackage{booktabs}%
\usepackage{algorithm}%
\usepackage{algorithmicx}%
\usepackage{algpseudocode}%
\usepackage{listings}%

%%%%

\raggedbottom
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\usepackage{amsthm}
\theoremstyle{plain}
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{remark}
\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
\newtheorem*{remark}{Remark}
\newtheorem*{solution}{Solution}
\newtheorem{refremark}{Remark}[section]
\newtheorem{refsolution}{Solution}[section]
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  pdftitle={Tail Flexibility in the Degrees of Preferential Attachment Networks},
  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
  pdfkeywords={networks, extremes},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title[Tail Flexibility in the Degrees of Preferential Attachment
Networks]{Tail Flexibility in the Degrees of Preferential Attachment
Networks}

% author setup
\author[1]{\fnm{Thomas William} \sur{Boughen}}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
% affil setup
\affil[1]{\orgdiv{School of Mathematics, Statistics and
Physics}, \orgname{Newcastle University}}

% abstract

\abstract{Devising the underlying generating mechanism of a real-life
network is difficult as, more often than not, only its snapshots are
available, but not its full evolution. One candidate for the generating
mechanism is preferential attachment which, in its simplest form,
results in a degree distribution that follows the power law.
Consequently, the growth of real-life networks that roughly display such
power-law behaviour is commonly modelled by preferential attachment.
However, the validity of the power law has been challenged by the
presence of alternatives with comparable performance, as well as the
recent findings that the right tail of the degree distribution is often
lighter than implied by the body, whilst still being heavy. In this
paper, we study a modified version of the model with a flexible
preference function that allows super/sub-linear behaviour whilst also
guaranteeing that the limiting degree distribution has a heavy tail. We
relate the distributions tail heaviness directly to the model
parameters, allowing direct inference of the parameters from the degree
distribution alone.}

% keywords
\keywords{networks,  extremes}

\begin{document}
\maketitle

\section{Introduction}\label{sec-intro}

Networks appear in many fields such as sociology, politics,
epidemiology, and economics. Statistical methods have been and continue
to be used to study networks that appear in these fields and beyond,
from the use of stochastic block models to detect communities within the
French political blogosphere \citep{Latouche11}, to using Exponential
Random Graph Models (ERGMs) to analyse the global trade network
\citep{Setayesh22}, and the use of mechanistic models in neuroscience to
investigate the patterns of connections in neural systems
\citep{Betzel17}. One such mechanistic model is the general preferential
attachment (PA) model where as a network grows and new vertices join
they connect to those already present in the network with weights
proportional to some preference function \(b\) of their degree. The
well-known Barabási-Albert (BA) model is a special case of this when
\(b\sim k\) {[}BA99{]}.

Mechanistic models, like the PA model, are in general only particularly
useful when simulating networks and not fitting to the real data as
usually only snapshots of real networks are available and not the entire
evolution. As such, one of the most informative statistics when
regarding a network is the degree distribution. When looking at the
degree distributions of real networks it is attractive to many to use
the power law to model the degree distribution, as the BA model
generates degrees that follow a power law, and therefore claim that
since the degree distribution is roughly power law then the underlying
growth mechanism may be preferential attachment such as the BA model.
However, it has been shown that a general preference function will not
always lead to a power law degree distribution \citep{krapivsky01} and
additionally it has been debated for decades whether most real data
actually follows a power law \citep{Voitalov_2019} or
not\citep{Broido_2019}.

Throughout this debate the use of methods from extremes has been absent
until relatively recently with the papers such as \citep{Voitalov_2019}
and {[}Lee{]}. Forgoing the use of extreme value methods essentially
downplays how deviations from the power law in the largest degrees may
affect the network given that they are often much more influential than
the vertices with smaller degrees.

Noting this, many have now started to use these extreme value methods
when analysing networks. \citep{wan19} showed that extreme value
estimation methods prove more robust than existing parametric approaches
especially when the data are corrupted or the model has been
misspecified, demonstrating that extreme value methods can be a valuable
tool when studying networks. \citep{wang2022random} investigates a PA
model with a affine preference function (similar to the BA model) with
the added behaviour of heterogeneous reciprocal edges, they show that
the distribution of in-degrees and out-degrees produced by such a model
is multivariate regularly varying and they also have hidden regular
variation. {[}Lee{]} using methods from discrete extremes modelled the
right tail of network degrees with a discretised variation of the
generalised Pareto distribution (GPD) usually dubbed the Integer GPD
(IGPD) where \(X|X>v \sim \mathrm{IGPD}(\xi,\sigma,v)\) is defined by
the survival:

\[
\Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
\] for \(v\in\mathbb Z^+, \sigma>0,\xi\in \mathbb R\).

The study of degree distributions, whether using extremes or not,
usually does not reveal any information about the preference function
even if PA was the underlying generating mechanism of the network. Here
we address this gap in the literature, more specifically investigating
if given the degree distribution of a network assumed to come from the
PA model can we directly infer the model parameters? We build upon the
foundation provided by \citep{rudas07} deriving the limiting degree
distribution in terms of the preference function and suggest a class of
preference functions that result in realistic tails that are heavy but
are lighter than implied by the power law.

\subsection{Structure of this paper}\label{structure-of-this-paper}

Section~\ref{sec-gpa} looks at the theoretical limiting behaviour of a
GPA model when \(m=1\) according to the results from \citep{rudas07},
going on to introduce a flexible class of preference functions that can
guarantee a heavy tailed degree distribution while still being flexible
in the body. Section~\ref{sec-rec} shows that the preference function
parameters can be fairly well recovered from the degree distributions of
networks simulated from the GPA model using various functions of the
class introduced in Section~\ref{sec-gpa}. Having shown that for
simulated data the model parameters are recoverable from the degree
distribution alone, Section~\ref{sec-real} builds upon this and attempts
to fit the model to degree distributions of real networks; modelling not
only the degree distribution itself but also the estimates for the
parameters of the preference function assuming the networks evolved
according to the GPA scheme. Ultimately, Section~\ref{sec-conc}
discusses the main results of the paper, pitfalls of this method and
future work.

\section{Preferential attachment model}\label{sec-gpa}

The network generative model that we will be focussing on in this paper
is dubbed General Preferential Attachment in \citep{rudas07} and is
defined as follows:

Starting at time \(t=0\) with an initial network of \(m\) vertices that
each have no edges, at times \(t=1,2,\ldots\) a new vertex is added to
the network bringing with it \(m\) directed edges (with the new vertex
as the source); the target for each of these edges are selected from the
vertices already in the network with weights proportional to some
function \(b\) of their degree where the preference function \(b\) is
chosen such that:

\begin{equation}\phantomsection\label{eq-condb1}{
b:\mathbb N \mapsto \mathbb R^+\setminus\{0\},
}\end{equation}

\begin{equation}\phantomsection\label{eq-condb2}{
\sum_{k=0}^\infty\frac{1}{b(k)} = \infty.
}\end{equation}

Special cases of this model include the Barabási-Albert (BA) model when
\(b(k) = k+\varepsilon\), which leads to a power-law degree distribution
with index 2 and the Uniform Attachment (UA) model where \(b(k)=c\)
leading to a degree distribution in the Gumbel maximum domain of
attraction.

Given conditions Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}
an expression for the survival of the limiting degree distribution can
be found in the case that \(m=1\); obtained by considering a branching
process that is equivalent to the growth of the network, as in
\citep{rudas07}. Theorem 1 from \citep{rudas07} states that for the tree
\(\Upsilon(t)\) at time \(t\):

\[
\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]dt
\] where \(\lambda^*\) satisfies \(\hat\rho(\lambda^*)=1\).

The limiting survival can be viewed as the limit of the empirical
proportion of vertices with degree over a threshold \(k\), that is:

\[
\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1}
\] which by the previously stated theorem can also be written as:

\[
\bar F(k) = \frac{\int_0^\infty e^{-\lambda^* t}\mathbb E\left[\mathbb I\left\{\text{deg}(x,\Upsilon(t))>k\right\}\right]dt}{\int_0^\infty e^{-\lambda^* t}dt} = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
\] Additionally, using the fact that \(f(k) = \bar F(k-1) - \bar F(k)\)
the probability mass function (p.m.f.) can be shown to be:

\[
f(k) = \frac{\lambda^*}{\lambda^* + b(k)}\prod_{i=0}^{k-1}\frac{b(i)}{\lambda^*+b(i)},\qquad k\in\mathbb N.
\] We are specifically interested in how the tail heaviness (\(\xi\) in
the IGPD) is affected by the preference function \(b\),
\citep{shimura12} introduces a quantity that will help in determining
what domain of attraction this discrete distribution belongs to, that
is:

For a distribution \(F\) with survival function \(\bar F\) and some
\(k\in\mathbb Z^+\) let:

\[
\Omega(F,k) = \left(\log\displaystyle\frac{\bar F (k+1)}{\bar F (k+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (k)}{\bar F (k+1)}\right)^{-1}
\]

\citep{shimura12} then states that if
\(\lim_{n\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
then \(F\) is heavy tailed with \(\bar F(k) \sim k^{-\alpha}\).
Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,k) = 0\) then the
distribution is light tailed. This allows us to show the following:

\begin{proposition}[]\protect\hypertarget{prp-omega}{}\label{prp-omega}

If \(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\) then \[
\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}
\] and

\[
\lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
\]

\end{proposition}

See \hyperref[appendix]{appendix} for the details of the proof.

Proposition~\ref{prp-omega} aligns with the result from
\citep{krapivsky01} demonstrating that a sub-linear preference function
will lead to a light tailed distribution. So, in order for the degree
distribution to be heavy tailed we need that the limit
\(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) exists and is positive. We can
in fact show that BA model produces a heavy tailed degree distribution
with index \(\xi=0.5\) by considering the preference function
\(b(k) = k + \varepsilon\), noting that the limit
\(\lim_{k\rightarrow\infty} b(k+1)-b(k)\) is simply
\(\lim_{k\rightarrow\infty} 1 = 1\) leaving the tail heaviness to be
\(1/\lambda^*\) which using \(\hat\rho\) can be found to be \(1/2\).

We can show that in order for the degree distribution to be heavy
tailed, the preference function must be asymptotically linear
i.e.~\(\lim_{k\rightarrow\infty}\frac{b(k)}{k} =  c>0\).

Consider the definition of heavy tails:

\begin{align*}
\lim_{k\rightarrow\infty}\Omega(F,n) &= c>0\\
\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*} &= c\\
\lim_{k\rightarrow\infty}[b(k+1)-b(k)] &= \lambda^*c\\
\lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{(k+1)-(k)} &= \lambda^*c\\
\end{align*}

by Stolz-Cesaro Theorem:

\[
\lim_{k\rightarrow\infty}\frac{b(k)}{k} = \lambda^*c > 0.
\]

Since \(b\) must be asymptotically linear in order to have positive tail
heaviness we propose the following preference function:

\[
b(k) = \begin{cases}
k^\alpha + \varepsilon,&k<k_0\\
k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
\end{cases}
\] for \(\alpha,\beta, \varepsilon>0\) and \(k_0\in\mathbb N\).

This preference function, as per Proposition~\ref{prp-omega}, will
produce a degree distribution with tail heaviness \(\beta/\lambda^*\)
guaranteeing a heavy tail. We study this preference function further in
the next section.

\section{Preferential Attachment with flexible heavy
tail}\label{sec-model}

Using a preference function with guaranteed linear behaviour in the
limit, allows for the inclusion of sub/super linear behaviour without
losing the heavy tails or ending up with a degenerate degree
distribution.

The limiting degree distribution resulting from using a preference
function of this form can be found to have survival:

\begin{equation}\phantomsection\label{eq-polysurv}{
\bar F(k) = \begin{cases}
\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0\\
\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0.
\end{cases}
}\end{equation}

with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where:

\[
\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon} 
\]

which must be solved numerically for most parameter choices. Also, note
that \(\lambda>\beta\).

Some examples of what the degree distribution looks like are shown below
in Figure~\ref{fig-polylinsurv}:

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}

}

\caption{\label{fig-polylinsurv}Theoretical survival distributions of
the limiting degree distributions, according to various combinations of
\((\alpha, \beta, \varepsilon)\) and \(k_0=20\) of the proposed
preferential attachment model.}

\end{figure}%

The survival (Equation~\ref{eq-polysurv}) can be connected to the IGPD
mentioned in Section~\ref{sec-intro} by using Stirling's approximation
of the gamma function to obtain:

\[
\frac{\Gamma(x+y)}{\Gamma(x)}\approx x^{y},
\] and utilising the expression from Equation~\ref{eq-polysurv}:

\begin{align*}
\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}
\end{align*}

\[
\bar F(k) 
\begin{cases}
=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
\end{cases}
\] meaning that for \(k\ge k_0\) the limiting degree distribution is
similar to
\(\text{IGPD}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\).

To assess how close of an approximation this is the theoretical
conditional survivals are shown in Figure~\ref{fig-approx_surv} in
colour and their IGPD approximations are shown in grey. The
approximation seems to hold up fairly well even for large degrees.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}

}

\caption{\label{fig-approx_surv}Theoretical conditional survivals (grey)
alongside their IGPD approximations (coloured).}

\end{figure}%

Since \(\beta>0\), the shape parameter of the IGPD is positive and thus
the distribution is heavy tailed. Additionally the value of the shape
parameter \(\xi\) is shown in Figure~\ref{fig-polyheat} for various
parameter choices:

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}

}

\caption{\label{fig-polyheat}Heat maps of \(\xi\) for various
combinations of the parameters of the proposed model.}

\end{figure}%

The darker regions on the heat maps correspond to a heavier tail and the
lighter to a lighter tail, the red dashed line shows combinations of
\(\alpha\) and \(\beta\) that produce a limiting degree distribution
with the same tail heaviness as the Barabási-Albert model, \(\xi=0.5\).

For a given network with degree count vector
\(\pmb n = (n_0, n_1, \ldots, n_M)\) and maximum degree \(M\) the
likelihood is given by:

\begin{align*}
L(\pmb x,\pmb n | \pmb \theta) = \left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ \times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
\end{align*}

where \(B(y,z)\) is the the beta function and \(l\ge0\) variable that
allows truncating the data such that the minimum degree is \(l\).

This likelihood allows for the parameters to be inferred using either a
frequentist or a Bayesian approach.

\section{Recovery}\label{sec-rec}

The goal of was to provide a possible preference function that is able
to explain the growth of real networks with varied tail behaviour in
their degree distributions. This section aims to show that the
parameters of the model in Section~\ref{sec-model} can be recovered from
the degree distribution of a network simulated from it.

The procedure for recovering the parameters begins with simulating a
network from the model with \(N=100,000\) vertices and \(m=1\) given
some set of parameters
\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
counts and using the likelihood from the previous section alongisde the
priors:

\begin{align*}
\alpha&\sim \text{Ga}(1,0.01),\\
\beta &\sim  \text{Ga}(1,0.01),\\
k_0 &\sim \text{U}(1,10,000),\\
\varepsilon &\sim \text{Ga(1,0.01)},
\end{align*}

to obtain a posterior distribution that can then be used in an adaptive
Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm to obtain
posterior samples. The results of this inference are shown in
Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2}. For these simulated
networks \(l=0\).

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}

}

\caption{\label{fig-rec1}Posterior estimates of survival function for
data simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

\end{figure}%

\begin{figure}

\centering{

\pandocbounded{\includegraphics[keepaspectratio]{images/mcmc_plot.jpg}}

}

\caption{\label{fig-rec2}: Posterior estimates of paramters for data
simulated from the proposed model with various combinations of
(\(\alpha\),\(\beta\),\(\varepsilon\)) and \(k_0=20\).}

\end{figure}%

Figure~\ref{fig-rec1} and Figure~\ref{fig-rec2} demonstrate that using
this methodology it is possible to recover the model parameters fairly
well from only the final degree distribution of a simulated network.
This indicates that the method may also be able to be applied to the
degree distributions of real networks, estimating the model parameters
assuming they evolved according to the GPA scheme.

\section{Application to Real Data}\label{sec-real}

Turning now to real data, the goal is to fit the model to the degree
distributions of real networks from various sources. {[}describe
networks being used{]}. Alongside fitting this model to the degree
distributions, we then compare the fit to that of an mixture
distribution that was used in {[}clement{]} and is similar to others
that have been used to model degree distributions {[}others{]}.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}

}

\caption{\label{fig-real1}Posterior estimates (solid red) of survival
for several real data sets and their 95\% credible intervals (dotted
red)}

\end{figure}%

Figure~\ref{fig-real1} displays the posterior estimates of the survival
function for various data sets, obtained from fitting the GPA model and
the Zipf-IGP mixture model. In most cases, the GPA model does not
necessarily provide an improvement in fit when compared to the Zipf-IGP
model but where the GPA model fits well we gain additional information
about the preference function assuming that the network evolved
according the the GPA scheme. Figure~\ref{fig-shapes} shows the
posterior of the shape parameter \(\xi\) obtained from the Zipf-IGP
model alongside the posterior of the equivalent shape parameter
\(\beta/\lambda^*\) obtained from fitting the GPA model. Generally, the
GPA model performs similarly to the Zipf-IGP when estimating the tail
behaviour of the degree distribution and where it doesn't it appears to
either be because it is fitting better at the tail than the Zipf-IGP
model or because of the threshold being estimated as too low forcing
almost all of the data to be modelled by the linear part of the GPA.
This again shows the effects that small degrees have on this model,
which is somewhat expected as the theory used for this model is for
trees and none of these real networks (nor many real networks) are.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}

}

\caption{\label{fig-shapes}Posterior estimates (solid red) of survival
for several real data sets and their 95\% credible intervals (dotted
red)}

\end{figure}%

Figure~\ref{fig-pa} shows the estimated preference function \(b\)
alongside the 95\% credible interval on a log-log plot. Although the
credible interval becomes very large for the largest degrees, this is
expected as not all of these networks had data in that region, for those
that do the credible interval is much narrower as is the case for
\texttt{sx-mathoverflow}. The most insightful conclusion we can draw
from these plots come from the shape of the preference function. There
appears to be two distinct shapes of preference function. The first
appears mostly flat (similar to UA) for the smallest degrees and then
after a threshold PA kicks in, some with this shape are
\texttt{pajek-erdos} and \texttt{sx-mathoverflow}. The second distinct
shape appears provide some clear PA behaviour that then slows down after
a certain point, examples of this are seen in the two infrastructure
networks \texttt{opsahl-openflights} and \texttt{topology}.

\begin{figure}

\centering{

\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pa-1.pdf}

}

\caption{\label{fig-pa}Posterior estimate for preference function
(solid) with 95\% credible interval (dashed) on log-log scale.}

\end{figure}%

\begin{figure}[H]

{\centering \pandocbounded{\includegraphics[keepaspectratio]{images/pars_plot.png}}

}

\caption{: Posterior estimates of paramters for real data.}

\end{figure}%

\section{Conclusion and Discussion}\label{sec-conc}

In this paper we introduced a flexible class of preference function that
when used under the GPA scheme (in the tree setting) is guaranteed to
generate a network with a heavy tailed degree distribution whilst
remaining flexible in the body. Using simulations from networks using
this class of preference function we showed that the model parameters
are quite easily estimated from the degrees alone at a snapshot in time.
Seeing this we applied this method to the degree distributions of real
networks, estimating their model parameters assuming they evolved in the
same way. Not only did this yield fairly good fits for the degree
distribution, similar to that of the Zipf-IGP, it came with the added
benefit of giving a posterior estimate for a preference function.

Obviously this method had its flaws in that the lowest degrees needed to
be truncated as they had a very large effect on the fit of the model as
a result of using theory developed for trees and applying it to general
networks. Hopefully, as the field progresses we will be able to apply
theory developed for general networks using a similar method to this,
allowing us to compare the results here something that is more accurate.

GPA is also a fairly simplistic one on its own despite having large
flexibility in the preference function allowing for edges to only be
added at times when a new node is introduced into the network and not
allowing for anything like vertex/edge death, reciprocal edges and
rewiring steps. \citep{deijfen2015} introduces some theory for GPA with
vertex death based on the work by \citep{rudas07}, but is unable to come
to an expression for the degree distribution. However, adding something
like fitness scores into the GPA model seems fairly simple to do even
going so far as making each of the parameters used in the class of
functions here node-wise instead of universal. For example having the
preference weight for a vertex be given by
\(b(k_i, \underline{\gamma_i})\) where \(\underline{\gamma_i}\sim G\) is
the vertex's fitness and \(k_i\) is the vertex's degree, it is
reasonable to expect that the degree distribution for a model like this
to be given by:

\[
\bar F(k) = \mathbb E_G\left[\prod_{i=0}^k \frac{b(i, \gamma)}{\lambda^* + b(i,\gamma)}\right],\qquad \gamma\sim G,\, k=0,1,2,\ldots
\]

If this were the case, the posterior distributions of parameters from
Section~\ref{sec-real} could be interpreted as the approximate
distributions from which a vertex's fitness values are drawn instead of
providing estimates of the universal parameter values. Additionally, it
allows the ability to set some parameters as universal and some as
vertex-wise for example one could let \(\alpha, \beta, k_0\) be the same
for all vertices but let each vertex in the network have its own ``zero
appeal''
\(\varepsilon_i\overset{\text{iid}}{\sim} \text{Ga}(\theta_1, \theta_2)\);
moving the model to being more realistic and accounting for the fact
that in many real networks the objects vertices represent have some
value intrinsic to themselves other than the number of connections they
have.

\section{Appendix}\label{appendix}

\subsection{Additional Results}\label{additional-results}

\subsubsection{Tail heaviness of general preferential attachment
model}\label{tail-heaviness-of-general-preferential-attachment-model}

Recall the limiting survival function:

\[
\bar F(k) = \prod_{i=0}^k \frac{b(i)}{\lambda^*+b(i)},
\]

The aim is to determine how this distribution behaves for different
choices of \(b\), specifically what maximum domain of attraction does
this distribution belong to and is it affected by the choice of
preference function \(b\). \citet{shimura12} introduces a quantity that
will help in determining what domain of attraction a discrete
distribution belongs to, that is:

For a distribution \(F\) with survival function \(\bar F\) and some
\(n\in\mathbb Z^+\) let:

\[
\Omega(F,n) = \left(\log\displaystyle\frac{\bar F (n+1)}{\bar F (n+2)}\right)^{-1} - \left(\log\displaystyle\frac{\bar F (n)}{\bar F (n+1)}\right)^{-1}
\]

\citep{shimura12} then states that if
\(\lim_{n\rightarrow\infty} \Omega(F,n) = 1/\alpha\) (\(\alpha>0\)),
then \(F\) is heavy tailed with \(\bar F(n) \sim n^{-\alpha}\).
Additionally, if \(\lim_{n\rightarrow\infty} \Omega(F,n) = 0\) then the
distribution is light tailed.

Consider a preference function \(b(\cdot)\) that fulfills:

\begin{equation}\phantomsection\label{eq-condb1}{
\lim_{n\rightarrow \infty} b(n) = \infty
}\end{equation}

Substituting in the form of \(\bar F(n)\) from \textbf{?@eq-surv} and
taking the limit, subject to Equation~\ref{eq-condb1}:

\begin{equation}\phantomsection\label{eq-omegares}{
\lim_{n\rightarrow\infty}\Omega(F,n) = \lim_{n\rightarrow\infty}\frac{b(n+2)-b(n+1)}{\lambda^*}.
}\end{equation} (see appendix for full proof).

Whilst there are many classes of functions that satisfy
Equation~\ref{eq-condb1} and Equation~\ref{eq-condb2}, the vast majority
of these result in \(\Omega(F,n) \rightarrow 0\), meaning that the
limiting degree distribution is light tailed. In fact, in order for the
limiting degree distribution to be heavy tailed, \(b\) must be
asymptotically linear (\(b(k)\sim k\)).

\subsection{Proofs}\label{proofs}

\subsubsection{Tail heaviness of GPA}\label{tail-heaviness-of-gpa}

Taking the form of the GPA degree survival: \[
\bar F(n) = \prod_{i=0}^n\frac{b(i)}{\lambda+b(i)}
\] and substituting into the formula for \(\Omega(F,n)\):

\begin{align*}
\Omega(F,n)&=\left(\log\frac{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{n}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{n+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
&=\left(\log\frac{\lambda+b(n+2)}{b(n+2)}\right)^{-1}-\left(\log\frac{\lambda+b(n+1)}{b(n+1)}\right)^{-1}\\
&=\left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}
\end{align*}

Clearly if \(b(n)=c\) or \(\lim_{n\rightarrow\infty}b(n)=c\) for some
\(c>0\) then \(\Omega(F,n)=0\). Now consider a non-constant \(b(n)\) and
re-write \(\Omega(F,n)\) as:

\begin{align*}
\Omega(F,n) &= \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}+\frac{b(n+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}+\frac{b(n+1)}{\lambda}  -\frac{b(n+1)}{\lambda}\\
&=\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}+\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}
\end{align*}

Then if \(\lim_{n\rightarrow\infty}b(n)=\infty\) it follows that:

\begin{align*}
\lim_{n\rightarrow\infty}\Omega(F,n) &= \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+2)}\right]\right)^{-1}-\frac{b(n+2)}{\lambda}\right\} - \lim_{n\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(n+1)}\right]\right)^{-1}-\frac{b(n+1)}{\lambda}\right\}\\
&\qquad+\lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{2}-\frac{1}{2} + \lim_{n\rightarrow\infty}\left(\frac{b(n+2)}{\lambda}-\frac{b(n+1)}{\lambda}\right)\\
&=\frac{1}{\lambda}\lim_{n\rightarrow\infty}\left[b(n+2)-b(n+1)\right]\qquad \square
\end{align*}

\subsubsection{Removing the infinite
sum}\label{removing-the-infinite-sum}

For a preference function of the form:

\[
b(k) = \begin{cases}
g(k),&k<k_0\\
g(k_0) + \beta(k-k_0), &k\ge k_0
\end{cases}
\] for \(\beta>0, k_0\in\mathbb N\) we have that

\begin{align*}
\hat\rho(\lambda) = \sum_{n=0}^\infty\prod_{i=0}^{n-1}\frac{b(i)}{\lambda+b(i)} &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \sum_{n=k_0+1}^\infty\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\prod_{i=k_0}^{n-1}\frac{g(k_0) + \beta(i-k_0)}{\lambda +g(k_0) + \beta(i-k_0)}
\end{align*}

Now using the fact that:

\[
\prod_{i=0}^n(x+yi) = x^{n+1}\frac{\Gamma(\frac{x}{y}+n+1)}{\Gamma(\frac{x}{y})}
\] and reindexing the product in the second sum

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}
\end{align*}

In order to simplify the infinite sum, consider:

\begin{align*}
\sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{y}(n+x,y)\\
&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
&= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}
\end{align*}

this infinite sum does not converge when \(x\le1\) as each term is
\(O(n^{-x})\). We can now use this in \(\hat\rho(\lambda)\):

\begin{align*}
\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\qquad \square
\end{align*}

\section{References}\label{references}

\renewcommand{\bibsection}{}
\bibliography{refs.bib}

\end{document}

### 2025-09-22 Files/presentation/pres.qmd
Restructuring

---
title: diaquarto 
author: Author Name
date: "01.11.2024"
subtitle: easy slides in typst
toc: true
format:
  diaquarto-typst: 
    layout: medium # small, medium, large
    ratio: 16/9 # any ratio possible 
    title-color: "154C79" # Hex code for the title color (without #)
---

# About *diatypst*

## Usage

To use this template, add it to your Quarto project with the following command:

```bash
quarto add skriptum/diatypst/diaquarto
```

Then, create a qmd file with the following YAML frontmatter:

```yaml
title: Untitled # required
author: Author Name
date: 01.07.2024
subtitle: easy slides in typst
format:
  diaquarto-typst: # required
    layout: medium 
    ratio: 16/9 
```

Then, insert your content.

- Level-one headings corresponds to new sections.
- Level-two headings corresponds to new slides.
- `---` is used to separate slides.

## Styling

This is the second slide, we test some more content here

> A blockquote

A table:

| A | B | C |
|---|---|---|
| 1 | 2 | 3 |

And finally a

Term
: with a Definition

### 2025-09-24 Files/paper.qmd
Fixing typos

--- 2025-09-22 paper.qmd
@@
 On the side against the claim is @Broido_2019 who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority in real networks.
-On the other side of this debate is @Voitalov_2019 who disagrees and claims that these networks are not nearly as rare and only appear to so as a result of an unrealistic expectation of a power law without deviations or noise.
+On the other side of this debate is @Voitalov_2019 who disagrees and claims that these networks are not nearly as rare and only appear to be so as a result of an unrealistic expectation of a power law without deviations or noise.
 Additional evidence of these deviations from a power law is shown by @Lee24 who demonstrate that a lot of networks are partially scale-free, in that the body of the degree distribution is often modelled well by a power law, while the tail is lighter than what is implied by the body, albeit still regularly varying.
@@
 @wang2022random returns to a linear preference function of the form $b(k) = k+\varepsilon$ but adds the possibility for reciprocal edges to be sent, resulting in the joint distribution of in-degree and out-degree being multivariate regularly varying and having the property of hidden regular variation.
-@rudas07 follows in the footsteps of @krapivsky01, by considering a PA tree and using theory from continuous branching processes, derives a limiting degree distribution in terms of the preference function $b(\cdot)$.
+@rudas07 followed in the footsteps of @krapivsky01, by considering a PA tree and using theory from continuous branching processes, deriving a limiting degree distribution in terms of the preference function $b(\cdot)$.
 Nevertheless, research in this area tends to only focus on the theoretical asymptotic results of network growth models with little analysis of real networks.
@@
 Using these results, we demonstrate how the tail of the degree distribution is affected by $b(\cdot)$, and subsequently propose a class of preference functions that is tail realistic for real networks.
-These analytical results enables the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to PA, to be inferred directly.
+These analytical results enable the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to PA, to be inferred directly.
 
@@
 @sec-model utilises the proposed preference function and illustrates numerically how the tail index of the degree distribution varies with the model parameters.
-The simulation study in @sec-sim demonstrates that the parameters can be recovered from fitting the model to only the degree distribution.
-@sec-real fits the model to some real data and provides posterior estimates for the preference function.
+The simulation study in Subsection -@sec-sim demonstrates that the parameters can be recovered from fitting the model to only the degree distribution.
+Subsection -@sec-real fits the model to some real data and provides posterior estimates for the preference function.
 @sec-conc provides a discussion of this paper and possible avenues for future work.
@@
 ::: {#rem-omega}
-If $b(k)\rightarrow <\infty$, then $\Omega(F, k) = 0$.
+If $\lim_{k\rightarrow\infty}b(k)<\infty$, then $\Omega(F, k) = 0$.
 :::
@@
 ::: {#cor-omega2}
-Let $\bar F (k)$ bet he limiting survival function of degree in a GPA network with preference function $b$ such that $b(k)\rightarrow\infty$. Then:
+Let $\bar F (k)$ be the limiting survival function of degrees in a GPA network with preference function $b$ such that $b(k)\rightarrow\infty$. Then:
 
@@
 @cor-omega2 is a direct consequence of @prp-omega and the results of @shimura12, and it aligns with previous findings in the literature.
-In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow}[b(k+1)-b(k)]=0$.
+In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$.
 Conversely, it also agrees with the limiting behavior of BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+\varepsilon$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $1/\lambda^*$, in which cases the Malthusian parameter $\lambda^*$ is 2.
@@
 $$
-\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda}.
+\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}.
 $$

### 2025-09-24 Files/paper.tex
Fixing references

--- 2025-09-22 paper.tex
@@
 of this debate is \citet{Voitalov_2019} who disagrees and claims that
-these networks are not nearly as rare and only appear to so as a result
-of an unrealistic expectation of a power law without deviations or
-noise. Additional evidence of these deviations from a power law is shown
-by \citet{Lee24} who demonstrate that a lot of networks are partially
-scale-free, in that the body of the degree distribution is often
-modelled well by a power law, while the tail is lighter than what is
-implied by the body, albeit still regularly varying. Nevertheless, most
-studies into the appropriateness of a power law for the degrees of real
-networks, the aforementioned references included, have been largely
+these networks are not nearly as rare and only appear to be so as a
+result of an unrealistic expectation of a power law without deviations
+or noise. Additional evidence of these deviations from a power law is
+shown by \citet{Lee24} who demonstrate that a lot of networks are
+partially scale-free, in that the body of the degree distribution is
+often modelled well by a power law, while the tail is lighter than what
+is implied by the body, albeit still regularly varying. Nevertheless,
+most studies into the appropriateness of a power law for the degrees of
+real networks, the aforementioned references included, have been largely
 descriptive in the sense that no information about the growth of the
@@
 in-degree and out-degree being multivariate regularly varying and having
-the property of hidden regular variation. \citet{rudas07} follows in the
-footsteps of \citet{krapivsky01}, by considering a PA tree and using
-theory from continuous branching processes, derives a limiting degree
+the property of hidden regular variation. \citet{rudas07} followed in
+the footsteps of \citet{krapivsky01}, by considering a PA tree and using
+theory from continuous branching processes, deriving a limiting degree
 distribution in terms of the preference function \(b(\cdot)\).
@@
 subsequently propose a class of preference functions that is tail
-realistic for real networks. These analytical results enables the
+realistic for real networks. These analytical results enable the
 likelihood of the degree distribution to be expressed in terms of the
@@
 illustrates numerically how the tail index of the degree distribution
-varies with the model parameters. The simulation study in
-Section~\ref{sec-sim} demonstrates that the parameters can be recovered
-from fitting the model to only the degree distribution.
-Section~\ref{sec-real} fits the model to some real data and provides
-posterior estimates for the preference function. Section~\ref{sec-conc}
-provides a discussion of this paper and possible avenues for future
-work.
+varies with the model parameters. The simulation study in Subsection
+\ref{sec-sim} demonstrates that the parameters can be recovered from
+fitting the model to only the degree distribution. Subsection
+\ref{sec-real} fits the model to some real data and provides posterior
+estimates for the preference function. Section~\ref{sec-conc} provides a
+discussion of this paper and possible avenues for future work.
 
@@
 \begin{refremark}
-If \(b(k)\rightarrow <\infty\), then \(\Omega(F, k) = 0\).
+If \(\lim_{k\rightarrow\infty}b(k)<\infty\), then \(\Omega(F, k) = 0\).
 
@@
 
-Let \(\bar F (k)\) bet he limiting survival function of degree in a GPA
+Let \(\bar F (k)\) be the limiting survival function of degrees in a GPA
 network with preference function \(b\) such that
@@
 \(b(k)=k^\alpha\) with \(\alpha<1\), we have
-\(\lim_{k\rightarrow}[b(k+1)-b(k)]=0\). Conversely, it also agrees with
-the limiting behavior of BA model, which produces a regularly degree
-distribution with tail index 2; for the preference function
+\(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\). Conversely, it also agrees
+with the limiting behavior of BA model, which produces a regularly
+degree distribution with tail index 2; for the preference function
 \(b(k)=k+\varepsilon\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\),
@@
 \[
-\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda}.
+\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}.
 \]

### 2025-09-24 Files/paper.qmd
Small abstract changes

@@
 abstract: |
-  Devising the underlying generating mechanism of a real-life network is difficult as, more often than not, only its snapshots are available, but not its full evolution.
+  Identifying the generating mechanism of a network is challenging as, more often than not, only snapshots are available, but not the full evolution.
 One candidate for the generating mechanism is preferential attachment which, in its simplest form, results in a degree distribution that follows the power law.
 Consequently, the growth of real-life networks that display such power-law behaviour is commonly modelled by preferential attachment.
-However, the validity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the right tail of the degree distribution is often lighter than implied by the body, whilst still being heavy.
-In this paper, we study a modified version of the model with a flexible preference function that allows super/sub-linear behaviour whilst also guaranteeing that the limiting degree distribution has a heavy tail.
-We relate the distributions tail index directly to the model parameters, allowing direct inference of the parameters from the degree distribution alone.
+The power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the tail of the degree distribution is often lighter than implied by the body, whilst still being regularly varying.
+We use methods from discrete extremes to study a version of the model with a flexible preference function that allows super/sub-linear behaviour whilst guaranteeing the limiting degree distribution is regularly varying.
+We relate the tail index directly to the model parameters and enable inference of parameters from the degree distribution alone.
 keywords: [networks, discrete extremes,power law, preferential attachment]

### 2025-09-24 Files/paper.tex
Small abstract changes

@@
 \theoremstyle{plain}
+\newtheorem{corollary}{Corollary}[section]
+\theoremstyle{plain}
 \newtheorem{proposition}{Proposition}[section]
-\theoremstyle{plain}
-\newtheorem{corollary}{Corollary}[section]
 \theoremstyle{remark}
@@
 
-\abstract{Devising the underlying generating mechanism of a real-life
-network is difficult as, more often than not, only its snapshots are
-available, but not its full evolution. One candidate for the generating
-mechanism is preferential attachment which, in its simplest form,
-results in a degree distribution that follows the power law.
-Consequently, the growth of real-life networks that display such
-power-law behaviour is commonly modelled by preferential attachment.
-However, the validity of the power law has been challenged by the
-presence of alternatives with comparable performance, as well as the
-recent findings that the right tail of the degree distribution is often
-lighter than implied by the body, whilst still being heavy. In this
-paper, we study a modified version of the model with a flexible
-preference function that allows super/sub-linear behaviour whilst also
-guaranteeing that the limiting degree distribution has a heavy tail. We
-relate the distributions tail index directly to the model parameters,
-allowing direct inference of the parameters from the degree distribution
-alone.}
+\abstract{Identifying the generating mechanism of a real-life network is
+difficult as, more often than not, only snapshots are available, but not
+the full evolution. One candidate for the generating mechanism is
+preferential attachment which, in its simplest form, results in a degree
+distribution that follows the power law. Consequently, the growth of
+real-life networks that display such power-law behaviour is commonly
+modelled by preferential attachment. Recently, the power law has been
+challenged by the presence of alternatives with comparable performance,
+as well as the recent findings that the tail of the degree distribution
+is often lighter than implied by the body, whilst still being regularly
+varying. We use methods from discrete extemes to study a version of the
+model with a flexible preference function that allows super/sub-linear
+behaviour whilst guaranteeing the limiting degree distribution is
+regularly varying. We relate the distributions tail index directly to
+the model parameters and infer their true values from the degree
+distribution alone.}
 

### 2025-09-24 Files/paper.tex
Previous, with the pdf

@@
 
-\abstract{Identifying the generating mechanism of a real-life network is
-difficult as, more often than not, only snapshots are available, but not
-the full evolution. One candidate for the generating mechanism is
+\abstract{Identifying the generating mechanism of a network is
+challenging as, more often than not, only snapshots are available, but
+not the full evolution. One candidate for the generating mechanism is
 preferential attachment which, in its simplest form, results in a degree
@@
 real-life networks that display such power-law behaviour is commonly
-modelled by preferential attachment. Recently, the power law has been
-challenged by the presence of alternatives with comparable performance,
-as well as the recent findings that the tail of the degree distribution
-is often lighter than implied by the body, whilst still being regularly
-varying. We use methods from discrete extemes to study a version of the
-model with a flexible preference function that allows super/sub-linear
+modelled by preferential attachment. The power law has been challenged
+by the presence of alternatives with comparable performance, as well as
+the recent findings that the tail of the degree distribution is often
+lighter than implied by the body, whilst still being regularly varying.
+We use methods from discrete extremes to study a version of the model
+with a flexible preference function that allows super/sub-linear
 behaviour whilst guaranteeing the limiting degree distribution is
-regularly varying. We relate the distributions tail index directly to
-the model parameters and infer their true values from the degree
+regularly varying. We relate the tail index directly to the model
+parameters and enable inference of parameters from the degree
 distribution alone.}

### 2025-09-27 Files/paper.qmd
Trying to implement shared counters on thmlike

@@
 Consequently, the growth of real-life networks that display such power-law behaviour is commonly modelled by preferential attachment.
-The power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the tail of the degree distribution is often lighter than implied by the body, whilst still being regularly varying.
-We use methods from discrete extremes to study a version of the model with a flexible preference function that allows super/sub-linear behaviour whilst guaranteeing the limiting degree distribution is regularly varying.
-We relate the tail index directly to the model parameters and enable inference of parameters from the degree distribution alone.
-keywords: [networks, discrete extremes,power law, preferential attachment]
+The ubiquity of the power law has been challenged by the presence of alternatives with comparable performance, as well as the recent findings that the tail of the degree distribution is often lighter than implied by the body, whilst still being regularly varying.
+In this paper, we propose a preferential attachment model with a flexible preference function.
+Using methods for discrete extremes, we characterise the tail behaviour of the limiting degree distribution directly by the form of the preference function.
+Directly relating the tail index to the model parameters enables them to be inferred when fitting to degree distributions alone, which is supported by simulation studies.
+Results of applications to real data are promising and comparable to alternatives.
 bibliography: refs.bib  
 keep-tex: true
-pdf-engine-opt: -outdir=''
-output-temp-dir: 'temporary_files'
 ---
@@
 
-As the degrees of networks are discrete values, we will use methods from discrete extreme value theory, in particular those by @shimura12, who provided theoretical guarantees for a discrete distribution to be regularly varying or not.
-Using these results, we demonstrate how the tail of the degree distribution is affected by $b(\cdot)$, and subsequently propose a class of preference functions that is tail realistic for real networks.
+While Voitalov et al.
+(2019) have pioneered using methods in extreme value theory to analyse degree distributions, they overlooked the inapplicability of standard tools for continuous extremes, due to the discrete nature of degrees.
+A key difficulty arises because many standard discrete distributions, such as the Poisson, geometric, or negative binomial, do not satisfy the conditions required to belong to a maximum domain of attraction (MDA).
+Shimura (2012) showed that for a discrete distribution $F$ must be long-tailed in order to be in an MDA, and introduced the notion of recoverability to the MDA, also referred to as the discrete MDA (Hitz et al., 2024).
+Using the theoretical guarantees of regular variation for a discrete distribution by Shimura (2012), we demonstrate how the tail of the degree distribution is directly influenced by $b(\cdot)$, and subsequently proposed a class of preference functions that not only imply regularly varying degree distributions but are also tail realistic for real networks.
 These analytical results enable the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to PA, to be inferred directly.
@@
 
-Now, let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
+Now, let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t)))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
 
@@
 
-We investigate how the tail behaviour of the discrete limiting degree distribution is influenced by the preference function $b(·)$.
-Since the limiting degree distribution is discrete, standard tools for analysing extremes of continuous distributions are not directly applicable.
-A key difficulty arises because many standard discrete distributions, such as the Poisson, Geometric, or Negative Binomial, do not satisfy the conditions required to belong to an MDA.
-@shimura12 showed that for a discrete random variable X to lie in an MDA, its distribution must be long-tailed and introduced the notion of recoverability to the MDA, sometimes referred to as the discrete MDA (see e.g.
-@hitz24).
-In particular, if a distribution is both long-tailed and recoverable, then it can be placed in the appropriate MDA.
+We investigate how the tail behaviour of the discrete limiting degree distribution is influenced by the preference function $b(·)$, using the characterisation by @shimura12.
 
-A central tool in this characterisation is the quantity $\Omega(F,k)$. For a discrete distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ define:
+A central tool is the quantity $\Omega(F,k)$. For a discrete distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ define:
 
@@
 In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$.
-Conversely, it also agrees with the limiting behavior of BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+\varepsilon$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $1/\lambda^*$, in which cases the Malthusian parameter $\lambda^*$ is 2.
+Conversely, it also agrees with the limiting behavior of BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+\varepsilon$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$
 

### 2025-09-27 Files/paper.tex
Trying to implement shared counters on thmlike

@@
 \theoremstyle{plain}
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{plain}
 \newtheorem{corollary}{Corollary}[section]
-\theoremstyle{plain}
-\newtheorem{proposition}{Proposition}[section]
 \theoremstyle{remark}
@@
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
-  pdfkeywords={networks, discrete extremes, power law, preferential
-attachment},
   colorlinks=true,
@@
 real-life networks that display such power-law behaviour is commonly
-modelled by preferential attachment. The power law has been challenged
-by the presence of alternatives with comparable performance, as well as
-the recent findings that the tail of the degree distribution is often
-lighter than implied by the body, whilst still being regularly varying.
-We use methods from discrete extremes to study a version of the model
-with a flexible preference function that allows super/sub-linear
-behaviour whilst guaranteeing the limiting degree distribution is
-regularly varying. We relate the tail index directly to the model
-parameters and enable inference of parameters from the degree
-distribution alone.}
+modelled by preferential attachment. The ubiquity of the power law has
+been challenged by the presence of alternatives with comparable
+performance, as well as the recent findings that the tail of the degree
+distribution is often lighter than implied by the body, whilst still
+being regularly varying. In this paper, we propose a preferential
+attachment model with a flexible preference function. Using methods for
+discrete extremes, we characterise the tail behaviour of the limiting
+degree distribution directly by the form of the preference function.
+Directly relating the tail index to the model parameters enables them to
+be inferred when fitting to degree distributions alone, which is
+supported by simulation studies. Results of applications to real data
+are promising and comparable to alternatives.}
 
 % keywords
-\keywords{networks,  discrete extremes,  power law,  preferential
-attachment}
 
@@
 
-As the degrees of networks are discrete values, we will use methods from
-discrete extreme value theory, in particular those by \citet{shimura12},
-who provided theoretical guarantees for a discrete distribution to be
-regularly varying or not. Using these results, we demonstrate how the
-tail of the degree distribution is affected by \(b(\cdot)\), and
-subsequently propose a class of preference functions that is tail
-realistic for real networks. These analytical results enable the
-likelihood of the degree distribution to be expressed in terms of the
-parameters of \(b(\cdot)\), which in turn allows the underlying
-mechanism of the network, assumed to grow according to PA, to be
-inferred directly.
+While Voitalov et al.~(2019) have pioneered using methods in extreme
+value theory to analyse degree distributions, they overlooked the
+inapplicability of standard tools for continuous extremes, due to the
+discrete nature of degrees. A key difficulty arises because many
+standard discrete distributions, such as the Poisson, geometric, or
+negative binomial, do not satisfy the conditions required to belong to a
+maximum domain of attraction (MDA). Shimura (2012) showed that for a
+discrete distribution \(F\) must be long-tailed in order to be in an
+MDA, and introduced the notion of recoverability to the MDA, also
+referred to as the discrete MDA (Hitz et al., 2024). Using the
+theoretical guarantees of regular variation for a discrete distribution
+by Shimura (2012), we demonstrate how the tail of the degree
+distribution is directly influenced by \(b(\cdot)\), and subsequently
+proposed a class of preference functions that not only imply regularly
+varying degree distributions but are also tail realistic for real
+networks. These analytical results enable the likelihood of the degree
+distribution to be expressed in terms of the parameters of \(b(\cdot)\),
+which in turn allows the underlying mechanism of the network, assumed to
+grow according to PA, to be inferred directly.
 
@@
 existing node \(x\) in \(\Upsilon(t)\) gives birth to a child with rate
-\(b(\mathrm{deg}(x, \Upsilon(t))\) independently of the other nodes
+\(b(\mathrm{deg}(x, \Upsilon(t)))\) independently of the other nodes
 where \(\mathrm{deg}(x, \Upsilon(t))\) is the degree of node \(x\) in
@@
 We investigate how the tail behaviour of the discrete limiting degree
-distribution is influenced by the preference function \(b(·)\). Since
-the limiting degree distribution is discrete, standard tools for
-analysing extremes of continuous distributions are not directly
-applicable. A key difficulty arises because many standard discrete
-distributions, such as the Poisson, Geometric, or Negative Binomial, do
-not satisfy the conditions required to belong to an MDA.
-\citet{shimura12} showed that for a discrete random variable X to lie in
-an MDA, its distribution must be long-tailed and introduced the notion
-of recoverability to the MDA, sometimes referred to as the discrete MDA
-(see e.g. \citet{hitz24}). In particular, if a distribution is both
-long-tailed and recoverable, then it can be placed in the appropriate
-MDA.
-
-A central tool in this characterisation is the quantity \(\Omega(F,k)\).
-For a discrete distribution \(F\) with survival function \(\bar F\) and
+distribution is influenced by the preference function \(b(·)\), using
+the characterisation by \citet{shimura12}.
+
+A central tool is the quantity \(\Omega(F,k)\). For a discrete
+distribution \(F\) with survival function \(\bar F\) and
 \(k\in\mathbb Z^+\) define:
@@
 \(b(k)=k+\varepsilon\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\),
-giving a tail index of \(1/\lambda^*\), in which cases the Malthusian
-parameter \(\lambda^*\) is 2.
+giving a tail index of \(\lambda^*=2\)
 

### 2025-09-27 Files/fix_theorems.py

@@ -1,0 +1,41 @@
+import sys
+
+# Get filename from command line
+texfile = sys.argv[1]
+
+# Read the original file lines
+with open(texfile, "r", encoding="utf-8") as f:
+    lines = f.readlines()
+
+# Prepare the text to prepend/replace
+text_to = (
+    "\\usepackage{amsthm}\n"
+    "\\theoremstyle{plain}\n"
+    "\\newtheorem{proposition}{Proposition}[section]\n"
+    "\\theoremstyle{plain}\n"
+    "\\newtheorem{corollary}[proposition]{Corollary}\n"
+    "\\theoremstyle{remark}\n"
+    "\\AtBeginDocument{\\renewcommand*{\\proofname}{Proof}}\n"
+    "\\newtheorem{remark}[proposition]{Remark}\n"
+    "\\newtheorem*{solution}{Solution}\n"
+    "\\newtheorem{refremark}[proposition]{Remark}\n"
+    "\\newtheorem{refsolution}{Solution}[section]\n"
+)
+
+pattern = "newtheorem"
+fixed_lines = []
+
+# Remove lines containing the pattern
+for line in lines:
+    if pattern not in line:
+        fixed_lines.append(line)
+
+# Join lines into a single string
+result = "".join(fixed_lines)
+
+# Replace the first occurrence of \usepackage{amsthm} with text_to
+result = result.replace(r"\usepackage{amsthm}", text_to, 1)
+
+# Write the fixed content to a new file
+with open("fixed_" + texfile, "w", encoding="utf-8") as f:
+    f.write(result)

### 2025-09-27 Files/render.py

@@ -1,0 +1,46 @@
+import subprocess
+import os
+import sys
+
+def run_command(cmd, work_dir):
+    """
+    Run a command with live output. Stops if the command fails.
+    """
+    print(f"\nRunning: {' '.join(cmd)}\n{'='*50}")
+    process = subprocess.Popen(
+        cmd,
+        cwd=work_dir,
+        stdout=subprocess.PIPE,
+        stderr=subprocess.STDOUT,
+        text=True
+    )
+
+    # Print output line by line
+    for line in process.stdout:
+        print(line, end="")
+
+    process.wait()
+    if process.returncode != 0:
+        raise RuntimeError(f"Command failed: {' '.join(cmd)}")
+
+def build_paper():
+    # Use the directory where this script is located
+    work_dir = os.path.dirname(os.path.abspath(__file__))
+
+    # List of commands
+    commands = [
+    ["quarto", "render", "paper.qmd"],  # stop at .tex
+    ["python", "fix_theorems.py", "paper.tex"],
+    ["xelatex", "-interaction=nonstopmode", "fixed_paper.tex"]
+]
+
+    try:
+        for cmd in commands:
+            run_command(cmd, work_dir)
+        print("\n✅ Build finished successfully!")
+    except RuntimeError as e:
+        print(f"\n❌ Build failed:\n{e}")
+        sys.exit(1)
+
+if __name__ == "__main__":
+    build_paper()

### 2025-09-28 Files/paper.qmd
Adding MSC codes

@@
   nature-pdf:
+    indent: true
     # journal.cite-style is included in the tex file but ignored by pandoc if 
@@
       - ref: 1
-  - name: Vianey Palacios Ramirez
+  - name: Vianey Palacios Ramírez
     affiliations:
@@
 Results of applications to real data are promising and comparable to alternatives.
+keywords: extremes, discrete extremes, power law, preferential attachment
+msc: 
+ - 05C82
+ - 60G70
 bibliography: refs.bib  
@@
 
-Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process. See Appendix [-@sec-proofs] for the details of the proof.
+Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process.See Appendix [-@sec-proofs] for the details of the proof.
 :::
@@
 If $\lim_{k\rightarrow\infty}b(k)<\infty$, then $\Omega(F, k) = 0$.
+Furthermore, $F$ is recovered to the Gumbel MDA but does not belong to the Gumbel MDA itself as it is not long-tailed.
+See Appendix [-@sec-proofs] for the proof.
 :::
@@
 
--   $\bar F(k)$ is light-tailed if and only if $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$.
+-   $\bar F(k)$ is light-tailed if and only if $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$. Further, $F$ is in the Gumbel MDA as it is also long-tailed.
 :::
@@
 \end{cases}
-$$ {#eq-pref} for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$. By @cor-omega2, the resulting degree distribution is regularly varying with tail index $\lambda^*/\beta$, since
-
-$$
-\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}.
-$$
+$$ {#eq-pref} for $\alpha,\beta, \varepsilon>0$ and $k_0\in\mathbb N$.
+By @cor-omega2, the resulting degree distribution is regularly varying with tail index $\lambda^*/\beta$, since $\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}$.
 
@@
 #| fig-width: 6
-#| fig-height: 6
+#| fig-asp: 0.7
 #| fig-cap: Example construction of the preference function in @eq-pref
 #| label: fig-ex
+library(ggbrace)
 x = seq(0,10,l=100)
@@
   geom_text(aes(x=k0, y=0, label='k[0]'), parse = TRUE, size=textsize)+
-  geom_text(aes(x=0.25, y=eps, label='epsilon'), parse=TRUE, size=textsize)+
   geom_text(aes(x=6.3, y=4.3, label='beta'), parse=TRUE, size=textsize)+
+  stat_brace(aes(x=c(0, 0), y=c(0, eps)), rotate=90, width=0.2)+
+  geom_text(aes(x=0.3, y=eps/2, label='epsilon'), parse=TRUE, size=textsize)+
   theme_bw() + xlab('Degree')+ylab('Preference') + ylim(0,6)+
@@
 
-# A PA Model with Flexible Regular Variation {#sec-model}
+# A Tail-realistic GPA model {#sec-model}
 
-In the previous section, we found that using an asymptotically linear preference function allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing regular variation of the degrees.
+In the previous section, we constructed an asymptotically linear preference function that allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing regular variation of the degrees.
 In this section, we demonstrate the flexibility of the preference function in @eq-pref, with regard to the tail behaviour of the limiting degree distribution.
-Using @eq-surv-origin, the limiting survival is
+The limiting survival if the proposed GPA model is:
 
@@
 
-with $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where
+with Malthusian parameter $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where
 
@@
 
-The analytic form of the survival function in ([-@eq-polysurv]), offers a natural connection to the discrete version of the generalised Pareto (GP) distribution , providing a link to a well established component of discrete extremes in the literature.
+An interesting property is that the analytic form of the survival function in ([-@eq-polysurv]), offers a natural connection to the discrete version of the generalised Pareto (GP) distribution , providing a link to a well established component of discrete extremes in the literature.
 Specifically, it is connected to the Integer GP (IGP) distribution seen in @Rohrbeck_2018 with conditional survival: $$
@@
 
-By @eq-polysurv and using Stirling's approximation:
+This connection is made explicitly below, let $\bar F (k|k\ge k_0)$ be the conditional limiting survival of the proposed GPA model, then:
 
@@
 
-Through the connection implied by @eq-igp-est, fitting the proposed model is almost equivalent to fitting the IGP distribution to the degrees and estimating its parameters.
-However, instead of only describing the shape of the degree distribution, we would also gain estimates for the shape of the preference function, thus gaining a direct understanding the mechanisms underlying the growth of the network.
+The connection implied by @eq-igp-est is interesting because it shows that fitting the proposed GPA model is almost equivalent to fitting the IGP distribution to the degrees and estimating its parameters, something that has already been done by @Lee24 with promising results.
+However, instead of only describing the shape of the degree distribution, we would learn about the shape of the preference function, thus gaining a direct understanding the mechanisms underlying the growth of the network.
 
@@
 
-To perform inference of the model parameters, we consider a network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$, where $M$ is the maximum degree.
+To learn about the growth mechanisms, we estimate the parameters of the proposed preference function.
+In order to do this we first define the likelihood.
+Consider a network with degree count vector $\pmb n = (n_0, n_1, \ldots, n_M)$, where $M$ is the maximum degree and $n_i$ being the number of nodes in the network with degree $i$.
 Using @eq-pmf-origin, the likelihood is:
@@
 This will allow the model to be fitted whilst ignoring the influence of the lower degrees (those less than $l$) as the model does not capture the behaviour well at the lower degrees, since @rudas07 only provides results for the case of a preferential attachment tree.
+
+In order to carry out the Bayesian inference of the parameters we must first assign priors, for the parameters defined on the positive real line ($\alpha, \beta,\varepsilon$) we use diffuse gamma priors and for $k_0$ we use a discrete uniform prior over a large and reasonable range of values.
 
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="90%"}
-#| fig-width: 8
-#| fig-height: 8
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+#| fig-width: 6
+#| fig-height: 6
 #| fig-cap: Empirical (black dots) and fitted (red line) survival functions for data simulated from the proposed model with various combinations of $(\alpha,\beta,\epsilon)$ and $k_0=20$.
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
 #| fig-width: 10
-#| fig-height: 10
-#| fig-cap: Empirical (black dots) and posterior medians (solid red) of the fitted survival function for several real data sets and their 95% credible intervals (dotted red).
+#| fig-asp: 0.7
+#| fig-cap: Empirical (black dots) and posterior medians of the fitted survival function of the proposed GPA model (solid red) for several real data sets and their 95% credible intervals (dotted red).
+Also shown are the posterior medians (solid blue) and 95% credible intervals (dotted blue) for a model from @Lee24.
+The corresponding thresholds are shown with short vertical lines towards the top of each plot.
 #| label: fig-real1
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="100%"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
 #| fig-width: 10
-#| fig-height: 10
+#| fig-asp: 0.7
 #| fig-cap: Posterior estimates of shape parameter of Zipf-IGP distribution (right) and the analagous quantity obtained from fitting the proposed model (left).

### 2025-09-28 Files/paper.tex
Adding MSC codes

@@
 }{}
-\makeatletter
-\@ifundefined{KOMAClassName}{% if non-KOMA class
-  \IfFileExists{parskip.sty}{%
-    \usepackage{parskip}
-  }{% else
-    \setlength{\parindent}{0pt}
-    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
-}{% if KOMA class
-  \KOMAoptions{parskip=half}}
-\makeatother
 % Make \paragraph and \subparagraph free-standing
@@
   pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
-  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
+  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
+  pdfkeywords={extremes, discrete extremes, power law, preferential
+attachment},
   colorlinks=true,
@@
 % author setup
-\author*[1]{\fnm{Thomas William} \sur{Boughen}}\email{t.w.boughen1@newcastle.ac.uk}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
+\author*[1]{\fnm{Thomas William} \sur{Boughen}}\email{t.w.boughen1@newcastle.ac.uk}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramírez}}
 % affil setup
@@
 % keywords
+\keywords{extremes, discrete extremes, power law, preferential
+attachment}
+
+\pacs[MSC 2020]{05C82, 60G70}
 
@@
 Here, \(\lambda^*\) is the Malthusian parameter of the corresponding
-branching process. See Appendix \ref{sec-proofs} for the details of the
+branching process.See Appendix \ref{sec-proofs} for the details of the
 proof.
@@
 If \(\lim_{k\rightarrow\infty}b(k)<\infty\), then \(\Omega(F, k) = 0\).
+Furthermore, \(F\) is recovered to the Gumbel MDA but does not belong to
+the Gumbel MDA itself as it is not long-tailed. See Appendix
+\ref{sec-proofs} for the proof.
 
@@
   \(\bar F(k)\) is light-tailed if and only if
-  \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\).
+  \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\). Further, \(F\) is in the
+  Gumbel MDA as it is also long-tailed.
 \end{itemize}
@@
 since
-
-\[
-\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}.
-\]
+\(\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}\).
 
@@
 
-\section{A PA Model with Flexible Regular Variation}\label{sec-model}
-
-In the previous section, we found that using an asymptotically linear
-preference function allows for the inclusion of sub/super-linear
+\section{A Tail-realistic GPA model}\label{sec-model}
+
+In the previous section, we constructed an asymptotically linear
+preference function that allows for the inclusion of sub/super-linear
 behaviour below the threshold, while simultaneously guaranteeing regular
@@
 flexibility of the preference function in Equation~\ref{eq-pref}, with
-regard to the tail behaviour of the limiting degree distribution. Using
-Equation~\ref{eq-surv-origin}, the limiting survival is
+regard to the tail behaviour of the limiting degree distribution. The
+limiting survival if the proposed GPA model is:
 
@@
 
-with \(\lambda^*\) satisfying \(\hat \rho(\lambda^*)=1\) where
+with Malthusian parameter \(\lambda^*\) satisfying
+\(\hat \rho(\lambda^*)=1\) where
 
@@
 
-The analytic form of the survival function in (\ref{eq-polysurv}),
-offers a natural connection to the discrete version of the generalised
-Pareto (GP) distribution , providing a link to a well established
-component of discrete extremes in the literature. Specifically, it is
-connected to the Integer GP (IGP) distribution seen in
-\citet{Rohrbeck_2018} with conditional survival: \[
+An interesting property is that the analytic form of the survival
+function in (\ref{eq-polysurv}), offers a natural connection to the
+discrete version of the generalised Pareto (GP) distribution , providing
+a link to a well established component of discrete extremes in the
+literature. Specifically, it is connected to the Integer GP (IGP)
+distribution seen in \citet{Rohrbeck_2018} with conditional survival: \[
 \Pr(X> x|X> v) = \left(\frac{\xi(x-v)}{\sigma} + 1\right)^{-1/\xi},\qquad x=v+1,v+2,\ldots
@@
 
-By Equation~\ref{eq-polysurv} and using Stirling's approximation:
+This connection is made explicitly below, let \(\bar F (k|k\ge k_0)\) be
+the conditional limiting survival of the proposed GPA model, then:
 
@@
 
-Through the connection implied by Equation~\ref{eq-igp-est}, fitting the
-proposed model is almost equivalent to fitting the IGP distribution to
-the degrees and estimating its parameters. However, instead of only
-describing the shape of the degree distribution, we would also gain
-estimates for the shape of the preference function, thus gaining a
-direct understanding the mechanisms underlying the growth of the
-network.
+The connection implied by Equation~\ref{eq-igp-est} is interesting
+because it shows that fitting the proposed GPA model is almost
+equivalent to fitting the IGP distribution to the degrees and estimating
+its parameters, something that has already been done by \citet{Lee24}
+with promising results. However, instead of only describing the shape of
+the degree distribution, we would learn about the shape of the
+preference function, thus gaining a direct understanding the mechanisms
+underlying the growth of the network.
 
@@
 
-To perform inference of the model parameters, we consider a network with
-degree count vector \(\pmb n = (n_0, n_1, \ldots, n_M)\), where \(M\) is
-the maximum degree. Using Equation~\ref{eq-pmf-origin}, the likelihood
-is:
+To learn about the growth mechanisms, we estimate the parameters of the
+proposed preference function. In order to do this we first define the
+likelihood. Consider a network with degree count vector
+\(\pmb n = (n_0, n_1, \ldots, n_M)\), where \(M\) is the maximum degree
+and \(n_i\) being the number of nodes in the network with degree \(i\).
+Using Equation~\ref{eq-pmf-origin}, the likelihood is:
 
@@
 
+In order to carry out the Bayesian inference of the parameters we must
+first assign priors, for the parameters defined on the positive real
+line (\(\alpha, \beta,\varepsilon\)) we use diffuse gamma priors and for
+\(k_0\) we use a discrete uniform prior over a large and reasonable
+range of values.
+
 \section{Applications}\label{applications}
@@
 
-\includegraphics[width=0.9\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
 
@@
 \caption{\label{fig-real1}Empirical (black dots) and posterior medians
-(solid red) of the fitted survival function for several real data sets
-and their 95\% credible intervals (dotted red).}
+of the fitted survival function of the proposed GPA model (solid red)
+for several real data sets and their 95\% credible intervals (dotted
+red). Also shown are the posterior medians (solid blue) and 95\%
+credible intervals (dotted blue) for a model from \citet{Lee24}. The
+corresponding thresholds are shown with short vertical lines towards the
+top of each plot.}
 
@@
 
-\includegraphics[width=1\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
+\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
 

### 2025-09-28 Files/scripts/data_fit_background.R

@@ -73,8 +73,8 @@ saveRDS(fits, 'modelfits.rds')
 
 
 # saveRDS(fits,'modelfits.rds')
-mixfits = readRDS('mixfits.rds')
-dat_list = readRDS('dat_list.rds')
+mixfits = readRDS('../results/mixfits.rds')
+dat_list = readRDS('../results/dat_list.rds')
 real_survs = ggpubr::ggarrange(plotlist = plots)
 real_survs
 saveRDS(real_survs,'real_survs.rds')
@@ -130,7 +130,7 @@ mixplt = readRDS('mix_survs.rds')
 
 comp_plot = ggpubr::ggarrange(mixplt, real_survs,labels=c('Zipf-IGP', 'GPA'), ncol=1)
 
-saveRDS(comp_plot, 'comp_plot.rds')
+saveRDS(comp_plot, '../results/comp_plot.rds')
 fits = readRDS('../results/modelfits.rds')
 plots = list()
 for(i in 1:length(fits)){
@@ -140,12 +140,12 @@ for(i in 1:length(fits)){
   ptrunc = sdf$surv[sdf$degree==min(fits[[i]]$dat[,1])]
   ptrunc_mix = sdf$surv[sdf$degree==min(mixfits[[i]]$data[,1])]
   plots[[i]] = ggplot() + scale_x_log10(limits = c(1, max(dat_list[[i]][,1]))) + scale_y_log10(limits=c(1/sum(dat_list[[i]][,2]),1)) +
-    geom_segment(aes(x = jitter(!!fits[[i]]$smps$k0[seq(1,length(fits[[i]]$smps$k0),by=20)]),yend=0.5,y=1, colour='GPA'), alpha=0.1)+
+    geom_segment(aes(x = jitter(!!fits[[i]]$smps$k0[seq(1,length(fits[[i]]$smps$k0),by=20)]),yend=0.5,y=1, colour='Proposed GPA'), alpha=0.1)+
     geom_segment(data=NULL,aes(x = jitter(mixfits[[i]]$pars$u[seq(1,length(mixfits[[i]]$pars$u),by=20)]),yend=0.5,y=1, colour='Zipf-IGP'), alpha=0.5)+
     geom_point(data = twbfn::deg_surv(counts_to_degs(dat_list[[i]])),aes(x=degree, y=surv))+
[…]

### 2025-09-28 Files/paper.qmd
Fixing plots

@@
   stat_brace(aes(x=c(0, 0), y=c(0, eps)), rotate=90, width=0.2)+
-  geom_text(aes(x=0.3, y=eps/2, label='epsilon'), parse=TRUE, size=textsize)+
+  geom_text(aes(x=0.5, y=eps/2, label='epsilon'), parse=TRUE, size=textsize)+
   theme_bw() + xlab('Degree')+ylab('Preference') + ylim(0,6)+

### 2025-09-29 Files/paper.qmd
Formatting closer to Extremes articles

@@
   nature-pdf:
+    fontsize: 10pt
     indent: true
@@
     keep-tex: true
-    equal-margins: true
     number-sections: true
@@
 Results of applications to real data are promising and comparable to alternatives.
-keywords: extremes, discrete extremes, power law, preferential attachment
+keywords: Extremes · Discrete extremes · Power law · Preferential attachment
 msc: 
@@
 @sec-conc provides a discussion of this paper and possible avenues for future work.
-
-\newpage
 
@@
 
-```{r, echo=FALSE, out.width="60%", warning=FALSE, message=FALSE, cache=TRUE}
+```{r, echo=FALSE, out.width="119mm", warning=FALSE, message=FALSE, cache=TRUE}
 #| fig-width: 6
@@
 In the previous section, we constructed an asymptotically linear preference function that allows for the inclusion of sub/super-linear behaviour below the threshold, while simultaneously guaranteeing regular variation of the degrees.
-In this section, we demonstrate the flexibility of the preference function in @eq-pref, with regard to the tail behaviour of the limiting degree distribution.
-The limiting survival if the proposed GPA model is:
+In this section, we demonstrate the flexibility of the proposed GPA model, that is using the preference function from @eq-pref in the GPA model, with regard to the tail behaviour of the limiting degree distribution.
+The limiting survival of the proposed GPA model is:
 
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
 #| fig-width: 8
@@
 
-An interesting property is that the analytic form of the survival function in ([-@eq-polysurv]), offers a natural connection to the discrete version of the generalised Pareto (GP) distribution , providing a link to a well established component of discrete extremes in the literature.
+An interesting property is that the analytic form of the survival function in ([-@eq-polysurv]), offers a natural connection to the discrete version of the generalised Pareto (GP) distribution, providing a link to a well established component of discrete extremes in the literature.
 Specifically, it is connected to the Integer GP (IGP) distribution seen in @Rohrbeck_2018 with conditional survival: $$
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
 #| fig-width: 8
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
 #| fig-width: 6
@@
 Figures [-@fig-rec1] and [-@fig-rec2] demonstrate the usefulness of the model, as we can recover the model parameters well from only the final degree distribution of the simulated network.
-This indicates that the method may also be applied to real networks, with the assumption that they evolved according to the GPA scheme.
+This indicates that the method may also be applied to real networks, with the assumption that they evolved according to the proposed GPA scheme.
 
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
 #| fig-width: 10
@@
 Also shown are the posterior medians (solid blue) and 95% credible intervals (dotted blue) for a model from @Lee24.
-The corresponding thresholds are shown with short vertical lines towards the top of each plot.
+The corresponding $k_0$ values are shown with short vertical lines towards the top of each plot.
 #| label: fig-real1
@@
 
-@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the GPA model and the Zipf-IGP mixture model from @Lee24.
-In most cases, the GPA model gives a similar fit to the Zipf-IGP model but where the GPA model fits well we gain additional information about the preference function, assuming that the network evolved according the the GPA scheme.
+@fig-real1 displays the posterior estimates of the survival function for various data sets, obtained from fitting the proposed GPA model and the Zipf-IGP mixture model from @Lee24.
+In most cases, the proposed GPA model gives a similar fit to the Zipf-IGP model but where the proposed GPA model fits well we gain additional information about the preference function, assuming that the network evolved according the the proposed GPA scheme.
 
-@fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the GPA model.
-Generally, the GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution.
-In the cases of substantial discrepancies, it is either because the GPA model fits the tail better than the Zipf-IGP model does, or because of the threshold being too low, forcing almost all of the data to be modelled by the linear part of the GPA.
+@fig-shapes shows the posterior of the shape parameter $\xi$ obtained from the Zipf-IGP model alongside the posterior of the equivalent shape parameter $\beta/\lambda^*$ obtained from fitting the proposed GPA model.
+Generally, the proposed GPA model performs similarly to the Zipf-IGP when estimating the tail behaviour of the degree distribution.
+In the cases of substantial discrepancies, it is either because the proposed GPA model fits the tail better than the Zipf-IGP model does, or because of the threshold being too low, forcing almost all of the data to be modelled by the linear part of the GPA.
 This again shows the effects that small degrees have on this model.
 
-![Posterior densities of the parameters of the proposed model fitted to real data.](images/pars_plot.jpg){#fig-parplot width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
+#| fig-width: 10
+#| fig-asp: 0.7
+#| fig-cap: Posterior densities of the parameters of the proposed model fitted to real data.
+#| label: fig-parplot
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+plt3 = readRDS('results/pars_plot.rds')
+
+plot(plt3)
+
+```
+
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
 #| fig-width: 10
@@
 #| label: fig-shapes
+#| fig-pos: 'H'
 
@@
 
-```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="80%"}
+```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
 #| fig-width: 9
-#| fig-height: 9
+#| fig-asp: 0.7
 #| fig-cap: Posterior median for the preference function (solid) with 95% credible interval (dashed) on log-log scale.
 #| label: fig-pa
+#| fig-pos: 'H'
 
@@
 \begin{align*}
-\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
-&=\left(\log\frac{\lambda+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda+b(k+1)}{b(k+1)}\right)^{-1}\\
-&=\left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}.
+\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda^{*}+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda^{*}+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda^{*}+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda^{*}+b(i)}}\right)^{-1}\\
+&=\left(\log\frac{\lambda^{*}+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda^{*}+b(k+1)}{b(k+1)}\right)^{-1}\\
+&=\left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}.
 \end{align*}
 
-Clearly if $\lim_{k\rightarrow\infty}b(k)=c$ for some $c>0$ then $\Omega(F,k)= 0 (= \lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}} [b(k+2)-b(k+1)])$.
-Now consider a non-constant $b(k)$ and re-write $\Omega(F,k)$ as:
+As it is assumed throughout that $b(k)$ is non-decreasing, we consider the two cases where $b(k)$ is bounded and unbounded.
+When $b(k)$ is bounded, by the monotone convergence theorem, $\lim_{k\rightarrow\infty}b(k) = c$ for some $c>0$, which implies that $\lim_{k\rightarrow\infty}\Omega(F,k)=\lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}}\left[b(k+2)-b(k+1)\right] = 0$.
+This proves the first part of @rem-omega.
+Now, using @eq-surv-origin,  $\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}\rightarrow\frac{c}{\lambda^{*}+c}<1$ as $k\rightarrow\infty$.
+This means that $F$ is not long-tailed, and therefore is only recovered to the Gumbel MDA.
+
+Now we consider the case where $b(k)$ is unbounded, re-write $\Omega(F,k)$ as:
 
 \begin{align*}
-\Omega(F,k) &= \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}+\frac{b(k+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}\\&+\frac{b(k+1)}{\lambda}  -\frac{b(k+1)}{\lambda}\\
-&=\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\&+\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}.
+\Omega(F,k) &= \left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda^{*}}+\frac{b(k+2)}{\lambda^{*}}\\&\qquad  -\left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}+\frac{b(k+1)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}\\
+&=\left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda^{*}}\right\}\\ &\qquad  - \left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda^{*}}\right\}\\&\qquad  +\frac{b(k+2)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}.
 \end{align*}
 
-Then if $\lim_{k\rightarrow\infty}b(k)=\infty$ it follows that:
+Then as $\lim_{k\rightarrow\infty}b(k)=\infty$ it follows that:
 
 \begin{align*}
-\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} \\&- \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\
-&\qquad+\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
-&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
-&=\frac{1}{\lambda}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad \square
+\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda^{*}}\right\} \\&\qquad - \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda^{*}}\right\}\\
+&\qquad +\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}\right)\\
+&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}\right)\\
+&=\frac{1}{\lambda^{*}}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad  \square
 \end{align*}
+
+The first two lines both have the limit $1/2$ as $\lim_{z\rightarrow{}0}\left[\log(1+z)\right]^{-1}-z^{-1} = 1/2$.
+
+Now, using @eq-surv-origin, $\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}=\frac{1}{\lambda^{*}/b(k+1) + 1}\rightarrow{}1$ as $k\rightarrow\infty$.
+This means that $F$ is long-tailed, and therefore is in an MDA.
 
@@
 \begin{align*}
-\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
-&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}.
+\hat\rho(\lambda) &= C_{g,k_0} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
+&= C_{g,k_0} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
+&=C_{g,k_0} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)},
 \end{align*}
+
+where $C_{g,k_0} = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)}$.
 
@@
 &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
-&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\,at\\
+&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\qquad at\\
 &=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
@@
 \begin{align*}
-\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}.\qquad \square
+\hat\rho(\lambda) &= C_{g,k_0} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\\&\qquad \times\left(\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-\frac{\Gamma\left(\frac{g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}\right)\\
+&=C_{g,k_0} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\Gamma\left(\frac{g(k_0)+\lambda}{\beta}\right)}{\left(\frac{\lambda}{\beta}-1\right)\Gamma\left(\frac{g(k_0)+\lambda}{\beta}-1\right)}-1\right)\\
+&=C_{g,k_0} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{\frac{g(k_0)+\lambda}{\beta}-1}{\frac{\lambda}{\beta}-1}-1\right)\\
+&=C_{g,k_0} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\left(\frac{g(k_0)+\lambda-\beta}{\lambda-\beta}-1\right)\\&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{g(k_0)}{\lambda-\beta}\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}.\qquad  \square
 \end{align*}

### 2025-09-29 Files/paper.tex
Formatting closer to Extremes articles

@@
   sn-basic,
+  10pt,
 ]{sn-jnl}
-
-\setlength\evensidemargin{\oddsidemargin}
 
@@
 \usepackage{listings}%
+\usepackage{setspace}%
 
@@
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
-  pdfkeywords={extremes, discrete extremes, power law, preferential
+  pdfkeywords={Extremes · Discrete extremes · Power law · Preferential
 attachment},
@@
 % abstract 
-
 \abstract{Identifying the generating mechanism of a network is
@@
 % keywords
-\keywords{extremes, discrete extremes, power law, preferential
+\keywords{Extremes · Discrete extremes · Power law · Preferential
 attachment}
 
-\pacs[MSC 2020]{05C82, 60G70}
+\pacs[\sffamily MSC 2020]{05C82· 60G70}
 
@@
 
-\newpage
-
 \section{Tail Behaviour of Preferential Attachment
@@
 
-\includegraphics[width=0.6\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-ex-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-ex-1.pdf}
 
@@
 variation of the degrees. In this section, we demonstrate the
-flexibility of the preference function in Equation~\ref{eq-pref}, with
-regard to the tail behaviour of the limiting degree distribution. The
-limiting survival if the proposed GPA model is:
+flexibility of the proposed GPA model, that is using the preference
+function from Equation~\ref{eq-pref} in the GPA model, with regard to
+the tail behaviour of the limiting degree distribution. The limiting
+survival of the proposed GPA model is:
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polylinsurv-1.pdf}
 
@@
 function in (\ref{eq-polysurv}), offers a natural connection to the
-discrete version of the generalised Pareto (GP) distribution , providing
+discrete version of the generalised Pareto (GP) distribution, providing
 a link to a well established component of discrete extremes in the
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-approx_surv-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-polyheat-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-rec1-1.pdf}
 
@@
 the method may also be applied to real networks, with the assumption
-that they evolved according to the GPA scheme.
+that they evolved according to the proposed GPA scheme.
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-real1-1.pdf}
 
@@
 credible intervals (dotted blue) for a model from \citet{Lee24}. The
-corresponding thresholds are shown with short vertical lines towards the
-top of each plot.}
+corresponding \(k_0\) values are shown with short vertical lines towards
+the top of each plot.}
 
@@
 Figure~\ref{fig-real1} displays the posterior estimates of the survival
-function for various data sets, obtained from fitting the GPA model and
-the Zipf-IGP mixture model from \citet{Lee24}. In most cases, the GPA
-model gives a similar fit to the Zipf-IGP model but where the GPA model
-fits well we gain additional information about the preference function,
-assuming that the network evolved according the the GPA scheme.
+function for various data sets, obtained from fitting the proposed GPA
+model and the Zipf-IGP mixture model from \citet{Lee24}. In most cases,
+the proposed GPA model gives a similar fit to the Zipf-IGP model but
+where the proposed GPA model fits well we gain additional information
+about the preference function, assuming that the network evolved
+according the the proposed GPA scheme.
 
@@
 equivalent shape parameter \(\beta/\lambda^*\) obtained from fitting the
-GPA model. Generally, the GPA model performs similarly to the Zipf-IGP
-when estimating the tail behaviour of the degree distribution. In the
-cases of substantial discrepancies, it is either because the GPA model
-fits the tail better than the Zipf-IGP model does, or because of the
-threshold being too low, forcing almost all of the data to be modelled
-by the linear part of the GPA. This again shows the effects that small
-degrees have on this model.
+proposed GPA model. Generally, the proposed GPA model performs similarly
+to the Zipf-IGP when estimating the tail behaviour of the degree
+distribution. In the cases of substantial discrepancies, it is either
+because the proposed GPA model fits the tail better than the Zipf-IGP
+model does, or because of the threshold being too low, forcing almost
+all of the data to be modelled by the linear part of the GPA. This again
+shows the effects that small degrees have on this model.
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{images/pars_plot.jpg}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-parplot-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-shapes-1.pdf}
 
@@
 
-\includegraphics[width=0.8\linewidth,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pa-1.pdf}
+\includegraphics[width=119mm,height=\textheight,keepaspectratio]{paper_files/figure-pdf/fig-pa-1.pdf}
 
@@
 \begin{align*}
-\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda+b(i)}}\right)^{-1}\\
-&=\left(\log\frac{\lambda+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda+b(k+1)}{b(k+1)}\right)^{-1}\\
-&=\left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}.
+\Omega(F,k)&=\left(\log\frac{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda^{*}+b(i)}}{\prod_{i=0}^{k+2}\frac{b(i)}{\lambda^{*}+b(i)}}\right)^{-1}-\left(\log\frac{\prod_{i=0}^{k}\frac{b(i)}{\lambda^{*}+b(i)}}{\prod_{i=0}^{k+1}\frac{b(i)}{\lambda^{*}+b(i)}}\right)^{-1}\\
+&=\left(\log\frac{\lambda^{*}+b(k+2)}{b(k+2)}\right)^{-1}-\left(\log\frac{\lambda^{*}+b(k+1)}{b(k+1)}\right)^{-1}\\
+&=\left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}.
 \end{align*}
 
-Clearly if \(\lim_{k\rightarrow\infty}b(k)=c\) for some \(c>0\) then
-\(\Omega(F,k)= 0 (= \lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}} [b(k+2)-b(k+1)])\).
-Now consider a non-constant \(b(k)\) and re-write \(\Omega(F,k)\) as:
+As it is assumed throughout that \(b(k)\) is non-decreasing, we consider
+the two cases where \(b(k)\) is bounded and unbounded. When \(b(k)\) is
+bounded, by the monotone convergence theorem,
+\(\lim_{k\rightarrow\infty}b(k) = c\) for some \(c>0\), which implies
+that
+\(\lim_{k\rightarrow\infty}\Omega(F,k)=\lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}}\left[b(k+2)-b(k+1)\right] = 0\).
+This proves the first part of Remark~\ref{rem-omega}. Now, using
+Equation~\ref{eq-surv-origin},
+\(\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}\rightarrow\frac{c}{\lambda^{*}+c}<1\)
+as \(k\rightarrow\infty\). This means that \(F\) is not long-tailed, and
+therefore is only recovered to the Gumbel MDA.
+
+Now we consider the case where \(b(k)\) is unbounded, re-write
+\(\Omega(F,k)\) as:
 
 \begin{align*}
-\Omega(F,k) &= \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}+\frac{b(k+2)}{\lambda}-\left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}\\&+\frac{b(k+1)}{\lambda}  -\frac{b(k+1)}{\lambda}\\
-&=\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} - \left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\&+\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}.
+\Omega(F,k) &= \left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda^{*}}+\frac{b(k+2)}{\lambda^{*}}\\&\qquad  -\left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}+\frac{b(k+1)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}\\
+&=\left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda^{*}}\right\}\\ &\qquad  - \left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda^{*}}\right\}\\&\qquad  +\frac{b(k+2)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}.
 \end{align*}
 
-Then if \(\lim_{k\rightarrow\infty}b(k)=\infty\) it follows that:
+Then as \(\lim_{k\rightarrow\infty}b(k)=\infty\) it follows that:
 
 \begin{align*}
-\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda}\right\} \\&- \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda}\right\}\\
-&\qquad+\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
-&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda}-\frac{b(k+1)}{\lambda}\right)\\
-&=\frac{1}{\lambda}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad \square
+\lim_{k\rightarrow\infty}\Omega(F,k) &= \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+2)}\right]\right)^{-1}-\frac{b(k+2)}{\lambda^{*}}\right\} \\&\qquad - \lim_{k\rightarrow\infty}\left\{ \left(\log\left[1+\frac{\lambda^{*}}{b(k+1)}\right]\right)^{-1}-\frac{b(k+1)}{\lambda^{*}}\right\}\\
+&\qquad +\lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}\right)\\
+&=\frac{1}{2}-\frac{1}{2} + \lim_{k\rightarrow\infty}\left(\frac{b(k+2)}{\lambda^{*}}-\frac{b(k+1)}{\lambda^{*}}\right)\\
+&=\frac{1}{\lambda^{*}}\lim_{k\rightarrow\infty}\left[b(k+2)-b(k+1)\right].\qquad  \square
 \end{align*}
+
+The first two lines both have the limit \(1/2\) as
+\(\lim_{z\rightarrow{}0}\left[\log(1+z)\right]^{-1}-z^{-1} = 1/2\).
+
+Now, using Equation~\ref{eq-surv-origin},
+\(\frac{\bar{F}(k+1)}{\bar{F}(k)}=\frac{b(k+1)}{\lambda^{*}+b(k+1)}=\frac{1}{\lambda^{*}/b(k+1) + 1}\rightarrow{}1\)
+as \(k\rightarrow\infty\). This means that \(F\) is long-tailed, and
+therefore is in an MDA.
 
@@
 \begin{align*}
-\hat\rho(\lambda) &= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
-&= \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
-&=\sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)}.
+\hat\rho(\lambda) &= C_{g,k_0} + \left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)\Gamma\left(\frac{g(k_0)}{\beta}\right)}\\
+&= C_{g,k_0} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=k_0+1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n-k_0\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n-k_0\right)}\\
+&=C_{g,k_0} + \frac{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}\right)}{\Gamma\left(\frac{g(k_0)}{\beta}\right)}\left(\prod_{i=0}^{k_0-1}\frac{g(i)}{\lambda+g(i)}\right)\sum_{n=1}^\infty\frac{\Gamma\left(\frac{g(k_0)}{\beta}+n\right)}{\Gamma\left(\frac{\lambda+g(k_0)}{\beta}+n\right)},
 \end{align*}
+
+where
+\(C_{g,k_0} = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{g(i)}{\lambda+g(i)}\).
 

### 2025-09-29 Files/fix_theorems.py

@@ -10,9 +10,9 @@ with open(texfile, "r", encoding="utf-8") as f:
 # Prepare the text to prepend/replace
 text_to = (
     "\\usepackage{amsthm}\n"
-    "\\theoremstyle{plain}\n"
+    "\\theoremstyle{thmstyleone}\n"
     "\\newtheorem{proposition}{Proposition}[section]\n"
-    "\\theoremstyle{plain}\n"
+    "\\theoremstyle{thmstyleone}\n"
     "\\newtheorem{corollary}[proposition]{Corollary}\n"
     "\\theoremstyle{remark}\n"
     "\\AtBeginDocument{\\renewcommand*{\\proofname}{Proof}}\n"

### 2025-09-29 Files/render.py

@@ -31,6 +31,8 @@ def build_paper():
     commands = [
     ["quarto", "render", "paper.qmd"],  # stop at .tex
     ["python", "fix_theorems.py", "paper.tex"],
+    ["xelatex","-interaction=nonstopmode", "fixed_paper.tex"],
+    ["bibtex", "fixed_paper"],
     ["xelatex", "-interaction=nonstopmode", "fixed_paper.tex"]
 ]
 

### 2025-09-30 Files/paper.qmd
Fixing references

@@
 Denote the density of the point process associated with the pure birth process corresponding to the growth of an individual node by $\rho(t)$, and its Laplace transform by $\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t$.
-Next, denote the Malthusian parameter of this process by $\lambda^*$, that is $\lambda^*$ satisfies $\hat\rho(\lambda^*) = 1$.
+We denote the Malthusian parameter of this process by $\lambda^*$, that is $\lambda^*$ satisfies $\hat\rho(\lambda^*) = 1$.
 
@@
 In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$.
-Conversely, it also agrees with the limiting behavior of BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+\varepsilon$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$
+Conversely, it also agrees with the limiting behaviour of BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+\varepsilon$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$
 

### 2025-09-30 Files/paper.tex
Fixing references

@@
 \(\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t\).
-Next, denote the Malthusian parameter of this process by \(\lambda^*\),
+We denote the Malthusian parameter of this process by \(\lambda^*\),
 that is \(\lambda^*\) satisfies \(\hat\rho(\lambda^*) = 1\).
@@
 \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\). Conversely, it also agrees
-with the limiting behavior of BA model, which produces a regularly
+with the limiting behaviour of BA model, which produces a regularly
 degree distribution with tail index 2; for the preference function

### 2025-09-30 Files/paper.qmd
Finalising edits

@@
 
-The popularity of using the power law for network degrees can be traced back to the preferential attachment (PA) model popularised by @Barabasi99.
-In the general model, as new nodes join the network, an existing node with degree $k$ gains edges at a rate proportional to $b(k)$, where $b(\cdot)$ is a non-negative preference function.
-@Barabasi99 showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 2.
+```{=tex}
+The popularity of using the power law for network degrees can be traced back to a preferential attachment (PA) model popularised by \cite{Barabasi99}.
+In the general model, general preferential attachment (GPA), as new nodes join the network, an existing node with degree $k$ gains edges at a rate proportional to $b(k)$, where $b(\cdot)$ is a non-negative preference function.
+\cite{Barabasi99} showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 2 — this model is commonly known as the Barab\'asi-Albert (BA) model.
 Subsequently, if a real network is shown to be scale-free, one can loosely justify PA as the underlying mechanism of its growth.
+```
 
@@
 
-This paper aims to address the gap between the applied and theoretical works, by asking if a network is assumed to come from a PA model, can we use the degree distribution alone to directly infer the parameters of the the preference function and learn about the growth mechanisms?
+This paper aims to address the gap between the applied and theoretical works, by asking if a network is assumed to come from a GPA model, can we use the degree distribution alone to directly infer the parameters of the the preference function and learn about the growth mechanisms?
 Moreover, proper consideration is given to the tail of the degree distribution, because otherwise the effects of the largest degrees, which correspond to the most influential nodes, deviating from the power law will be discounted.
 
-While Voitalov et al.
-(2019) have pioneered using methods in extreme value theory to analyse degree distributions, they overlooked the inapplicability of standard tools for continuous extremes, due to the discrete nature of degrees.
+While @Voitalov_2019 have pioneered using methods in extreme value theory to analyse degree distributions, they overlooked the inapplicability of standard tools for continuous extremes, due to the discrete nature of degrees.
 A key difficulty arises because many standard discrete distributions, such as the Poisson, geometric, or negative binomial, do not satisfy the conditions required to belong to a maximum domain of attraction (MDA).
-Shimura (2012) showed that for a discrete distribution $F$ must be long-tailed in order to be in an MDA, and introduced the notion of recoverability to the MDA, also referred to as the discrete MDA (Hitz et al., 2024).
-Using the theoretical guarantees of regular variation for a discrete distribution by Shimura (2012), we demonstrate how the tail of the degree distribution is directly influenced by $b(\cdot)$, and subsequently proposed a class of preference functions that not only imply regularly varying degree distributions but are also tail realistic for real networks.
-These analytical results enable the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to PA, to be inferred directly.
+@shimura12 showed that a discrete distribution $F$ must be long-tailed in order to be in an MDA, and introduced the notion of recoverability to the MDA, also referred to as the discrete MDA [@hitz24].
+Using the theoretical guarantees of regular variation for a discrete distribution by @shimura12, we demonstrate how the tail of the degree distribution is directly influenced by $b(\cdot)$, and subsequently propose a class of preference functions that not only imply regularly varying degree distributions but are also tail realistic for real networks.
+These analytical results enable the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to GPA, to be inferred directly.
 
-The remainder of the paper is as follows: @sec-tail gives a detailed description of the PA model alongside the theoretical results for the survival function of the limiting degree distribution, with a focus on the tail behaviour in terms of the preference function $b(\cdot)$.
+The remainder of the paper is as follows: @sec-tail gives a detailed description of the GPA model alongside the theoretical results for the survival function of the limiting degree distribution, with a focus on the tail behaviour in terms of the preference function $b(\cdot)$.
 A class of asymptotically linear preference functions will be introduced and shown to guarantee regular variation in the degree distribution while remaining flexible up until a threshold.
 @sec-model utilises the proposed preference function and illustrates numerically how the tail index of the degree distribution varies with the model parameters.
-The simulation study in Subsection -@sec-sim demonstrates that the parameters can be recovered from fitting the model to only the degree distribution.
-Subsection -@sec-real fits the model to some real data and provides posterior estimates for the preference function.
+The simulation study in @sec-sim demonstrates that the parameters can be recovered from fitting the model to only the degree distribution.
+@sec-real fits the model to some real data and provides posterior estimates for the preference function.
 @sec-conc provides a discussion of this paper and possible avenues for future work.
 
-# Tail Behaviour of Preferential Attachment Model {#sec-tail}
+# Tail Behaviour of GPA Model {#sec-tail}
 
@@
 
-Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges from the new vertex; the target for each of these edges are selected from the vertices already in the network with weights proportional to some non-decreasing preference function $b(\cdot)$ of their degree, where $b: \mathbb N \mapsto \mathbb R^+\setminus\{0\}$ is such that:
+Starting at time $t=0$ with an initial network of $m$ vertices that each have no edges, at times $t=1,2,\ldots$ a new vertex is added to the network bringing with it $m$ directed edges from the new vertex; the targets for each of these edges are selected from the vertices already in the network with weights proportional to some non-decreasing preference function $b(\cdot)$ of their degree, where $b: \mathbb N \mapsto \mathbb R^+\setminus\{0\}$ is such that:
 
@@
 ```{=tex}
-Special cases of this model include the Barab\'asi-Albert (BA) model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
+Special cases of this model include the BA model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
 ```
@@
 
-Now, let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(t)=\{\emptyset\}$ and $\Upsilon(t)=G$ where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t)))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
+Now, let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(0)=\{\emptyset\}$ and $\Upsilon(t)$ is a graph where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t)))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
 
@@
 $$
-\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda \int_0^\infty e^{-\lambda t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
+\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
 $$ {#eq-survlim}
@@
 
-@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Frechet MDA.
+@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Fréchet MDA.
 If instead, $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is recoverable to the Gumbel MDA; in this cases we say that the distribution is light-tailed.
@@
 
-Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process.See Appendix [-@sec-proofs] for the details of the proof.
+Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process. See Appendix [-@sec-proofs] for the details of the proof.
 :::
@@
 In particular it is consistent with @krapivsky01, who showed that a sub-linear preference function leads to a light-tailed distributions, since for $b(k)=k^\alpha$ with $\alpha<1$, we have $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0$.
-Conversely, it also agrees with the limiting behaviour of BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+\varepsilon$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$
+Conversely, it also agrees with the limiting behaviour of the BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+1$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$.
 
@@
 
-which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$.
+which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$. See Appendix [-@sec-proofs] for the derivations of @eq-rho.
 
@@
 #| fig-height: 4
-#| fig-cap: The limiting survival, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed PA model.
+#| fig-cap: The limiting survival, according to various combinations of $(\alpha, \beta, \varepsilon)$ and $k_0=20$ of the proposed GPA model.
 #| label: fig-polylinsurv
@@
 
-\begin{align*}
+$$
+\begin{aligned}
 L(\pmb n | \pmb \theta,l) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
-\end{align*}\label{eq-lh}
-
+\end{aligned}
+$${#eq-llh}
 where $\text{B}(\cdot,\cdot)$ is the the beta function, $\pmb \theta = (\alpha, \varepsilon, k_0,\beta)$, and $l\ge0$ is a quantity that allows truncating the data such that the minimum degree is $l$.
@@
 
-This subsection aims to show that the parameters of the model (and therefore the preference function) in @sec-model can be recovered from simulating a network from the model, and fitting it to the observed degree distribution, using the likelihood in \eqref{eq-lh}.
+This subsection aims to show that the parameters of the proposed model (and therefore the preference function) in @sec-model can be recovered from simulating a network from the model, and fitting it to the observed degree distribution, using the likelihood in @eq-llh.
 
@@
 
-where Gamma($a$, $b$) is the gamma distribution with shape $a$ and rate $b$, and U($a$, $b$) is uniform distribution with lower and upper bounds $a$ and $b$, to obtain a posterior distribution, up to the proportionality constant.
+where Gamma($a$, $b$) is the gamma distribution with shape $a$ and rate $b$, and U($a$, $b$) is discrete uniform distribution with lower and upper bounds $a$ and $b$, to obtain a posterior distribution, up to the proportionality constant.
 Posterior samples can then be obtained by an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC) algorithm.
@@
 \sum_{n=0}^\infty\frac{\Gamma(n+x)}{\Gamma(n+x+y)} &=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty \text{B}(n+x,y)\\
-&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}at\\
-&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n\qquad at\\
-&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}at\\
+&=\frac{1}{\Gamma(y)}\sum_{n=0}^\infty\int_0^1t^{n+x-1}(1-t)^{y-1}\text{d}t\\
+&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\sum_{n=0}^\infty t^n \text{d}t\\
+&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}\text{d}t\\
 &=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
-&=\frac{1}{\Gamma(y)}\text{y}(x,y-1)\\
+&=\frac{1}{\Gamma(y)}\text{B}(x,y-1)\\
 &= \frac{\Gamma(x)}{(y-1)\Gamma(x+y-1)}.

### 2025-09-30 Files/paper.tex
Finalising edits

@@
 
-The popularity of using the power law for network degrees can be traced
-back to the preferential attachment (PA) model popularised by
-\citet{Barabasi99}. In the general model, as new nodes join the network,
-an existing node with degree \(k\) gains edges at a rate proportional to
-\(b(k)\), where \(b(\cdot)\) is a non-negative preference function.
-\citet{Barabasi99} showed that, when \(b(k) = k + 1\), in the limit the
-resulting degree distribution is regularly varying with index 2.
-Subsequently, if a real network is shown to be scale-free, one can
-loosely justify PA as the underlying mechanism of its growth.
+The popularity of using the power law for network degrees can be traced back to a preferential attachment (PA) model popularised by \cite{Barabasi99}.
+In the general model, general preferential attachment (GPA), as new nodes join the network, an existing node with degree $k$ gains edges at a rate proportional to $b(k)$, where $b(\cdot)$ is a non-negative preference function.
+\cite{Barabasi99} showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 2 — this model is commonly known as the Barab\'asi-Albert (BA) model.
+Subsequently, if a real network is shown to be scale-free, one can loosely justify PA as the underlying mechanism of its growth.
 
@@
 This paper aims to address the gap between the applied and theoretical
-works, by asking if a network is assumed to come from a PA model, can we
-use the degree distribution alone to directly infer the parameters of
+works, by asking if a network is assumed to come from a GPA model, can
+we use the degree distribution alone to directly infer the parameters of
 the the preference function and learn about the growth mechanisms?
@@
 
-While Voitalov et al.~(2019) have pioneered using methods in extreme
+While \citet{Voitalov_2019} have pioneered using methods in extreme
 value theory to analyse degree distributions, they overlooked the
@@
 negative binomial, do not satisfy the conditions required to belong to a
-maximum domain of attraction (MDA). Shimura (2012) showed that for a
+maximum domain of attraction (MDA). \citet{shimura12} showed that a
 discrete distribution \(F\) must be long-tailed in order to be in an
 MDA, and introduced the notion of recoverability to the MDA, also
-referred to as the discrete MDA (Hitz et al., 2024). Using the
-theoretical guarantees of regular variation for a discrete distribution
-by Shimura (2012), we demonstrate how the tail of the degree
+referred to as the discrete MDA \citep{hitz24}. Using the theoretical
+guarantees of regular variation for a discrete distribution by
+\citet{shimura12}, we demonstrate how the tail of the degree
 distribution is directly influenced by \(b(\cdot)\), and subsequently
-proposed a class of preference functions that not only imply regularly
+propose a class of preference functions that not only imply regularly
 varying degree distributions but are also tail realistic for real
@@
 which in turn allows the underlying mechanism of the network, assumed to
-grow according to PA, to be inferred directly.
+grow according to GPA, to be inferred directly.
 
 The remainder of the paper is as follows: Section~\ref{sec-tail} gives a
-detailed description of the PA model alongside the theoretical results
+detailed description of the GPA model alongside the theoretical results
 for the survival function of the limiting degree distribution, with a
@@
 illustrates numerically how the tail index of the degree distribution
-varies with the model parameters. The simulation study in Subsection
-\ref{sec-sim} demonstrates that the parameters can be recovered from
-fitting the model to only the degree distribution. Subsection
-\ref{sec-real} fits the model to some real data and provides posterior
-estimates for the preference function. Section~\ref{sec-conc} provides a
-discussion of this paper and possible avenues for future work.
-
-\section{Tail Behaviour of Preferential Attachment
-Model}\label{sec-tail}
+varies with the model parameters. The simulation study in
+Section~\ref{sec-sim} demonstrates that the parameters can be recovered
+from fitting the model to only the degree distribution.
+Section~\ref{sec-real} fits the model to some real data and provides
+posterior estimates for the preference function. Section~\ref{sec-conc}
+provides a discussion of this paper and possible avenues for future
+work.
+
+\section{Tail Behaviour of GPA Model}\label{sec-tail}
 
@@
 the network bringing with it \(m\) directed edges from the new vertex;
-the target for each of these edges are selected from the vertices
+the targets for each of these edges are selected from the vertices
 already in the network with weights proportional to some non-decreasing
@@
 
-Special cases of this model include the Barab\'asi-Albert (BA) model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
+Special cases of this model include the BA model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
 
@@
 Now, let \(\Upsilon(t)\) be the tree determined by \(\zeta(t)\) as
-follows: \(\Upsilon(t)=\{\emptyset\}\) and \(\Upsilon(t)=G\) where each
-existing node \(x\) in \(\Upsilon(t)\) gives birth to a child with rate
-\(b(\mathrm{deg}(x, \Upsilon(t)))\) independently of the other nodes
-where \(\mathrm{deg}(x, \Upsilon(t))\) is the degree of node \(x\) in
-the tree \(\Upsilon(t)\) at time \(t\), denote by
+follows: \(\Upsilon(0)=\{\emptyset\}\) and \(\Upsilon(t)\) is a graph
+where each existing node \(x\) in \(\Upsilon(t)\) gives birth to a child
+with rate \(b(\mathrm{deg}(x, \Upsilon(t)))\) independently of the other
+nodes where \(\mathrm{deg}(x, \Upsilon(t))\) is the degree of node \(x\)
+in the tree \(\Upsilon(t)\) at time \(t\), denote by
 \(\Upsilon(t)_{\downarrow x}\) the tree when treating node \(x\) as the
@@
 \begin{equation}\phantomsection\label{eq-survlim}{
-\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda \int_0^\infty e^{-\lambda t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
+\lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
 }\end{equation}
@@
 then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\), and
-hence belongs to the Frechet MDA. If instead,
+hence belongs to the Fréchet MDA. If instead,
 \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then the distribution is
@@
 Here, \(\lambda^*\) is the Malthusian parameter of the corresponding
-branching process.See Appendix \ref{sec-proofs} for the details of the
+branching process. See Appendix \ref{sec-proofs} for the details of the
 proof.
@@
 \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=0\). Conversely, it also agrees
-with the limiting behaviour of BA model, which produces a regularly
+with the limiting behaviour of the BA model, which produces a regularly
 degree distribution with tail index 2; for the preference function
-\(b(k)=k+\varepsilon\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\),
-giving a tail index of \(\lambda^*=2\)
+\(b(k)=k+1\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\), giving a
+tail index of \(\lambda^*=2\).
 
@@
 which has to be solved numerically for most parameter choices. Also,
-note that \(\lambda>\beta\).
+note that \(\lambda>\beta\). See Appendix \ref{sec-proofs} for the
+derivations of Equation~\ref{eq-rho}.
 
@@
 various combinations of \((\alpha, \beta, \varepsilon)\) and \(k_0=20\)
-of the proposed PA model.}
+of the proposed GPA model.}
 
@@
 
-\begin{align*}
+\begin{equation}\phantomsection\label{eq-llh}{
+\begin{aligned}
 L(\pmb n | \pmb \theta,l) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
-\end{align*}\label{eq-lh}
-
-where \(\text{B}(\cdot,\cdot)\) is the the beta function,
-\(\pmb \theta = (\alpha, \varepsilon, k_0,\beta)\), and \(l\ge0\) is a
-quantity that allows truncating the data such that the minimum degree is
-\(l\). This will allow the model to be fitted whilst ignoring the
-influence of the lower degrees (those less than \(l\)) as the model does
-not capture the behaviour well at the lower degrees, since
-\citet{rudas07} only provides results for the case of a preferential
-attachment tree.
+\end{aligned}
+}\end{equation} where \(\text{B}(\cdot,\cdot)\) is the the beta
+function, \(\pmb \theta = (\alpha, \varepsilon, k_0,\beta)\), and
+\(l\ge0\) is a quantity that allows truncating the data such that the
+minimum degree is \(l\). This will allow the model to be fitted whilst
+ignoring the influence of the lower degrees (those less than \(l\)) as
+the model does not capture the behaviour well at the lower degrees,
+since \citet{rudas07} only provides results for the case of a
+preferential attachment tree.
 
@@
 
-This subsection aims to show that the parameters of the model (and
-therefore the preference function) in Section~\ref{sec-model} can be
-recovered from simulating a network from the model, and fitting it to
-the observed degree distribution, using the likelihood in \eqref{eq-lh}.
+This subsection aims to show that the parameters of the proposed model
+(and therefore the preference function) in Section~\ref{sec-model} can
+be recovered from simulating a network from the model, and fitting it to
+the observed degree distribution, using the likelihood in
+Equation~\ref{eq-llh}.
 
@@
 where Gamma(\(a\), \(b\)) is the gamma distribution with shape \(a\) and
-rate \(b\), and U(\(a\), \(b\)) is uniform distribution with lower and
-upper bounds \(a\) and \(b\), to obtain a posterior distribution, up to
-the proportionality constant. Posterior samples can then be obtained by
-an adaptive Metropolis-Hastings Markov chain Monte Carlo (MCMC)
-algorithm. For these simulated networks \(l=0\).
+rate \(b\), and U(\(a\), \(b\)) is discrete uniform distribution with
+lower and upper bounds \(a\) and \(b\), to obtain a posterior
+distribution, up to the proportionality constant. Posterior samples can
+then be obtained by an adaptive Metropolis-Hastings Markov chain Monte
+Carlo (MCMC) algorithm. For these simulated networks \(l=0\).
 

### 2025-09-30 Files/render.py

@@ -12,6 +12,7 @@ def run_command(cmd, work_dir):
         cwd=work_dir,
         stdout=subprocess.PIPE,
         stderr=subprocess.STDOUT,
+        encoding="utf-8",
         text=True
     )
 

### 2025-10-03 Files/paper.qmd
Update .gitignore

@@
 
-```{r, echo=FALSE, warning=FALSE, cache=TRUE, out.width="119mm"}
+```{r, echo=FALSE, warning=FALSE, cache=FALSE, out.width="119mm"}
 #| fig-width: 8
@@
     S(pars$k0[i], polylin(pars$a[i], pars$eps[i]),pars$lambda[i], pars$k0[i],pars$b[i])
-  gp_surv[i] = mev::pgp(pars$x[i],pars$k0[i]-1,((pars$k0[i]+1)^pars$a[i] +pars$eps[i]-1 )/pars$lambda[i],pars$b[i]/pars$lambda[i],lower.tail = F)
+  gp_surv[i] = mev::pgp(pars$x[i],pars$k0[i],((pars$k0[i])^pars$a[i] +pars$eps[i]-1 +pars$b[i])/pars$lambda[i],pars$b[i]/pars$lambda[i],lower.tail = F)
 }
@@
 ggplot(data = pars) + geom_line(aes(x=(x),y=surv, linetype=as.character(b),colour =as.character(b)), lwd=1) +
-  geom_line(data = pars[pars$x>=pars$k0,],aes(x=(x+1),y=gp_surv, linetype=as.character(b)),colour ='grey')+
+  geom_line(data = pars[pars$x>=pars$k0,],aes(x=(x),y=gp_surv, linetype=as.character(b)),colour ='grey')+
   scale_x_log10()+

### 2025-10-03 Files/paper.qmd
Continued changes

@@
 Amid the recent rise of interest in networks, there has been a debate on whether most real networks are scale-free.
-Claiming a real network is scale-free is equivalent to saying that its degree distribution follows a power law, that is the fraction of nodes with degree $k$ is proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with tail index $\alpha$.
+Claiming a real network is scale-free is equivalent to saying that the tail of the degree distribution follows a power law, that is the fraction of nodes with degree greater than $k$ is proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with tail index $\alpha$.
 On the side against the claim is @Broido_2019 who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority in real networks.
-On the other side of this debate is @Voitalov_2019 who disagrees and claims that these networks are not nearly as rare and only appear to be so as a result of an unrealistic expectation of a power law without deviations or noise.
+@Voitalov_2019 disagree and claim that these networks are not nearly as rare and only appear so as a result of an unrealistic expectation of a power law without deviations or noise.
 Additional evidence of these deviations from a power law is shown by @Lee24 who demonstrate that a lot of networks are partially scale-free, in that the body of the degree distribution is often modelled well by a power law, while the tail is lighter than what is implied by the body, albeit still regularly varying.
-Nevertheless, most studies into the appropriateness of a power law for the degrees of real networks, the aforementioned references included, have been largely descriptive in the sense that no information about the growth of the networks is revealed.
 
@@
 ```
+
+Nevertheless, most studies into the appropriateness of a power law for the degrees of real networks, the aforementioned references included, have been largely descriptive in the sense that no information about the growth of the networks is revealed.
 
@@
 
-While @Voitalov_2019 have pioneered using methods in extreme value theory to analyse degree distributions, they overlooked the inapplicability of standard tools for continuous extremes, due to the discrete nature of degrees.
+While @Voitalov_2019 have pioneered using methods in extreme value theory to analyse degree distributions, while their approach is good they overlooked the inapplicability of standard tools for continuous extremes, due to the discrete nature of degrees.
 A key difficulty arises because many standard discrete distributions, such as the Poisson, geometric, or negative binomial, do not satisfy the conditions required to belong to a maximum domain of attraction (MDA).
-@shimura12 showed that a discrete distribution $F$ must be long-tailed in order to be in an MDA, and introduced the notion of recoverability to the MDA, also referred to as the discrete MDA [@hitz24].
-Using the theoretical guarantees of regular variation for a discrete distribution by @shimura12, we demonstrate how the tail of the degree distribution is directly influenced by $b(\cdot)$, and subsequently propose a class of preference functions that not only imply regularly varying degree distributions but are also tail realistic for real networks.
+@shimura12 showed that a discrete distribution $F$ must be long-tailed,that is $\lim_{x\rightarrow\infty}\bar F(x+1)/\bar F(x) = 1$, in order to be in an MDA, and introduced the notion of recoverability to the MDA, also sometimes referred to as the discrete MDA [@hitz24].
+Using the theoretical guarantees of being in a discrete MDA for a discrete distribution by @shimura12, we demonstrate how the tail of the degree distribution is directly influenced by $b(\cdot)$, and subsequently propose a class of preference functions that not only imply regularly varying degree distributions but are also tail realistic for real networks.
 These analytical results enable the likelihood of the degree distribution to be expressed in terms of the parameters of $b(\cdot)$, which in turn allows the underlying mechanism of the network, assumed to grow according to GPA, to be inferred directly.
@@
 @sec-model utilises the proposed preference function and illustrates numerically how the tail index of the degree distribution varies with the model parameters.
-The simulation study in @sec-sim demonstrates that the parameters can be recovered from fitting the model to only the degree distribution.
+The simulation study in @sec-sim demonstrates that the parameters can be recovered from fitting the model to only the degree distribution and as a consequence recover the underlying growth mechanism.
 @sec-real fits the model to some real data and provides posterior estimates for the preference function.
@@
 ```{=tex}
-Special cases of this model include the BA model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
+Special cases of this model include the BA model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution with a tail that is not regularly varying.
 ```
@@
 
-Now, let $\Upsilon(t)$ be the tree determined by $\zeta(t)$ as follows: $\Upsilon(0)=\{\emptyset\}$ and $\Upsilon(t)$ is a graph where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t)))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
+Now, we construct the tree $\Upsilon(t)$ determined by $\zeta(t)$ as follows: $\Upsilon(0)=\{\emptyset\}$ and $\Upsilon(t)$ is a graph where each existing node $x$ in $\Upsilon(t)$ gives birth to a child with rate $b(\mathrm{deg}(x, \Upsilon(t)))$ independently of the other nodes where $\mathrm{deg}(x, \Upsilon(t))$ is the degree of node $x$ in the tree $\Upsilon(t)$ at time $t$, denote by $\Upsilon(t)_{\downarrow x}$ the tree when treating node $x$ as the root.
 
-Theorem 1 from @rudas07 states that for the tree $\Upsilon(t)$ at time $t$ and a characteristic function of the tree $\varphi(\cdot)$ :
+The limiting survival can be viewed as the limit over time of the empirical proportion of vertices with degree over a threshold $k\in\mathbb N$, that is:
+
+$$
+\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1},
+$$
+
+For the tree $\Upsilon(t)$ defined as above, Theorem 1 from @rudas07 states that
 
@@
 $$ {#eq-survlim}
+for some characteristic function of the tree $\varphi(\cdot)$.
 
-The limiting survival can be viewed as the limit of the empirical proportion of vertices with degree over a threshold $k\in\mathbb N$, that is:
-
-$$
-\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1},
-$$ which can also be written using @eq-survlim as:
+Using this result, and the characteristic function $\varphi(\Upsilon(t)) = \mathbb I\{\text{deg}(x,\Upsilon(t))>k\}$, we can write the limiting survival as:
 
@@
 
-We investigate how the tail behaviour of the discrete limiting degree distribution is influenced by the preference function $b(·)$, using the characterisation by @shimura12.
+We now investigate how the tail behaviour of the discrete limiting degree distribution is influenced by the preference function $b(·)$, using the characterisation by @shimura12.
 
-A central tool is the quantity $\Omega(F,k)$. For a discrete distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ define:
+A central tool of this characterisation is the quantity $\Omega(F,k)$. For a discrete distribution $F$ with survival function $\bar F$ and $k\in\mathbb Z^+$ define:
 
@@
 
-@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Fréchet MDA.
+@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Fréchet MDA with tail index $\alpha$.
 If instead, $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is recoverable to the Gumbel MDA; in this cases we say that the distribution is light-tailed.
@@
 ::: {#prp-omega}
-If $\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}$ with $b(k) \rightarrow \infty$ as $k\rightarrow \infty$, then $$
+Let $\bar F (k)$ be the limiting survival function of degrees in a GPA network with preference function $b$ such that $b(k)\rightarrow\infty$. Then:
+$$
 \lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
@@
 b(k) = \begin{cases}
-k^\alpha + \varepsilon,&k<k_0,\\
-k_0^\alpha + \varepsilon + \beta(k-k_0), &k\ge k_0
+k^\alpha + \varepsilon,&k\le k_0,\\
+k_0^\alpha + \varepsilon + \beta(k-k_0), &k> k_0
 \end{cases}
@@
 By @cor-omega2, the resulting degree distribution is regularly varying with tail index $\lambda^*/\beta$, since $\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}$.
+
+Shown in @fig-ex is an example of one such preference function of this form, the value of $\varepsilon$ controls the baseline appeal of a node with zero degree and $k_0$ controls the point at which the behaviour changes, below this point the curvature of the function is controlled by the value of $\alpha$ and above $k_0$ the gradient of the linear section is contolled by $\beta$.
 
@@
 \bar F(k) = \begin{cases}
-\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
-\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k\ge k_0,
+\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k\le k_0,\\
+\left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right)\frac{\Gamma(\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)},&k > k_0,
 \end{cases}
@@
 \begin{align*}
-\bar F(k|k\ge k_0) &= \frac{\Gamma\left(\frac{\lambda^* + k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(\frac{k_0^\alpha + \varepsilon}{\beta}\right)}\times\frac{\Gamma\left(k-k_0  +1 + \frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0  +1 + \frac{\lambda^*+ k_0^\alpha + \varepsilon}{\beta}\right)}\\
-&\approx\left(\frac{k_0^\alpha+\varepsilon}{\beta}\right)^{\lambda^*/\beta}\left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
-&=\left(\frac{k_0^\alpha+\varepsilon}{k_0^\alpha+\varepsilon + \beta}\right)^{\lambda^*/\beta}\left(\frac{\beta(k-k_0)}{\beta + k_0^\alpha+\varepsilon} + 1\right)^{-\lambda^*/\beta}\\
-&=\left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^{*}/\beta}.
+\bar F(k|k>k_0) &= \frac{\bar F(k)}{\bar F(k_0)}\\
+&= \frac{\lambda^* + k_0^\alpha + \varepsilon}{k_0^\alpha + \varepsilon}\times \frac{\Gamma((\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left((k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)}\\
+&=\frac{\Gamma(1 + (\lambda^*+k_0^\alpha + \varepsilon)/\beta)}{\Gamma\left(1+(k_0^\alpha + \varepsilon)/\beta\right)} \frac{\Gamma\left(k-k_0 + 1 +\frac{k_0^\alpha + \varepsilon}{\beta}\right)}{\Gamma\left(k-k_0 + 1 +\frac{\lambda^* +k_0^\alpha + \varepsilon}{\beta}\right)}\\
+&\approx \left(1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{\lambda^*/\beta} \left(k-k_0+1+\frac{k_0^\alpha + \varepsilon}{\beta}\right)^{-\lambda^*/\beta}\\
+&=\left(\frac{\beta(k-k_0)}{k_0^{\alpha}+\varepsilon+\beta} + 1\right)^{-\lambda^{*}/\beta}
 \end{align*}
@@
 \begin{cases}
-=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
-\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
+=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k\le k_0,\\
+\approx \left(\prod_{i=0}^{k_0}\frac{i^\alpha + \varepsilon}{\lambda^* + i^\alpha + \varepsilon}\right)\left(\frac{\beta(k-k_0)}{k_0^{\alpha}+\varepsilon+\beta} + 1\right)^{-\lambda^{*}/\beta},&k > k_0,
 \end{cases}
-$$ {#eq-igp-est} meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)$ distribution.
+$$ {#eq-igp-est} meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon+\beta}{\lambda^*},k_0-1\right)$ distribution.
 
@@
 The connection implied by @eq-igp-est is interesting because it shows that fitting the proposed GPA model is almost equivalent to fitting the IGP distribution to the degrees and estimating its parameters, something that has already been done by @Lee24 with promising results.
-However, instead of only describing the shape of the degree distribution, we would learn about the shape of the preference function, thus gaining a direct understanding the mechanisms underlying the growth of the network.
+However, instead of only describing the shape of the degree distribution, we would learn about the shape of the preference function, thus gaining a direct understanding of the mechanisms underlying the growth of the network.
 
@@
 
-## Proof of @prp-omega
+## Preliminary
 
@@
 
+## Proof of @rem-omega
+
 As it is assumed throughout that $b(k)$ is non-decreasing, we consider the two cases where $b(k)$ is bounded and unbounded.
-When $b(k)$ is bounded, by the monotone convergence theorem, $\lim_{k\rightarrow\infty}b(k) = c$ for some $c>0$, which implies that $\lim_{k\rightarrow\infty}\Omega(F,k)=\lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}}\left[b(k+2)-b(k+1)\right] = 0$.
+When $b(k)$ is bounded, by the monotone convergence theorem, $\lim_{k\rightarrow\infty}b(k) = c$ for some $c>0$, which implies that $\lim_{k\rightarrow\infty}\Omega(F,k)= 0$.
 This proves the first part of @rem-omega.
@@
 This means that $F$ is not long-tailed, and therefore is only recovered to the Gumbel MDA.
+
+## Proof of @prp-omega
 

### 2025-10-03 Files/paper.tex
Continued changes

@@
 whether most real networks are scale-free. Claiming a real network is
-scale-free is equivalent to saying that its degree distribution follows
-a power law, that is the fraction of nodes with degree \(k\) is
-proportional to \(k^{-\alpha}\), and therefore has a regularly varying
-tail with tail index \(\alpha\). On the side against the claim is
-\citet{Broido_2019} who compared the fits of a power law model against
-that of several non-scale-free models to nearly a thousand networks,
-only to find strong evidence for scale-freeness in four percent and weak
-evidence in over half of the networks, thus claiming that scale-free
-networks do not make up a majority in real networks. On the other side
-of this debate is \citet{Voitalov_2019} who disagrees and claims that
-these networks are not nearly as rare and only appear to be so as a
-result of an unrealistic expectation of a power law without deviations
-or noise. Additional evidence of these deviations from a power law is
-shown by \citet{Lee24} who demonstrate that a lot of networks are
-partially scale-free, in that the body of the degree distribution is
-often modelled well by a power law, while the tail is lighter than what
-is implied by the body, albeit still regularly varying. Nevertheless,
-most studies into the appropriateness of a power law for the degrees of
-real networks, the aforementioned references included, have been largely
-descriptive in the sense that no information about the growth of the
-networks is revealed.
+scale-free is equivalent to saying that the tail of the degree
+distribution follows a power law, that is the fraction of nodes with
+degree greater than \(k\) is proportional to \(k^{-\alpha}\), and
+therefore has a regularly varying tail with tail index \(\alpha\). On
+the side against the claim is \citet{Broido_2019} who compared the fits
+of a power law model against that of several non-scale-free models to
+nearly a thousand networks, only to find strong evidence for
+scale-freeness in four percent and weak evidence in over half of the
+networks, thus claiming that scale-free networks do not make up a
+majority in real networks. \citet{Voitalov_2019} disagree and claim that
+these networks are not nearly as rare and only appear so as a result of
+an unrealistic expectation of a power law without deviations or noise.
+Additional evidence of these deviations from a power law is shown by
+\citet{Lee24} who demonstrate that a lot of networks are partially
+scale-free, in that the body of the degree distribution is often
+modelled well by a power law, while the tail is lighter than what is
+implied by the body, albeit still regularly varying.
 
@@
 Subsequently, if a real network is shown to be scale-free, one can loosely justify PA as the underlying mechanism of its growth.
+
+Nevertheless, most studies into the appropriateness of a power law for
+the degrees of real networks, the aforementioned references included,
+have been largely descriptive in the sense that no information about the
+growth of the networks is revealed.
 
@@
 While \citet{Voitalov_2019} have pioneered using methods in extreme
-value theory to analyse degree distributions, they overlooked the
-inapplicability of standard tools for continuous extremes, due to the
-discrete nature of degrees. A key difficulty arises because many
-standard discrete distributions, such as the Poisson, geometric, or
-negative binomial, do not satisfy the conditions required to belong to a
-maximum domain of attraction (MDA). \citet{shimura12} showed that a
-discrete distribution \(F\) must be long-tailed in order to be in an
-MDA, and introduced the notion of recoverability to the MDA, also
-referred to as the discrete MDA \citep{hitz24}. Using the theoretical
-guarantees of regular variation for a discrete distribution by
-\citet{shimura12}, we demonstrate how the tail of the degree
-distribution is directly influenced by \(b(\cdot)\), and subsequently
-propose a class of preference functions that not only imply regularly
-varying degree distributions but are also tail realistic for real
-networks. These analytical results enable the likelihood of the degree
-distribution to be expressed in terms of the parameters of \(b(\cdot)\),
-which in turn allows the underlying mechanism of the network, assumed to
-grow according to GPA, to be inferred directly.
+value theory to analyse degree distributions, while their approach is
+good they overlooked the inapplicability of standard tools for
+continuous extremes, due to the discrete nature of degrees. A key
+difficulty arises because many standard discrete distributions, such as
+the Poisson, geometric, or negative binomial, do not satisfy the
+conditions required to belong to a maximum domain of attraction (MDA).
+\citet{shimura12} showed that a discrete distribution \(F\) must be
+long-tailed,that is
+\(\lim_{x\rightarrow\infty}\bar F(x+1)/\bar F(x) = 1\), in order to be
+in an MDA, and introduced the notion of recoverability to the MDA, also
+sometimes referred to as the discrete MDA \citep{hitz24}. Using the
+theoretical guarantees of being in a discrete MDA for a discrete
+distribution by \citet{shimura12}, we demonstrate how the tail of the
+degree distribution is directly influenced by \(b(\cdot)\), and
+subsequently propose a class of preference functions that not only imply
+regularly varying degree distributions but are also tail realistic for
+real networks. These analytical results enable the likelihood of the
+degree distribution to be expressed in terms of the parameters of
+\(b(\cdot)\), which in turn allows the underlying mechanism of the
+network, assumed to grow according to GPA, to be inferred directly.
 
@@
 Section~\ref{sec-sim} demonstrates that the parameters can be recovered
-from fitting the model to only the degree distribution.
+from fitting the model to only the degree distribution and as a
+consequence recover the underlying growth mechanism.
 Section~\ref{sec-real} fits the model to some real data and provides
@@
 
-Special cases of this model include the BA model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution that is not regularly varying.
+Special cases of this model include the BA model when $b(k) = k+1$, which in the limit of $t\rightarrow \infty$ leads to a power-law degree distribution with tail index 2, and the Uniform Attachment (UA) model where $b(k)=c$ leading to a degree distribution with a tail that is not regularly varying.
 
@@
 
-Now, let \(\Upsilon(t)\) be the tree determined by \(\zeta(t)\) as
+Now, we construct the tree \(\Upsilon(t)\) determined by \(\zeta(t)\) as
 follows: \(\Upsilon(0)=\{\emptyset\}\) and \(\Upsilon(t)\) is a graph
@@
 
-Theorem 1 from \citet{rudas07} states that for the tree \(\Upsilon(t)\)
-at time \(t\) and a characteristic function of the tree
-\(\varphi(\cdot)\) :
+The limiting survival can be viewed as the limit over time of the
+empirical proportion of vertices with degree over a threshold
+\(k\in\mathbb N\), that is:
+
+\[
+\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1},
+\]
+
+For the tree \(\Upsilon(t)\) defined as above, Theorem 1 from
+\citet{rudas07} states that
 
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
-}\end{equation}
-
-The limiting survival can be viewed as the limit of the empirical
-proportion of vertices with degree over a threshold \(k\in\mathbb N\),
-that is:
-
-\[
-\bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1},
-\] which can also be written using Equation~\ref{eq-survlim} as:
+}\end{equation} for some characteristic function of the tree
+\(\varphi(\cdot)\).
+
+Using this result, and the characteristic function
+\(\varphi(\Upsilon(t)) = \mathbb I\{\text{deg}(x,\Upsilon(t))>k\}\), we
+can write the limiting survival as:
 
@@
 
-We investigate how the tail behaviour of the discrete limiting degree
-distribution is influenced by the preference function \(b(·)\), using
-the characterisation by \citet{shimura12}.
-
-A central tool is the quantity \(\Omega(F,k)\). For a discrete
-distribution \(F\) with survival function \(\bar F\) and
+We now investigate how the tail behaviour of the discrete limiting
+degree distribution is influenced by the preference function \(b(·)\),
+using the characterisation by \citet{shimura12}.
+
+A central tool of this characterisation is the quantity \(\Omega(F,k)\).
+For a discrete distribution \(F\) with survival function \(\bar F\) and
 \(k\in\mathbb Z^+\) define:
@@
 then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\), and
-hence belongs to the Fréchet MDA. If instead,
+hence belongs to the Fréchet MDA with tail index \(\alpha\). If instead,
 \(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then the distribution is
@@
 
-If \(\bar F(k) = \prod_{i=0}^k\frac{b(i)}{\lambda^* + b(i)}\) with
-\(b(k) \rightarrow \infty\) as \(k\rightarrow \infty\), then \[
+Let \(\bar F (k)\) be the limiting survival function of degrees in a GPA
+network with preference function \(b\) such that
+\(b(k)\rightarrow\infty\). Then: \[
 \lim_{k\rightarrow\infty}\Omega(F,k) = \lim_{k\rightarrow\infty}\frac{b(k+1)-b(k)}{\lambda^*}.
@@
 \(\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}\).
+
+Shown in Figure~\ref{fig-ex} is an example of one such preference
+function of this form, the value of \(\varepsilon\) controls the
+baseline appeal of a node with zero degree and \(k_0\) controls the
+point at which the behaviour changes, below this point the curvature of
+the function is controlled by the value of \(alpha\) and above \(k_0\)
+the gradient of the linear section is contolled by \(\beta\).
 
@@
 \begin{cases}
-=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k<k_0,\\
-\approx \left(\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon}\right) \left(\frac{\beta(k+1-k_0)}{k_0^{\alpha}+\varepsilon} + 1\right)^{-\lambda^*/\beta},&k\ge k_0,
+=\prod_{i=0}^{k}\frac{i^\alpha + \varepsilon}{\lambda^*+i^\alpha + \varepsilon},&k\le k_0,\\
+\approx \left(\prod_{i=0}^{k_0}\frac{i^\alpha + \varepsilon}{\lambda^* + i^\alpha + \varepsilon}\right)\left(\frac{\beta(k-k_0)}{k_0^{\alpha}+\varepsilon+\beta} + 1\right)^{-\lambda^{*}/\beta},&k > k_0,
 \end{cases}
@@
 distribution (for large \(k_0^\alpha\)) is approximated by the
-\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon}{\lambda^*},k_0-1\right)\)
+\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon+\beta}{\lambda^*},k_0-1\right)\)
 distribution.
@@
 the degree distribution, we would learn about the shape of the
-preference function, thus gaining a direct understanding the mechanisms
-underlying the growth of the network.
+preference function, thus gaining a direct understanding of the
+mechanisms underlying the growth of the network.
 
@@
 
-\subsection{\texorpdfstring{Proof of
-Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{proof-of-prp-omega}
+\subsection{Preliminary}\label{preliminary}
 
@@
 
+\subsection{\texorpdfstring{Proof of
+Remark~\ref{rem-omega}}{Proof of Remark~}}\label{proof-of-rem-omega}
+
 As it is assumed throughout that \(b(k)\) is non-decreasing, we consider
@@
 \(\lim_{k\rightarrow\infty}b(k) = c\) for some \(c>0\), which implies
-that
-\(\lim_{k\rightarrow\infty}\Omega(F,k)=\lim_{k\rightarrow\infty}\frac{1}{\lambda^{*}}\left[b(k+2)-b(k+1)\right] = 0\).
-This proves the first part of Remark~\ref{rem-omega}. Now, using
+that \(\lim_{k\rightarrow\infty}\Omega(F,k)= 0\). This proves the first
+part of Remark~\ref{rem-omega}. Now, using
 Equation~\ref{eq-surv-origin},
@@
 therefore is only recovered to the Gumbel MDA.
+
+\subsection{\texorpdfstring{Proof of
+Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{proof-of-prp-omega}
 

### 2025-10-03 Files/fixed_paper.tex
Update .gitignore
Updating pdf

--- 2025-09-30 Files/paper.tex
@@
 \usepackage{amsmath,amssymb,amsfonts}%
-\usepackage{amsthm}%
+\usepackage{amsthm}
+\theoremstyle{thmstyleone}
+\newtheorem{proposition}{Proposition}[section]
+\theoremstyle{thmstyleone}
+\newtheorem{corollary}[proposition]{Corollary}
+\theoremstyle{remark}
+\AtBeginDocument{\renewcommand*{\proofname}{Proof}}
+\newtheorem{remark}[proposition]{Remark}
+\newtheorem*{solution}{Solution}
+\newtheorem{refremark}[proposition]{Remark}
+\newtheorem{refsolution}{Solution}[section]
+%
 \usepackage{mathrsfs}%
@@
 \theoremstyle{plain}
-\newtheorem{corollary}{Corollary}[section]
 \theoremstyle{plain}
-\newtheorem{proposition}{Proposition}[section]
 \theoremstyle{remark}
 \AtBeginDocument{\renewcommand*{\proofname}{Proof}}
-\newtheorem*{remark}{Remark}
-\newtheorem*{solution}{Solution}
-\newtheorem{refremark}{Remark}[section]
-\newtheorem{refsolution}{Solution}[section]
 \makeatother
@@
 \(\lim_{k\rightarrow\infty}\Omega(F,k) = \frac{\beta}{\lambda^*}\).
+
+Shown in Figure~\ref{fig-ex} is an example of one such preference
+function of this form, the value of \(\varepsilon\) controls the
+baseline appeal of a node with zero degree and \(k_0\) controls the
+point at which the behaviour changes, below this point the curvature of
+the function is controlled by the value of \(\alpha\) and above \(k_0\)
+the gradient of the linear section is contolled by \(\beta\).
 

### 2025-10-03 Files/paper.tex
Updating pdf

@@
 point at which the behaviour changes, below this point the curvature of
-the function is controlled by the value of \(alpha\) and above \(k_0\)
+the function is controlled by the value of \(\alpha\) and above \(k_0\)
 the gradient of the linear section is contolled by \(\beta\).

### 2025-10-03 Files/render.py

@@ -31,7 +31,7 @@ def build_paper():
     # List of commands
     commands = [
     ["quarto", "render", "paper.qmd"],  # stop at .tex
-    ["python", "fix_theorems.py", "paper.tex"],
+    ["python3", "fix_theorems.py", "paper.tex"],
     ["xelatex","-interaction=nonstopmode", "fixed_paper.tex"],
     ["bibtex", "fixed_paper"],
     ["xelatex", "-interaction=nonstopmode", "fixed_paper.tex"]

### 2025-10-03 Files/fixed_paper.tex
Fixing small issues

@@
 distribution follows a power law, that is the fraction of nodes with
-degree greater than \(k\) is proportional to \(k^{-\alpha}\), and
-therefore has a regularly varying tail with tail index \(\alpha\). On
-the side against the claim is \citet{Broido_2019} who compared the fits
-of a power law model against that of several non-scale-free models to
-nearly a thousand networks, only to find strong evidence for
-scale-freeness in four percent and weak evidence in over half of the
-networks, thus claiming that scale-free networks do not make up a
-majority in real networks. \citet{Voitalov_2019} disagree and claim that
-these networks are not nearly as rare and only appear so as a result of
-an unrealistic expectation of a power law without deviations or noise.
-Additional evidence of these deviations from a power law is shown by
-\citet{Lee24} who demonstrate that a lot of networks are partially
-scale-free, in that the body of the degree distribution is often
-modelled well by a power law, while the tail is lighter than what is
-implied by the body, albeit still regularly varying.
+degree greater than \(k\) is eventually proportional to \(k^{-\alpha}\),
+and therefore has a regularly varying tail with tail index \(\alpha\).
+On the side against the claim, that most real networks are scale-free,
+is \citet{Broido_2019} who compared the fits of a power law model
+against that of several non-scale-free models to nearly a thousand
+networks, only to find strong evidence for scale-freeness in four
+percent and weak evidence in over half of the networks, thus claiming
+that scale-free networks do not make up a majority of real networks.
+\citet{Voitalov_2019} disagree and claim that these networks are not
+nearly as rare and only appear so as a result of an unrealistic
+expectation of a power law without deviations or noise. Additional
+evidence of these deviations from a power law is shown by \citet{Lee24}
+who demonstrate that a lot of networks are partially scale-free, in that
+the body of the degree distribution is often modelled well by a power
+law, while the tail is lighter than what is implied by the body, albeit
+still regularly varying.
 
@@
 generalisations --- \citet{krapivsky01} considered
-\(b(k) = (k+1)^\alpha\), and showed that the degree distribution is not
-regularly varying (and therefore not following the power law) when
-\(0<\alpha<1\), and when \(\alpha>1\) a finite number of nodes end up
-with all edges after a certain point resulting in a degenerate degree
-distribution. \citet{wang2022random} returns to a linear preference
-function of the form \(b(k) = k+\varepsilon\) but adds the possibility
-for reciprocal edges to be sent, resulting in the joint distribution of
-in-degree and out-degree being multivariate regularly varying and having
-the property of hidden regular variation. \citet{rudas07} followed in
-the footsteps of \citet{krapivsky01}, by considering a PA tree and using
-theory from continuous branching processes, deriving a limiting degree
-distribution in terms of the preference function \(b(\cdot)\).
-Nevertheless, research in this area tends to only focus on the
-theoretical asymptotic results of network growth models with little
-analysis of real networks.
+\(b(k) = (k+1)^\alpha\), and showed that the tail of the degree
+distribution is not regularly varying (and therefore not following the
+power law) when \(0<\alpha<1\), and when \(\alpha>1\) a finite number of
+nodes end up with all edges after a certain point resulting in a
+degenerate degree distribution. \citet{wang2022random} return to a
+linear preference function of the form \(b(k) = k+\varepsilon\) but add
+the possibility for reciprocal edges to be sent, resulting in the joint
+distribution of in-degree and out-degree being multivariate regularly
+varying and having the property of hidden regular variation.
+\citet{rudas07} followed in the footsteps of \citet{krapivsky01}, by
+considering a GPA tree and using theory from continuous branching
+processes, deriving a limiting degree distribution in terms of the
+preference function \(b(\cdot)\). Unfortunately, research in this area
+tends to only focus on the theoretical asymptotic results of network
+growth models with little analysis of real networks.
 
@@
 \citet{shimura12} showed that a discrete distribution \(F\) must be
-long-tailed,that is
+long-tailed, that is
 \(\lim_{x\rightarrow\infty}\bar F(x+1)/\bar F(x) = 1\), in order to be
@@
 \bar F(k) = \lim_{t\rightarrow\infty}\frac{\sum_{x\in\Upsilon(t)}\mathbb I\left\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\right\}}{\sum_{x\in\Upsilon(t)} 1},
-\]
-
-For the tree \(\Upsilon(t)\) defined as above, Theorem 1 from
+\] for the tree \(\Upsilon(t)\) defined as above, Theorem 1 from
 \citet{rudas07} states that
@@
 \lim_{t\rightarrow\infty}\frac{1}{|\Upsilon(t)|}\sum_{x\in\Upsilon(t)}\varphi(\Upsilon(t)_{\downarrow x}) = \lambda^* \int_0^\infty e^{-\lambda^* t}\mathbb E\left[\varphi(\Upsilon(t))\right]\text{d}t.
-}\end{equation} for some characteristic function of the tree
-\(\varphi(\cdot)\).
+}\end{equation}
 
 Using this result, and the characteristic function
-\(\varphi(\Upsilon(t)) = \mathbb I\{\text{deg}(x,\Upsilon(t))>k\}\), we
-can write the limiting survival as:
+\(\varphi(\Upsilon(t)_{\downarrow x}) = \mathbb I\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\}\),
+we can write the limiting survival as:
 
@@
 \end{cases}
-}\end{equation}
-
-with Malthusian parameter \(\lambda^*\) satisfying
+}\end{equation} with Malthusian parameter \(\lambda^*\) satisfying
 \(\hat \rho(\lambda^*)=1\) where
@@
 \begin{equation}\phantomsection\label{eq-rho}{
-\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}
-}\end{equation}
-
-which has to be solved numerically for most parameter choices. Also,
-note that \(\lambda>\beta\). See Appendix \ref{sec-proofs} for the
-derivations of Equation~\ref{eq-rho}.
+\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon},
+}\end{equation} which has to be solved numerically for most parameter
+choices. Also, note that \(\lambda>\beta\). See Appendix
+\ref{sec-proofs} for the derivations of Equation~\ref{eq-rho}.
 
@@
 \end{cases}
-}\end{equation} meaning that for \(k\ge k_0\) the limiting degree
+}\end{equation} meaning that for \(k> k_0\) the limiting degree
 distribution (for large \(k_0^\alpha\)) is approximated by the
-\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon+\beta}{\lambda^*},k_0-1\right)\)
+\(\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon+\beta}{\lambda^*},k_0\right)\)
 distribution.
@@
 b(k) = \begin{cases}
-g(k),&k<k_0,\\
-g(k_0) + \beta(k-k_0), &k\ge k_0,
+g(k),&k\le k_0,\\
+g(k_0) + \beta(k-k_0), &k> k_0,
 \end{cases}
@@
 &=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-1}\frac{1}{1-t}\text{d}t\\
-&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}at\\
+&=\frac{1}{\Gamma(y)}\int_0^1 t^{x-1}(1-t)^{y-2}\text{d}t\\
 &=\frac{1}{\Gamma(y)}\text{B}(x,y-1)\\

### 2025-10-03 Files/paper.qmd
Updating pdf
Fixing small issues

@@
 keep-tex: true
+pdf-engine: xelatex
 ---
@@
 Amid the recent rise of interest in networks, there has been a debate on whether most real networks are scale-free.
-Claiming a real network is scale-free is equivalent to saying that the tail of the degree distribution follows a power law, that is the fraction of nodes with degree greater than $k$ is proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with tail index $\alpha$.
-On the side against the claim is @Broido_2019 who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority in real networks.
+Claiming a real network is scale-free is equivalent to saying that the tail of the degree distribution follows a power law, that is the fraction of nodes with degree greater than $k$ is eventually proportional to $k^{-\alpha}$, and therefore has a regularly varying tail with tail index $\alpha$.
+On the side against the claim, that most real networks are scale-free, is @Broido_2019 who compared the fits of a power law model against that of several non-scale-free models to nearly a thousand networks, only to find strong evidence for scale-freeness in four percent and weak evidence in over half of the networks, thus claiming that scale-free networks do not make up a majority of real networks.
 @Voitalov_2019 disagree and claim that these networks are not nearly as rare and only appear so as a result of an unrealistic expectation of a power law without deviations or noise.
@@
 
-The model from @Barabasi99 provided the foundations for various generalisations — @krapivsky01 considered $b(k) = (k+1)^\alpha$, and showed that the degree distribution is not regularly varying (and therefore not following the power law) when $0<\alpha<1$, and when $\alpha>1$ a finite number of nodes end up with all edges after a certain point resulting in a degenerate degree distribution.
-@wang2022random returns to a linear preference function of the form $b(k) = k+\varepsilon$ but adds the possibility for reciprocal edges to be sent, resulting in the joint distribution of in-degree and out-degree being multivariate regularly varying and having the property of hidden regular variation.
-@rudas07 followed in the footsteps of @krapivsky01, by considering a PA tree and using theory from continuous branching processes, deriving a limiting degree distribution in terms of the preference function $b(\cdot)$.
-Nevertheless, research in this area tends to only focus on the theoretical asymptotic results of network growth models with little analysis of real networks.
+The model from @Barabasi99 provided the foundations for various generalisations — @krapivsky01 considered $b(k) = (k+1)^\alpha$, and showed that the tail of the degree distribution is not regularly varying (and therefore not following the power law) when $0<\alpha<1$, and when $\alpha>1$ a finite number of nodes end up with all edges after a certain point resulting in a degenerate degree distribution.
+@wang2022random return to a linear preference function of the form $b(k) = k+\varepsilon$ but add the possibility for reciprocal edges to be sent, resulting in the joint distribution of in-degree and out-degree being multivariate regularly varying and having the property of hidden regular variation.
+@rudas07 followed in the footsteps of @krapivsky01, by considering a GPA tree and using theory from continuous branching processes, deriving a limiting degree distribution in terms of the preference function $b(\cdot)$.
+Unfortunately, research in this area tends to only focus on the theoretical asymptotic results of network growth models with little analysis of real networks.
 
@@
 A key difficulty arises because many standard discrete distributions, such as the Poisson, geometric, or negative binomial, do not satisfy the conditions required to belong to a maximum domain of attraction (MDA).
-@shimura12 showed that a discrete distribution $F$ must be long-tailed,that is $\lim_{x\rightarrow\infty}\bar F(x+1)/\bar F(x) = 1$, in order to be in an MDA, and introduced the notion of recoverability to the MDA, also sometimes referred to as the discrete MDA [@hitz24].
+@shimura12 showed that a discrete distribution $F$ must be long-tailed, that is $\lim_{x\rightarrow\infty}\bar F(x+1)/\bar F(x) = 1$, in order to be in an MDA, and introduced the notion of recoverability to the MDA, also sometimes referred to as the discrete MDA [@hitz24].
 Using the theoretical guarantees of being in a discrete MDA for a discrete distribution by @shimura12, we demonstrate how the tail of the degree distribution is directly influenced by $b(\cdot)$, and subsequently propose a class of preference functions that not only imply regularly varying degree distributions but are also tail realistic for real networks.
@@
 $$
-
-For the tree $\Upsilon(t)$ defined as above, Theorem 1 from @rudas07 states that
+for the tree $\Upsilon(t)$ defined as above, Theorem 1 from @rudas07 states that
 
@@
 $$ {#eq-survlim}
-for some characteristic function of the tree $\varphi(\cdot)$.
 
-Using this result, and the characteristic function $\varphi(\Upsilon(t)) = \mathbb I\{\text{deg}(x,\Upsilon(t))>k\}$, we can write the limiting survival as:
+Using this result, and the characteristic function $\varphi(\Upsilon(t)_{\downarrow x}) = \mathbb I\{\text{deg}(x,\Upsilon(t)_{\downarrow x})>k\}$, we can write the limiting survival as:
 
@@
 $$ {#eq-polysurv}
-
 with Malthusian parameter $\lambda^*$ satisfying $\hat \rho(\lambda^*)=1$ where
@@
 $$
-\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon}
+\hat\rho(\lambda) = \sum_{n=0}^{k_0}\prod_{i=0}^{n-1}\frac{i^\alpha + \varepsilon}{\lambda+i^\alpha + \varepsilon} + \left(\frac{k_0^\alpha + \varepsilon}{\lambda-\beta}\right)\prod_{i=0}^{k_0-1}\frac{i^\alpha + \varepsilon}{\lambda + i^\alpha + \varepsilon},
 $$ {#eq-rho}
-
 which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$. See Appendix [-@sec-proofs] for the derivations of @eq-rho.
@@
 \end{cases}
-$$ {#eq-igp-est} meaning that for $k\ge k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon+\beta}{\lambda^*},k_0-1\right)$ distribution.
+$$ {#eq-igp-est} meaning that for $k> k_0$ the limiting degree distribution (for large $k_0^\alpha$) is approximated by the $\text{IGP}\left(\frac{\beta}{\lambda^*}, \frac{k_0^\alpha + \varepsilon+\beta}{\lambda^*},k_0\right)$ distribution.
 

### 2025-10-03 Files/fixed_paper.tex
Updating title page

@@
 
-\pacs[\sffamily MSC 2020]{05C82· 60G70}
+\pacs[\sffamily AMS 2000 Subject Classification]{05C80· 60J85· 62G32}
 
@@
 
-While \citet{Voitalov_2019} have pioneered using methods in extreme
-value theory to analyse degree distributions, while their approach is
-good they overlooked the inapplicability of standard tools for
-continuous extremes, due to the discrete nature of degrees. A key
-difficulty arises because many standard discrete distributions, such as
-the Poisson, geometric, or negative binomial, do not satisfy the
-conditions required to belong to a maximum domain of attraction (MDA).
-\citet{shimura12} showed that a discrete distribution \(F\) must be
-long-tailed, that is
+\citet{Voitalov_2019} have pioneered using methods in extreme value
+theory to analyse degree distributions and while their approach is good
+they overlooked the inapplicability of standard tools for continuous
+extremes, due to the discrete nature of degrees. A key difficulty arises
+because many standard discrete distributions, such as the Poisson,
+geometric, or negative binomial, do not satisfy the conditions required
+to belong to a maximum domain of attraction (MDA). \citet{shimura12}
+showed that a discrete distribution \(F\) must be long-tailed, that is
 \(\lim_{x\rightarrow\infty}\bar F(x+1)/\bar F(x) = 1\), in order to be
@@
 \begin{aligned}
-L(\pmb n | \pmb \theta,l) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i}
+L(\pmb n | \pmb \theta,l) = &\left(\frac{\lambda^*}{\lambda^*+\varepsilon}\right)^{n_0}\left(\prod_{j=l}^{k_0-1}\frac{j^\alpha +\varepsilon}{\lambda^* + j^\alpha +\varepsilon}\right)^{\left(\sum_{i\ge k_0}n_{i}\right)} \\ &\times \prod_{l \le i<k_0}\left(\frac{\lambda^*}{\lambda^* +i^\alpha + \varepsilon } \prod_{j=l}^{k_0-1}\frac{j^\alpha + \varepsilon}{\lambda^* + j^\alpha + \varepsilon}\right)^{n_i}\\ &\times \prod_{i\ge k_0}\left(\frac{\text{B}(i-k_0 + (k_0^\alpha + \varepsilon)/\beta,1+\lambda^*/\beta)}{\text{B}((k_0^\alpha + \varepsilon)/\beta,\lambda^*/\beta)}\right)^{n_i},
 \end{aligned}

### 2025-10-03 Files/paper.qmd
Updating title page

@@
 msc: 
- - 05C82
- - 60G70
+ - 05C80
+ - 60J85
+ - 62G32
 bibliography: refs.bib  
@@
 
-While @Voitalov_2019 have pioneered using methods in extreme value theory to analyse degree distributions, while their approach is good they overlooked the inapplicability of standard tools for continuous extremes, due to the discrete nature of degrees.
+@Voitalov_2019 have pioneered using methods in extreme value theory to analyse degree distributions and while their approach is good they overlooked the inapplicability of standard tools for continuous extremes, due to the discrete nature of degrees.
 A key difficulty arises because many standard discrete distributions, such as the Poisson, geometric, or negative binomial, do not satisfy the conditions required to belong to a maximum domain of attraction (MDA).

### 2025-10-03 Files/fixed_paper.tex
Final changes for preprint

@@
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
-  pdfkeywords={Extremes · Discrete extremes · Power law · Preferential
-attachment},
+  pdfkeywords={Network degrees · Discrete extremes · Power law ·
+Preferential attachment},
   colorlinks=true,
@@
 % keywords
-\keywords{Extremes · Discrete extremes · Power law · Preferential
+\keywords{Network degrees · Discrete extremes · Power law · Preferential
 attachment}
@@
 Markovian pure birth process, with \(\zeta(0)=0\) and birth rates
-depending on a non-negative function \(b(\cdot)\):
-
-\[
+depending on a non-negative function \(b(\cdot)\): \[
 \Pr(\zeta(t+\text{d}t)=k+1|\zeta(t)=k) = b(k)\text{d}t + o(\text{d}t).
-\]
-
-Denote the density of the point process associated with the pure birth
-process corresponding to the growth of an individual node by
+\] Denote the density of the point process associated with the pure
+birth process corresponding to the growth of an individual node by
 \(\rho(t)\), and its Laplace transform by
@@
 Here, \(\lambda^*\) is the Malthusian parameter of the corresponding
-branching process. See Appendix \ref{sec-proofs} for the details of the
-proof.
+branching process. See Appendix \ref{sec-prpproof} for the details of
+the proof.
 
@@
 the Gumbel MDA itself as it is not long-tailed. See Appendix
-\ref{sec-proofs} for the proof.
+\ref{sec-remproof} for the proof.
 
@@
 \(b(k)=k+1\), \(\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1\), giving a
-tail index of \(\lambda^*=2\).
+tail index of \(\lambda^*=2\). We can also obtain the afforementioned
+results for UA, taking \(b(k)=c\) it is clear that
+\(\lim_{k\rightarrow\infty}b(k)<\infty\) and therefore by
+Remark~\ref{rem-omega} we have tha
+\(\lim_{k\rightarrow\infty}\Omega(F,k)=0\) meaning \(F\)'s tail is not
+regularly varying and is in the Gumbel domain as \(b(k)\) is also
+long-tailed.
 
@@
 choices. Also, note that \(\lambda>\beta\). See Appendix
-\ref{sec-proofs} for the derivations of Equation~\ref{eq-rho}.
+\ref{sec-rhoproof} for the derivations of Equation~\ref{eq-rho}.
 
@@
 
-In order to carry out the Bayesian inference of the parameters we must
-first assign priors, for the parameters defined on the positive real
-line (\(\alpha, \beta,\varepsilon\)) we use diffuse gamma priors and for
-\(k_0\) we use a discrete uniform prior over a large and reasonable
-range of values.
+The methodology for the Bayesian inference of the parameters is detailed
+in Section~\ref{sec-sim}.
 
@@
 some set of parameters
-\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\), obtaining the degree
-counts and using the likelihood from the previous section alongside the
-priors:
+\(\pmb\theta = (\alpha, \beta, \varepsilon, k_0)\). In order to carry
+out the Bayesian inference of the parameters we must first assign
+priors, for the parameters defined on the positive real line
+(\(\alpha, \beta,\varepsilon\)) we use diffuse gamma priors and for
+\(k_0\) we use a discrete uniform prior over a large and reasonable
+range of values, specifically we use the priors below:
 
@@
 \end{itemize}
+
+The parameter inferences that follow were obtained using the same priors
+and MCMC sampling method as detailed in Section~\ref{sec-sim}.
 
@@
 \subsection{\texorpdfstring{Proof of
-Remark~\ref{rem-omega}}{Proof of Remark~}}\label{proof-of-rem-omega}
+Remark~\ref{rem-omega}}{Proof of Remark~}}\label{sec-remproof}
 
@@
 \(\lim_{k\rightarrow\infty}b(k) = c\) for some \(c>0\), which implies
-that \(\lim_{k\rightarrow\infty}\Omega(F,k)= 0\). This proves the first
-part of Remark~\ref{rem-omega}. Now, using
+that
+\(\lim_{k\rightarrow\infty}\Omega(F,k)=\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}= 0\).
+This proves the first part of Remark~\ref{rem-omega}. Now, using
 Equation~\ref{eq-surv-origin},
@@
 \subsection{\texorpdfstring{Proof of
-Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{proof-of-prp-omega}
+Proposition~\ref{prp-omega}}{Proof of Proposition~}}\label{sec-prpproof}
 
@@
 \subsection{\texorpdfstring{Derivation of
-Equation~\ref{eq-rho}}{Derivation of Equation~}}\label{derivation-of-eq-rho}
+Equation~\ref{eq-rho}}{Derivation of Equation~}}\label{sec-rhoproof}
 

### 2025-10-03 Files/paper.qmd
Final changes for preprint

@@
 Results of applications to real data are promising and comparable to alternatives.
-keywords: Extremes · Discrete extremes · Power law · Preferential attachment
+keywords: Network degrees · Discrete extremes · Power law · Preferential attachment
 msc: 
@@
 Consider a continuous time branching process $\zeta(t)$ driven by a Markovian pure birth process, with $\zeta(0)=0$ and birth rates depending on a non-negative function $b(\cdot)$:
-
 $$
@@
 $$
-
 Denote the density of the point process associated with the pure birth process corresponding to the growth of an individual node by $\rho(t)$, and its Laplace transform by $\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t$.
@@
 
-Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process. See Appendix [-@sec-proofs] for the details of the proof.
+Here, $\lambda^*$ is the Malthusian parameter of the corresponding branching process. See Appendix [-@sec-prpproof] for the details of the proof.
 :::
@@
 Furthermore, $F$ is recovered to the Gumbel MDA but does not belong to the Gumbel MDA itself as it is not long-tailed.
-See Appendix [-@sec-proofs] for the proof.
+See Appendix [-@sec-remproof] for the proof.
 :::
@@
 Conversely, it also agrees with the limiting behaviour of the BA model, which produces a regularly degree distribution with tail index 2; for the preference function $b(k)=k+1$, $\lim_{k\rightarrow\infty}[b(k+1)-b(k)]=1$, giving a tail index of $\lambda^*=2$.
+We can also obtain the afforementioned results for UA, taking $b(k)=c$ it is clear that $\lim_{k\rightarrow\infty}b(k)<\infty$ and therefore by @rem-omega we have tha $\lim_{k\rightarrow\infty}\Omega(F,k)=0$ meaning $F$'s tail is not regularly varying and is in the Gumbel domain as $b(k)$ is also long-tailed.
 
@@
 #| fig-width: 6
-#| fig-asp: 0.7
+#| fig-asp: 0.55
 #| fig-cap: Example construction of the preference function in @eq-pref
@@
 $$ {#eq-rho}
-which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$. See Appendix [-@sec-proofs] for the derivations of @eq-rho.
+which has to be solved numerically for most parameter choices. Also, note that $\lambda>\beta$. See Appendix [-@sec-rhoproof] for the derivations of @eq-rho.
 
@@
 
-In order to carry out the Bayesian inference of the parameters we must first assign priors, for the parameters defined on the positive real line ($\alpha, \beta,\varepsilon$) we use diffuse gamma priors and for $k_0$ we use a discrete uniform prior over a large and reasonable range of values.
+The methodology for the Bayesian inference of the parameters is detailed in @sec-sim.
 
@@
 
-The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$, obtaining the degree counts and using the likelihood from the previous section alongside the priors:
+The procedure for recovering the parameters begins with simulating a network from the model with $N=100,000$ vertices and $m=1$ given some set of parameters $\pmb\theta = (\alpha, \beta, \varepsilon, k_0)$.
+In order to carry out the Bayesian inference of the parameters we must first assign priors, for the parameters defined on the positive real line ($\alpha, \beta,\varepsilon$) we use diffuse gamma priors and for $k_0$ we use a discrete uniform prior over a large and reasonable range of values, specifically we use the priors below:
 
@@
 -   `topology`: network of connections between autonomous systems of the Internet
+
+The parameter inferences that follow were obtained using the same priors and MCMC sampling method as detailed in @sec-sim.
 
@@
 
-## Proof of @rem-omega
+## Proof of @rem-omega {#sec-remproof}
 
 As it is assumed throughout that $b(k)$ is non-decreasing, we consider the two cases where $b(k)$ is bounded and unbounded.
-When $b(k)$ is bounded, by the monotone convergence theorem, $\lim_{k\rightarrow\infty}b(k) = c$ for some $c>0$, which implies that $\lim_{k\rightarrow\infty}\Omega(F,k)= 0$.
+When $b(k)$ is bounded, by the monotone convergence theorem, $\lim_{k\rightarrow\infty}b(k) = c$ for some $c>0$, which implies that $\lim_{k\rightarrow\infty}\Omega(F,k)=\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}-\left(\log\left[1+\frac{\lambda^{*}}{c}\right]\right)^{-1}= 0$.
 This proves the first part of @rem-omega.
@@
 
-## Proof of @prp-omega
+## Proof of @prp-omega {#sec-prpproof}
 
@@
 
-## Derivation of @eq-rho
+## Derivation of @eq-rho {#sec-rhoproof}
 

### 2025-10-03 Files/fixed_paper.tex
Changing to pdflatex

@@
 \usepackage{setspace}%
+\usepackage{mathtools}%
 
@@
   pdftitle={Learning growth mechanisms of tail realistic preferential attachment models from network degree distributions},
-  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramírez},
+  pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
   pdfkeywords={Network degrees · Discrete extremes · Power law ·
@@
 % author setup
-\author*[1]{\fnm{Thomas William} \sur{Boughen}}\email{t.w.boughen1@newcastle.ac.uk}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramírez}}
+\author*[1]{\fnm{Thomas William} \sur{Boughen}}\email{t.w.boughen1@newcastle.ac.uk}\author[1]{\fnm{Clement} \sur{Lee}}\author[1]{\fnm{Vianey Palacios} \sur{Ramirez}}
 % affil setup
@@
 In the general model, general preferential attachment (GPA), as new nodes join the network, an existing node with degree $k$ gains edges at a rate proportional to $b(k)$, where $b(\cdot)$ is a non-negative preference function.
-\cite{Barabasi99} showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 2 — this model is commonly known as the Barab\'asi-Albert (BA) model.
+\cite{Barabasi99} showed that, when $b(k) = k + 1$, in the limit the resulting degree distribution is regularly varying with index 2 --- this model is commonly known as the Barab\'asi-Albert (BA) model.
 Subsequently, if a real network is shown to be scale-free, one can loosely justify PA as the underlying mechanism of its growth.
@@
 \(\rho(t)\), and its Laplace transform by
-\(\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t\).
+\(\hat \rho(\lambda)\coloneqq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t\).
 We denote the Malthusian parameter of this process by \(\lambda^*\),
@@
 
-\citet{shimura12} established that if
-\(\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha\) (\(\alpha>0\)),
-then \(F\) is regularly varying with \(\bar F(k) \sim k^{-\alpha}\), and
-hence belongs to the Fréchet MDA with tail index \(\alpha\). If instead,
-\(\lim_{k\rightarrow\infty} \Omega(F,k) = 0\) then the distribution is
-recoverable to the Gumbel MDA; in this cases we say that the
-distribution is light-tailed. Moreover, if the distribution is also
-long-tailed, we can conclude that \(F\) itself lies in the Gumbel MDA.
+@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Fr\'echet MDA with tail index $\alpha$.
+If instead, $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is recoverable to the Gumbel MDA; in this cases we say that the distribution is light-tailed.
+Moreover, if the distribution is also long-tailed, we can conclude that $F$ itself lies in the Gumbel MDA.
 
@@
 \item
-  \texttt{opsahl−openflights}: network containing flights between
+  \texttt{opsahl-openflights}: network containing flights between
   airports of the world.
 \item
-  \texttt{pajek-erdos}: co-authorship network around Paul Erdős
+  \texttt{pajek-erdos}:
+  \(\text{co-authorship network around Paul Erd\H{o}s}\)
 \item

### 2025-10-03 Files/paper.qmd
Changing to pdflatex

@@
       - ref: 1
-  - name: Vianey Palacios Ramírez
+  - name: Vianey Palacios Ramirez
     affiliations:
@@
 keep-tex: true
-pdf-engine: xelatex
+pdf-engine: pdflatex
 ---
@@
 $$
-Denote the density of the point process associated with the pure birth process corresponding to the growth of an individual node by $\rho(t)$, and its Laplace transform by $\hat \rho(\lambda)\coloneq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t$.
+Denote the density of the point process associated with the pure birth process corresponding to the growth of an individual node by $\rho(t)$, and its Laplace transform by $\hat \rho(\lambda)\coloneqq \int_0^\infty e^{-\lambda t}\rho(t)\text{d}t$.
 We denote the Malthusian parameter of this process by $\lambda^*$, that is $\lambda^*$ satisfies $\hat\rho(\lambda^*) = 1$.
@@
 
-@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Fréchet MDA with tail index $\alpha$.
+```{=tex}
+@shimura12 established that if $\lim_{k\rightarrow\infty} \Omega(F,k) = 1/\alpha$ ($\alpha>0$), then $F$ is regularly varying with $\bar F(k) \sim k^{-\alpha}$, and hence belongs to the Fr\'echet MDA with tail index $\alpha$.
 If instead, $\lim_{k\rightarrow\infty} \Omega(F,k) = 0$ then the distribution is recoverable to the Gumbel MDA; in this cases we say that the distribution is light-tailed.
 Moreover, if the distribution is also long-tailed, we can conclude that $F$ itself lies in the Gumbel MDA.
-
+```
 This framework offers a natural approach for analysing the tail behaviour of the limiting degree distribution.
@@
 -   `moreno-blogs-blogs`: network of URLs found on the first pages of individual blogs
--   `opsahl−openflights`: network containing flights between airports of the world.
--   `pajek-erdos`: co-authorship network around Paul Erdős
+-   `opsahl-openflights`: network containing flights between airports of the world.
+-   `pajek-erdos`: $\text{co-authorship network around Paul Erd\H{o}s}$
 -   `reactome`: network of protein–protein interactions in humans

### 2025-10-03 Files/render.py

@@ -32,9 +32,9 @@ def build_paper():
     commands = [
     ["quarto", "render", "paper.qmd"],  # stop at .tex
     ["python3", "fix_theorems.py", "paper.tex"],
-    ["xelatex","-interaction=nonstopmode", "fixed_paper.tex"],
+    ["pdflatex","-interaction=nonstopmode", "fixed_paper.tex"],
     ["bibtex", "fixed_paper"],
-    ["xelatex", "-interaction=nonstopmode", "fixed_paper.tex"]
+    ["pdflatex", "-interaction=nonstopmode", "fixed_paper.tex"]
 ]
 
     try:

### 2025-10-04 Files/fixed_paper.tex
Arxiv version

@@
   pdfauthor={Thomas William Boughen; Clement Lee; Vianey Palacios Ramirez},
-  pdfkeywords={Network degrees · Discrete extremes · Power law ·
+  pdfkeywords={Network degrees - Discrete extremes - Power law -
 Preferential attachment},
@@
 % keywords
-\keywords{Network degrees · Discrete extremes · Power law · Preferential
+\keywords{Network degrees - Discrete extremes - Power law - Preferential
 attachment}

### 2025-10-04 Files/paper.qmd
Arxiv version

@@
 Results of applications to real data are promising and comparable to alternatives.
-keywords: Network degrees · Discrete extremes · Power law · Preferential attachment
+keywords: Network degrees - Discrete extremes - Power law - Preferential attachment
 msc: 
